---
author: "Marc Elosua-Bayes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
params: 
    sample_id: "Default!"
    donor_id: "Default!"
title: "`r sprintf('%s T cells SC mapping', params$donor_id)`"
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message = FALSE, warning = FALSE, cache = FALSE,
                      cache.lazy = FALSE)
options(width = 1200)
```

## Introduction
In this script we are going to map T-cell subtypes onto the Visium slides.

## Libraries
```{r}
library(Seurat)
library(ggpubr)
library(cowplot)
library(dplyr)
library(ggplot2)
library(stringr)
library(readr)
library(SPOTlight)
```

## Setting parameters
Loading necessary paths and parameters
```{r}
set.seed(123)
source(here::here("misc/paths.R"))
source(here::here("utils/bin.R"))

dir.create(path = here::here(sprintf("%s/%s", decon_32, plt_dir)),
           showWarnings = FALSE,
           recursive = TRUE)

dir.create(path =here::here(sprintf("%s/%s", decon_32, robj_dir)),
           showWarnings = FALSE,
           recursive = TRUE)
```

Extract sample id and get Donor ID
```{r}
sample_id <- params$sample_id
# sample_id <- "esvq52_nluss5"
donor_id <- id_df[id_df$gem_id == sample_id, "BCLL_ID"]
```

## Load data
We have 4 different datasets that we are going to analyze separately.
The spatial data comes from the script *03-clustering/03-clustering.Rmd* while the sc data can be found in Ramon's scRNAseq analysis: **/scratch/devel/rmassoni/tonsil_atlas_private/2-DOWNSTREAM_PROCESSING/results/R_objects/processed_seurat_objects/processed_seurat_objects/tonsil_integrated_with_harmony_scrublet_annotated.rds**. We will also load the metadata of a l2 annotation of CD4 so that we can have the annotation. \
In each Rmarkdown notebook we analyze 1 dataset. \
```{r}
sp_obj <- readRDS(
  file = here::here(glue::glue("{clust_32}/{robj_dir}/clust_se_{sample_id}.RDS")))

sc_obj <- readRDS(
  file = here::here(glue::glue("{decon_32}/{robj_dir}/tonsil_integrated_with_harmony_scrublet_annotated.rds")))

metadata_cd4 <- readRDS(
  file = here::here(glue::glue("{decon_32}/{robj_dir}/CD4+_T_metadata.rds"))) %>%
  tibble::column_to_rownames("barcode")
```

## Analysis
We are going to use SPOTlight to deconvolute the spatial spots and map the cell types.

We will start by adding the labels to the clusters as Juan Nieto proposes
```{r}
metadata_cd4 <- metadata_cd4 %>%
  dplyr::mutate(
    seurat_clusters = as.character(seurat_clusters),
    l2_annotation = dplyr::case_when(
      seurat_clusters == 0 ~ "CD4 follicular",
      seurat_clusters == 1 ~ "CD4 CM",
      seurat_clusters == 2 ~ "CD4 naive",
      seurat_clusters == 3 ~ "CD4 EM/reg/h17",
      seurat_clusters == 4 ~ "CD8",
      seurat_clusters == 5 ~ "Doublet",
      seurat_clusters == 6 ~ "Doublet",
      seurat_clusters == 7 ~ "Follicular proliferating T cells",
      seurat_clusters == 8 ~ "Doublet",
      seurat_clusters == 9 ~ "Unkown-9",
      TRUE ~ seurat_clusters)
  ) %>%
  tibble::column_to_rownames("cell_barcode")

metadata_cd4 %>%
  dplyr::count(l2_annotation)
```

Next we subset the the CD4+ T cells and adding the relevant metadata...
```{r}
sc_obj <- sc_obj[, sc_obj$annotation_level1 == "CD4+ T"]

sc_obj@meta.data <- metadata_cd4

sc_obj <- sc_obj[, sc_obj$l2_annotation != "Doublet"]
```

Since the normalization was carried out with the entire dataset we are going to do normalization again within this subset
```{r}
sc_obj <- sc_obj %>%
  Seurat::NormalizeData(.,
                normalization.method = "LogNormalize",
                scale.factor = 10000,
                verbose = FALSE) %>%
  Seurat::FindVariableFeatures(.,
                               nfeatures = 3000,
                               verbose = FALSE) %>% 
  Seurat::ScaleData(.,
                    verbose = FALSE)
```

1st we need to determine the marker genes for each high-level cluster
```{r}
if (file.exists(here::here(glue::glue("{decon_32}/{robj_dir}/markers_cd4.RDS")))) {
  
  markers <- readRDS(file = here::here(glue::glue("{decon_32}/{robj_dir}/markers_cd4.RDS")))
  
} else {
  #### Extract marker genes from each cluster ####
  Seurat::Idents(object = sc_obj) <- sc_obj@meta.data$l2_annotation
  markers <- Seurat::FindAllMarkers(
    object = sc_obj, 
    assay = "RNA",
    slot = "data",
    verbose = TRUE, 
    only.pos = TRUE,
    max.cells.per.ident = 200)
  
  saveRDS(object = markers,
          file = here::here(glue::glue("{decon_32}/{robj_dir}/markers_cd4.RDS")))
}
```

2nd run deconvolution
```{r}
Seurat::Idents(object = sc_obj) <- as.character(sc_obj@meta.data$l2_annotation)
# sc_obj@meta.data <- sc_obj@meta.data %>% tibble::column_to_rownames("cell_barcode")

spotlight_ls <- spotlight_deconvolution(
  se_sc = sc_obj,
  counts_spatial = sp_obj@assays$Spatial@counts,
  clust_vr = "l2_annotation",
  cluster_markers = markers,
  cl_n = 75,
  hvg = 3000,
  ntop = NULL,
  transf = "uv",
  method = "nsNMF",
  min_cont = 0)
```

## Save data
```{r}
decon_fn <- here::here(glue::glue("{decon_32}/{robj_dir}/spotlight_ls_cd4_{sample_id}.rds"))
if (! file.exists(decon_fn) ) {
  saveRDS(
    object = spotlight_ls,
    file = here::here(glue::glue("{decon_32}/{robj_dir}/spotlight_ls_cd4_{sample_id}.rds")))
} else {
  spotlight_ls <- readRDS(file = here::here(glue::glue("{decon_32}/{robj_dir}/spotlight_ls_cd4_{sample_id}.rds")))
}

```

3rd Check Topic profiles
```{r fig.height=8, fig.width=8}
nmf_mod_ls <- spotlight_ls[[1]]
nmf_mod <- nmf_mod_ls[[1]]

h <- NMF::coef(nmf_mod)
rownames(h) <- paste("Topic", 1:nrow(h), sep = "_")
topic_profile_plts <- SPOTlight::dot_plot_profiles_fun(
  h = h,
  train_cell_clust = nmf_mod_ls[[2]])

topic_profile_plts[[2]] + theme(axis.text.x = element_text(angle = 90), 
                                axis.text = element_text(size = 12))
```



## Visualize deconvolution results
```{r}
decon_mtrx <- spotlight_ls[[2]]
decon_mtrx <- decon_mtrx[, colnames(decon_mtrx) != "res_ss"]

# Set as 0 cell types predicted to be under 1 % of the spot
decon_mtrx[decon_mtrx < 0.07] <- 0
```

We are going to add the deconvolution to the Seurat object.
```{r}
metadata <- cbind(sp_obj@meta.data, decon_mtrx)
sp_obj@meta.data <- metadata
```

Look at cells topic profile
```{r}
basis_spotlight <- data.frame(NMF::basis(spotlight_ls[[1]][[1]]))

colnames(basis_spotlight) <- unique(stringr::str_wrap(train_labs, width = 30))

basis_spotlight[basis_spotlight < 0.0000001] <- 0

DT::datatable(basis_spotlight)
```

#### Cell type location
Look at the location of each cell type in each slice separately
```{r fig.height=25, fig.width=25}
# Iterate over cell types
ct <- colnames(decon_mtrx)
ct_plt_ls <- lapply(ct, function(i) {
  tmp_plt <- Seurat::SpatialFeaturePlot(object = sp_obj,
                                        features = i,
                                        alpha = c(0, 1)) +
    scale_fill_gradientn(
      colors = heat.colors(10, rev = TRUE)) +
    scale_alpha(range = c(0, 1)) +
    labs(title = stringr::str_wrap(string = i,
                                   width = 25),
         fill = "") +
    theme(plot.title = element_text(hjust = 0.5,
                                    size = 20,
                                    face = "bold"))
  
  if (sum(sp_obj@meta.data[, i]) == 0) {
    tmp_plt <- suppressMessages(tmp_plt + scale_alpha(range = c(0,0)))
  }
  
  return(tmp_plt)
})

plt_arr <- cowplot::plot_grid(
  plotlist = ct_plt_ls,
  axis = "trbl",
  align = "hv",
  nrow = 4,
  ncol = 3)

plt_arr

cowplot::save_plot(
  filename = here::here(glue::glue("{decon_32}/{plt_dir}/
                                   cell_type_location_cd4_{donor_id}.pdf")),
  plot = plt_arr,
  base_height = 25,
  base_width = 18)
```


## Session Info
```{r}
sessionInfo()
```

