---
author: "Marc Elosua-Bayes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
params: 
    sample_id: "Default!"
    donor_id: "Default!"
title: "`r sprintf('%s L1 SC mapping visualization', params$donor_id)`"
editor_options: 
  chunk_output_type: console
---

---
title: `r params$donor_id`
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message = FALSE, warning = FALSE, cache = FALSE,
                      cache.lazy = FALSE)
options(width = 1200)
```

## Introduction
In this script we are going to visualize level 1 cell types onto the Visium slides.

## Libraries
```{r}
library(Seurat)
library(ggpubr)
library(cowplot)
library(dplyr)
library(ggplot2)
library(stringr)
library(readr)
library(SPOTlight)
```

## Setting parameters
Loading necessary paths and parameters
```{r}
set.seed(123)
source(here::here("misc/paths.R"))
source(here::here("utils/bin.R"))

dir.create(path = here::here(sprintf("%s/%s", decon_32, plt_dir)),
           showWarnings = FALSE,
           recursive = TRUE)

dir.create(path =here::here(sprintf("%s/%s", decon_32, robj_dir)),
           showWarnings = FALSE,
           recursive = TRUE)
```

Extract sample id and get Donor ID
```{r}
sample_id <- params$sample_id
# sample_id <- "esvq52_nluss5"
donor_id <- id_df[id_df$gem_id == sample_id, "BCLL_ID"]
```

## Load data
We have 4 different datasets that we are going to analyze separately.
The spatial data comes from the script *03-clustering/03-clustering.Rmd* while the SPOTlight model was generated in the script *level1-mapping.Rmd*.
**col_df_l1** contains the color for each cell-type and the relations between modified and unmodified names, this is generated in *misc/col_df_l1.Rmd*.
```{r}
se_obj <- readRDS(file = here::here(glue::glue("{clust_32}/{robj_dir}/clust_se_{sample_id}.RDS")))

spotlight_ls <- readRDS(file = here::here(glue::glue("{decon_32}/{robj_dir}/spotlight_ls_l1_{sample_id}.rds")))

col_df <- readRDS(file = here::here("misc/col_df_l1.rds"))
ct <- col_df$cell_type
```

## Analysis
Change old names to ones unmodified without dots
```{r}
decon_mtrx <- spotlight_ls[[2]]
decon_mtrx <- decon_mtrx[, colnames(decon_mtrx) != "res_ss"]

# Set as 0 cell types predicted to be under 1 % of the spot
decon_mtrx[decon_mtrx < 0.07] <- 0

new_colnames <- data.frame(ct_name = colnames(decon_mtrx),
                           stringsAsFactors = FALSE) %>%
  dplyr::left_join(col_df, by = c("ct_name" = "cell_type_mod")) %>%
  dplyr::pull(cell_type)

colnames(decon_mtrx) <- new_colnames
```

We are going to add the deconvolution to the Seurat object.
```{r}
metadata <- cbind(se_obj@meta.data, decon_mtrx)
se_obj@meta.data <- metadata
```

Save the seurat object with the Cell type scores
```{r}
saveRDS(
  object = se_obj,
  file = here::here(glue::glue("{decon_32}/{robj_dir}/se_deconv_{sample_id}.RDS")))
```

### Visualization

#### Topic profiles
```{r fig.height=8, fig.width=8}
h <- NMF::coef(spotlight_ls[[1]][[1]])

# Extract coefficient labels
train_labs <- data.frame(ct_name = spotlight_ls[[1]][[2]]) %>%
  dplyr::left_join(col_df, by = c("ct_name" = "cell_type")) %>%
  pull(ct_name)

rownames(h) <- paste("Topic", 1:nrow(h), sep = " ")
profile_plt <- SPOTlight::dot_plot_profiles_fun(
  h = h,
  train_cell_clust = stringr::str_wrap(train_labs, width = 30))

gen_topic <- profile_plt[[2]] +
    labs(
      title = "Topic profiles by cell type",
      x = "",
      y = "") +
    scale_x_discrete(limits = unique(str_wrap(train_labs, width = 30))) +
    theme(axis.text = element_text(size = 22),
          axis.text.x = element_text(hjust = 1, vjust = 0.5),
          legend.title = element_text(size = 20, face = "bold"),
          legend.text = element_text(size = 18),
          # plot.title = element_blank()
          plot.title = element_text(size = 30)
          )

gen_topic

cowplot::save_plot(
  filename = here::here(glue::glue("{decon_32}/{plt_dir}/ct_profiles_{donor_id}.svg")),
  plot = gen_topic,
  base_height = 15,
  base_width = 15)
```


Look at cells topic profile
```{r}
basis_spotlight <- data.frame(NMF::basis(spotlight_ls[[1]][[1]]))

colnames(basis_spotlight) <- unique(stringr::str_wrap(train_labs, width = 30))

basis_spotlight[basis_spotlight < 0.0000001] <- 0

DT::datatable(basis_spotlight)
```

#### Cell type location
Look at the location of each cell type in each slice separately
```{r fig.height=25, fig.width=25}
# Iterate over cell types
ct_plt_ls <- lapply(ct, function(i) {
  tmp_plt <- Seurat::SpatialFeaturePlot(object = se_obj,
                                        features = i,
                                        alpha = c(0, 1)) +
    scale_fill_gradientn(
      colors = heat.colors(10, rev = TRUE)) +
    scale_alpha(range = c(0, 1)) +
    labs(title = stringr::str_wrap(string = i,
                                   width = 25),
         fill = "") +
    theme(plot.title = element_text(hjust = 0.5,
                                    size = 20,
                                    face = "bold"))
  
  if (sum(se_obj@meta.data[, i]) == 0) {
    tmp_plt <- suppressMessages(tmp_plt + scale_alpha(range = c(0,0)))
  }
  
  return(tmp_plt)
})

plt_arr <- cowplot::plot_grid(
  plotlist = ct_plt_ls,
  axis = "trbl",
  align = "hv",
  nrow = 4,
  ncol = 3)

plt_arr

cowplot::save_plot(
  filename = here::here(glue::glue("{decon_32}/{plt_dir}/
                                   cell_type_location_{donor_id}.pdf")),
  plot = plt_arr,
  base_height = 25,
  base_width = 18)
```

#### Spatial scatterpie
```{r fig.height=8, fig.width=8}
sct_plt1 <- SPOTlight::scatterpie_plot(se_obj = se_obj,
                                       cell_types_all = ct,
                                       pie_scale = 0.5) +
          scale_fill_manual(values = col_df[col_df$cell_type %in% ct, "ct_col"],
                            breaks = ct) +
          coord_fixed(ratio = 1) +
          guides(fill = guide_legend(ncol = 1))
# sct_plt1

cowplot::save_plot(filename = here::here(glue::glue("{decon_32}/{plt_dir}/
                                                    spatial_{donor_id}_scatterpie.svg")),
                   plot = sct_plt1,
                   base_width = 12,
                   base_height = 9)
```

We can also take a look at the spatial scatterpie by looking at cell types which are not present throughout the entire tissue.
```{r fig.height=8, fig.width=8, eval = FALSE}
# Subset cell types from metadata
metadata_subset <- se_obj@meta.data[, ct]

# Create masks
keep_0.9 <- colSums(metadata_subset > 0) < 0.7 * ncol(se_obj)
keep_g0 <- colSums(metadata_subset > 0) > 0

# Select cell types fullfiling the conditions
ct_var <- colnames(metadata_subset)[keep_0.9 & keep_g0]

sct_int_plt <- SPOTlight::scatterpie_plot(se_obj = se_obj,
                                       cell_types_all = ct_var,
                                       pie_scale = 0.4) +
          scale_fill_manual(values = col_df[col_df$cell_type %in% ct_var, "ct_col"],
                            breaks = ct_var,
                            labels = stringr::str_wrap(string = ct_var,
                                                       width = 25)) +
          coord_fixed(ratio = 1) +
          theme(legend.position = "top",
                legend.text = element_text(size = 12),
                legend.title = element_blank()) +
          guides(fill = guide_legend(nrow = 3, byrow = TRUE))sct_int_plt

# sct_int_plt

cowplot::save_plot(filename = here::here(glue::glue("{decon_32}/{plt_dir}/
                                                    {donor_id}_scatterpie_subset.svg")),
                   plot = sct_int_plt,
                   base_width = 9,
                   base_height = 9)
```

#### Cell-Cell correlation plot
By looking at correlations we want to assess if there are cell types which are correlating in space. So when one is present is another one also present?

We will be using `ggcorrplot` to show these correlation plots a bit nicer:
http://www.sthda.com/english/wiki/ggcorrplot-visualization-of-a-correlation-matrix-using-ggplot2

Get correlation matrix between all cell types
```{r fig.height=12, fig.width=12}
# mat is a matrix object
cor.mtest <- function(mat, conf.level = 0.95){
    mat <- as.matrix(mat)
    n <- ncol(mat)
    p.mat <- lowCI.mat <- uppCI.mat <- matrix(NA, n, n)
    diag(p.mat) <- 0
    diag(lowCI.mat) <- diag(uppCI.mat) <- 1
    for(i in 1:(n-1)){
        for(j in (i+1):n){
            tmp <- cor.test(mat[,i], mat[,j], conf.level = conf.level)
            p.mat[i,j] <- p.mat[j,i] <- tmp$p.value
            lowCI.mat[i,j] <- lowCI.mat[j,i] <- tmp$conf.int[1]
            uppCI.mat[i,j] <- uppCI.mat[j,i] <- tmp$conf.int[2]
        }
    }
    return(list(p.mat, lowCI.mat, uppCI.mat))
}


decon_df <- se_obj@meta.data %>%
  dplyr::select(all_of(ct))

# Remove cell types not predicted to be on the tissue
decon_df <- decon_df[, colSums(decon_df) > 0]
decon_cor <- cor(decon_df)

# Compute correlation P-value
p.mat <- cor.mtest(decon_df, 0.95)
colnames(decon_cor) <- stringr::str_wrap(string = colnames(decon_cor), width = 30)
rownames(decon_cor) <- stringr::str_wrap(string = rownames(decon_cor), width = 30)

cor_plt <- ggcorrplot::ggcorrplot(corr = decon_cor,
           p.mat = p.mat[[1]],
           hc.order = TRUE,
           type = "full",
           insig = "blank",
           lab = TRUE,
           outline.col = "lightgrey",
           method = "square",
           # colors = c("#4477AA", "white", "#BB4444"))
           colors = c("#6D9EC1", "white", "#E46726"),
           title = glue::glue("{donor_id} cell-cell proportion correlation\nSample"),
           legend.title = "Correlation\n(Pearson)") +
   theme(
    plot.title = element_text(size = 22, hjust = 0.5, face = "bold"),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 15),
    axis.text.x = element_text(angle = 90),
    axis.text = element_text(size = 18, vjust = 0.5))
cor_plt

cowplot::save_plot(plot = cor_plt,
                 filename = here::here(glue::glue("{decon_32}/{plt_dir}/
                                                  correlation_plot_{donor_id}.svg")),
                 base_width = 16,
                 base_height = 12)

# Now without labels
cor_plt <- ggcorrplot::ggcorrplot(corr = decon_cor,
           p.mat = p.mat[[1]],
           hc.order = TRUE,
           type = "full",
           insig = "blank",
           lab = FALSE,
           outline.col = "lightgrey",
           method = "square",
           # colors = c("#4477AA", "white", "#BB4444"))
           colors = c("#6D9EC1", "white", "#E46726"),
           title = glue::glue("{donor_id} cell-cell proportion correlation\nSample"),
           legend.title = "Correlation\n(Pearson)") +
  theme(
    plot.title = element_text(size = 25, hjust = 0.5, face = "bold"),
    legend.text = element_text(size = 18),
    legend.title = element_text(size = 22),
    axis.text.x = element_text(angle = 90, size = 18, vjust = 0.5),
    axis.text.y = element_text(vjust = 0.5, size = 18))

cor_plt

cowplot::save_plot(plot = cor_plt,
                 filename = here::here(glue::glue("{decon_32}/{plt_dir}/
                                                  correlation_plot_{donor_id}_labeless.svg")),
                 base_width = 18,
                 base_height = 15)
```

#### Cell-Cell scatter correlation
```{r fig.height=15, fig.width=15}
# Correlation panel
panel.cor <- function(x, y){
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- round(cor(x, y), digits=2)
    txt <- paste0("R = ", r)
    cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = 2)
}
# Customize upper panel
upper.panel<-function(x, y){
  points(x,y, pch = 20)
}
# Create the plots
pairs(decon_mtrx,
      lower.panel = panel.cor,
      upper.panel = upper.panel,
      cex.labels = 2)
```

## Save to Shiny app
Lastly we will extract the metadata and counts and save them so we can use them in th in-house shinny app
`seurat2shiny` function can be found [here](https://github.com/Single-Cell-Genomics-Group-CNAG-CRG/shiny-annotation/blob/main/seurat2shiny.R)
```{r echo = FALSE}
seurat2shiny = function(
    object                         ,
    tech      = c("sc", "sp")      ,
    assay     = object@active.assay,
    slot      = "data"             ,
    reduction = "umap"             ,
    image     = NULL               ,
    asfactors = NULL               ,
    save      = FALSE               ,
    path      = "."                  # path = getwd()
) {
    suppressMessages( library(Seurat) );

    # Input check.
    if ( ! is(object, "Seurat") )
        stop("'object' is not a Seurat object.");

    if ( ! assay %in% Seurat::Assays(object) )
        stop("'assay' not in the Seurat object's available assays.");

    if ( tech == "sc" & ! (reduction %in% names(object@reductions)) )
        stop("'reduction' not in the Seurat object's available reductions.");

    if ( ! slot %in% c("counts", "data", "scale.data") )
        stop("'slot' not in the Seurat object's available slots.");
    
    if ( ! tech %in% c("sc", "sp") )
        stop("tech must be sc or sp.");
    
    
    # Check Which technology it is processing
    if (tech == "sc") {
        # Extract 2D coordinates.
        embeds <- as.data.frame(object@reductions[[reduction]]@cell.embeddings);
        names(embeds) <- c("coord_x", "coord_y");
    } else if (tech == "sp") {
        # If the image is null select the first one
        if (is.null(image)) {
            image <- names(object@images)[1]
            warning(sprintf("image is not set, we will use %s", image))
        } 
        
        embeds <- data.frame(object@images[[image]]@coordinates[, c("imagerow", "imagecol")])
        colnames(embeds) <- c("coord_y", "coord_x");
        
        # Inverse coord_y
        embeds$coord_y <- - embeds$coord_y
    }
    

    # Join metadata with coordinates.
    metadata <- object@meta.data;

    for (col in asfactors) {
        metadata[[col]] <- as.factor(metadata[[col]]);
    };

    metadata <- merge(x = metadata, y = embeds, by = "row.names");
    names(metadata)[1] <-  "barcode"; # names(metadata)[names(metadata) == "Row.names"] = "barcode";
    rownames(metadata) <- metadata$barcode

    # Extract expression data.
    # expression = as.matrix( Seurat::GetAssayData(object = object, slot = slot, assay = assay) );
    expression = Seurat::GetAssayData(object = object, slot = slot, assay = assay);

    if ( ! identical( as.character(metadata$barcode), colnames(expression) ) )
        warning("Cells in metadata and expression matrix do not match.");

    if (save) {
        saveRDS( object = metadata  , file = paste0(path, "/metadata.rds"  ) );
        saveRDS( object = expression, file = paste0(path, "/expression.rds") );
    };

    invisible(
        list(metadata = metadata, expression = expression)
    );
}
```

Extract elements
```{r}
se_obj@images
shiny_ls <- seurat2shiny(
  object = se_obj,
  tech = "sp",
  assay = "Spatial",
  slot = "data",
  image = sample_id)

saveRDS(object = shiny_ls[[1]],
        file = here::here(glue::glue("{decon_32}/{robj_dir}/metadata_{donor_id}.rds")))

saveRDS(object = shiny_ls[[2]],
        file = here::here(glue::glue("{decon_32}/{robj_dir}/expression_{donor_id}.rds")))

```

## Session Info
```{r}
sessionInfo()
```



