---
author: "Marc Elosua-Bayes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
params: 
    sample_id: "Default!"
    donor_id: "Default!"
title: "`r sprintf('%s L1 SC mapping', params$donor_id)`"
editor_options: 
  chunk_output_type: console
---

---
title: `r params$donor_id`
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message = FALSE, warning = FALSE, cache = FALSE,
                      cache.lazy = FALSE)
options(width = 1200)
```

## Introduction
In this script we are going to map level 1 cell types onto the Visium slides.

## Libraries
```{r}
library(Seurat)
library(ggpubr)
library(cowplot)
library(dplyr)
library(ggplot2)
library(stringr)
library(readr)
library(SPOTlight)
```

## Setting parameters
Loading necessary paths and parameters
```{r}
set.seed(123)
source(here::here("misc/paths.R"))
source(here::here("utils/bin.R"))

dir.create(path = here::here(sprintf("%s/%s", decon_32, plt_dir)),
           showWarnings = FALSE,
           recursive = TRUE)

dir.create(path =here::here(sprintf("%s/%s", decon_32, robj_dir)),
           showWarnings = FALSE,
           recursive = TRUE)
```

Extract sample id and get Donor ID
```{r}
sample_id <- params$sample_id
# sample_id <- "esvq52_nluss5"
donor_id <- id_df[id_df$gem_id == sample_id, "BCLL_ID"]
```

## Load data
We have 4 different datasets that we are going to analyze separately.
The spatial data comes from the script *03-clustering/03-clustering.Rmd* while the sc data can be found in Ramon's scRNAseq analysis: **/scratch/devel/rmassoni/tonsil_atlas_private/2-DOWNSTREAM_PROCESSING/results/R_objects/processed_seurat_objects/processed_seurat_objects/tonsil_integrated_with_harmony_scrublet_annotated.rds**. \
In each Rmarkdown notebook we analyze 1 dataset. \
```{r}
se_obj <- readRDS(file = here::here(glue::glue("{clust_32}/{robj_dir}/clust_se_{sample_id}.RDS")))

sc_obj <- readRDS(file = "/scratch/devel/rmassoni/tonsil_atlas_private/2-DOWNSTREAM_PROCESSING/results/R_objects/processed_seurat_objects/processed_seurat_objects/tonsil_integrated_with_harmony_scrublet_annotated.rds")
```

## Analysis
We are going to use SPOTlight to deconvolute the spatial spots and map the cell types.
```{r}
sc_obj@meta.data %>%
  dplyr::count(annotation_level1)
```

1st we need to determine the marker genes for each high-level cluster
```{r}
if (file.exists(here::here(glue::glue("{decon_32}/{robj_dir}/markers_l1.RDS")))) {
  
  markers <- readRDS(file = here::here(glue::glue("{decon_32}/{robj_dir}/markers_l1.RDS")))
  
} else {
  #### Extract marker genes from each cluster ####
  Seurat::Idents(object = sc_obj) <- sc_obj@meta.data$annotation_level1
  markers <- Seurat::FindAllMarkers(
    object = sc_obj, 
    assay = "RNA",
    slot = "data",
    verbose = TRUE, 
    only.pos = TRUE,
    max.cells.per.ident = 200)
  
  saveRDS(object = markers,
          file = here::here(glue::glue("{decon_32}/{robj_dir}/markers_l1.RDS")))
}
```

2nd run deconvolution
```{r}
spotlight_ls <- spotlight_deconvolution(
  se_sc = sc_obj,
  counts_spatial = se_obj@assays$Spatial@counts,
  clust_vr = "annotation_level1",
  cluster_markers = markers,
  cl_n = 75,
  hvg = 3000,
  ntop = NULL,
  transf = "uv",
  method = "nsNMF",
  min_cont = 0)
```

3rd Check Topic profiles
```{r fig.height=8, fig.width=8}
nmf_mod_ls <- spotlight_ls[[1]]
nmf_mod <- nmf_mod_ls[[1]]

h <- NMF::coef(nmf_mod)
rownames(h) <- paste("Topic", 1:nrow(h), sep = "_")
topic_profile_plts <- SPOTlight::dot_plot_profiles_fun(
  h = h,
  train_cell_clust = nmf_mod_ls[[2]])

topic_profile_plts[[2]] + theme(axis.text.x = element_text(angle = 90), 
                                axis.text = element_text(size = 12))
```

## Save data
```{r}
saveRDS(object = spotlight_ls,
        file = here::here(glue::glue("{decon_32}/{robj_dir}/spotlight_ls_l1_{sample_id}.rds")))
```

## Session Info
```{r}
sessionInfo()
```

