---
author: "Marc Elosua-Bayes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
params: 
    sample_id: "Default!"
    donor_id: "Default!"
title: "`r sprintf('%s L1 Cell-Type clustering', params$donor_id)`"
editor_options: 
  chunk_output_type: console
---

---
title: `r params$donor_id`
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message = FALSE, warning = FALSE, cache = FALSE,
                      cache.lazy = FALSE)
options(width = 1200)
```

## Introduction
In this script we are going to attempt to cluster the spots by cell type composition. We are going to follow different approaches to assess which is the best one


## Libraries
```{r}
library(Seurat)
library(ggpubr)
library(cowplot)
library(dplyr)
library(ggplot2)
library(stringr)
library(readr)
library(SPOTlight)
```

## Setting parameters
Loading necessary paths and parameters
```{r}
set.seed(123)
source(here::here("misc/paths.R"))
source(here::here("utils/bin.R"))

dir.create(path = here::here(sprintf("%s/%s", decon_32, plt_dir)),
           showWarnings = FALSE,
           recursive = TRUE)

dir.create(path =here::here(sprintf("%s/%s", decon_32, robj_dir)),
           showWarnings = FALSE,
           recursive = TRUE)
```

Extract sample id and get Donor ID
```{r}
sample_id <- params$sample_id
sample_id <- "esvq52_nluss5"
donor_id <- id_df[id_df$gem_id == sample_id, "BCLL_ID"]
```

## Load data
We have 4 different datasets that we are going to analyze separately.
The spatial data comes from the script *03-clustering/03-clustering.Rmd* while the SPOTlight model was generated in the script *level1-mapping.Rmd*.
**col_df_l1** contains the color for each cell-type and the relations between modified and unmodified names, this is generated in *misc/col_df_l1.Rmd*.
```{r}
se_obj <- readRDS(file = here::here(glue::glue("{clust_32}/{robj_dir}/clust_se_{sample_id}.RDS")))

spotlight_ls <- readRDS(file = here::here(glue::glue("{decon_32}/{robj_dir}/spotlight_ls_l1_{sample_id}.rds")))

col_df <- readRDS(file = here::here("misc/col_df_l1.rds"))
ct <- col_df$cell_type
```

## Analysis
Change old names to ones unmodified without dots
```{r}
decon_mtrx <- spotlight_ls[[2]]
decon_mtrx <- decon_mtrx[, colnames(decon_mtrx) != "res_ss"]

# Set as 0 cell types predicted to be under 1 % of the spot
decon_mtrx[decon_mtrx < 0.07] <- 0

new_colnames <- data.frame(ct_name = colnames(decon_mtrx),
                           stringsAsFactors = FALSE) %>%
  dplyr::left_join(col_df, by = c("ct_name" = "cell_type_mod")) %>%
  dplyr::pull(cell_type)

colnames(decon_mtrx) <- new_colnames
```

We are going to add the deconvolution to the Seurat object.
```{r}
metadata <- cbind(se_obj@meta.data, decon_mtrx)
se_obj@meta.data <- metadata
```

### K-means clustering
[guided tutorial](https://www.datanovia.com/en/lessons/k-means-clustering-in-r-algorith-and-practical-examples/)
k-means clustering consists of defining clusters so that the total intra-cluster variation (known as total within-cluster variation) is minimized.

The standard algorithm is the Hartigan-Wong algorithm (Hartigan and Wong 1979), which defines the total within-cluster variation as the sum of squared distances Euclidean distances between items and the corresponding centroid.
$W(C_{k}) = \sum_{x_{i}\epsilon_{k}}(x_{i}-\mu_{k})^2$

*How do we estimate the optimal number of clusters?*
The idea is to compute k-means clustering using different values of clusters k. Next, the wss (within sum of square) is drawn according to the number of clusters. The location of a bend (knee) in the plot is generally considered as an indicator of the appropriate number of clusters.

The R function `fviz_nbclust` in `factoextra` package provides a convenient solution to estimate the optimal number of clusters.
```{r}
install.packages("factoextra")
factoextra::fviz_nbclust(
  x = decon_mtrx,
  kmeans,
  method = "wss") +
  geom_vline(xintercept = 4, linetype = 2) +
  geom_vline(xintercept = 8, linetype = 2) +
  labs(subtitle = "Elbow method")

```

We can see there are 2 elbows when k is 4 and 8, we are going to carry out the clustering at both resolutions.
```{r}
k3 <- kmeans(decon_mtrx, 3, iter.max = 10, nstart = 5)

k8 <- kmeans(decon_mtrx, 8, iter.max = 20, nstart = 5)
```

Function to print stacked barplots of the cell types contributing to each cluster:
```{r fig.height=8, fig.width=8}
kmeans_barplot <- function(mtrx) {
  
  data.frame(mtrx, check.names = FALSE) %>%
    tibble::rownames_to_column("Cluster") %>%
    tidyr::pivot_longer(
      cols = colnames(decon_mtrx),
      names_to = "Cell type",
      values_to = "Loadings") %>%
    ggplot(
      .,
      aes(
        x = Cluster,
        y = Loadings,
        fill = `Cell type`)) + 
      geom_bar(position = "stack", stat = "identity") +
    ggpubr::theme_pubr()
}

kmeans_barplot(mtrx = k3$centers)
kmeans_barplot(mtrx = k8$centers)
```

Add the clusterings to the spatial seurat object
```{r}
k_df <- data.frame(k3$cluster, k8$cluster)
se_obj@meta.data <- dplyr::bind_cols(se_obj@meta.data, k_df)
```

Plot k-means clusterings in space
```{r}
Seurat::SpatialDimPlot(
  object = se_obj,
  group.by = c("Spatial_snn_res.0.1",
               "Spatial_snn_res.0.75",
               "k3.cluster",
               "k8.cluster"))
```

## Save to Shiny app
Lastly we will extract the metadata and counts and save them so we can use them in th in-house shinny app
`seurat2shiny` function can be found [here](https://github.com/Single-Cell-Genomics-Group-CNAG-CRG/shiny-annotation/blob/main/seurat2shiny.R)
```{r echo = FALSE}
seurat2shiny = function(
    object                         ,
    tech      = c("sc", "sp")      ,
    assay     = object@active.assay,
    slot      = "data"             ,
    reduction = "umap"             ,
    image     = NULL               ,
    asfactors = NULL               ,
    save      = FALSE               ,
    path      = "."                  # path = getwd()
) {
    suppressMessages( library(Seurat) );

    # Input check.
    if ( ! is(object, "Seurat") )
        stop("'object' is not a Seurat object.");

    if ( ! assay %in% Seurat::Assays(object) )
        stop("'assay' not in the Seurat object's available assays.");

    if ( tech == "sc" & ! (reduction %in% names(object@reductions)) )
        stop("'reduction' not in the Seurat object's available reductions.");

    if ( ! slot %in% c("counts", "data", "scale.data") )
        stop("'slot' not in the Seurat object's available slots.");
    
    if ( ! tech %in% c("sc", "sp") )
        stop("tech must be sc or sp.");
    
    
    # Check Which technology it is processing
    if (tech == "sc") {
        # Extract 2D coordinates.
        embeds <- as.data.frame(object@reductions[[reduction]]@cell.embeddings);
        names(embeds) <- c("coord_x", "coord_y");
    } else if (tech == "sp") {
        # If the image is null select the first one
        if (is.null(image)) {
            image <- names(object@images)[1]
            warning(sprintf("image is not set, we will use %s", image))
        } 
        
        embeds <- data.frame(object@images[[image]]@coordinates[, c("imagerow", "imagecol")])
        colnames(embeds) <- c("coord_y", "coord_x");
        
        # Inverse coord_y
        embeds$coord_y <- - embeds$coord_y
    }
    

    # Join metadata with coordinates.
    metadata <- object@meta.data;

    for (col in asfactors) {
        metadata[[col]] <- as.factor(metadata[[col]]);
    };

    metadata <- merge(x = metadata, y = embeds, by = "row.names");
    names(metadata)[1] <-  "barcode"; # names(metadata)[names(metadata) == "Row.names"] = "barcode";
    rownames(metadata) <- metadata$barcode

    # Extract expression data.
    # expression = as.matrix( Seurat::GetAssayData(object = object, slot = slot, assay = assay) );
    expression = Seurat::GetAssayData(object = object, slot = slot, assay = assay);

    if ( ! identical( as.character(metadata$barcode), colnames(expression) ) )
        warning("Cells in metadata and expression matrix do not match.");

    if (save) {
        saveRDS( object = metadata  , file = paste0(path, "/metadata.rds"  ) );
        saveRDS( object = expression, file = paste0(path, "/expression.rds") );
    };

    invisible(
        list(metadata = metadata, expression = expression)
    );
}
```

Extract elements
```{r eval = FALSE}
se_obj@images
shiny_ls <- seurat2shiny(
  object = se_obj,
  tech = "sp",
  assay = "Spatial",
  slot = "data",
  image = sample_id)

saveRDS(object = shiny_ls[[1]],
        file = here::here(glue::glue("{decon_32}/{robj_dir}/metadata_{donor_id}.rds")))

saveRDS(object = shiny_ls[[2]],
        file = here::here(glue::glue("{decon_32}/{robj_dir}/expression_{donor_id}.rds")))
```

## Session Info
```{r}
sessionInfo()
```



