---
author: "Ramon Massoni-Badosa"
date: "2021/03/15"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
title <- paste("Visualize integration (level 2):", cell_type, sep = " ")
```

---
title: "`r title`"
---

TODO: function "feature_plot_doublets" discretizes any quantiative variable and changes the scale. Not a problem here, since it's purely exploratory. However, to prevent this, instead of passing a vector of viridis colors to FeaturePlot, run it normally and then + scale_color_viridis_c(option = "magma")


# Introduction

The objectives of this notebook are:

- Visualize how well we could remove technical variability associated with the assay (3', 5' or multiome). We should see a high degree of intermixing between the 3 assays.
- Visualize how well we could preserve the biological variability. We will use the annotation of King et al., as a positive control, so that the data integration should preserve the cell type annotations.
- Plot several markers of doublets: lineage markers, pDNN, scrublet predictions, number of features, etc. The broader objective of level 2 is to eliminate most remaining doublets and poor-quality cells, as we will discuss in future notebooks. These visualizations will allow us to explore this. We will also plot their location in the UMAP obtained at level 1.


# Pre-processing

## Load packages

```{r}
library(Seurat)
library(tidyverse)
```


## Parameters

```{r}
# Paths
path_to_obj <- str_c(
  here::here("scRNA-seq/results/R_objects/level_2/"),
  cell_type,
  "/",
  cell_type,
  "_integrated_level_2.rds",
  sep = ""
)


# Functions
source(here::here("scRNA-seq/bin/utils.R"))


# Colors
color_palette <- c("black", "gray", "red", "yellow", "violet", "green4",
                   "blue", "gold", "antiquewhite", "chocolate1", "coral2",
                   "blueviolet", "brown1", "darkmagenta", "deepskyblue1",
                   "dimgray", "deeppink1", "green", "lightgray", "hotpink1",
                   "indianred4", "khaki", "mediumorchid2")
```


## Load data

```{r}
# Seurat object
seurat <- readRDS(path_to_obj)
seurat
```


# Assess integration

## Removal of technical variability

```{r}
p_assay <- plot_split_umap(seurat, var = "assay")
p_assay
```


## Preservation of biological variability

```{r}
p_king <- plot_annotation_king(seurat, color_palette = color_palette) +
  guides(colour = guide_legend(override.aes = list(size = 2.5)))
p_king
```


# Spot potential doublets

## Lineage markers

```{r}
canonical_markers <- c("CD79A", "CD79B", "CD3D", "CD3E", "NKG7", "LYZ",
                       "IGHD", "IGHM", "IGHA1", "IGHG1", "FDCSP", "PTCRA",
                       "XBP1", "TOP2A", "KRT19", "SPRR3", "DNTT", "VPREB1")
canonical_markers_gg <- purrr::map(canonical_markers, function(x) {
  p <- feature_plot_doublets(seurat_obj = seurat, feature = x)
  p
})
names(canonical_markers_gg) <- canonical_markers
canonical_markers_gg
```


## pDNN and scrublet prediction

```{r}
# pDNN
pDNN_vars <- c("pDNN_hashing", "pDNN_scrublet", "pDNN_union")
pDNN_gg <- purrr::map(pDNN_vars, function(x) {
  p <- plot_pDNN(seurat_obj = seurat, pDNN_var = x, pt_size = 0.2)
  p
})
pDNN_gg


# Scrublet
seurat$scrublet_predicted_doublet[seurat$scrublet_predicted_doublet == "True"] <- "TRUE"
seurat$scrublet_predicted_doublet[seurat$scrublet_predicted_doublet == "False"] <- "FALSE"
DimPlot(seurat, group.by = "scrublet_predicted_doublet")
```


## QC metrics

```{r}
qc_vars <- c(
  "nCount_RNA",
  "nFeature_RNA",
  "pct_mt",
  "pct_ribosomal"
)
qc_gg <- purrr::map(qc_vars, function(x) {
  p <- FeaturePlot(seurat, features = x, cols = viridis::viridis(10))
  p
})
qc_gg
```


## UMAP level 1

```{r}
umap_level_1 <- FeatureScatter(
  seurat,
  "UMAP_1_level_1",
  "UMAP_2_level_1",
  group.by = "seurat_clusters"
)
umap_level_1 <- umap_level_1 +
  theme(
    legend.position = "none",
    plot.title = element_blank()
  )
umap_level_1
```


## Cell cycle scoring

```{r}
s_gg <- feature_plot_doublets(seurat_obj = seurat, feature = "S.Score")
g2m_gg <- feature_plot_doublets(seurat_obj = seurat, feature = "G2M.Score")
s_gg
g2m_gg
```


# Session Information

```{r}
sessionInfo()
```
