---
title: "CD4 T: Level 5"
author: "Ramon Massoni-Badosa"
date: "2021/05/13"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```

NOTE: from level 4/5 onwards I will execute most steps locally, as the datasets are not as large anymore.


# Introduction

Here, we will work on the level 5 of the CD4 T cell, which entails:

* Removal of 3 clusters: activated CD4 T cells, metabolic/poor-quality cells, and mitochondrial+ T cells.
* Subcluster Follicular Th CXCL13+CBLB+ into 2 clusters
* Recalculate the UMAP.
* Subset the Th17, IL2RA+FOXP3+ Treg and Memory T to further stratify them (3P only).


## Load packages

```{r}
library(Seurat)
library(SeuratWrappers)
library(harmony)
library(tidyverse)
```


## Parameters

```{r}
# Paths
path_to_obj <- here::here("scRNA-seq/results/R_objects/level_4/CD4_T/CD4_T_integrated_level_4.rds")
# path_to_obj <- "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/results/R_objects/level_4/CD4_T/CD4_T_integrated_level_4.rds"
path_to_save <- here::here("scRNA-seq/results/R_objects/level_5/CD4_T/paper/CD4_T_subseted_annotated_level_5.rds")
# path_to_save <- "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/results/R_objects/level_5/CD4_T/paper/CD4_T_subseted_annotated_level_5.rds"
path_to_save_th <- "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/results/R_objects/level_5/CD4_T/paper/CD4_T_Th_level_5_annotated.rds"

# Colors
color_palette <-  c("#1CFFCE", "#90AD1C", "#C075A6", "#85660D", "#5A5156", "#AA0DFE",   
                    "#F8A19F", "#F7E1A0", "#1C8356", "#FEAF16", "#822E1C", "#C4451C",   
                    "#1CBE4F", "#325A9B", "#F6222E", "#FE00FA", "#FBE426", "#16FF32", 
                    "black",   "#3283FE", "#B00068", "#DEA0FD", "#B10DA1", "#E4E1E3",   
                    "#90AD1C", "#FE00FA", "#85660D", "#3B00FB", "#822E1C", "coral2", 
                    "#1CFFCE", "#1CBE4F", "#3283FE", "#FBE426", "#F7E1A0", "#325A9B",   
                    "#2ED9FF", "#B5EFB5", "#5A5156", "#DEA0FD", "#FEAF16", "#683B79",   
                    "#B10DA1", "#1C7F93", "#F8A19F", "dark orange", "#FEAF16", "#FBE426",  
                    "Brown")

# Functions
source(here::here("scRNA-seq/bin/utils.R"))
# source("~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/bin/utils.R")
# source("~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/bin/SLOcatoR_functions.R")
source(here::here("scRNA-seq/bin/SLOcatoR_functions.R"))


# Additional function
plot_subcluster <- function(seurat_obj, pattern) {
  p <- seurat_obj@reductions$umap@cell.embeddings %>%
    as.data.frame() %>%
    mutate(cluster = seurat_obj$annotation_level_3) %>%
    filter(str_detect(cluster, pattern)) %>%
    ggplot(aes(UMAP_1, UMAP_2, color = cluster)) +
      geom_point(size = 0.1) +
      theme_classic()
  p
}
```


## Load data

```{r}
# Seurat object
seurat <- readRDS(path_to_obj)
seurat
DimPlot(seurat, cols = color_palette, pt.size = 0.2)
```


# Subcluster

```{r}
seurat <- FindSubCluster(
  seurat,
  cluster = "Follicular Th CXCL13+CBLB+",
  subcluster.name = "annotation_level_5",
  resolution = 0.1,
  graph.name = "RNA_snn"
)
DimPlot(seurat, group.by = "annotation_level_5", cols = color_palette)
```

## Markers

```{r}
Idents(seurat) <- "annotation_level_5"
clusters_interest <- seurat$annotation_level_5 %>%
  unique() %>%
  str_subset("^Follicular Th CXCL13") %>%
  sort()
markers <- purrr::map(clusters_interest, function(x) {
  group_1 <- clusters_interest[which(clusters_interest == x)]
  group_2 <- clusters_interest[which(clusters_interest != x)]
  print(group_1)
  print(group_2)
  df <- FindMarkers(
    seurat,
    ident.1 = group_1,
    ident.2 = group_2,
    only.pos = TRUE,
    logfc.threshold = 0.5,
    verbose = TRUE
  )
  df <- df %>%
    rownames_to_column(var = "gene") %>%
    arrange(desc(avg_log2FC))
  df
})
names(markers) <- clusters_interest
# path_save_markers <- "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/3-clustering/5-level_5/CD4_T/tmp/markers_Tfh_CXCL13_new.xlsx"
path_save_markers <- here::here("scRNA-seq/3-clustering/5-level_5/CD4_T/tmp/markers_Tfh_CXCL13_new.xlsx")
openxlsx::write.xlsx(x = markers, file = path_save_markers, overwrite = TRUE)
```


# Subset

```{r}
clusters_to_exclude <- c("activated CD4 T", "Mitochondrial+ T cells",
                         "metabolic/poor-quality")
mask <- !(Idents(seurat) %in% clusters_to_exclude)
table(mask)
selected_cells <- colnames(seurat)[mask]
seurat <- subset(seurat, cells = selected_cells)
```


Let us exclude all the cells from Royal London (discussed elsewhere):

```{r}
seurat <- subset(seurat, hospital != "Royal London")
```


# Reprocess

```{r}
hvg <- find_assay_specific_features(seurat)
seurat <- integrate_assays(
  seurat,
  assay_specific = TRUE,
  shared_hvg = hvg,
  assay_var = "assay",
  n_dim = 30
)
```


# Recompute UMAP

```{r}
seurat <- RunUMAP(seurat, reduction = "harmony", dims = 1:30)
DimPlot(seurat, cols = color_palette)
```


# Subset Th

```{r}
clusters_interest_2 <- "Th17"
th <- subset(seurat, idents = clusters_interest_2)
th <- subset(th, assay == "3P")
th <- th %>%
  FindVariableFeatures() %>%
  ScaleData() %>%
  RunPCA() %>%
  RunHarmony(group.by.vars = "donor_id", reduction = "pca", dims = 1:30) %>%
  RunUMAP(reduction = "harmony", dims = 1:30)

markers_interest <- c("IFNG", "GATA3", "IL26", "IL21", "CCL3", "IL4")
feat_plots <- purrr::map(markers_interest, function(x) {
  p <- FeaturePlot(th, features = x) +
    scale_color_viridis_c(option = "magma")
  p
})
feat_plots
DimPlot(th)


# Cluster
th <- FindNeighbors(th, dims = 1:30, reduction = "harmony")
th <- FindClusters(th, resolution = 0.65)
DimPlot(th)


# Markers
markers_th <- FindAllMarkers(th, only.pos = TRUE, logfc.threshold = 0.5)
```


# Annotation (as of 5th Nov, 2021)

All CD4 T cells:

```{r}
seurat$annotation_paper <- case_when(
  seurat$annotation_level_5 == "Naive" ~ "Naive",
  seurat$annotation_level_5 == "Central Mem PASK+" ~ "CM Pre-non-Tfh",
  seurat$annotation_level_5 == "Central Mem PASK-" ~ "CM PreTfh",
  seurat$annotation_level_5 == "Memory T cells" ~ "T-Trans-Mem",
  seurat$annotation_level_5 == "Follicular Th CXCL13+CBLB+_1" ~ "T-Eff-Mem",
  seurat$annotation_level_5 == "Th17" ~ "T-helper",
  seurat$annotation_level_5 == "Follicular Th CXCL13+CBLB+_2" ~ "Tfh T:B border",
  seurat$annotation_level_5 == "Follicular Th CXCL13+CBLB+_0" ~ "Tfh-LZ-GC",
  seurat$annotation_level_5 == "Follicular Th TOX2+" ~ "Tfh-LZ-GC",
  seurat$annotation_level_5 == "Follicular Th CXCR5+" ~ "GC-Tfh-SAP",
  seurat$annotation_level_5 == "CD200+TOX+" ~ "GC-Tfh-0X40",
  seurat$annotation_level_5 == "CD4 Non-Tfh KLRB1+ " ~ "Tfh-Mem",
  seurat$annotation_level_5 == "IL2RA+FOXP3+ Treg" ~ "Eff-Tregs",
  seurat$annotation_level_5 == "Treg IKZF2+HPGD+" ~ "non-GC-Tf-regs",
  seurat$annotation_level_5 == "naive Treg IKZF2+" ~ "GC-Tf-regs",
)
ordered_levels <- c("Naive", "CM Pre-non-Tfh", "CM PreTfh", "T-Trans-Mem",
                    "T-Eff-Mem", "T-helper", "Tfh T:B border", "Tfh-LZ-GC",
                    "GC-Tfh-SAP", "GC-Tfh-0X40", "Tfh-Mem", "Eff-Tregs", 
                    "non-GC-Tf-regs", "GC-Tf-regs")
seurat$annotation_paper <- factor(
  seurat$annotation_paper,
  levels = ordered_levels
)
Idents(seurat) <- "annotation_paper"
```


T helper subsets: since we found a cluster of Th22 cells, we will recluster to find it.

```{r}
th <- FindSubCluster(
  th,
  cluster = "3",
  graph.name = "RNA_snn",
  subcluster.name = "Th22", 
  resolution = 0.25
)
th$annotation_paper <- case_when(
  th$Th22 == "0" ~ "Th2",
  th$Th22 == "1" ~ "Th0",
  th$Th22 == "2" ~ "Th17",
  th$Th22 == "3_0" ~ "Th1",
  th$Th22 == "3_1" ~ "Th22",
  th$Th22 == "4" ~ "Th1/Th17"
)
ordered_levels_th <- c("Th0", "Th1", "Th2", "Th17", "Th1/Th17", "Th22")
th$annotation_paper <- factor(
  th$annotation_paper,
  levels = ordered_levels_th
)
Idents(th) <- "annotation_paper"
DimPlot(th, cols = color_palette)
```

Find all markers

```{r}
# All CD4 T cells
markers_all_cd4 <- FindAllMarkers(seurat, only.pos = TRUE, logfc.threshold = 0.6)
markers_all_cd4 <- markers_all_cd4 %>%
  group_by(cluster) %>%
  arrange(desc(avg_log2FC), group_by = TRUE) %>%
  ungroup()
markers_all_dfs <- purrr::map(unique(markers_all_cd4$cluster), function(x) {
  df <- markers_all_cd4[markers_all_cd4$cluster == x, ]
  df <- df[, c(7, 1:6)]
  df
})
names(markers_all_dfs) <- unique(markers_all_cd4$cluster)
markers_all_dfs <- markers_all_dfs[levels(seurat$annotation_paper)]


# Th cells
markers_th <- FindAllMarkers(th, only.pos = TRUE, logfc.threshold = 0.45)
markers_th <- markers_th %>%
  group_by(cluster) %>%
  arrange(desc(avg_log2FC), group_by = TRUE) %>%
  ungroup()
markers_th_dfs <- purrr::map(unique(markers_th$cluster), function(x) {
  df <- markers_th[markers_th$cluster == x, ]
  df <- df[, c(7, 1:6)]
  df
})
names(markers_th_dfs) <- unique(markers_th$cluster)
markers_th_dfs <- markers_th_dfs[levels(th$annotation_paper)]
names(markers_th_dfs)[names(markers_th_dfs) == "Th1/Th17"] <- "Th1_Th17"
```



# Save

```{r}
# Save Seurat objects
saveRDS(seurat, path_to_save)
saveRDS(th, path_to_save_th)


# Save markers
# path_save_markers <- "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/3-clustering/5-level_5/CD4_T/tmp/CD4_T_markers_annotated_level_5.xlsx"
names(markers_all_dfs)[names(markers_all_dfs) == "Tfh T:B border"] <- "Tfh T-B border"
path_save_markers <- here::here("scRNA-seq/3-clustering/5-level_5/CD4_T/tmp/CD4_T_markers_annotated_level_5.xlsx")
openxlsx::write.xlsx(
  x = markers_all_dfs,
  file = path_save_markers,
  overwrite = TRUE
)
# path_to_save_markers_th <- "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/3-clustering/5-level_5/CD4_T/tmp/markers_Th_subsets_annotated.xlsx"
path_to_save_markers_th <- here::here("scRNA-seq/3-clustering/5-level_5/CD4_T/tmp/markers_Th_subsets_annotated.xlsx")
openxlsx::write.xlsx(
  x = markers_th_dfs,
  file = path_to_save_markers_th,
  overwrite = TRUE
)


# Save UMAP
umap_level_5 <- DimPlot(seurat, cols = color_palette)
# path_to_save_umap1 <- "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/3-clustering/5-level_5/CD4_T/tmp/CD4_T_umap_level_5_2_annotated.png"
path_to_save_umap1 <- here::here("scRNA-seq/3-clustering/5-level_5/CD4_T/tmp/CD4_T_umap_level_5_2_annotated.png")
ggsave(
  filename = path_to_save_umap1,
  plot = umap_level_5,
  width = 14,
  height = 12,
  units = "cm"
)
umap_level_5_th <- DimPlot(th, cols = color_palette)
# path_to_save_umap2 <- "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/3-clustering/5-level_5/CD4_T/tmp/CD4_Th_umap_annotated.png"
path_to_save_umap2 <- here::here("scRNA-seq/3-clustering/5-level_5/CD4_T/tmp/CD4_Th_umap_annotated.png")
ggsave(
  filename = path_to_save_umap2,
  plot = umap_level_5_th,
  width = 14,
  height = 12,
  units = "cm"
)
```


# Session information

```{r}
sessionInfo()
```

