---
title: "CD4 T: Integrate scRNA-seq and CITE-seq"
author: "Ramon Massoni-Badosa"
date: "2021/05/13"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```

NOTE: from level 4/5 onwards I will execute most steps locally, as the datasets are not as large anymore.


# Introduction

Here, we will (1) integrate scRNA-seq and CITE-seq data for CD4 T cells, (2) transfer the UMAP coordinates and the annotation, and (3) find the protein markers for each cell type.


## Load packages

```{r}
library(Seurat)
library(SeuratWrappers)
library(harmony)
library(tidyverse)
library(caret)
library(class)
```


## Parameters

```{r}
# path_to_cd4 <- "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/results/R_objects/level_5/CD4_T/paper/CD4_T_subseted_annotated_level_5.rds"
path_to_cd4 <- here::here("scRNA-seq/results/R_objects/level_5/CD4_T/paper/CD4_T_subseted_annotated_level_5.rds")
# path_to_cite_seq <- "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/CITE-seq/results/seurat_object_cite_seq_seurat_wnn.rds"
path_to_cite_seq <- here::here("CITE-seq/results/seurat_object_cite_seq_seurat_wnn.rds")
path_to_save_cite_seq <- here::here("scRNA-seq/results/R_objects/level_5/CD4_T/paper/CD4_T_subseted_annotated_level_5_cite_seq.rds")
path_to_save_prot_markers <- here::here("scRNA-seq/3-clustering/5-level_5/CD4_T/tmp/CD4_T_markers_annotated_level_5_cite_seq.xlsx")
path_to_save_umap_cite_seq <- here::here("scRNA-seq/3-clustering/5-level_5/CD4_T/tmp/CD4_T_umap_level_5_2_annotated_cite_seq.png")


# Colors
color_palette <-  c("#1CFFCE", "#90AD1C", "#C075A6", "#85660D", "#5A5156", "#AA0DFE",   
                    "#F8A19F", "#F7E1A0", "#1C8356", "#FEAF16", "#822E1C", "#C4451C",   
                    "#1CBE4F", "#325A9B", "#F6222E", "#FE00FA", "#FBE426", "#16FF32", 
                    "black",   "#3283FE", "#B00068", "#DEA0FD", "#B10DA1", "#E4E1E3",   
                    "#90AD1C", "#FE00FA", "#85660D", "#3B00FB", "#822E1C", "coral2", 
                    "#1CFFCE", "#1CBE4F", "#3283FE", "#FBE426", "#F7E1A0", "#325A9B",   
                    "#2ED9FF", "#B5EFB5", "#5A5156", "#DEA0FD", "#FEAF16", "#683B79",   
                    "#B10DA1", "#1C7F93", "#F8A19F", "dark orange", "#FEAF16", "#FBE426",  
                    "Brown")


# Source functions
# source("~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/bin/SLOcatoR_functions.R")
source(here::here("scRNA-seq/bin/SLOcatoR_functions.R"))
```


## Read data

```{r}
reference <- readRDS(path_to_cd4)
cite_seq <- readRDS(path_to_cite_seq)
query <- subset(cite_seq, idents = "2")
rm(cite_seq)
```


# Project CITE-seq data

```{r}
reference$UMAP1 <- Embeddings(reference, "umap")[, "UMAP_1"]
reference$UMAP2 <- Embeddings(reference, "umap")[, "UMAP_2"]
reference_diet <- DietSeurat(reference)
query_diet <- DietSeurat(query)
reference_diet$type <- "reference"
query_diet$type <- "query"
query_diet$assay <- "CITEseq"
integrated <- merge(x = reference_diet, y = query_diet)
rm(query_diet, reference_diet, reference)
integrated <- NormalizeData(integrated)
hvg <- find_assay_specific_features(
  integrated,
  assay_var = "assay",
  n_features = 5000
)
integrated <- integrate_assays(
  integrated,
  assay_specific = TRUE,
  n_dim = 30,
  shared_hvg = hvg
)
integrated <- RunUMAP(integrated, reduction = "harmony", dims = 1:30)
DimPlot(integrated, split.by = "assay", group.by = "assay")
data_sets <- split_training_and_test_sets(
  integrated,
  split_var = "type",
  referece_label = "reference",
  query_label = "query",
  reduction = "harmony",
  n_dims = 30
)
annotation_data  <- transfer_label(
  seurat_obj = integrated,
  training_set = data_sets$training_set,
  test_set = data_sets$test_set,
  k = 15,
  response_var = "annotation_paper"
)
umap_coords <- transfer_umap_coords(
  seurat_obj = integrated,
  training_set = data_sets$training_set,
  test_set = data_sets$test_set,
  k = 15,
  umap1_var = "UMAP1",
  umap2_var = "UMAP2"
)
integrated$UMAP1[umap_coords$query_cells] <- umap_coords$UMAP1
integrated$UMAP2[umap_coords$query_cells] <- umap_coords$UMAP2
integrated@reductions$umap@cell.embeddings[, "UMAP_1"] <- integrated$UMAP1
integrated@reductions$umap@cell.embeddings[, "UMAP_2"] <- integrated$UMAP2
integrated$annotation_paper[annotation_data$query_cells] <- annotation_data$annotation
integrated$annotation_prob <- NA
integrated$annotation_prob[annotation_data$query_cells] <- annotation_data$annotation_prob
DimPlot(integrated, group.by = "annotation_paper", split.by = "assay", cols = color_palette)
```


# Find markers

```{r}
# Split
integrated_rna_multi <- subset(integrated, assay != "CITEseq")
integrated_cite_seq <- subset(integrated, assay == "CITEseq")


# Markers protein
Idents(integrated_cite_seq) <- "annotation_paper"
DefaultAssay(integrated_cite_seq) <- "ADT"
sorted_levels <- c("Naive", "CM Pre-non-Tfh", "CM PreTfh", "T-Trans-Mem",
                    "T-Eff-Mem", "T-helper", "Tfh T:B border", "Tfh-LZ-GC",
                    "GC-Tfh-SAP", "GC-Tfh-0X40", "Tfh-Mem", "Eff-Tregs", 
                    "non-GC-Tf-regs", "GC-Tf-regs")
markers_prot <- FindAllMarkers(
  integrated_cite_seq,
  only.pos = TRUE
)
markers_prot_dfs <- purrr::map(sorted_levels, function(x) {
  df <- markers_prot[markers_prot$cluster == x, ]
  df
})
names(markers_prot_dfs) <- sorted_levels
```


# Save

```{r}
# Seurat object
saveRDS(integrated_cite_seq, path_to_save_cite_seq)


# UMAP
umap_cite_seq <- DimPlot(integrated_cite_seq, cols = color_palette)
ggsave(
  filename = path_to_save_umap_cite_seq,
  plot = umap_cite_seq,
  width = 14,
  height = 12,
  units = "cm"
)


# Markers
names(markers_prot_dfs)[names(markers_prot_dfs) == "Tfh T:B border"] <- "Tfh T-B border"
openxlsx::write.xlsx(
  x = markers_prot_dfs,
  file = path_to_save_prot_markers,
  overwrite = TRUE
)
```


# Session Information

```{r}
sessionInfo()
```

