---
author: "Ramon Massoni-Badosa"
date: "2021/03/15"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
title <- paste(cell_type, ": Compositional analysis (age and sex)", sep = "")
```

---
title: "`r title`"
---


# Introduction

In this notebook, we will perform a basic compositional analysis to find the difference in cell type composition across age and sex groups. We will consider the clusters defined at level 3.


# Pre-processing

## Load packages

```{r}
library(Seurat)
library(tidyverse)
library(gridExtra)
library(cowplot)
```


## Parameters

```{r}
# Paths
path_to_level_3 <- here::here("scRNA-seq/results/R_objects/level_3/")
path_to_level_3_cell_type <- str_c(path_to_level_3, cell_type, sep = "")
path_to_obj <- str_c(
  path_to_level_3_cell_type,
  "/",
  cell_type,
  "_clustered_level_3.rds",
  sep = ""
)
path_to_tmp <- here::here("scRNA-seq/4-compositional_analysis/tmp/")
path_to_tmp2 <- str_c(
  path_to_tmp,
  cell_type,
  sep = ""
)
dir.create(path_to_tmp, showWarnings = FALSE)
dir.create(path_to_tmp2, showWarnings = FALSE)
path <- str_c(
  path_to_tmp2,
  "/",
  cell_type,
  sep = ""
)
path_to_save_age <- str_c(path, "_age_driven_compositional_analysis.pdf")
path_to_save_sex <- str_c(path, "_sex_driven_compositional_analysis.pdf")
path_to_annotation <- str_c(here::here("annotation/level_3/"), cell_type, sep = "")
path_to_save_age_annotation <- str_c(path_to_annotation, "/", cell_type, "_age_driven_compositional_analysis.pdf")
path_to_save_sex_annotation <- str_c(path_to_annotation, "/", cell_type, "_sex_driven_compositional_analysis.pdf")


# Point sizes
pt_sizes <- c(
  NBC_MBC = 0.15,
  GCBC = 0.15,
  CD4_T = 0.25,
  Cytotoxic = 0.4,
  PC = 0.5,
  myeloid = 0.6,
  FDC = 0.6,
  PDC = 1,
  epithelial = 1
)
pt_size <- pt_sizes[cell_type]


# Colors
color_palette <- c("black", "gray", "red", "yellow", "plum4", "green4",
                   "blue", "mediumorchid2", "coral2", "blueviolet",
                   "indianred4", "deepskyblue1", "dimgray", "deeppink1",
                   "green", "lightgray", "hotpink1", "gold", "brown",
                   "mediumvioletred", "mediumaquamarine")


# Source functions
source(here::here("scRNA-seq/bin/utils.R"))
```


## Load data

```{r}
seurat <- readRDS(path_to_obj)
seurat
```


# Age

## UMAPs

```{r fig.wide=TRUE}
seurat$age_group <- factor(
  seurat$age_group,
  levels = c("kid", "young_adult", "old_adult")
)
umap_age <- plot_split_umap(seurat_obj = seurat, var = "age_group")
umap_age
```


## Stacked bar plot 1

```{r}
proportions_df_age1 <- find_proportions_df(
  seurat_obj = seurat,
  x = "age_group",
  fill = "seurat_clusters"
)
DT::datatable(proportions_df_age1)
stacked_barplot_age1 <- plot_stacked_barplot(
  proportions_df_age1,
  x = "age_group",
  fill = "seurat_clusters",
  colors = color_palette
)
stacked_barplot_age1
```


## Stacked bar plot 2

```{r}
proportions_df_age2 <- find_proportions_df(
  seurat_obj = seurat,
  x = "seurat_clusters",
  fill = "age_group"
)
DT::datatable(proportions_df_age2)
stacked_barplot_age2 <- plot_stacked_barplot(
  proportions_df_age2,
  x = "seurat_clusters",
  fill = "age_group",
  colors = color_palette
)
stacked_barplot_age2
```


# Sex

## UMAPs

```{r fig.wide=TRUE}
umap_sex <- plot_split_umap(seurat_obj = seurat, var = "sex")
umap_sex
```


## Stacked bar plot 1

```{r}
proportions_df_sex1 <- find_proportions_df(
  seurat_obj = seurat,
  x = "sex",
  fill = "seurat_clusters"
)
DT::datatable(proportions_df_sex1)
stacked_barplot_sex1 <- plot_stacked_barplot(
  proportions_df_sex1,
  x = "sex",
  fill = "seurat_clusters",
  colors = color_palette
)
stacked_barplot_sex1
```


## Stacked bar plot 2

```{r}
proportions_df_sex2 <- find_proportions_df(
  seurat_obj = seurat,
  x = "seurat_clusters",
  fill = "sex"
)
DT::datatable(proportions_df_sex2)
stacked_barplot_sex2 <- plot_stacked_barplot(
  proportions_df_sex2,
  x = "seurat_clusters",
  fill = "sex",
  colors = color_palette
)
stacked_barplot_sex2
```


# Save

```{r}
# Age
age_plot <- plot_grid(
  umap_age,
  stacked_barplot_age1,
  stacked_barplot_age2,
  nrow = 3,
  ncol = 1,
  rel_heights = c(1, 1, 1)
)
ggsave(
  filename = path_to_save_age,
  plot = age_plot,
  height = 27,
  width = 19,
  units = "cm"
)
ggsave(
  filename = path_to_save_age_annotation,
  plot = age_plot,
  height = 27,
  width = 19,
  units = "cm"
)


# Sex
sex_plot <- plot_grid(
  umap_sex,
  stacked_barplot_sex1,
  stacked_barplot_sex2,
  nrow = 3,
  ncol = 1,
  rel_heights = c(1, 1, 1)
)
ggsave(
  filename = path_to_save_sex,
  plot = sex_plot,
  height = 29,
  width = 21,
  units = "cm"
)
ggsave(
  filename = path_to_save_sex_annotation,
  plot = sex_plot,
  height = 29,
  width = 21,
  units = "cm"
)
```


# Session Information

```{r}
sessionInfo()
```

