---
title: "Merge coordinates and annotation for figure 1"
author: "Ramon Massoni-Badosa"
date: "2021/08/01"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```



# Introduction

The purpose of this notebook is to define a dataframe with two column (cell barcodes and annotation) of all cells and clusters that we will include in the final dataset (cluster freeze). The cells to include come from three different sources:

1. Cells profiled with scRNA-seq or multiome cells from the objects in advanced levels of the analysis for overrepresented cell types (T, B).
2. Cells profiled with scRNA-seq from the objects in advanced levels of the analysis for underrepresented cell types (non-lymphoid). Remember that for these cells, the integration between multiome and RNA was masking biological heterogeneity, so we decided to just focus on scRNA-seq for the annotation and transfer the labels later on.
3. Cells profiled with multiome from the objects in initial levels of the analysis for underrepresented cell types (non-lymphoid).

We will fetch the UMAP coords and annotation for (1) and (2) from the Seurat objects we saved in the previous notebook (00-harmonize_seurat_objects.Rmd). For the third case, we will upload intermediate Seurat objects and early levels of the analysis.


## Load packages

```{r}
library(Seurat)
library(harmony)
library(tidyverse)
library(here)
```


## Define parameters

All the parameters (paths, colors, variables) are encoded in the file "utils_final_clusters.R":

```{r}
source(here("scRNA-seq/bin/utils_final_clusters.R"))
```


# Non-lymphoid


## Myeloid


```{r}
# Read
myeloid <- readRDS(path_to_save_myeloid)
myeloid_multi <- readRDS(path_to_multiome_myeloid)
DimPlot(myeloid, group.by = "annotation_20220215", cols = color_palette)
DimPlot(myeloid, group.by = "annotation_figure_1", cols = color_palette)
table(myeloid_multi$assay)
DimPlot(myeloid_multi, cols = color_palette, split.by = "assay")


# Set up dataframe to plot figure 1
myeloid_multi <- subset(myeloid_multi, assay == "multiome")
myeloid_multi$barcode <- colnames(myeloid_multi)
myeloid_multi$annotation_figure_1 <- NA
myeloid_multi$annotation_20220215 <- "myeloid_multiome"
myeloid_multi$UMAP_1_20220215 <- NA
myeloid_multi$UMAP_2_20220215 <- NA
fig1_df <- myeloid@meta.data
condition1 <- all(selected_cols %in% colnames(myeloid_multi@meta.data))
condition2 <- all(myeloid_multi$assay == "multiome")
if (condition1 & condition2) {
  fig1_df <- bind_rows(fig1_df, myeloid_multi@meta.data[, selected_cols])
} else {
  stop("Conditions not met!")
}


# Remove
rm(myeloid, myeloid_multi)
```


## Epithelial

```{r}
# Read
epithelial <- readRDS(path_to_save_epithelial)
epithelial_multi <- readRDS(path_to_multiome_epithelial)
DimPlot(epithelial, group.by = "annotation_20220215", cols = color_palette)
DimPlot(epithelial, group.by = "annotation_figure_1", cols = color_palette)
table(epithelial_multi$assay)
DimPlot(epithelial_multi, cols = color_palette, split.by = "assay")


# Set up dataframe to plot figure 1
epithelial_multi <- subset(epithelial_multi, assay == "multiome")
epithelial_multi$barcode <- colnames(epithelial_multi)
epithelial_multi$annotation_figure_1 <- "epithelial"
epithelial_multi$annotation_20220215 <- "epithelial_multiome"
epithelial_multi$UMAP_1_20220215 <- NA
epithelial_multi$UMAP_2_20220215 <- NA
condition1 <- all(selected_cols %in% colnames(epithelial_multi@meta.data))
condition2 <- all(epithelial_multi$assay == "multiome")
if (condition1 & condition2) {
  fig1_df <- bind_rows(
    fig1_df,
    epithelial@meta.data,
    epithelial_multi@meta.data[, selected_cols]
  )
} else {
  stop("Conditions not met!")
}


# Remove
rm(epithelial, epithelial_multi)
```


## PDC

```{r}
# Read
pdc <- readRDS(path_to_save_pdc)
pdc_multi <- readRDS(path_to_multiome_pdc)
DimPlot(pdc, group.by = "annotation_20220215", cols = color_palette)
DimPlot(pdc, group.by = "annotation_figure_1", cols = color_palette)
DimPlot(pdc_multi, cols = color_palette, split.by = "assay")
table(pdc$hospital)
table(pdc_multi$hospital)
pdc <- subset(pdc, hospital != "Royal London")
pdc_multi <- subset(pdc_multi, hospital != "Royal London")


# Set up dataframe to plot figure 1
pdc_multi <- subset(pdc_multi, assay == "multiome")
pdc_multi$barcode <- colnames(pdc_multi)
pdc_multi$annotation_figure_1 <- "PDC"
pdc_multi$annotation_20220215 <- "PDC_multiome"
pdc_multi$UMAP_1_20220215 <- NA
pdc_multi$UMAP_2_20220215 <- NA
condition1 <- all(selected_cols %in% colnames(pdc_multi@meta.data))
condition2 <- all(pdc_multi$assay == "multiome")
if (condition1 & condition2) {
  fig1_df <- bind_rows(
    fig1_df,
    pdc@meta.data,
    pdc_multi@meta.data[, selected_cols]
  )
} else {
  stop("Conditions not met!")
}


# Remove
rm(pdc, pdc_multi)
```


## FDC

```{r}
# Read
fdc <- readRDS(path_to_save_fdc)
fdc_multi <- readRDS(path_to_multiome_fdc)
fdc_proliferative <- readRDS(path_to_fdc_proliferative)
DimPlot(fdc, group.by = "annotation_20220215", cols = color_palette)
DimPlot(fdc, group.by = "annotation_figure_1", cols = color_palette)
table(fdc_multi$assay)
DimPlot(fdc_multi, cols = color_palette, split.by = "assay")
table(fdc_proliferative$assay)
table(fdc$hospital)
fdc <- subset(fdc, hospital != "Royal London")


# Set up dataframe to plot figure 1
fdc_multi <- subset(fdc_multi, assay == "multiome")
fdc_multi$barcode <- colnames(fdc_multi)
fdc_multi$annotation_figure_1 <- "FDC"
fdc_multi$annotation_20220215 <- "FDC_multiome"
fdc_proliferative$annotation_figure_1 <- "cycling FDC"
fdc_proliferative$annotation_20220215 <- "cycling FDC"
fdc_multi$UMAP_1_20220215 <- NA
fdc_multi$UMAP_2_20220215 <- NA
fdc_proliferative$UMAP_1_20220215 <- NA
fdc_proliferative$UMAP_2_20220215 <- NA
condition1 <- all(selected_cols %in% colnames(fdc_multi@meta.data))
condition2 <- all(fdc_multi$assay == "multiome")
condition3 <- all(selected_cols %in% colnames(fdc_proliferative@meta.data))
if (condition1 & condition2 & condition3) {
  fig1_df <- bind_rows(
    fig1_df,
    fdc@meta.data,
    fdc_multi@meta.data[, selected_cols],
    fdc_proliferative@meta.data[, selected_cols],
  )
} else {
  stop("Conditions not met!")
}


# Remove
rm(fdc, fdc_multi, fdc_proliferative)
```


# Precursor lymphocytes

## preB

Note that, because we have very few preB cells, we couldn't stritify them into meaningful clusters. Thus, all of them are labeled as "preB":

```{r}
# Read
preB <- readRDS(path_to_save_preB)
DimPlot(preB, group.by = "annotation_20220215", cols = color_palette)
DimPlot(preB, group.by = "annotation_figure_1", cols = color_palette)
table(preB$assay)
table(preB$hospital)
DimPlot(preB, cols = color_palette, split.by = "hospital")
preB <- subset(preB, hospital != "Royal London")


# Set up dataframe to plot figure 1
condition1 <- all(selected_cols %in% colnames(preB@meta.data))
if (condition1) {
  fig1_df <- bind_rows(fig1_df, preB@meta.data[, selected_cols])
} else {
  stop("Conditions not met!")
}


# Remove
rm(preB)
```


## preT

```{r}
# Read
preT <- readRDS(path_to_save_preT)
DimPlot(preT, group.by = "annotation_20220215", cols = color_palette)
DimPlot(preT, group.by = "annotation_figure_1", cols = color_palette)
table(preT$assay)
table(preT$hospital)
DimPlot(preT, cols = color_palette, split.by = "hospital")
preT <- subset(preT, hospital != "Royal London")


# Set up dataframe to plot figure 1
condition1 <- all(selected_cols %in% colnames(preT@meta.data))
if (condition1) {
  fig1_df <- bind_rows(fig1_df, preT@meta.data[, selected_cols])
} else {
  stop("Conditions not met!")
}


# Remove
rm(preT)
```


# NK/ILC

All the cell types hereafter included multiome cells, so we only load the last object with all cells.

```{r}
# Read
ilc_nk <- readRDS(path_to_save_ilc_nk)
DimPlot(ilc_nk, group.by = "annotation_20220215", cols = color_palette)
DimPlot(ilc_nk, group.by = "annotation_figure_1", cols = color_palette)
table(ilc_nk$assay)
table(ilc_nk$hospital)
DimPlot(ilc_nk, cols = color_palette, split.by = "hospital")


# Set up dataframe to plot figure 1
condition1 <- all(selected_cols %in% colnames(ilc_nk@meta.data))
if (condition1) {
  fig1_df <- bind_rows(fig1_df, ilc_nk@meta.data[, selected_cols])
} else {
  stop("Conditions not met!")
}


# Remove
rm(ilc_nk)
```


# CD8 T cells

```{r}
# Read
cd8 <- readRDS(path_to_save_cd8)
DimPlot(cd8, group.by = "annotation_20220215", cols = color_palette)
DimPlot(cd8, group.by = "annotation_figure_1", cols = color_palette)
table(cd8$assay)
table(cd8$hospital)
DimPlot(cd8, cols = color_palette, split.by = "hospital")


# Set up dataframe to plot figure 1
condition1 <- all(selected_cols %in% colnames(cd8@meta.data))
if (condition1) {
  fig1_df <- bind_rows(fig1_df, cd8@meta.data[, selected_cols])
} else {
  stop("Conditions not met!")
}


# Remove
rm(cd8)
```


# CD4 T cells

```{r}
# Read
cd4 <- readRDS(path_to_save_cd4)
cd4_proli <- readRDS(path_to_proliferative_cd4_t)
DimPlot(cd4, group.by = "annotation_20220215", cols = color_palette)
DimPlot(cd4, group.by = "annotation_figure_1", cols = color_palette)
table(cd4$assay)
table(cd4$hospital)
DimPlot(cd4, cols = color_palette, split.by = "hospital")
table(cd4_proli$assay)
table(cd4_proli$hospital)
DimPlot(cd4_proli, cols = color_palette, split.by = "hospital")


# Set up dataframe to plot figure 1
cd4_proli$annotation_figure_1 <- "cycling T"
cd4_proli$annotation_20220215 <- "cycling T"
cd4_proli$UMAP_1_20220215 <- NA
cd4_proli$UMAP_2_20220215 <- NA
condition1 <- all(selected_cols %in% colnames(cd4@meta.data))
condition2 <- all(selected_cols %in% colnames(cd4_proli@meta.data))
if (condition1 & condition2) {
  fig1_df <- bind_rows(
    fig1_df,
    cd4@meta.data,
    cd4_proli@meta.data[, selected_cols]
  )
} else {
  stop("Conditions not met!")
}


# Remove
rm(cd4, cd4_proli)
```


# Plasma cells (PC)

The challenge with all the Seurat objects coming from the B cell lineage is that it also includes cells that belong to other cell types/states. For instance in the PC object, we have cells from GCBC and NBC/MBC. Hence, we will subset them to avoid duplication:


```{r}
# Read
pc <- readRDS(path_to_save_pc)
DimPlot(pc, group.by = "annotation_20220215", cols = color_palette)
DimPlot(pc, group.by = "annotation_figure_1", cols = color_palette)
pc <- subset(pc, annotation_figure_1 == "PC")
table(pc$assay)
table(pc$hospital)
DimPlot(pc, cols = color_palette, split.by = "hospital")


# Set up dataframe to plot figure 1
condition1 <- all(selected_cols %in% colnames(pc@meta.data))
if (condition1) {
  fig1_df <- bind_rows(fig1_df, pc@meta.data)
} else {
  stop("Conditions not met!")
}


# Remove
rm(pc)
```


# NBC/MBC

```{r}
# Read
nbc_mbc <- readRDS(path_to_save_nbc_mbc)
DimPlot(nbc_mbc, group.by = "annotation_20220215", cols = color_palette)
DimPlot(nbc_mbc, group.by = "annotation_figure_1", cols = color_palette)
nbc_mbc <- subset(nbc_mbc, annotation_level_1 == "NBC_MBC")
table(nbc_mbc$assay)
table(nbc_mbc$hospital)
DimPlot(nbc_mbc, cols = color_palette, split.by = "hospital")


# Set up dataframe to plot figure 1
condition1 <- all(selected_cols %in% colnames(nbc_mbc@meta.data))
if (condition1) {
  fig1_df <- bind_rows(fig1_df, nbc_mbc@meta.data)
} else {
  stop("Conditions not met!")
}


# Remove
rm(nbc_mbc)
```


# GCBC

```{r}
# Read
gcbc <- readRDS(path_to_save_gcbc)
DimPlot(gcbc, group.by = "annotation_20220215", cols = color_palette)
DimPlot(gcbc, group.by = "annotation_figure_1", cols = color_palette)
table(gcbc$assay)
table(gcbc$hospital)
DimPlot(gcbc, cols = color_palette, split.by = "hospital")


# Set up dataframe to plot figure 1
condition1 <- all(selected_cols %in% colnames(gcbc@meta.data))
if (condition1) {
  fig1_df <- bind_rows(fig1_df, gcbc@meta.data)
} else {
  stop("Conditions not met!")
}


# Remove
rm(gcbc)
```


# Sanity checks

```{r}
# Duplicated cells?
any(duplicated(fig1_df$barcode))


# Lingering cells from King et al.?
any(fig1_df$hospital == "Royal London")
```


# Save

```{r}
saveRDS(fig1_df, path_to_save_umap_df_rds)
write_delim(
  fig1_df,
  file = path_to_save_umap_df_csv,
  delim = ";",
  col_names = TRUE
)
```


# Session information

```{r}
sessionInfo()
```

