---
title: "Precursor T cells analysis"
author: "Ramon Massoni-Badosa"
date: "`r format(Sys.time(), '%B %Y')`"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```


# Introduction

The objective of this notebook is to reanalyze the precursor T cells to validate their annotation. We will analyze them jointly with the precursor B cells and plasmacytoid dendritic cells (PDC). The 3 cell types clustered together in level 1.


# Pre-processing

## Load packages

```{r}
library(Seurat)
library(tidyverse)
library(here)
library(pals)
library(DT)
```


## Define parameters

```{r}
path_to_preT <- here("scRNA-seq/results/R_objects/final_clusters/20220215/20220215_preT_seurat_obj.rds")
path_to_preB <- here("scRNA-seq/results/R_objects/final_clusters/20220215/20220215_preB_seurat_obj.rds")
path_to_pdc <- here("scRNA-seq/results/R_objects/final_clusters/20220215/20220215_PDC_seurat_obj.rds")
path_to_save <- here("data/raw_data_figures/seurat_obj_pdc_preB_preT_fig1.rds")

cols <- pals::polychrome()
names(cols) <- NULL
```


## Read data

```{r}
preT <- readRDS(path_to_preT)
preB <- readRDS(path_to_preB)
pdc <- readRDS(path_to_pdc)
```


# Merge and reprocess

```{r}
merged <- merge(x = preT, y = c(preB, pdc))
merged <- merged %>%
  FindVariableFeatures() %>%
  ScaleData() %>%
  RunPCA() %>%
  RunUMAP(dims = 1:15, reduction = "pca")
DimPlot(merged, group.by = "annotation_figure_1", cols = cols)
DimPlot(merged, group.by = "annotation_20220215", cols = cols)
```


# Markers

```{r}
# Markers all
Idents(merged) <- "annotation_20220215"
markers_all <- FindAllMarkers(merged, only.pos = TRUE)
markers_all <- markers_all[markers_all$p_val_adj < 0.001, ]
DT::datatable(markers_all, options = list(scrollX = TRUE))


# Precursor-specific markers
markers_precursor <- FindMarkers(
  merged,
  ident.1 = c("preT", "preB"),
  only.pos = TRUE
)
markers_precursor <- markers_precursor[markers_precursor$p_val_adj < 0.001, ]
DT::datatable(markers_precursor, options = list(scrollX = TRUE))


# Interferon+ PDC-specific genes
markers_ifn <- FindMarkers(
  merged,
  ident.1 = "IFN1+ PDC",
  ident.2 = "PDC",
  only.pos = TRUE
)
markers_ifn <- markers_ifn[markers_ifn$p_val_adj < 0.001, ]
DT::datatable(markers_ifn, options = list(scrollX = TRUE))
```


# Save

```{r}
saveRDS(merged, path_to_save)
```


# Session Information

```{r}
sessionInfo()
```

