---
title: 'GCBC: scATAC annotation level 4'
author: "Paula Soler-Vila"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```

# Introduction

The objective of this notebook is to refine the clustering annotation done at level 3. This refinement is the result of a manual curation carried out by specialists to remove poor quality cells, misclassified cells or clusters with very few cells.

# Pre-processing

## Load packages

```{r}
library(Seurat)
library(Signac)
library(tidyverse)
library(reshape2)
library(ggpubr)
library(harmony)
library(plotly)

set.seed(1234)
```


## Parameters
```{r}
cell_type = "GCBC"

# Paths
path_to_obj <- str_c(
  here::here("scATAC-seq/results/R_objects/level_3/"),
  cell_type,
  "/",
  cell_type,
  "_integrated_level_3.rds",
  sep = ""
)

path_to_obj_RNA <- str_c(
  here::here("scRNA-seq/3-clustering/4-level_4/"),
  cell_type,
    "/",
  cell_type,
  "_clustered_clean_level_4_with_final_clusters.rds",
  sep = ""
)

# Functions
source(here::here("scRNA-seq/bin/utils.R"))


# Colors
color_palette <-  c("#1CFFCE", "#90AD1C", "#C075A6", "#85660D", 
                    "#5A5156", "#AA0DFE", "#F8A19F", "#F7E1A0", 
                    "#1C8356", "#FEAF16", "#822E1C", "#C4451C", 
                    "#1CBE4F", "#325A9B", "#F6222E", "#FE00FA", 
                    "#FBE426", "#16FF32", "black", "#3283FE",
                    "#B00068", "#DEA0FD", "#B10DA1", "#E4E1E3",
                    "#90AD1C", "#FE00FA", "#85660D", "#3B00FB",
                    "#822E1C", "coral2", "#1CFFCE", "#1CBE4F",
                    "#3283FE", "#FBE426", "#F7E1A0", "#325A9B",   
                    "#2ED9FF", "#B5EFB5", "#5A5156", "#DEA0FD",
                    "#FEAF16", "#683B79", "#B10DA1", "#1C7F93", 
                    "#F8A19F", "dark orange", "#FEAF16", "#FBE426",  
                    "Brown")

path_to_level_4 <- here::here("scATAC-seq/results/R_objects/level_4/GCBC/")
path_to_save <- str_c(path_to_level_4, "GCBC_integrated_level_4.rds")
```

## Load data
```{r}
# Seurat object
seurat <- readRDS(path_to_obj)
seurat

seurat_RNA <- readRDS(path_to_obj_RNA)
seurat_RNA
```

## Visualization of the data
```{r fig.wide=TRUE}
DimPlot(seurat,
  pt.size = 0.1, split.by = "assay") 

DimPlot(seurat,
  pt.size = 0.1, split.by = "age_group") 

p1 <- DimPlot(seurat,
  pt.size = 0.1) + NoLegend()

p2 <- DimPlot(seurat_RNA,
  pt.size = 0.1,cols = color_palette)

p1 + p2
```

# Spot potential doublets & poor-quality cells from Multiome experiment

## From scRNA assay
```{r}
tonsil_RNA_annotation <- seurat_RNA@meta.data %>%
  rownames_to_column(var = "cell_barcode") %>%
  dplyr::filter(assay == "multiome") %>%
  dplyr::select("cell_barcode", "cluster8_subcluster")
```

## From scATAC assay
```{r}
tonsil_ATAC_cell_barcode <- seurat@meta.data %>%
  rownames_to_column(var = "cell_barcode") %>%
  dplyr::filter(assay == "multiome") %>%
  dplyr::select("cell_barcode")
```

```{r}
possible_doublets_ATAC <- setdiff(tonsil_ATAC_cell_barcode$cell_barcode,tonsil_RNA_annotation$cell_barcode)
seurat$quality_cells <- ifelse(colnames(seurat) %in% possible_doublets_ATAC, "Poor-quality", "Good-quality")
```

###  Visualize the projection of the problematic cells onto the scATAC-seq UMAP

```{r fig.wide=TRUE}
DimPlot(
  seurat, 
  split.by = "quality_cells")
```


## Scrublet prediction

```{r}
DimPlot(seurat,split.by = "scrublet_predicted_doublet_atac")
table(seurat$scrublet_predicted_doublet_atac)
```

## QC metrics

```{r}
qc_vars <- c(
  "nCount_peaks",
  "nFeature_peaks",
  "nucleosome_signal",
  "TSS.enrichment"
)
qc_gg <- purrr::map(qc_vars, function(x) {
  p <- FeaturePlot(seurat, features = x, max.cutoff = "q95")
  p
})
qc_gg
```

## Chromatin Signature

```{r}
qc_vars <- c("NBC.MBC", "GCBC", "PC", "CD4.T", "Cytotoxic")
qc_gg <- purrr::map(qc_vars, function(x) {
  p <- FeaturePlot(seurat, feature = x, 
                   max.cutoff = 4, min.cutoff = -4) + 
    scale_color_viridis_c(option = "magma")
  p
})
qc_gg
```

# Clustering 

```{r}
resolutions <- c(0.1, 0.12, 0.2, 0.3 , 0.5, 0.75)
seurat <- FindClusters(seurat, resolution = resolutions)
vars <- str_c("peaks_macs_snn_res.", resolutions)
umap_clusters <- purrr::map(vars, function(x) {
  p <- DimPlot(seurat, group.by = x, cols = color_palette)
  p
})
umap_clusters
```


# 3D representation of the clusters
```{r}
# Re-run UMAPs that you have accurate calculations for all UMAP(s)
yourseuratobject <- RunUMAP(reduction = 'harmony', seurat,
                            dims = 2:40,  
                            n.components = 3L)

# Extract tSNE information from Seurat Object
umap_1 <- yourseuratobject[["umap"]]@cell.embeddings[,1]
umap_2 <- yourseuratobject[["umap"]]@cell.embeddings[,2]
umap_3 <- yourseuratobject[["umap"]]@cell.embeddings[,3]

# Prepare a dataframe for cell plotting
plot.data <- FetchData(object = yourseuratobject, vars = c("UMAP_1", "UMAP_2", "UMAP_3", "peaks_macs_snn_res.0.5"))

# Make a column of row name identities (these will be your cell/barcode names)
plot.data$label <- paste(rownames(plot.data))

fig <- plot_ly(data = plot.data, 
        x = ~UMAP_1, y = ~UMAP_2, z = ~UMAP_3, 
        color = ~peaks_macs_snn_res.0.5, 
        colors = c("gray10","gray20","gray30",
                   "gray40","gray50","gray60",
                   "red","gray70","gray80"), 
        type = "scatter3d", 
        mode = "markers", 
        marker = list(size = 3, width=2), # controls size of points
        text=~label, #This is that extra column we made earlier for which we will use for cell ID
        hoverinfo="text") #When you visualize your plotly object, hovering your mouse pointer over a point shows cell names
fig
```


## Proportion of doublets per cluster based on Scrublet

```{r}
doublet_clusters <- purrr::map(vars, function(x) {
  df1 <- data.frame(table(seurat@meta.data[,x], seurat@meta.data$scrublet_predicted_doublet_atac))
  colnames(df1) <- c("Cluster", "Scrublet","Cells")
  p <- ggbarplot(df1, "Cluster", "Cells",
  fill = "Scrublet", color = "Scrublet",
  label = TRUE,
  position = position_dodge(0.9))
  p
})
doublet_clusters
```

## Proportion of doublets per cluster based on Multiome

```{r}
doublet_clusters <- purrr::map(vars, function(x) {
  df1 <- data.frame(table(seurat@meta.data[,x], seurat$quality_cells))
  colnames(df1) <- c("Cluster", "Quality","Cells")
  p <- ggbarplot(df1, "Cluster", "Cells",
  fill = "Quality",
  label = TRUE,
  position = position_dodge(0.9))
  p
})
doublet_clusters
```

## UMAP level 1

```{r}
umap_clusters_level1 <- purrr::map(vars, function(x) {
  p <- FeatureScatter(seurat, 
                      "UMAP_1_level_1",
                      "UMAP_2_level_1", group.by = x, cols = color_palette)
  p
})
umap_clusters_level1
```


## UMAP level 1

```{r}
umap_level_1 <- FeatureScatter(
  seurat,
  "UMAP_1_level_1",
  "UMAP_2_level_1",
  group.by = "annotation_level_1"
)
umap_level_1 <- umap_level_1 +
  theme(
    #legend.position = "none",
    plot.title = element_blank()
  )
umap_level_1
```

## Visualization before removing problematic cells

```{r}
cluster4 <- subset(seurat, peaks_macs_snn_res.0.3 == 4)
DimPlot(seurat, cells.highlight = colnames(cluster4))
```

# Exclude problematic clusters and poor quality cells.

```{r}
seurat <- subset(seurat, 
                 quality_cells == "Good-quality" &
                 peaks_macs_snn_res.0.3 != 4)
seurat
```

## Visualization after removing problematic cells
```{r fig.wide=TRUE}
DimPlot(seurat)
```

# Integration

```{r fig.wide=TRUE}
seurat <- seurat %>%
  RunTFIDF() %>%
  FindTopFeatures(min.cutoff = 10) %>%
  RunSVD()

#DepthCor(seurat)
#seurat <- RunUMAP(object = seurat, reduction = 'lsi', dims = 2:30)
#DimPlot(seurat)

seurat <- RunHarmony(
  object = seurat,
  dims = 2:30,
  group.by.vars = 'gem_id',
  reduction = 'lsi',
  assay.use = 'peaks_macs',
  project.dim = FALSE,
  max.iter.harmony = 20
)

seurat <- RunUMAP(seurat, reduction = "harmony", dims = 2:30)

DimPlot(seurat,  split.by = "assay")
DimPlot(seurat,  split.by = "age_group")
```



# Save
```{r}
saveRDS(seurat, path_to_save)
```

# Session Information

```{r}
sessionInfo()
```
