---
title: "Normalization and Batch Effect Correction after the new Peak calling"
author: "Paula Soler-Vila"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```


# Introduction

The new peak calling round applied on the previous notebook significantly increased the number of the features identified in our dataset. Therefore, we must need to repeat the standard downstream analysis, including data normalization, dimensionality reduction analysis and batch correction to account for this change in the number of features detected.


# Pre-processing

## Load packages

```{r}
library(Seurat)
library(SeuratWrappers)
library(Signac)
library(harmony)
library(tidyverse)
library(plotly)
library(plyr)

set.seed(333)
```

## Parameters

```{r parameters}
cell_type = "GCBC"

color_palette <-  c("#1CFFCE", "#90AD1C", "#C075A6", "#85660D", 
                    "#5A5156", "#AA0DFE", "#F8A19F", "#F7E1A0", 
                    "#1C8356", "#FEAF16", "#822E1C", "#C4451C", 
                    "#1CBE4F", "#325A9B", "#F6222E", "#FE00FA", 
                    "#FBE426", "#16FF32", "black", "#3283FE",
                    "#B00068", "#DEA0FD", "#B10DA1", "#E4E1E3",
                    "#90AD1C", "#FE00FA", "#85660D", "#3B00FB",
                    "#822E1C", "coral2", "#1CFFCE", "#1CBE4F",
                    "#3283FE", "#FBE426", "#F7E1A0", "#325A9B",   
                    "#2ED9FF", "#B5EFB5", "#5A5156", "#DEA0FD",
                    "#FEAF16", "#683B79", "#B10DA1", "#1C7F93", 
                    "#F8A19F", "dark orange", "#FEAF16", "#FBE426",  
                    "Brown")

# Paths
path_to_obj <- paste0(
  here::here("scATAC-seq/results/R_objects/level_4/"),
  cell_type,
  "/",
  cell_type,
  "_annotated_peak_calling_level_4.rds",
  sep = ""
)

path_to_save <- paste0(
  here::here("scATAC-seq/results/R_objects/level_4/"),
  cell_type,
  "/",
  cell_type,
  "_integration_peak_calling_level_4.rds",
  sep = ""
)
```

## Load data
```{r}
seurat <- readRDS(path_to_obj)
seurat
```

# Visualize UMAP without batch effect correction
```{r}
# Normalization, dimensionality reduction 
seurat <- seurat %>%
  RunTFIDF() %>% 
  FindTopFeatures(min.cutoff = 10) %>%
  RunSVD() %>%
  RunUMAP(reduction = "lsi", dims = 2:30)

DepthCor(seurat)
```

```{r}
DimPlot(seurat, 
        cols = color_palette,
        group.by = "level_5", 
        pt.size = 0.1)
```

```{r}
# Visualize UMAP's confounders
confounders <- c("library_name", "sex", "age_group", "hospital", "assay")
umaps_before_integration <- purrr::map(confounders, function(x) {
  p <- DimPlot(seurat, group.by = x, pt.size = 0.1)
  p
})
names(umaps_before_integration) <- confounders
print("UMAP colored by library, sex, age group, cell hashing status, sampling center and assay:")
umaps_before_integration[1:length(umaps_before_integration)]
```

# Run and visualize Harmony’s integration
```{r}
seurat <- RunHarmony(
  object = seurat,
  dims = 2:30,
  group.by.vars = 'gem_id',
  reduction = 'lsi',
  assay.use = 'peaks_redefined',
  project.dim = FALSE,
  max.iter.harmony = 20
)
    
# Non-linear dimension reduction and clustering
seurat <- RunUMAP(seurat, dims = 2:30, reduction = 'harmony')
```

```{r}
DimPlot(seurat, 
    cols = color_palette,
    group.by = "level_5", 
    pt.size = 0.1)
```

```{r}
DimPlot(seurat, 
    cols = c("darkblue","darkblue","darkblue",
               "darkblue","lightblue","lightblue",
               "red","red","red","gray","gray"),
   group.by  = "level_5", 
    pt.size = 0.1)
```

```{r}
# Visualize UMAP's confounders
umaps_after_integration <- purrr::map(confounders, function(x) {
  p <- DimPlot(seurat, group.by = x, pt.size = 0.1)
  p
})
names(umaps_after_integration) <- confounders
print("UMAP colored by library, sex, age group, cell hashing status, sampling center and assay:")
umaps_after_integration[1:length(umaps_before_integration)]
```

```{r}
#
# Re-run UMAPs that you have accurate calculations for all UMAP(s)
seurat <- RunUMAP(reduction = 'harmony', 
                            seurat,
                            dims = 2:30,  
                            n.components = 3L)

# Extract tSNE information from Seurat Object
umap_1 <- seurat[["umap"]]@cell.embeddings[,1]
umap_2 <- seurat[["umap"]]@cell.embeddings[,2]
umap_3 <- seurat[["umap"]]@cell.embeddings[,3]

# Prepare a dataframe for cell plotting
plot.data <- FetchData(object = seurat, vars = c("UMAP_1", "UMAP_2", "UMAP_3", "level_5"))

# Make a column of row name identities (these will be your cell/barcode names)
plot.data$label <- paste(rownames(plot.data))

fig <- plot_ly(data = plot.data, 
        x = ~UMAP_1, y = ~UMAP_2, z = ~UMAP_3, 
        color = ~level_5, 
        colors = c("darkblue","darkblue","darkblue",
                   "darkblue","lightblue","lightblue",
                   "red","red","red","gray","gray"),
        type = "scatter3d", 
        mode = "markers", 
        marker = list(size = 2, width=1), # controls size of points
        text=~label, #This is that extra column we made earlier for which we will use for cell ID
        hoverinfo="text") #When you visualize your plotly object, hovering your mouse pointer over a point shows cell names

fig
```

# Save
```{r}
# Save integrated Seurat object
saveRDS(seurat, path_to_save)
```

# Session Information

```{r}
sessionInfo()
```
