---
title: "Normalization and Batch Effect Correction (multiome)"
author: "Ramon Massoni-Badosa"
date: "2021/01/29"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```


# Introduction

In previous notebooks and scripts, we integrated our dataset with the one of [Hamish et al.](https://immunology.sciencemag.org/content/6/56/eabe6291), and merge it with the RNA fraction of the multiome dataset. Here, we will correct for batch effects with [Harmony](https://www.nature.com/articles/s41592-019-0619-0), and visualize the intermixing of different potential confounders pre- and post-integration.

In addition, we will save the dimensionality reduction (PCA) matrices before and after integration to further quantify the effect of the aforementioned confounders in high dimensional space. 


# Pre-processing

## Load packages

```{r}
library(Seurat)
library(SeuratWrappers)
library(harmony)
library(tidyverse)
```


## Parameters

```{r}
# Paths
path_to_obj <- here::here("scRNA-seq/results/R_objects/tonsil_rna_merged_with_multiome_missing_integration.rds")
# path_to_obj <- here::here("scRNA-seq/results/R_objects/tonsil_rna_merged_with_multiome_missing_integration_downsampled.rds")
path_to_save_obj <- here::here("scRNA-seq/results/R_objects/tonsil_rna_integrated_with_multiome.rds")
path_tmp_dir <- here::here("scRNA-seq/2-QC/5-batch_effect_correction/3-data_integration_multiome/tmp/")
path_to_save_dimred_uncorrect <- str_c(path_tmp_dir, "batch_uncorrected_pca.rds", sep = "")
path_to_save_dimred_correct <- str_c(path_tmp_dir, "batch_corrected_pca.rds", sep = "")
path_to_save_confounders_df <- str_c(path_tmp_dir, "confounders_df.rds", sep = "") 
path_to_save_supp_df <- here::here("data/raw_data_figures/umaps_rna_batch_correction.csv")
```


## Load data

```{r}
tonsil <- readRDS(path_to_obj)
tonsil
```


# Visualize UMAP without batch effect correction

Since we noticed that the 3 matrices (scRNA-seq matrices, King et al., and multiome) have different sets of features (genes), we will find assay-specific variable genes. Then, we will use the intersection to improve the batch-effect correction.

```{r fig.wide=TRUE}
# Find shared highly variable genes (HGV)
tonsil_list <- SplitObject(tonsil, split.by = "assay")
tonsil_list <- purrr::map(tonsil_list, FindVariableFeatures, nfeatures = 5000)
hvg <- purrr::map(tonsil_list, VariableFeatures)
shared_hvg <- Reduce(intersect, hvg)
print(length(shared_hvg))
print(class(shared_hvg))
print(head(shared_hvg))
print(tail(shared_hvg))
rm(tonsil_list)


# Process Seurat object
tonsil <- tonsil%>% 
  NormalizeData(normalization.method = "LogNormalize", scale.factor = 1e4) %>%
  ScaleData(features = shared_hvg) %>% 
  RunPCA(features = shared_hvg) %>%
  RunUMAP(reduction = "pca", dims = 1:30)


# Visualize UMAP
confounders <- c("library_name", "sex", "age_group", "is_hashed",
                 "hospital", "assay")
umaps_before_integration <- purrr::map(confounders, function(x) {
  p <- DimPlot(tonsil, group.by = x, pt.size = 0.1)
  p
})
names(umaps_before_integration) <- confounders
print("UMAP colored by GEM:")
umaps_before_integration$library_name + NoLegend()
print("UMAP colored by sex, age group, cell hashing status, sampling center and assay:")
umaps_before_integration[2:length(umaps_before_integration)]


# Dataframe to generate supplementary figures
unintegrated_df <- data.frame(
  barcode = colnames(tonsil),
  assay = tonsil$assay,
  UMAP_1 = tonsil@reductions$umap@cell.embeddings[, "UMAP_1"],
  UMAP_2 = tonsil@reductions$umap@cell.embeddings[, "UMAP_2"],
  annotation_king = tonsil$cell_type
)
unintegrated_df$is_integrated <- "unintegrated"
```


# Run and visualize Harmony's integration

```{r fig.wide=TRUE}
tonsil <- tonsil %>%
  RunHarmony(reduction = "pca", dims = 1:30, group.by.vars = "gem_id") %>%
  RunUMAP(reduction = "harmony", dims = 1:30)
umaps_after_integration <- purrr::map(confounders, function(x) {
  p <- DimPlot(tonsil, group.by = x, pt.size = 0.1)
  p
})
names(umaps_after_integration) <- confounders
print("UMAP colored by GEM:")
umaps_after_integration$library_name + NoLegend()
print("UMAP colored by sex, age group, cell hashing status, sampling center and assay:")
umaps_after_integration[2:length(umaps_before_integration)]


# Dataframe to generate supplementary figures
integrated_df <- data.frame(
  barcode = colnames(tonsil),
  assay = tonsil$assay,
  UMAP_1 = tonsil@reductions$umap@cell.embeddings[, "UMAP_1"],
  UMAP_2 = tonsil@reductions$umap@cell.embeddings[, "UMAP_2"],
  annotation_king = tonsil$cell_type
)
integrated_df$is_integrated <- "integrated"
supp_df <- bind_rows(list(unintegrated_df, integrated_df))
```


# Save

```{r}
# If it doesn't exist create temporal directory
dir.create(path_tmp_dir, showWarnings = FALSE) 


# Save integrated Seurat object
saveRDS(tonsil, path_to_save_obj)


# Save PCA matrices to compute the Local Inverse Simpson Index (LISI)
confounders_df <- tonsil@meta.data[, confounders]
saveRDS(confounders_df, path_to_save_confounders_df)
saveRDS(
  tonsil@reductions$pca@cell.embeddings[, 1:30],
  path_to_save_dimred_uncorrect
)
saveRDS(
  tonsil@reductions$harmony@cell.embeddings[, 1:30],
  path_to_save_dimred_correct
)


# Save dataframe to plot supplementary
write_delim(supp_df, path_to_save_supp_df, delim = ";", col_names = TRUE)
```


# Session Information

```{r}
sessionInfo()
```
