---
title: "1. Finding co-accessible modules in CD4-T using Cicero"
author: "Paula Soler-Vila"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, out.width = "100%", fig.align = "center",
  message = FALSE, warning = FALSE
)
options(width = 1200)
```

# Objective

The objective of this notebook is to use [Cicero][https://cole-trapnell-lab.github.io/cicero-release/docs_m3/#constructing-cis-regulatory-networks] to build a genome-wide cis-regulatory map.

# Pre-processing

## Load packages

```{r comment=FALSE}
library(Signac)
library(Seurat)
library(SeuratWrappers)
library(ggplot2)
library(patchwork)
library(cicero)
library(EnsDb.Hsapiens.v86)
library(ggpubr)
library(lattice)
library(tidyr)
library(MASS)
library(factoextra)
library(plyr)
library(reshape2)
library(pheatmap)
library(stringr)
```

## Parameters
```{r parameters}
cell_type <- "CD4_T"

path_to_obj <- str_c(
  here::here("scATAC-seq/results/R_objects/level_5/"),
  cell_type,
  "/04.",
  cell_type,
  "_integration_peak_calling_level_5.rds",
  sep = ""
)

color_palette<- c("black", "gray", "red", "yellow", "violet", "green4",
                  "blue", "chocolate1", "coral2", "blueviolet",
                  "brown1", "darkmagenta", "deepskyblue1", "dimgray",
                  "deeppink1", "green", "lightgray", "hotpink1",
                  "indianred4", "khaki", "mediumorchid2")

path_to_save <- str_c(
  here::here("scATAC-seq/results/Cicero/"),
  cell_type,"/",
  sep = ""
)
```

## Load data
Cicero uses the normalized scATAC data integrated by Harmony. 
We are going to start the analysis by visualizing the main clusters defined in level 4 thanks to the scRNAseq that will become our idents.

```{r fig.wide=TRUE}
seurat <- readRDS(path_to_obj)
seurat

Idents(seurat) <- seurat$annotation_paper

DimPlot(
  seurat, 
  cols = color_palette,
  pt.size = 0.1
)
```


# Get annotations

First, we need to extract the lenght of the chromosomes (be aware that we are using from EnsDb the hg38 as the reference human assembly)

```{r}
annotations <- GetGRangesFromEnsDb(ensdb = EnsDb.Hsapiens.v86)
seqlevelsStyle(annotations) <- "UCSC"
genome(annotations) <- "hg38"

# get the chromosome sizes from the Seurat object
genome <- seqlengths(annotations)

# convert chromosome sizes to a dataframe
genome.df <- data.frame("chr" = names(genome), "length" = genome)

# Removing the chrM
genome.df <- genome.df[1:nrow(genome.df)-1,]
genome.df
```

# CICERO

## Convert to CellDataSet format and make the cicero object.
```{r}
seurat.cds <- as.cell_data_set(x = seurat)
seurat.cds

seurat.cicero <- make_cicero_cds(seurat.cds, reduced_coordinates = reducedDims(seurat.cds)$HARMONY)
seurat.cicero
```

## Run main function 
The run_cicero function is a wrapper that runs the primary functions of the Cicero pipeline with default parameters such as:

1. estimate_distance_parameter. This function calculates the distance penalty parameter based on small random windows of the genome.
2. generate_cicero_models. This function uses the distance parameter determined above and uses graphical LASSO to calculate the co-accessibility scores of overlapping windows of the genome using a distance-based penalty.
3. assemble_connections. This function takes as input the output of generate_cicero_models and reconciles the overlapping models to create the final list of co-accessibility scores.

```{r}
conns <- run_cicero(seurat.cicero, 
                    genomic_coords = genome.df, 
                    sample_num = 100)

write.table(x = conns, 
            file = paste0(path_to_save, 
                          "level_5_conns.tsv"),
            row.names = F, quote = F)
```

## Find co-accessible networks
Now that we’ve found pairwise co-accessibility scores for each peak, we can now group these pairwise connections (using Louvain-based clustering algorithm) into larger co-accessible networks using the generate_ccans() function from Cicero. CCANS are modules of sites that are highly co-accessible with one another.

```{r}
ccans <- generate_ccans(conns)
print(paste("Number of chromatin networks:",max(ccans$CCAN)))

ccans_sorted <- ccans[order(-ccans$CCAN),] 
head(ccans_sorted)

write.table(x = ccans_sorted, 
file = paste0(path_to_save, "level_5_ccans.tsv"),
row.names = F, quote = F)
```

## Add links to a Seurat object
```{r}
conns <- read.table(paste0(path_to_save,"level_5_conns.tsv"), 
                              header = T)
ccans <- read.table(paste0(path_to_save,"level_5_ccans.tsv"),
                              header = T)

links <- ConnectionsToLinks(conns = conns, ccans = ccans)
Links(seurat) <- links

write.table(x = links, 
file = paste0(path_to_save, "level_5_links.tsv"),
row.names = F, quote = F)
```

```{r fig.wide=TRUE}
CoveragePlot(
object = seurat,
region = "chr8-127322183-128417210")
```

# Relationship between CCANS and level 1 assigment.

For each cluster, we are going to extract the normalized value of the cells that have a peak in each specific CCAN.
We are going to calculate the mean value from all the cells to check if there is any bias toward a certain cell type.

```{r}
#ccans_split = ccans %>% separate(Peak, c("chr","start","end"))
#ccans_split$end = as.numeric(ccans_split$end)
#ccans_split$start = as.numeric(ccans_split$start)

# For each module, we are going to extract the peaks tha conform it
CCANS_sum = c()
clusters = c()
modules = c()

for (module in unique(ccans_sorted$CCAN)){
peaks <- ccans_sorted[ccans_sorted$CCAN == module,]$Peak

for (clust in unique(seurat@meta.data$annotation_paper)){
cells_cluster = row.names(seurat@meta.data[seurat@meta.data$annotation_paper == clust,])
peaks_data_cluster <- seurat@assays$peaks@data[peaks,cells_cluster]
CCANS_sum = c(CCANS_sum,mean(peaks_data_cluster))
clusters = c(clusters,clust)
modules = c(modules, rep(module,length(clust)))
}
}

CCANS_per_cluster = data.frame(clusters,CCANS_sum,modules)
colnames(CCANS_per_cluster) = c("clusters","CCANS_means","module")
CCANS_per_cluster$module = as.factor(CCANS_per_cluster$module)
CCANS_per_cluster$clusters = as.factor(CCANS_per_cluster$clusters)
head(CCANS_per_cluster)

#write.table(x = CCANS_per_cluster, 
#file = paste0(path_to_save, "level_1_ccansxcluster.tsv"),
#row.names = F, quote = F)
```

## Conversion into matrix and compute the PCA
Now, we can visualize how the different cell types can be related or unrelated by looking at the co-accessible comunities defined by Cicero.

```{r}
CCANS_per_cluster_mat = acast(CCANS_per_cluster, clusters~module, value.var='CCANS_means')
res.pca <- prcomp(CCANS_per_cluster_mat, scale = TRUE)
```

```{r fig.wide=TRUE}
options(repr.plot.width=6, repr.plot.height=8)
fviz_pca_ind(res.pca,
repel = TRUE     # Avoid text overlapping
)
```

```{r fig.wide=TRUE}
options(repr.plot.width=15, repr.plot.height=6)
out = pheatmap(scale(CCANS_per_cluster_mat),cutree_cols = 3, cutree_rows = 2,
clustering_distance_rows = "euclidean",
clustering_distance_cols = "euclidean",
clustering_method = "ward.D2")
```


```{r}
sessionInfo()
```
