---
title: "QC metrics Spreadsheet"
author: "Marc Elosua Bayes"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  BiocStyle::html_document:
    toc: yes
    toc_float: yes
    number_sections: yes
    df_print: paged
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      out.width = "100%",
                      fig.align = "center", 
                      message = FALSE,
                      warning = FALSE,
                      cache = TRUE)
```

## Introduction

BCLLATLAS_32 is a project within the context of the BCLL\@las which aims to generate a B cell atlas, its lineages, trajectories and spatial localization on the tissue.
The objective of this notebook is to perform a basic quality control (QC) analysis of the mapping performed with [spaceranger 1.1.0](https://support.10xgenomics.com/spatial-gene-expression/software/pipelines/latest/using/count).

Reference script from Massonix: https://github.com/massonix/tonsil_atlas/blob/master/2-DOWNSTREAM_PROCESSING/2-QC/01-qc_cellranger_mapping.Rmd

## Load packages

```{r}
library(ggpubr)
library(ggrepel)
library(DT)
library(dplyr)
library(readr)
library(ggplot2)
library(stringr)
library(gt)
```

## Parameters
Loading necessary paths and parameters
```{r}
source(here::here("misc/paths.R"))

"{qc}/{plt_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(path = .,
             showWarnings = FALSE,
             recursive = TRUE)

"{qc}/{robj_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(path = .,
             showWarnings = FALSE,
             recursive = TRUE)
```


## Load data
We have 8 different datasets, 1 for each capture area from 2 Visium slides.
```{r}
qcmetric_df <- lapply(id_sp_df$gem_id, function(id) {
  tmp_df <- "{spaceranger}/{id}/outs/metrics_summary.csv" %>%
    glue::glue() %>%
    here::here() %>%
    readr::read_csv(file = .) %>%
    dplyr::mutate(
        sample_id = id_sp_df[id_sp_df$gem_id == id, ] %>% dplyr::pull(donor_id),
        subproject = id_sp_df[id_sp_df$gem_id == id, ] %>% dplyr::pull(subproject)
        ) %>%
    dplyr::rename(
        gem_id = `Sample ID`,
        donor_id = sample_id) %>%
    dplyr::select(subproject, gem_id, donor_id, dplyr::everything())
  
  return(tmp_df)
}) %>% dplyr::bind_rows()


DT::datatable(qcmetric_df)
```

Save to CSV
```{r}
"{qc}/{robj_dir}/Supplementary_QC_table_spatial.csv" %>%
    glue::glue() %>%
    here::here() %>%
    write_csv(x = qcmetric_df, file = .)
```

# Session Info

```{r}
sessionInfo()
```