---
title: "NBC_MBC: Subset (level 4)"
author: "Ramon Massoni-Badosa"
date: "2021/05/13"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```


# Introduction

The purpose of this notebook is to exclude technical clusters. In addition, we will exclude the following cells:

- BCLL-2-T profiled with multiome: we have observed in the QC that this was an outlier sample, and some clusters were exclusively composed of it.
- Dataset from [King et al](https://immunology.sciencemag.org/content/6/56/eabe6291): they have provided (i) an initial annotation and (ii) a positive control of biological preservation. However, some clusters were specific of this dataset, so we will also exclude it.


## Load packages

```{r}
library(Seurat)
library(tidyverse)
```


## Parameters

```{r}
# Paths
path_to_obj <- here::here("scRNA-seq/results/R_objects/level_3/NBC_MBC/NBC_MBC_annotated_level_3.rds")
path_to_level_4 <- here::here("scRNA-seq/results/R_objects/level_4/NBC_MBC/")
dir.create(path_to_level_4, recursive = TRUE, showWarnings = FALSE)
path_to_save <- str_c(path_to_level_4, "NBC_MBC_subsetted_level_4.rds")


# Colors
color_palette <-  c("#1CFFCE", "#90AD1C", "#C075A6", "#85660D", "#5A5156", "#AA0DFE",   
                    "#F8A19F", "#F7E1A0", "#1C8356", "#FEAF16", "#822E1C", "#C4451C",   
                    "#1CBE4F", "#325A9B", "#F6222E", "#FE00FA", "#FBE426", "#16FF32", 
                    "black",   "#3283FE", "#B00068", "#DEA0FD", "#B10DA1", "#E4E1E3",   
                    "#90AD1C", "#FE00FA", "#85660D", "#3B00FB", "#822E1C", "coral2", 
                    "#1CFFCE", "#1CBE4F", "#3283FE", "#FBE426", "#F7E1A0", "#325A9B",   
                    "#2ED9FF", "#B5EFB5", "#5A5156", "#DEA0FD", "#FEAF16", "#683B79",   
                    "#B10DA1", "#1C7F93", "#F8A19F", "dark orange", "#FEAF16", "#FBE426",  
                    "Brown")

# Functions
source(here::here("scRNA-seq/bin/utils.R"))
```


## Load data

```{r}
# Seurat object
seurat <- readRDS(path_to_obj)
seurat
Idents(seurat) <- "annotation_level_3"
DimPlot(seurat, cols = color_palette, pt.size = 0.2)
```


# Exclude poor-quality cells

```{r}
seurat <- subset(seurat, subset = annotation_level_3 != "technical")
seurat <- subset(seurat, subset = assay != "5P")
keep_cells <- colnames(seurat)[!(seurat$assay == "multiome" & seurat$donor_id == "BCLL-2-T")]
seurat <- subset(seurat, cells = keep_cells)
seurat
DimPlot(seurat, cols = color_palette, pt.size = 0.2)
```



# Save

```{r}
saveRDS(seurat, path_to_save)
```


# Session Info

```{r}
sessionInfo()
```

