---
title: "Doublet Evaluation"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: Sonal 
output:
  html_document:
    toc: yes
    toc_float: yes
    number_sections: yes
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(Seurat)
library(dplyr)
library(ggplot2)
library(scales)
library(stringr)
```

## Define Parameters

```{r}
# Paths
path_to_save_obj <- "../results"
path_to_saved_post_QC_obj <- str_c(
  path_to_save_obj,
  "seurat_object_cite_seq_postQC.rds",
  sep = "/"
)
path_to_save_seurat_assess_doublets_obj <- str_c(
  path_to_save_obj,
  "seurat_object_cite_seq_seurat_eval_doublets.rds",
  sep = "/"
)
saved_cell_cycle_obj <- str_c(
  "../data/cycle.rda",
  sep = "/"
)
```

## Read saved seurat object post QC

```{r}
filtered_bcll.combined <- readRDS(path_to_saved_post_QC_obj)
```

# Normalize

To normalize by sequencing depth, we will divide each count by the library size of the cell (total number of UMI) and log-transform it, similarly to other high-quality atlases, like the [thymus](https://science.sciencemag.org/content/367/6480/eaay3224) and the [heart](https://www.nature.com/articles/s41586-020-2797-4) atlas.

```{r}
filtered_bcll.combined <- NormalizeData(
  filtered_bcll.combined,
  normalization.method = "LogNormalize",
  scale.factor = 1e4
)
filtered_bcll.combined[["RNA"]]@data[1:10, 1:10]
```


# Visualize UMAP without batch effect correction

```{r fig.wide=TRUE}
filtered_bcll.combined <- filtered_bcll.combined %>%
  FindVariableFeatures(nfeatures = 3000) %>%
  ScaleData() %>% 
  RunPCA() %>%
  RunUMAP(reduction = "pca", dims = 1:30)
```

## Plot UMAP with respect to donor_id

```{r}
DimPlot(filtered_bcll.combined, group.by = "donor_id", pt.size = 0.1) + NoLegend()
```

## Plot UMAP with respect to subproject

```{r}
DimPlot(filtered_bcll.combined, group.by = "subproject", pt.size = 0.1)
```

## Plot UMAP with respect to gem_id

```{r}
DimPlot(filtered_bcll.combined, group.by = "gemid", pt.size = 0.1)
```

## Plot UMAP with respect to seurat

```{r}
DimPlot(filtered_bcll.combined, group.by = "scrublet_predicted_doublet", pt.size = 0.1)
```

```{r}
FeaturePlot(filtered_bcll.combined, features = "scrublet_doublet_scores")
```

## Check vireo doublets

```{r}
DimPlot(filtered_bcll.combined, group.by = "genotype_based_doublet_flag", pt.size = 0.1)
```

## Unassigned to any donor 

```{r}
DimPlot(filtered_bcll.combined, group.by = "genotype_based_unassigned_flag", pt.size = 0.1)
```

## Check doublet feature scatter

## Doublets defined by scrublet

```{r}
filtered_bcll.combined_scrublet_doublet = subset(filtered_bcll.combined, subset = scrublet_predicted_doublet == 'True')
p1 <- FeatureScatter(filtered_bcll.combined_scrublet_doublet, feature1 = "CD4", feature2 = "CD8", group.by
 = "donor_id")
p2 <- FeatureScatter(filtered_bcll.combined_scrublet_doublet, feature1 = "CD19", feature2 = "CD3E", group.by
 = "donor_id")
p3 <- FeatureScatter(filtered_bcll.combined_scrublet_doublet, feature1 = "CD79B", feature2 = "CD3E", group.by
 = "donor_id")
p1
p2
p3
```

## Check genotype doublet feature scatter

```{r}
filtered_bcll.combined_genotype_doublet = subset(filtered_bcll.combined, subset = genotype_based_doublet_flag == 'T')
p1 <- FeatureScatter(filtered_bcll.combined_genotype_doublet, feature1 = "CD4", feature2 = "CD8", group.by
 = "donor_id")
p2 <- FeatureScatter(filtered_bcll.combined_genotype_doublet, feature1 = "CD19", feature2 = "CD3E", group.by
 = "donor_id")
p3 <- FeatureScatter(filtered_bcll.combined_genotype_doublet, feature1 = "CD79B", feature2 = "CD3E", group.by
 = "donor_id")
p1
p2
p3
```

## Check genotype unassigned feature scatter

```{r}
filtered_bcll.combined_genotype_unassigned = subset(filtered_bcll.combined, subset = genotype_based_unassigned_flag == 'T')
p1 <- FeatureScatter(filtered_bcll.combined_genotype_unassigned, feature1 = "CD4", feature2 = "CD8", group.by
 = "donor_id")
p2 <- FeatureScatter(filtered_bcll.combined_genotype_unassigned, feature1 = "CD19", feature2 = "CD3E", group.by
 = "donor_id")
p3 <- FeatureScatter(filtered_bcll.combined_genotype_unassigned, feature1 = "CD79B", feature2 = "CD3E", group.by
 = "donor_id")
p1
p2
p3
```


## Check for Cell cycle stages variability between clusters

```{r}
load(saved_cell_cycle_obj)
filtered_bcll.combined <- CellCycleScoring(filtered_bcll.combined, 
                                 g2m.features = g2m_genes, 
                                 s.features = s_genes)
```

## Cell Cycle 

```{r}
# Plot the PCA colored by cell cycle phase
DimPlot(filtered_bcll.combined,
        reduction = "pca",
        group.by= "Phase",
        split.by = "Phase")
```

```{r}
DimPlot(filtered_bcll.combined, group.by = 'Phase', pt.size = 0.1)
```

## Check cells containing both TCR and BCR

## Plot UMAP with respect to TCR

```{r}
DimPlot(filtered_bcll.combined, group.by = "tcr_flag", pt.size = 0.1)
```

## Plot UMAP with respect to BCR

```{r}
DimPlot(filtered_bcll.combined, group.by = "bcr_flag", pt.size = 0.1)
```

# Session Info
```{r}
sessionInfo()
```