---
title: "Project CITE-seq onto RNA-seq (level 1)"
author: "Ramon Massoni-Badosa"
date: "`r format(Sys.time(), '%B %Y')`"
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center',
                      message=FALSE, warning = FALSE)
options(width = 1200)
```


# Introduction

This script predicts transfers the annotation (level 1) and UMAP coordinates from scRNA-seq to CITE-seq.


# Pre-processing

## Load packages

```{r}
library(Seurat)
#library(SeuratWrappers)
library(harmony)
library(caret)
library(class)
library(tidyverse)
```


## Parameters

```{r}
# Parameters and thresholds
optimal_k <- 8
# Source functions
source("SLOcatoR_functions.R")
# Paths
path_to_query <- "../results/seurat_object_cite_seq_seurat_wnn.rds"
path_to_reference <- "../results/tonsil_atlas_all_cells_20210930.rds"
path_to_save_cite_seq_df <- "../results/figure_1_umap_df_cite_seq.tsv"
path_to_save_integrated <- "../results/integrated_cite_seq_rna.rds"
```


## Read data

```{r}
query <- readRDS(path_to_query)
reference <- readRDS(path_to_reference)
```


# Merge query and reference

```{r}
DefaultAssay(query) <- "RNA"
query[["ADT"]] <- NULL
reference$UMAP_1_level_1 <- reference@reductions$umap@cell.embeddings[, "UMAP_1"]
reference$UMAP_2_level_1 <- reference@reductions$umap@cell.embeddings[, "UMAP_2"]
reference@meta.data$UMAP_1_level_1 <- reference$UMAP_1_level_1
reference@meta.data$UMAP_2_level_1 <- reference$UMAP_2_level_1
reference@meta.data$subproject <- reference@meta.data$donor_id 
reference <- DietSeurat(reference)
selected_cols_ref <- c("donor_id","gem_id", "assay", "nCount_RNA","nFeature_RNA", "pct_mt","annotation_level_1",     "scrublet_doublet_scores", "scrublet_predicted_doublet", "UMAP_1_level_1", "UMAP_2_level_1")
reference@meta.data <- reference@meta.data[, selected_cols_ref]
reference$type <- "reference"
colnames(query@meta.data)[colnames(query@meta.data) == "gemid"] <- "gem_id" 
colnames(query@meta.data)[colnames(query@meta.data) == "percent.mt"] <- "pct_mt" 
selected_cols_query <- c("donor_id","gem_id", "nCount_RNA","nFeature_RNA","pct_mt","scrublet_doublet_scores", "scrublet_predicted_doublet")
query@meta.data <- query@meta.data[, selected_cols_query]
query$assay <- "CITE-seq"
query$type <- "query"
query@meta.data$annotation_level_1 <- "unannotated"
query$UMAP_1_level_1 <- NA
query$UMAP_2_level_1 <- NA
query@meta.data$UMAP_1_level_1 <- query$UMAP_1_level_1
query@meta.data$UMAP_2_level_1 <- query$UMAP_2_level_1
integrated <- merge(x = reference, y = query)
Idents(integrated) <- "assay"
```


# Find assay-specific highly variable genes

```{r}
shared_hvg <- find_assay_specific_features(integrated, n_features = 5500)
print(length(shared_hvg))
```


# Integrate with Harmony

```{r}
integrated <- integrate_assays(
  integrated,
  assay_specific = TRUE,
  assay_var = "assay",
  shared_hvg = shared_hvg
)
integrated <- RunUMAP(integrated, dims = 1:30, reduction = "harmony")
DimPlot(integrated, split.by = "assay", group.by = "assay")
VlnPlot(integrated, "nFeature_RNA", group.by = "assay", pt.size = 0)
```


# Create training and test sets

```{r}
data_sets <- split_training_and_test_sets(
  integrated,
  split_var = "type",
  referece_label = "reference",
  query_label = "query",
  reduction = "harmony",
  n_dims = 30
)
```


# Find optimal K

```{r}
optimal_k_df <- find_optimal_k(
  seurat_obj = integrated,
  training_set = data_sets$training_set,
  response_var = "annotation_level_1",
  return_plot = TRUE
)
optimal_k_df$df
optimal_k_df$plot
```


# Transfer label

```{r}
annotation_data  <- transfer_label(
  seurat_obj = integrated,
  training_set = data_sets$training_set,
  test_set = data_sets$test_set,
  k = optimal_k,
  response_var = "annotation_level_1"
)
```


# Predict UMAP coords

```{r}
umap_coords <- transfer_umap_coords(
  seurat_obj = integrated,
  training_set = data_sets$training_set,
  test_set = data_sets$test_set,
  k = optimal_k,
  umap1_var = "UMAP_1_level_1",
  umap2_var = "UMAP_2_level_1"
)
ggplot(umap_coords, aes(UMAP1, UMAP2)) +
  geom_point() +
  theme_classic()
```


# Save

```{r}
# Homogenize columns
integrated$annotation_prob <- NA
integrated$annotation_prob[annotation_data$query_cells] <- annotation_data$annotation_prob
integrated$annotation_level_1[annotation_data$query_cells] <- annotation_data$annotation
integrated$UMAP_1_level_1[umap_coords$query_cells] <- umap_coords$UMAP1
integrated$UMAP_2_level_1[umap_coords$query_cells] <- umap_coords$UMAP2
annotation_cite_df <- left_join(
  annotation_data,
  umap_coords,
  by = "query_cells"
)
# Save
saveRDS(integrated, path_to_save_integrated)
saveRDS(annotation_cite_df, path_to_save_cite_seq_df)
```


# Session Information

```{r}
sessionInfo()
```