---
author: "Marc Elosua-Bayes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
title: "BCLL-10-T T cell markers mapping"
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message = FALSE, warning = FALSE, cache = FALSE,
                      cache.lazy = FALSE)
options(width = 1200)
```

## Introduction
In this script we are going to look at marker genes identifying different CD4-related cell types from the tonsil.
We aim to identify the different stages of CD4 T cells since they enter as naive until they make their way into the GC being Tfh.
The main papers of reference for this analysisare i) [**T Follicular Helper Cell Biology: A Decade of Discovery and Diseases**](https://doi.org/10.1016/j.immuni.2019.04.011) and ii) [Bcl6-Mediated Transcriptional Regulation of Follicular Helper T cells (TFH)](https://doi.org/10.1016/j.it.2021.02.002).

## Libraries
```{r}
library(Seurat)
library(ggpubr)
library(cowplot)
library(dplyr)
library(ggplot2)
library(stringr)
library(readr)
```

## Setting parameters
Loading necessary paths and parameters
```{r}
set.seed(123)
source(here::here("misc/paths.R"))
source(here::here("utils/bin.R"))

"{cd4}/{plt_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(path = ,
             showWarnings = FALSE,
             recursive = TRUE)

"{cd4}/{robj_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(path = ,
             showWarnings = FALSE,
             recursive = TRUE)
```

Extract sample id and get Donor ID
```{r}
sample_id <- params$sample_id
# sample_id <- "esvq52_nluss5"
donor_id <- id_sp_df[id_sp_df$gem_id == sample_id, ] %>% dplyr::pull(donor_id)
```

## Load data
The data used in this Rmarkdown document comes from **03-clustering_integration.Rmd** where the data was integrated.
```{r}
merged_se <- "{clust}/{robj_dir}/integrated_spatial.rds" %>%
  glue::glue() %>%
  here::here() %>%
  readRDS(file = .)
```

Load MAGIC data from the script *MAGIC_denoising.Rmd*
```{r}
magic_df <- "{cd4}/{robj_dir}/MAGIC-mtrx.rds" %>%
  glue::glue() %>%
  here::here() %>%
  readRDS(file = .)

# create a new assay to store ADT information
magic_assay <- CreateAssayObject(counts = as.matrix(magic_df))

# add this assay to the previously created Seurat object
merged_se[["MAGIC_Spatial"]] <- magic_assay

Seurat::DefaultAssay(merged_se) <- "MAGIC_Spatial"
```

subset object sample of interest
```{r}
se_sub <- subset(merged_se, subset = gem_id == "esvq52_nluss5")
se_sub@images <- se_sub@images[Seurat::Images(se_sub) == "esvq52_nluss5"]
```


## Analysis


## Session Info
```{r}
sessionInfo()
```
