---
author: "Marc Elosua-Bayes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
title: "CD4 deconvolution assessment"
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message = FALSE, warning = FALSE, cache = FALSE,
                      cache.lazy = FALSE)
options(width = 1200)
```

## Introduction
In this script we are going to map T-cell subtypes onto the Visium slides.

## Libraries
```{r}
library(Seurat)
library(ggpubr)
library(cowplot)
library(dplyr)
library(ggplot2)
library(stringr)
library(readr)
library(SPOTlight)
library(SPATA2)
```

## Setting parameters
Loading necessary paths and parameters
```{r}
source(here::here("misc/paths.R"))
source(here::here("utils/bin.R"))

"{cd4}/{plt_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(path = ,
             showWarnings = FALSE,
             recursive = TRUE)

"{cd4}/{robj_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(path = ,
             showWarnings = FALSE,
             recursive = TRUE)

set.seed(123)
```

Extract sample id and get Donor ID
```{r}
# sample_id <- params$sample_id
sample_id <- "esvq52_nluss5"
donor_id <- id_sp_df[id_sp_df$gem_id == sample_id, ] %>% dplyr::pull(donor_id)
```

Set gene dictionary
```{r}
"{cd4}/gene_dict.R" %>%
  glue::glue() %>%
  here::here() %>%
  source(file = .)
```

## Load data
We have 8 different datasets that we are going to analyze separately.
The spatial data comes from the script *03-clustering/03-clustering_integration.Rmd* while the sc data can be found in Ramon's scRNAseq analysis: **/scratch/devel/rmassoni/tonsil_atlas_private/2-DOWNSTREAM_PROCESSING/results/R_objects/processed_seurat_objects/processed_seurat_objects/tonsil_integrated_with_harmony_scrublet_annotated.rds**. \
```{r}
sp_obj <- "misc/{robj_dir}/20220215_tonsil_atlas_spatial_seurat_obj.rds" %>%
  glue::glue() %>%
  here::here() %>%
  readRDS(file = .)

# Load SPOTlight data
spotlight_ls <- "{cd4}/{robj_dir}/spotlight_ls_cd4_new.rds" %>%
  glue::glue() %>% 
  here::here() %>%
  readRDS(file = .)
```

Load MAGIC data from the script *MAGIC_denoising.Rmd*
```{r eval=FALSE}
magic_df <- "{cd4}/{robj_dir}/MAGIC-mtrx.rds" %>%
  glue::glue() %>%
  here::here() %>%
  readRDS(file = .)

# create a new assay to store ADT information
magic_assay <- CreateAssayObject(counts = as.matrix(magic_df))

# add this assay to the previously created Seurat object
sp_obj[["MAGIC_Spatial"]] <- magic_assay

Seurat::DefaultAssay(sp_obj) <- "MAGIC_Spatial"
```

Add colors to cell types
```{r}
col_vec <- c(
  "Naive" = "gray88",
  # "CM Pre-non-Tfh" = "gray71",
  # "CM PreTfh" = "gray59",
  "CM PreTfh/Pre-non-Tfh" = "gray59",
  "Tfh T:B border" = "#67a9cf",
  "Tfh-LZ-GC" = "#3690c0",
  "GC-Tfh-SAP" = "#02818a",
  "GC-Tfh-0X40"= "#016c59",
  "Tfh-Mem" = "#014636",
  "T-Trans-Mem" = "#fd8d3c",
  "T-Eff-Mem" = "#e31a1c",
  "T-helper" = "#800026",
  "Eff-Tregs" = "#df65b0",
  "non-GC-Tf-regs" = "#e7298a",
  "GC-Tf-regs" = "#ce1256"
)

(nm_df <- data.frame(col_vec))
nm_df$plt_nm <- rownames(nm_df)
nm_df$mod_nm <- stringr::str_replace_all(
  string = rownames(nm_df),
  pattern = "[[:punct:]]|[[:space:]]",
  replacement = ".")
```

## Analysis

### Preprocess data
```{r}
decon_mtrx <- spotlight_ls[[2]]
decon_mtrx <- decon_mtrx[, colnames(decon_mtrx) != "res_ss"]

# Set as 0 cell types predicted to be under 3 % of the spot
decon_mtrx[decon_mtrx < 0.03] <- 0
```

Change column names
```{r}
new_cn <- data.frame(mod_nm = colnames(decon_mtrx)) %>%
  dplyr::left_join(nm_df, by = "mod_nm") %>%
  dplyr::mutate(plt_nm = dplyr::if_else(is.na(plt_nm), mod_nm, plt_nm)) %>%
  dplyr::distinct() %>%
  dplyr::pull(plt_nm)

colnames(decon_mtrx) <- new_cn
```

We are going to add the deconvolution to the Seurat object.
```{r}
sp_obj@meta.data <- cbind(sp_obj@meta.data, decon_mtrx)
```

Subset sample of interest
```{r}
sp_sub <- sp_obj[, sp_obj@meta.data$gem_id == sample_id]
sp_sub@images <- sp_sub@images[Seurat::Images(sp_sub) == sample_id]
```

### Look at SPOTlight results
Check Topic profiles
```{r fig.height=8, fig.width=8}
nmf_mod_ls <- spotlight_ls[[1]]
nmf_mod <- nmf_mod_ls[[1]]

h <- NMF::coef(nmf_mod)
rownames(h) <- paste("Topic", 1:nrow(h), sep = "_")
topic_profile_plts <- SPOTlight::dot_plot_profiles_fun(
  h = h,
  train_cell_clust = nmf_mod_ls[[2]])

topic_profile_plts[[2]] +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90), 
                 axis.text = ggplot2::element_text(size = 12))
```

Look at all cells profiles
```{r fig.height=20, fig.width=20}
topic_profile_plts[[1]] +
  ggplot2::theme(
    axis.text.y = ggplot2::element_blank(),
    axis.text.x = ggplot2::element_blank(),
    axis.title = ggplot2::element_blank())
```

Look at cells topic profile
```{r}
basis_spotlight <- data.frame(NMF::basis(spotlight_ls[[1]][[1]]))

train_labs <- spotlight_ls[[1]][[2]]
colnames(basis_spotlight) <- unique(stringr::str_wrap(train_labs, width = 30))

basis_spotlight[basis_spotlight < 0.0000001] <- 0

DT::datatable(basis_spotlight, filter = "top")
```

#### Cell type location
Look at the location of each cell type in each slice separately
```{r fig.height=25, fig.width=25}
# Iterate over cell types
ct <- colnames(decon_mtrx)

# Iterate over images
lapply(names(sp_obj@images), function(nm) {
  print(nm)
  nm_donor <- id_sp_df %>% dplyr::filter(gem_id == nm) %>% dplyr::pull(donor_id)
  # Iterate over cell types
  ct_plt_ls <- lapply(ct, function(i) {
    tmp_plt <- Seurat::SpatialFeaturePlot(
      object = sp_obj,
      features = i,
      alpha = c(0, 1),
      images = nm) +
      ggplot2::scale_fill_gradientn(
        colors = heat.colors(10, rev = TRUE)) +
      ggplot2::scale_alpha(range = c(0, 1)) +
      ggplot2::labs(title = stringr::str_wrap(string = i,
                                     width = 25),
           fill = "") +
      ggplot2::theme(
        plot.title = ggplot2::element_text(
          hjust = 0.5,
          size = 20,
          face = "bold"))
    
    if (sum(sp_sub@meta.data[, i]) == 0) {
      tmp_plt <- suppressMessages(tmp_plt + ggplot2::scale_alpha(range = c(0,0)))
    }
    
    return(tmp_plt)
  })
  
  (plt_arr <- cowplot::plot_grid(
    plotlist = ct_plt_ls,
    axis = "trbl",
    align = "hv",
    nrow = 5,
    ncol = 5))
  
  "{cd4}/{plt_dir}/cell_type_location_cd4_{nm_donor}_new.pdf" %>%
    glue::glue() %>%
    here::here() %>%
    cowplot::save_plot(
      filename = .,
      plot = plt_arr,
      base_height = 25,
      base_width = 25)
  })
```


### Cell type - Gene
Look at correlation between cell types and markers
```{r}
sp_sub <- sp_obj[, sp_obj@meta.data$gem_id == sample_id]
sp_sub@images <- sp_sub@images[Seurat::Images(sp_sub) == sample_id]
```

```{r}
scatterplot2 <- function(se_obj, x, y, color) {
  df <- data.frame(
    feat1 = se_obj@assays$MAGIC_Spatial@data[x, ],
    feat2 = se_obj@meta.data[, y],
    color = se_obj@meta.data[, color]
    )
  
  ggplot2::ggplot(df,
                  ggplot2::aes(x = feat1,
                               y = feat2)) +
    ggplot2::geom_point(ggplot2::aes(color = color)) +
    ggplot2::geom_smooth(method = "loess", se = FALSE) +
    # ggplot2::geom_smooth(method = lm, se = FALSE) +
    ggplot2::theme_classic() +
    ggplot2::labs(
      x = x,
      y = glue::glue("Proportion of\n{y}"),
      color = color)

}
```

#### BCL6 

```{r eval=FALSE, fig.width=25, fig.height=25}
bcl6_ls <- lapply(ct, function(ii) {
  scatterplot2(
    se_obj = sp_sub,
    x = "BCL6",
    y = ii,
    color = "annotation")
})

patchwork::wrap_plots(bcl6_ls, ncol = 5) +
  patchwork::plot_layout(guides = "collect") &
  ggplot2::theme(legend.position = "right")
```


#### PRDM1 

```{r eval=FALSE, fig.width=25, fig.height=25}
prdm1_ls <- lapply(ct, function(ii) {
  scatterplot2(
    se_obj = sp_sub,
    x = "PRDM1",
    y = ii,
    color = "annotation")
})

patchwork::wrap_plots(prdm1_ls, ncol = 5) +
  patchwork::plot_layout(guides = "collect") &
  ggplot2::theme(legend.position = "right")
```

#### TCF7 

```{r eval=FALSE, fig.width=25, fig.height=25}
tcf7_ls <- lapply(ct, function(ii) {
  scatterplot2(
    se_obj = sp_sub,
    x = "TCF7",
    y = ii,
    color = "annotation")
})

patchwork::wrap_plots(tcf7_ls, ncol = 5) +
  patchwork::plot_layout(guides = "collect") &
  ggplot2::theme(legend.position = "right")
```

#### LEF1 

```{r eval=FALSE, fig.width=25, fig.height=25}
lef1_ls <- lapply(ct, function(ii) {
  scatterplot2(
    se_obj = sp_sub,
    x = "LEF1",
    y = ii,
    color = "annotation")
})

patchwork::wrap_plots(lef1_ls, ncol = 5) +
  patchwork::plot_layout(guides = "collect") &
  ggplot2::theme(legend.position = "right")
```

#### CXCR5 

```{r eval=FALSE, fig.width=25, fig.height=25}
cxcr5_ls <- lapply(ct, function(ii) {
  scatterplot2(
    se_obj = sp_sub,
    x = "CXCR5",
    y = ii,
    color = "annotation")
})

patchwork::wrap_plots(cxcr5_ls, ncol = 5) +
  patchwork::plot_layout(guides = "collect") &
  ggplot2::theme(legend.position = "right")
```

### T cell markers
```{r fig.width=25, fig.height=10}
Seurat::SpatialFeaturePlot(
  object = sp_obj,
  features = c("CD3D", "CD4", "CD8A", "CD8B"),
  alpha = c(0, 1),
  images = "esvq52_nluss5",
  slot = "data") |
    Seurat::SpatialFeaturePlot(
    object = sp_obj,
    features = c("Cytotoxic", "Naive"),
    alpha = c(0, 1),
    images = "esvq52_nluss5")

```

Correlation between markers
```{r eval=FALSE, fig.width=15, fig.height=10}
(scatterplot(se_obj = sp_sub, x = "CD4", y = "CD8A", color = "annotation") |
  scatterplot(se_obj = sp_sub, x = "CD4", y = "CD3D", color = "annotation")) /
  (scatterplot(se_obj = sp_sub, x = "CD8A", y = "CD3D", color = "annotation") | 
  scatterplot(se_obj = sp_sub, x = "CD4", y = "CXCR5", color = "annotation")) +
  patchwork::plot_layout(guides = "collect") &
  ggplot2::theme(legend.position = "right")
```

### Boxplots
Prepare data for boxplots
```{r}
metadata_long <- sp_sub@meta.data %>% 
  # tidyr::pivot_longer(cols = c("annotation"),
  #                     names_to = "stratification_id",
  #                     values_to = "stratification_val") %>%
  tidyr::pivot_longer(cols = all_of(ct), names_to = "ct_key", values_to = "ct_val") %>%
  # dplyr::left_join(col_df, by = c("ct_key" = "ct_name")) %>%
  dplyr::mutate(ct_val = dplyr::if_else(ct_val > 0.001, ct_val, 0))
```

Box plot of cell type proportion between stratified regions showing the unadjusted ANOVA Pvalue

```{r fig.width=30, fig.height=20}
keep_ct <- metadata_long %>%
  dplyr::group_by(ct_key) %>%
  dplyr::summarise(prop_sum = sum(ct_val)) %>% 
  dplyr::filter(prop_sum > 0) %>%
  dplyr::pull(ct_key)

(bplt <- metadata_long %>%
  dplyr::filter(ct_key %in% keep_ct) %>%
  dplyr::mutate(
    ct_key = stringr::str_wrap(string = ct_key,
                                   width = 30)) %>%
  # dplyr::mutate(stratification_val = factor(stratification_val, 
  #                                           levels = c("Fibrotic", "HER2+/ESR1+",
  #                                                      "HER2+/ESR1-", "HER2-/ESR1+",
  #                                                      "HER2-/ESR1-"))) %>%
  # dplyr::filter(! plt_name %in% c("CD8 EM-like", "CD4 naive-like")) %>%
  ggpubr::ggboxplot(data = .,
                    x = "annotation_20220215",
                    y = "ct_val",
                    facet.by = "ct_key",
                    color = "annotation_20220215",
                    fill = "annotation_20220215",
                    add = "jitter",
                    scales = "free",
                    repel = TRUE,
                    outlier.shape = NA,
                    alpha = 0.6,
                    palette = "Set1",
                    ncol = 6) +
  ggplot2::theme(
    strip.text = ggplot2::element_text(size = 18, face = "bold"),
    axis.text.y = ggplot2::element_text(size = 16),
    axis.title.y = ggplot2::element_text(size = 20),
    # axis.text.x = element_text(size = 12, angle = 90, vjust = 0.5, hjust = 0.5),
    axis.text.x = ggplot2::element_blank(),
    legend.text = ggplot2::element_text(size = 18),
    legend.title = ggplot2::element_blank(),
    strip.background = ggplot2::element_blank()) +
  ggplot2::labs(
    y = "Proportion",
    color = "Regions",
    fill = "Regions"))
# bplt <- bplt +
#   ggpubr::stat_compare_means(method = "anova", size = 6) +
#   ggplot2::scale_y_continuous(
#     expand = expansion(mult = c(0, 0.1)),
#     labels = function(x) sprintf("%.2f", x))

"{cd4}/{plt_dir}/strat_bplot_{donor_id}.pdf" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    filename = .,
    plot = bplt,
    base_height = 20,
    base_width = 25)
```

### Genes of interest
```{r fig.width=12, fig.height=5}
Seurat::DefaultAssay(sp_obj) <- "Spatial"

Seurat::SpatialFeaturePlot(
  object = sp_obj,
  features = c("KLRB1", "PRDM1"),
  ncol = 2,
  images = "esvq52_nluss5",
  alpha = c(0, 1))
```

### Integrated UMAP
```{r fig.width=12, fig.height=5}
Seurat::DimPlot(
  object = sp_obj,
  group.by = c("annotation", "sample_id"))
```

Next we want to look at the piecharts in the UMAP
```{r fig.width=25, fig.height=20, eval=FALSE}
# Loading libraries
suppressMessages(require(ggplot2))
suppressMessages(require(dplyr))
suppressMessages(require(tibble))
suppressMessages(require(grid))


metadata_ds <- data.frame(sp_obj@meta.data)

colnames(metadata_ds) <- colnames(sp_obj@meta.data)
cell_types_all <- ct
cell_types_interest <- cell_types_all

# If not all cell types are in the cell types of interest we only want to keep those spots which have at least one of the cell types of interest
if (!all(cell_types_all %in% cell_types_interest)) {

  metadata_ds <- metadata_ds %>%
    tibble::rownames_to_column("barcodeID") %>%
    dplyr::mutate(rsum = base::rowSums(.[, cell_types_interest,
                                         drop = FALSE])) %>%
    dplyr::filter(rsum != 0) %>%
    dplyr::select("barcodeID") %>%
    dplyr::left_join(metadata_ds %>% tibble::rownames_to_column("barcodeID"),
                     by = "barcodeID") %>%
    tibble::column_to_rownames("barcodeID")
}

## Preprocess data
umap_coord <- data.frame(sp_obj@reductions$umap@cell.embeddings[, c("UMAP_1", "UMAP_2")]) %>%
  tibble::rownames_to_column("barcodeID") %>%
  dplyr::inner_join(metadata_ds %>% tibble::rownames_to_column("barcodeID"),
                    by = "barcodeID")

## Plot spatial scatterpie plot
(umap_pie_plt <- ggplot2::ggplot() +
    scatterpie::geom_scatterpie(
      data = umap_coord,
      ggplot2::aes(
        x = UMAP_1,
        y = UMAP_2),
      # Fill the pie chart
      cols = cell_types_all,
      pie_scale = 0.15,
      # Remove black outline
      color = NA) +
    ggplot2::scale_fill_manual(
      breaks = col_df$plt_nm,
      values = col_df$color))

"{cd4}/{plt_dir}/UMAP_piehcart_integrated.pdf" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    filename = .,
    plot = umap_pie_plt,
    base_height = 20,
    base_width = 25)
```

### Cell type correlation matrix
```{r}
# se_sub <- subset(merged_se, subset = gem_id == "esvq52_nluss5")
# se_sub
# se_sub@images <- se_sub@images[Seurat::Images(se_sub) == "esvq52_nluss5"]

(cor_mtrx_ct <- SCrafty::correlation_heatmap( 
  se = sp_obj,
  feats = ct,
  assay = "MAGIC_Spatial",
  slot = "data") +
   ggplot2::labs(
     title = "Integrated cell-type correlation matrix"))

"{cd4}/{plt_dir}/magic_cor-mtrx_cell-type_integrated.pdf" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    filename = .,
    plot = cor_mtrx_ct,
    base_height = 9,
    base_width = 10)
```

### Cell type trajectories
Lastly we also want to take a look at how the cell-type vary over the trajectories previously defined.

Load drawn trajectories on slide *esvq52_nluss5*
```{r}
spata_sub <- "{anot}/{robj_dir}/spata-esvq52_nluss5.rds" %>%
  glue::glue() %>%
  here::here() %>%
  readRDS(file = .)
```

Add cell type proportions to it:
```{r}
decon_mtrx_t <- t(decon_mtrx)
colnames(decon_mtrx_t) <- colnames(sp_obj)

spata_sub <- SPATA2::addExpressionMatrix(
  object = spata_sub,
  expr_mtr = decon_mtrx_t,
  mtr_name = "decomp")

# Set as active assay
spata_sub <- SPATA2::setActiveExpressionMatrix(
  object = spata_sub,
  mtr_name = "decomp")
```

Look at all the genes in gene vec in the heat map
```{r fig.width=20, fig.height=15}
hm_colors <- viridis::inferno(n = 100)

SPATA2::plotTrajectoryHeatmap(
  object = spata_sub,
  trajectory_name = "Tfh-migration",
  variables = ct,
  arrange_rows = "maxima",
  colors = hm_colors,
  show_rownames = TRUE,
  split_columns = TRUE, 
  smooth_span = 0.5)

SPATA2::plotTrajectoryHeatmap(
  object = spata_sub,
  trajectory_name = "Tfh-migration-quiescent",
  variables = ct,
  arrange_rows = "maxima",
  colors = hm_colors,
  show_rownames = TRUE,
  split_columns = TRUE, 
  smooth_span = 0.5)
```

## Session Info
```{r}
sessionInfo()
```
