---
author: "Marc Elosua-Bayes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
params: 
    sample_id: "Default!"
    donor_id: "Default!"
title: "`r sprintf('%s L1 RCTD SC mapping', params$donor_id)`"
editor_options: 
  chunk_output_type: console
---

---
title: `r params$donor_id`
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message = FALSE, warning = FALSE, cache = FALSE,
                      cache.lazy = FALSE)
options(width = 1200)
```

## Introduction
In this script we are going to look at marker genes identifying different CD4-related cell types from the tonsil.
We aim to identify the different stages of CD4 T cells since they enter as naive until they make their way into the GC being Tfh.
The main papers of reference for this analysisare i) [**T Follicular Helper Cell Biology: A Decade of Discovery and Diseases**](https://doi.org/10.1016/j.immuni.2019.04.011) and ii) [Bcl6-Mediated Transcriptional Regulation of Follicular Helper T cells (TFH)](https://doi.org/10.1016/j.it.2021.02.002).

## Libraries
```{r}
library(Seurat)
library(ggpubr)
library(cowplot)
library(dplyr)
library(ggplot2)
library(stringr)
library(readr)
```

## Setting parameters
Loading necessary paths and parameters
```{r}
set.seed(123)
source(here::here("misc/paths.R"))
source(here::here("utils/bin.R"))

"{cd4}/{plt_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(path = ,
             showWarnings = FALSE,
             recursive = TRUE)

"{cd4}/{robj_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(path = ,
             showWarnings = FALSE,
             recursive = TRUE)
```

Extract sample id and get Donor ID
```{r}
sample_id <- params$sample_id
# sample_id <- "esvq52_nluss5"
donor_id <- id_sp_df[id_sp_df$gem_id == sample_id, ] %>% dplyr::pull(donor_id)
```

## Load data
We have 8 different datasets that we are going to analyze separately.
The spatial data comes from the script *03-clustering/03-clustering.Rmd*
In each Rmarkdown notebook we analyze 1 dataset. \
```{r}
sp_obj <- "{clust}/{robj_dir}/clust_se_{sample_id}.RDS" %>%
  glue::glue() %>%
  here::here() %>%
  readRDS(file = )
```

## Analysis

```{r}
Seurat::SpatialFeaturePlot(
  object = sp_obj,
  features = c("HLA-DMA", "HLA-DMB", "HLA-DOA", "HLA-DOB", "HLA-DPA1", "HLA-DPB1", "HLA-DQA1", "HLA-DQA2", "HLA-DQB1", "HLA-DQB2", "HLA-DRA", "HLA-DRB1", "HLA-DRB3", "HLA-DRB4", "HLA-DRB5"))
```


BTLA4, CD200, PDCD1, ICOS -> differentiated Tfh
```{r}
Seurat::SpatialFeaturePlot(
  object = sp_obj,
  features = c("BTLA", "CD200", "PDCD1", "ICOS"),
  alpha = c(0, 1))
```


```{r}
Seurat::DimPlot(object = merged_se2,
                group.by = "Spatial_snn_res.0.1",
                split.by = "age")

Seurat::SpatialDimPlot(
  object = merged_se2,
  group.by = "Spatial_snn_res.0.3",
  image = "esvq52_nluss5")

Seurat::SpatialFeaturePlot(
  object = merged_se2,
  features = c("CD40", "CD40LG", "IL21", "IL4", "PRDM1", "BTLA", "CD200", "CD4"),
  image = "esvq52_nluss5",
  alpha = c(0, 1))

Seurat::VlnPlot(object = merged_se2[, merged_se2@meta.data$gem_id == "esvq52_nluss5"],
  features = c("CD40", "CD40LG", "IL21", "IL4", "PRDM1", "BTLA", "CD200", "ICOSLG", "CD4"),
  group.by =  "Spatial_snn_res.0.3")
```

```{r}
Seurat::SpatialFeaturePlot(
  object = merged_se2,
  features = "BTLA",
  image = "esvq52_nluss5")
```

