---
title: "Trajectory Inference with TSCAN (PC)"
author: "William Blevins and Ramon Massoni-Badosa"
date: "`r format(Sys.time(), '%B %Y')`"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```


# Introduction

Here, we will perform a trajectory analysis with the final annotation of the plasma cells (PC). We will use the [TSCAN](https://github.com/LTLA/TSCAN) package and follow [this tutorial](http://bioconductor.org/books/3.14/OSCA.advanced/trajectory-analysis.html).


# Pre-processing

## Load packages

```{r}
library(Seurat)
library(scater)
library(TSCAN)
library(clusterProfiler)
library(org.Hs.eg.db)
library(ComplexHeatmap)
library(tidyverse)
```


## Define parameters

```{r}
# Paths
# path_to_obj <- "~/Desktop/PhD/PC_with_csMBC_and_AIDcells_analysis/data/PC_level_5_seurat_object_clean_with_new_csMBC_and_aid_and_with_annotated_clusters.rds"
path_to_obj <- "~/Desktop/diogenes/PC_with_csMBC_and_AIDcells_analysis/data/PC_level_5_seurat_object_clean_with_new_csMBC_and_aid_and_with_annotated_clusters.rds"


# Colors
color_palette <-  c("#1CFFCE", "#90AD1C", "#C075A6", "#85660D", "#5A5156", "#AA0DFE",   
                    "#F8A19F", "#F7E1A0", "#1C8356", "#FEAF16", "#822E1C", "#C4451C",   
                    "#1CBE4F", "#325A9B", "#F6222E", "#FE00FA", "#FBE426", "#16FF32", 
                    "black",   "#3283FE", "#B00068", "#DEA0FD", "#B10DA1", "#E4E1E3",   
                    "#90AD1C", "#FE00FA", "#85660D", "#3B00FB", "#822E1C", "coral2", 
                    "#1CFFCE", "#1CBE4F", "#3283FE", "#FBE426", "#F7E1A0", "#325A9B",   
                    "#2ED9FF", "#B5EFB5", "#5A5156", "#DEA0FD", "#FEAF16", "#683B79",   
                    "#B10DA1", "#1C7F93", "#F8A19F", "dark orange", "#FEAF16", "#FBE426",  
                    "Brown")


# Thresholds
alpha <- 0.05
GO_max_total_genes <- 250
GO_min_enriched_genes <- 3
GO_p_adj_threshold <- 0.01
GO_min_odds_ratio <- 2.5
max_gs_size <- 250
min_gs_size <- 10
simplify_cutoff <- 0.75
```


## Read data

```{r}
pc <- readRDS(path_to_obj)
DimPlot(pc, group.by = "names_level_5_clusters", cols = color_palette)
```


# Infer pseudotime

```{r}
# Convert to SingleCellExperiment
pc_sce <- as.SingleCellExperiment(pc)
colLabels(pc_sce) <- pc_sce$names_level_5_clusters


# Calculate minimum spanning tree (MST)
by_cluster <- aggregateAcrossCells(pc_sce, ids = colLabels(pc_sce))
centroids <- reducedDim(by_cluster, "HARMONY")[, 1:30]
mst <- createClusterMST(centroids, clusters = NULL)
mst
line_data <- reportEdges(
  by_cluster,
  mst = mst,
  clusters = NULL,
  use.dimred = "UMAP"
)
plotUMAP(pc_sce, colour_by = "names_level_5_clusters") + 
  geom_line(data = line_data, mapping = aes(x = UMAP_1, y = UMAP_2, group = edge))


# Predict pseudotime
map_tscan <- mapCellsToEdges(
  reducedDim(pc_sce, "HARMONY")[, 1:30],
  mst = mst,
  clusters = colLabels(pc_sce)
)
tscan_pseudo <- orderCells(map_tscan, mst, start = "LZ-GC B cells")
head(tscan_pseudo)
common_pseudo <- averagePseudotime(tscan_pseudo) 
plotUMAP(pc_sce, colour_by = I(common_pseudo)) +
  geom_line(data=line_data, mapping=aes(x=UMAP_1, y=UMAP_2, group=edge))
```


# Trajectory-based DEA

```{r}
# Fit a model for each gene
pseudotime_igha <- pathStat(tscan_pseudo)[, "Mature IgA+ 2"]
plotUMAP(pc_sce, colour_by = I(pseudotime_igha)) +
  geom_line(data=line_data, mapping=aes(x=UMAP_1, y=UMAP_2, group=edge))
pseudo <- testPseudotime(pc_sce, pseudotime = pseudotime_igha)
pseudo$symbol <- rownames(pseudo)
pseudo[order(pseudo$p.value), ]


# Make a copy of SCE to include pseudotimes to colData
pc_sce2 <- pc_sce
pc_sce2$pseudotime_main_trajectory <- pseudotime_igha
pseudo <- as.data.frame(pseudo)
sorted_up <- pseudo[pseudo$logFC > 0 & pseudo$FDR < 1e-4, ]
sorted_down <- pseudo[pseudo$logFC < 0 & pseudo$FDR < 1e-4, ]
sorted_up <- sorted_up[order(sorted_up$logFC, decreasing = TRUE), ]
sorted_down <- sorted_down[order(sorted_down$logFC, decreasing = FALSE), ]
rowData(pc_sce2)$symbol <- rownames(pc_sce2)


# Plot relevant TF
tf <- c("BCL6", "PRDM1", "PRDM2", "XBP1", "IRF4", "IRF1", "IRF8")
plotExpression(
  pc_sce2,
  features = tf,
  swap_rownames = "symbol",
  x = "pseudotime_main_trajectory",
  colour_by = "names_level_5_clusters"
) +
  theme(legend.position = "none")
plotExpression(
  pc_sce2,
  features = str_subset(sorted_up$symbol, "^IRF"),
  swap_rownames = "symbol",
  x = "pseudotime_main_trajectory",
  colour_by = "names_level_5_clusters"
) +
  geom_smooth() +
  theme(legend.position = "none")
```

```{r}
tfs_down <- c("BCL6", "PAX5", "IRF8", "SPIB", "BACH2", "CIITA", "SPI1", "ID3")
tfs_up <- c("FOS", "XBP1", "PRDM1", "IRF4", "POU2AF1", "ZBTB20")
tfs <- c(tfs_down, tfs_up)
tf_dfs <- purrr::map(tfs, function(x) {
  df <- data.frame(
    pseudotime = pc_sce2$pseudotime_main_trajectory,
    expression = scale(logcounts(pc_sce)[x, ], center = TRUE, scale = TRUE)[, 1],
    gene = x
  )
  df
})
tf_df <- bind_rows(tf_dfs)
tf_df$gene <- factor(tf_df$gene, levels = tfs)
tf_df$direction <- ifelse(tf_df$gene %in% tfs_up, "activated", "repressed")
(tf_gg <- tf_df %>%
    ggplot(aes(pseudotime, expression, color = direction)) +
    geom_smooth() +
    facet_wrap(~gene, ncol = 4, dir = "v")) +
    scale_color_manual(values = c("#e9003a", "#285ba1")) +
    theme(legend.title = element_blank())
```


## Gene Ontology enrichment analysis (GO)

```{r}
set.seed(1234)
top_list <- list(
  down = sorted_down$symbol[1:500], 
  up = sorted_up$symbol[1:500], 
  universe = rownames(sorted)
)
top_entrez <- purrr::map(top_list, function(x) {
  entrez <- AnnotationDbi::select(
    x = org.Hs.eg.db, 
    keys = x, 
    keytype = "SYMBOL",
    columns = "ENTREZID"
  )$ENTREZID
  entrez <- entrez[!is.na(entrez)]
  entrez
})
top_go <- purrr::map(top_entrez[c("down", "up")], function(x) {
  out <- enrichGO(
    gene = x,
    universe = top_entrez$universe,
    OrgDb = org.Hs.eg.db,
    ont = "BP",
    pAdjustMethod = "BH",
    pvalueCutoff = 0.01,
    qvalueCutoff = 0.05,
    readable = TRUE
  )
  out
})
selected_terms_down <- c(
  "MHC protein complex assembly",
  "B cell activation"
)
selected_terms_up <- c(
  "response to endoplasmic reticulum stress",
  "ubiquitin-dependent ERAD pathway",
  "protein N-linked glycosylation",
  "Golgi vesicle transport",
  "protein folding",
  "COPII-coated vesicle budding",
  "regulation of B cell activation"
)


# Plot
go_df_down <- top_go$down@result[top_go$down@result$Description %in% selected_terms_down, ]
go_df_up <- top_go$up@result[top_go$up@result$Description %in% selected_terms_up, ]
go_df_down$direction <- "down"
go_df_up$direction <- "up"
go_df <- bind_rows(go_df_down, go_df_up)
DT::datatable(go_df)
go_df <- go_df %>% arrange(direction, p.adjust)
go_df$Description <- factor(go_df$Description, levels = rev(go_df$Description))
go_gg <- go_df %>% 
  ggplot(aes(Description, -1 * log10(p.adjust), color = direction)) +
    geom_segment(aes(x = Description, xend = Description, y = 0, yend = -1 * log10(p.adjust))) +
    geom_point() +
    scale_color_manual("", values = c("dodgerblue3", "firebrick3")) +
    labs(x = "", y = "-log10 (adjusted p-value)", color = "") +
    theme_light() +
    coord_flip()
```


Heatmap

```{r}
# Select genes of interest (GOI)
goi <- str_split(go_df$geneID, "/")
names(goi) <- go_df$Description
goi_dfs <- purrr::map2(goi, names(goi), function(l, x) {
  df <- data.frame(
    gene = l,
    term = x
  )
  df
})
goi_df <- bind_rows(goi_dfs)
```

Plot heatmap

```{r}
# Define matrix and sort cells by increasing pseudotime
pc_sce3 <- pc_sce2[, !is.na(pc_sce2$pseudotime_main_trajectory)]
pc_sce3 <- pc_sce3[, order(pc_sce3$pseudotime_main_trajectory)]
input_mat <- logcounts(pc_sce3)[goi_df$gene, ]
input_mat <- t(scale(t(input_mat), center = TRUE, scale = TRUE))


# Create column annotation
column_annot_df <- data.frame(
  cluster = pc_sce3$names_level_5_clusters,
  pseudotime = pc_sce3$pseudotime_main_trajectory
)
clusters <- unique(pc_sce3$names_level_5_clusters)
colors_clusters <- color_palette[1:length(clusters)]
names(colors_clusters) <- clusters
breaks_colors <- c(
  min(pc_sce3$pseudotime_main_trajectory),
  (min(pc_sce3$pseudotime_main_trajectory) + max(pc_sce3$pseudotime_main_trajectory)) / 2,
  max(pc_sce3$pseudotime_main_trajectory)
)
col_fun = colorRamp2(breaks_colors, c("#f4e653", "#498b8c", "#3f145a"))
col_annot <- HeatmapAnnotation(
  df = column_annot_df,
  col = list(
    cluster = colors_clusters,
    pseudotime = col_fun
  ),
  annotation_legend_param = list(
    labels_gp = gpar(fontsize = 7)
  )
)


# Create row annotation
row_annot_df <- goi_df[, "term", drop = FALSE]
colors_row_annot <- c("#a6cee3","#1f78b4","#b2df8a", "#33a02c", "#fb9a99",
                      "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6")
names(colors_row_annot) <- unique(row_annot_df$term)
row_annot <- rowAnnotation(
  df = row_annot_df,
  col = list(term = colors_row_annot),
  annotation_legend_param = list(
    labels_gp = gpar(fontsize = 7)
  )
)


# Plot and save heatmap 
heatmap <- Heatmap(
  as.matrix(input_mat),
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  show_column_names = FALSE,
  row_names_gp = gpar(fontsize = 4),
  top_annotation = col_annot,
  right_annotation = row_annot,
  heatmap_legend_param = list(
    labels_gp = gpar(fontsize = 7)
  )
)
in2mm <- 25.4
pdf("~/Desktop/diogenes/PC_with_csMBC_and_AIDcells_analysis/pc_heatmap_pseudo_time.pdf", width = 210/in2mm, height = 260/in2mm)
heatmap
dev.off()
```


TODO tomorrow: generate heatmap with ComplexHeatmap:

- Order cells by pseudotime
- Group genes by GO term
- Include GO annotations in rows
- Include cluster info in columns

+ generate lollipop plot GO




