---
title: "Calling Peaks"
author: "Paula Soler-Vila"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, out.width = "100%", fig.align = "center",
  message = FALSE, warning = FALSE
)
options(width = 1200)
```

# Introduction
Looking at the scATAC-seq coverage plots, it's possible to vislumbrate that, as a consequence of the merging process, some peaks show a considerable increase in size. This can create a problem for certain analysis such as motif enrichment analysis. In order to solve it, we decided to do a new round of peak calling using MACS2 for each group of cells annotated at level 1, helping to identify specifics peaks for rarer populations.  

# Pre-processing

## Load packages

```{r libraries} 
library(Seurat)
library(Signac)
library(reshape)
library(ggplot2)
library(GenomicRanges)
library(EnsDb.Hsapiens.v86)
library(stringr)
library(plotly)
```

## Parameters

```{r parameters}
cell_type = "NBC_MBC"

# Paths
path_to_obj <- paste0(
  here::here("scATAC-seq/results/R_objects/level_4/"),
  cell_type,
  "/",
  cell_type,
  "_annotated_level_4.rds",
  sep = ""
)

path_to_save <- paste0(
  here::here("scATAC-seq/results/R_objects/level_4/"),
  cell_type,
  "/",
  cell_type,
  "_annotated_peak_calling_level_4.rds",
  sep = ""
)

cols_nbc <- c(
  "NBC" = "#dcf0f4",
  "NBC early activation" = "#95d1de", 
  "NBC IFN-activated" = "#52828D",
  "NBC CD229+" = "#8C76B0", 
  "DZ-early S phase" = "#006FC4", 
  
  "Early GC-commited NBC" = "#75a5b0",
  "GC-commited NBC" = "#71ccb7",  
  
  "preGC" = "#69c291",
  "Proliferative NBC" = "#205b67",
  "GC DZ Noproli"  = "#588e1a", 

  "Early MBC" = "#eae2c6", 
  "ncsMBC" = "#dbcc9b", 
  "ncsMBC FCRL4+/FCRL5+" = "#bca146",
  "csMBC" = "#72663f",  
  "csMBC FCRL4+/FCRL5+" = "#a3925a", 
  "MBC FCRL5+" = "#352f1d"
)
```

## Functions

```{r functions}
peaks_quantification <- function(seurat_filtered, new_peaks){ 
    
    annotation <- GetGRangesFromEnsDb(ensdb = EnsDb.Hsapiens.v86)
    #seqlevelsStyle(annotation) <- "UCSC"
    ucsc.levels <- str_replace(string=paste("chr",seqlevels(annotation),sep=""), pattern="chrMT", replacement="chrM")
    seqlevels(annotation) <- ucsc.levels
    genome(annotation) <- "hg38"

    macs2_counts <- FeatureMatrix(
    fragments = Fragments(seurat_filtered),
    features = new_peaks,
    cells = colnames(seurat_filtered))
    
    seurat_filtered[["peaks_redefined"]] <- CreateChromatinAssay(
      counts = macs2_counts, 
     # genome = "hg38",
      fragments = Fragments(seurat_filtered),
      annotation = annotation)
    
    return(seurat_filtered)
}
```

# Load data

```{r}
seurat <- readRDS(path_to_obj)

DimPlot(
  seurat,
  cols = cols_nbc,
  label = F,
  group.by = "names_level_5_clusters_eta",
  pt.size = 0.1)
```


# New round of Peak Calling using MACS2 on each annotated cell type at level 5.

```{r}
peaks_level5 <- CallPeaks(
  object = seurat,
  group.by = "names_level_5_clusters_eta",
  macs2.path = "/Users/pauli/opt/anaconda3/envs/Tonsil_atlas/bin/macs2")
```

```{r}
# remove peaks on nonstandard chromosomes and in genomic blacklist regions
peaks_level5 <- keepStandardChromosomes(peaks_level5, pruning.mode = "coarse")
peaks_level5 <- subsetByOverlaps(x = peaks_level5, 
                         ranges = blacklist_hg38_unified, 
                         invert = TRUE)
```

# Quantification of the new peak counts in the entire dataset

```{r}
seurat <- peaks_quantification(seurat, peaks_level5)
```

```{r}
# remove the initial peaks annotation
DefaultAssay(seurat) <- 'peaks_redefined'
seurat[['peaks_macs']] <- NULL
seurat

saveRDS(seurat,path_to_save)
```

# Coverage plot
```{r}
CoveragePlot(seurat, 
             ranges = peaks_level5,
             region = "chr8-127736231-127742951")
```

# Session Information
```{r session}
sessionInfo()
```
