---
title: "Plasma Cell Annotation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=False, message=FALSE, warning=FALSE}
# Libraries
library(readxl)
library(dplyr)
library(ggplot2)
library(Seurat)
library(ggpubr)
library(harmony)
library(knitr)
library(kableExtra)
library(clusterProfiler)
library(biomaRt)
library(writexl)
library(scales)
```

```{r}
save.image(file = "level_3_tonsil.RData")
```
## Data exploration and quality control
```{r}
setwd("/Users/spalominoe/Documents/Single-Cell/TONSIL/Plasma Cell/New Data 2/")

data <- readRDS("PC_clustered_level_3_with_pre_freeze.rds")
```

```{r}
# Colors
color_palette <- c("black", "gray", "red", "yellow", "green4",
                   "blue", "mediumorchid2", "coral2", "blueviolet",
                   "indianred4", "deepskyblue1", "dimgray", "deeppink1",
                   "green", "lightgray", "hotpink1", "gold", "brown")
```

#### Proposed UMAPs
```{r}
# by cluster
DimPlot(data, group.by="seurat_clusters", label=TRUE) 
```
# Proposed annotation (1st iteration)
```{r}
data@meta.data$level_3 <- as.character(data@meta.data$seurat_clusters)

data@meta.data$level_3[data@meta.data$level_3=="0"] <- "IgG mature PC"
data@meta.data$level_3[data@meta.data$level_3=="1"] <- "GC Derived precursor 1"
data@meta.data$level_3[data@meta.data$level_3=="2"] <- "IgA/IgG mature PC"
data@meta.data$level_3[data@meta.data$level_3=="3"] <- "GC Derived precursor 2"
data@meta.data$level_3[data@meta.data$level_3=="4"] <- "GC Derived precursor 3"
data@meta.data$level_3[data@meta.data$level_3=="5"] <- "IgG plasmablast"
data@meta.data$level_3[data@meta.data$level_3=="6"] <- "Proliferating plasmablast"
data@meta.data$level_3[data@meta.data$level_3=="7"] <- "IgM"
data@meta.data$level_3[data@meta.data$level_3=="8"] <- "IgA adult"
data@meta.data$level_3[data@meta.data$level_3=="9"] <- "IgD"
data@meta.data$level_3[data@meta.data$level_3=="10"] <- "GC Derived precursor 0"
data@meta.data$level_3[data@meta.data$level_3=="11"] <- "Histone like"
data@meta.data$level_3[data@meta.data$level_3=="12"] <- "Tecnichal 1"
data@meta.data$level_3[data@meta.data$level_3=="13"] <- "Tecnichal 2"
data@meta.data$level_3[data@meta.data$level_3=="14"] <- "Doublets"

DimPlot(data, reduction = "umap", group.by = "level_3",label = TRUE)
```

# Proportion of cell in relation to the annotation pre-freeze
```{r}
proportions_df_cells <- find_proportions_df(
  data,
  x = "seurat_clusters",
  fill = "annotation_pre_freeze"
)

DT::datatable(proportions_df_cells)
stacked_barplot_cells <- plot_stacked_barplot(
  proportions_df_cells,
  x = "seurat_clusters",
  fill = "annotation_pre_freeze",
  colors = color_palette
)
stacked_barplot_cells
```

```{r}
# annotation pre freeze
DimPlot(data, group.by="annotation_pre_freeze") 
```

```{r}
# by age group
DimPlot(data, group.by="age_group") 
```

```{r}
# by sex
DimPlot(data, group.by="sex") 
```

```{r}
# by assay
DimPlot(data, group.by="assay") 
```

#### Cell cycle analysis visualization
```{r}
s_gg <- FeaturePlot(data, features = "S.Score") +
  scale_color_viridis_c(option = "magma")
s_gg
```

```{r}
g2m_gg <- FeaturePlot(data, features = "G2M.Score") +
  scale_color_viridis_c(option = "magma")
g2m_gg
```

#### QC metrics
```{r}
# Visualize QC metrics-1
VlnPlot(data,c("nCount_RNA","nFeature_RNA"))

FeaturePlot(data,c("nFeature_RNA","nCount_RNA"),cols = c("lightgray","blue"), reduction="umap")
```

```{r}
# Visualize QC metrics-2
FeaturePlot(data,c("pct_mt","pct_ribosomal", "nCount_RNA","nFeature_RNA"),cols = c("lightgray","blue"), reduction="umap")

VlnPlot(data,c("pct_mt","pct_ribosomal"))
```

```{r}
# Find markers
markers <- FindAllMarkers(
  data,
  logfc.threshold = 0.5,
  test.use = "wilcox",
  only.pos = TRUE,
  verbose = TRUE
)
markers <- markers %>%
  group_by(cluster) %>%
  arrange(desc(avg_log2FC), group_by = TRUE) %>%
  ungroup()
```
# B cell lineage markers
```{r}
# B cell lineage markers
FeaturePlot(data, c("REL", "PAX5", "BACH2", "IRF8"), cols = c("lightgray","blue"), reduction="umap", label = TRUE)
```
# Plasma cell lineage markers
```{r}
# Plasma cell lineage markers
FeaturePlot(data, c("IRF4", "XBP1", "PRDM1", "BLIMP1"), cols = c("lightgray","blue"), reduction="umap", label = TRUE)
```
# Immunoglobulines distribution
```{r}
# Igs
ig <- markers[grepl("^IG", markers$gene),]

FeaturePlot(data, c("IGHA1","IGHA2", "IGHG1", "IGHG2", "IGHG3", "IGHG4","IGHGP", "IGHD", "IGHM" ), cols = c("lightgray","blue"), reduction="umap", label = TRUE)
```
# GC Derive Precursors: 5 - 10 - 1 - 3 - 4
Comparison of markers between these clusters:
```{r}
cl5vs10 <- FindMarkers(data, ident.1 = 5, ident.2 = 10, min.pct = 0.5, logfc.threshold = 0.5)
cl5vs4 <- FindMarkers(data, ident.1 = 5, ident.2 = 4, min.pct = 0.5, logfc.threshold = 0.5)
cl5vs1 <- FindMarkers(data, ident.1 = 5, ident.2 = 1, min.pct = 0.5, logfc.threshold = 0.5)
cl5vs3 <- FindMarkers(data, ident.1 = 5, ident.2 = 3, min.pct = 0.5, logfc.threshold = 0.5)

######
cl10vs4 <- FindMarkers(data, ident.1 = 10, ident.2 = 4, min.pct = 0.5, logfc.threshold = 0.5)
cl10vs3 <- FindMarkers(data, ident.1 = 10, ident.2 = 3, min.pct = 0.5, logfc.threshold = 0.5)
cl10vs1 <- FindMarkers(data, ident.1 = 10, ident.2 = 1, min.pct = 0.5, logfc.threshold = 0.5)

#####
cl4vs3 <- FindMarkers(data, ident.1 = 4, ident.2 = 3, min.pct = 0.5, logfc.threshold = 0.5)
cl4vs1 <- FindMarkers(data, ident.1 = 4, ident.2 = 1, min.pct = 0.5, logfc.threshold = 0.5)

#####
cl3vs1 <- FindMarkers(data, ident.1 = 3, ident.2 = 1, min.pct = 0.5, logfc.threshold = 0.5)
```

```{r}
# CLUSTER 5 IgG-positive plasmablasts precursors 
FeaturePlot(data, c("CIITA", "AFF3", "SPIB", "PAX5", "FCMR", "TRIM22","REL" ,"BANK1","MS4A1" ), cols = c("lightgray","blue"), reduction="umap")
```

```{r}
# CLUSTER 10 GC Derive Precursors 0
FeaturePlot(data, c("CD19", "CD38", "CD53","LRMP", "CD74", "HMCES", "BCL7A"), cols = c("lightgray","blue"), reduction="umap")
```

```{r}
# CLUSTER 1 GC Derive Precursors 1
FeaturePlot(data, c("CD19", "CD38", "CD53","LRMP", "CD74", "HMCES", "BCL7A"), cols = c("lightgray","blue"), reduction="umap")
```

```{r}
# CLUSTER 3 GC Derive Precursors 2
FeaturePlot(data, c("IGHG1", "CD20", "CD79B"), cols = c("lightgray","blue"), reduction="umap")
```

```{r}
# CLUSTER 4 GC Derive Precursors 3
FeaturePlot(data, c("IGHG1", "CD20", "CD79B"), cols = c("lightgray","blue"), reduction="umap")
```


#### let's explore clusters 6,7,8,9 (son los que me tocan)
```{r}
#CLUSTER 6
FeaturePlot(data, c("MKI67", "PCNA","EZH2","CKS2", "FDPS"), cols = c("lightgray","blue"), reduction="umap")

FeaturePlot(data, c("BCL6", "BACH2", "PAX5" ,"XBP1", "IRF4", "PRDM1", "BLIMP1"), cols = c("lightgray","blue"), reduction="umap")

FeaturePlot(data, c("CKS2", "EZH2", "PCNA"), cols = c("lightgray","blue"), reduction="umap")
```

```{r}
#CLUSTER 7
FeaturePlot(data, c("IGHM","HLA-A"), cols = c("lightgray","blue"), reduction="umap")

FeaturePlot(data, c("IGHG1", "IGHG2", "IGHG3"), cols = c("lightgray","blue"), reduction="umap")


FeaturePlot(data, c("ZNF804A",	"LINC01480",	"EML6",	"ITPR2",	"DUSP2",	"CCDC88A",		"IGFLR1",	"HCST",	"FCRL3",	"ARHGAP15"), cols = c("lightgray","blue"), reduction="umap")
```

```{r}
#CLUSTER 8
cl8 <- subset(data, subset = seurat_clusters=="8")
FeaturePlot(cl8, c("IGHM","IGHA1", "IGHA2", "IGHD","JCHAIN","IGLC1","IGLC2" ), cols = c("lightgray","blue"), reduction="umap")

DimPlot(cl8, group.by="age_group")
DimPlot(cl8, group.by="assay")
```

```{r}
#CLUSTER 9
FeaturePlot(data, c("IGHD", "IGHM"), cols = c("lightgray","blue"), reduction="umap")
cl9 <- subset(data, subset = seurat_clusters=="9")

DimPlot(data, group.by="seurat_clusters", label=TRUE) 

DimPlot(cl9, group.by="seurat_clusters") 
```

## cluster 2,8,0
```{r}
DimPlot(subset(x = data, idents = c("0" ,"2" ,"8")), group.by="seurat_clusters", label=TRUE)


FeaturePlot(subset(x = data, idents = c("0" ,"2" ,"8")),c("pct_mt","pct_ribosomal", "nCount_RNA","nFeature_RNA"),cols = c("lightgray","blue"), reduction="umap")
```

```{r}
#subclustering cluster 2
data <- FindSubCluster(
  data,
  cluster = "2",
  graph.name = "RNA_snn",
  resolution = 0.1,
  subcluster.name = "annotation_level_4"
)
Idents(data) <- "annotation_level_4"

table(data@meta.data$annotation_level_4)

DimPlot(subset(x = data, idents = c("0" ,"2_0","2_1" ,"8")), group.by="annotation_level_4", label=TRUE)
```
```{r}
doubts <- subset(x = data, idents = c("0" ,"2_0","2_1" ,"8"))

# Find markers
markers_doubts <- FindAllMarkers(
  doubts,
  logfc.threshold = 0.5,
  test.use = "wilcox",
  only.pos = TRUE,
  verbose = TRUE
)

markers_doubts <- markers_doubts %>%
  group_by(cluster) %>%
  arrange(desc(avg_log2FC), group_by = TRUE) %>%
  ungroup()

table(markers_doubts$cluster)
```

```{r}
FeaturePlot(subset(x = data, idents = c("2_0","2_1")), c("IGHG1", "IGHG2","IGHA1", "IGHA2"),cols = c("lightgray","blue"), reduction="umap", label=TRUE)

DimPlot(subset(x = data, idents = c("2_0","2_1")), group.by = "age_group", reduction="umap", label=TRUE)
```

```{r}
FeaturePlot(ig_doubts, c("IGHG","IGHG1","IGHG3", "IGHG2", "IGHG4"), cols = c("lightgray","blue"), reduction="umap", label=TRUE)
```

```{r}
FeaturePlot(ig_doubts, c("IGHA","IGHA1","IGHA2"), cols = c("lightgray","blue"), reduction="umap", label=TRUE)
```

```{r}
cl2_0vs0 <- FindMarkers(ig_doubts, ident.1 = "2_0", ident.2 = c(0), min.pct = 0.5, logfc.threshold = 0.5)
```

```{r}
cl2_0vs2_1 <- FindMarkers(ig_doubts, ident.1 = "2_0", ident.2 = "2_1", min.pct = 0.5, logfc.threshold = 0.5)
```

```{r}
cl2_0vs8 <- FindMarkers(ig_doubts, ident.1 = "2_0", ident.2 = "8", min.pct = 0.5, logfc.threshold = 0.5)
```

```{r}
cl2_1vs8 <- FindMarkers(ig_doubts, ident.1 = "2_1", ident.2 = 8, min.pct = 0.5, logfc.threshold = 0.5)
head(cl2_1vs8, n = 5)

cl2_1vs8$gene <- rownames(cl2_1vs8)
write.xlsx(cl2_1vs8, 'cl2_1vs8.xlsx', row.names=TRUE)
```

```{r}
cl0vs8 <- FindMarkers(ig_doubts, ident.1 = 0, ident.2 = 8, min.pct = 0.5, logfc.threshold = 0.5)
head(cl0vs8, n = 5)

cl0vs8$gene <- rownames(cl0vs8)
write.xlsx(cl0vs8, 'cl0vs8.xlsx', row.names=TRUE)
```

```{r}
#subclustering cluster IgM
data <- FindSubCluster(
  data,
  cluster = "7",
  graph.name = "RNA_snn",
  resolution = 0.1,
  subcluster.name = "annotation_level_4"
)
Idents(data) <- "annotation_level_4"

table(data@meta.data$annotation_level_4)

DimPlot(subset(x = data, idents = c("7_0","7_1","7_2")), group.by="annotation_level_4", label=TRUE)
```
```{r}
data@meta.data$annotation_level_4[data@meta.data$annotation_level_4=="0"] <- "IgG mature PC"
data@meta.data$annotation_level_4[data@meta.data$annotation_level_4=="1"] <- "GC Derived precursor 1"
data@meta.data$annotation_level_4[data@meta.data$annotation_level_4=="2_0"] <- "IgA mature PC"

data@meta.data$annotation_level_4[data@meta.data$annotation_level_4=="2_1"] <- "IgG mature PC"

data@meta.data$annotation_level_4[data@meta.data$annotation_level_4=="3"] <- "GC Derived precursor 2"
data@meta.data$annotation_level_4[data@meta.data$annotation_level_4=="4"] <- "GC Derived precursor 3"
data@meta.data$annotation_level_4[data@meta.data$annotation_level_4=="5"] <- "IgG plasmablast"
data@meta.data$annotation_level_4[data@meta.data$annotation_level_4=="6"] <- "Proliferating plasmablast"
data@meta.data$annotation_level_4[data@meta.data$annotation_level_4=="7"] <- "IgM"
data@meta.data$annotation_level_4[data@meta.data$annotation_level_4=="8"] <- "IgA mature PC"
data@meta.data$annotation_level_4[data@meta.data$annotation_level_4=="9"] <- "IgD"
data@meta.data$annotation_level_4[data@meta.data$annotation_level_4=="10"] <- "GC Derived precursor 0"
data@meta.data$annotation_level_4[data@meta.data$annotation_level_4=="11"] <- "Histone like"
data@meta.data$annotation_level_4[data@meta.data$annotation_level_4=="12"] <- "Tecnichal 1"
data@meta.data$annotation_level_4[data@meta.data$annotation_level_4=="13"] <- "Tecnichal 2"
data@meta.data$annotation_level_4[data@meta.data$annotation_level_4=="14"] <- "Doublets"

DimPlot(data, group.by = "annotation_level_4")

saveRDS(data, file = "PC_level3_annotated.rds")
```
### Cleaning in relation to ncount, nfeatures, mt and rb 
```{r}
# There is no "explicit" way to select the thresholds, so let's explore:
## mt
ggscatterhist(data@meta.data, x = "nCount_RNA", y = "nFeature_RNA", color = "pct_mt", 
                  size = 1, alpha = 0.6, 
                  margin.params = list(fill="black",color = "black", size = 0.2))

## rb
ggscatterhist(data@meta.data, x = "nCount_RNA", y = "nFeature_RNA",color = "pct_ribosomal", 
              size = 1, alpha = 0.6, 
              margin.params = list(fill="black",color = "black", size = 0.2))

```

```{r}
plot(data@meta.data$nCount_RNA, data@meta.data$nFeature_RNA,pch=16,bty="n")
abline(h=c(1000,7000),v=c(2500,5e04),col="red",lty=2)
```


####===================================
####===================================
####===================================
#### Recompute thresholds
```{r}
data2 <- subset(data, subset = nFeature_RNA > 1000 & nFeature_RNA < 5000 & nCount_RNA > 2500 & nCount_RNA < 5e04 & pct_mt < 8 & pct_ribosomal <10)

FeaturePlot(data2,c("nFeature_RNA","nCount_RNA", "pct_mt", "pct_ribosomal"),cols = c("lightgray","blue"), reduction="umap")
```

### Remove technical-doublets clusters  12,13,14
```{r}
data2 <- subset(x = data2, idents = c("12" ,"13" ,"14"), invert = TRUE)
DimPlot(data2, group.by="seurat_clusters", label=TRUE) 
```

#### Normalize data, and adjust 3P and multiome
```{r}
seurat_list <- SplitObject(data2, split.by = "assay")
  seurat_list <- seurat_list[c("3P", "multiome")]
  seurat_list <- purrr::map(
    seurat_list,
    FindVariableFeatures,
    nfeatures = 5000
  )
  
  hvg <- purrr::map(seurat_list, VariableFeatures)
  shared_hvg <- intersect(hvg$`3P`, hvg$multiome)
  
ElbowPlot(data2)
```

```{r}
data2@meta.data$assay <- as.factor(data2@meta.data$assay)

data3 <- data2 %>%
      ScaleData(features = shared_hvg) %>%
      RunPCA(features = shared_hvg) %>%
      RunHarmony(group.by.vars = "assay", reduction ="pca", dims = 1:12)
```

```{r}
data3 <- RunUMAP(data3, reduction = "harmony", dims = 1:12)
data3 <- FindNeighbors(data3, reduction = "harmony", dims = 1:12)
```

```{r}
DimPlot(data3, group.by= "assay", reduction="umap")

DimPlot(data3, group.by= "gem_id", reduction="umap")
```

```{r}
data3 <- FindClusters(data3, resolution = 0.7)
DimPlot(data3, reduction = "umap")

data3 <- FindVariableFeatures(data3)

```

```{r}
list <- SplitObject(data, split.by = "level_3")

list <- SplitObject(data, split.by = "annotation_level_4")

```

```{r}
tonsil.anchors <- FindTransferAnchors(reference = data, query = data3,
                                        dims = 1:12, project.query = T)

predictions <- TransferData(anchorset = tonsil.anchors, refdata = data$annotation_level_4,
                            dims = 1:12)

data3 <- AddMetaData(data3, metadata = predictions)
```

### See the generated plot with predicted ID
```{r}
DimPlot(data3, group.by="predicted.id", label=TRUE) 

data_predict.id <- data3
```

```{r}
DimPlot(data3, group.by="seurat_clusters", label=TRUE) 
```

```{r}
DimPlot(data3, group.by="assay", label=TRUE) 
```

```{r}
DimPlot(data3, group.by="age_group", label=TRUE) 
```


```{r}
proportions_df_cells <- find_proportions_df(
  data3,
  x = "seurat_clusters",
  fill = "predicted.id"
)

DT::datatable(proportions_df_cells)
stacked_barplot_cells <- plot_stacked_barplot(
  proportions_df_cells,
  x = "seurat_clusters",
  fill = "predicted.id",
  colors = color_palette
)
stacked_barplot_cells
```
```{r}
data3@meta.data$seurat_clusters2 <- as.character(data3@meta.data$predicted.id)
```

```{r}
proportions_df_cells <- find_proportions_df(
  data3,
  x = "seurat_clusters2",
  fill = "predicted.id"
)

DT::datatable(proportions_df_cells)
stacked_barplot_cells <- plot_stacked_barplot(
  proportions_df_cells,
  x = "seurat_clusters2",
  fill = "predicted.id",
  colors = color_palette
)
stacked_barplot_cells
```

```{r}
# Find markers
Idents(data3) <- "predicted.id"

markers3 <- FindAllMarkers(
  data3,
  logfc.threshold = 0.5,
  test.use = "wilcox",
  only.pos = TRUE,
  verbose = TRUE
)
markers3 <- markers3 %>%
  group_by(cluster) %>%
  arrange(desc(avg_log2FC), group_by = TRUE) %>%
  ungroup()
```

```{r}
IgA_IgG <- FindMarkers(data3, ident.1 = "IgA/IgG mature PC", ident.2 = "IgA adult", min.pct = 0.5, logfc.threshold = 0.5)
head(IgA_IgG, n = 5)

IgA_IgG$gene <- rownames(IgA_IgG)

write.xlsx(IgA_IgG, 'IgA_IgG.xlsx')
```

### Marker exploration
```{r}
# Plasma Cell linage
FeaturePlot(data3, c("IRF4", "XBP1", "PRDM1"), cols = c("lightgray","blue"), reduction="umap")
```

```{r}
# B cell lineage markers
FeaturePlot(data3, c("REL", "PAX5", "BACH2", "IRF8"), cols = c("lightgray","blue"), reduction="umap")
```
```{r}
# Immunoglobulines
FeaturePlot(data3, c("IGHM", "IGHD", "IGHA1", "IGHA2"), cols = c("lightgray","blue"), reduction="umap")
```

```{r}
# Immunoglobulines
FeaturePlot(data3, c("IGHG1", "IGHG2", "IGHG3", "IGHG4"), cols = c("lightgray","blue"), reduction="umap")
```
