---
title: "Naive and Memory B cells level 4 to level 5"
author: "Sergio Aguilar Fernández"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  BiocStyle::html_document:
    df_print: paged
    toc: true
    toc_float: true
    number_sections: false
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center',
                      message=FALSE, warning = FALSE)
```


## Introduction

In this Rmd we are going to clean the level 4 of the __*Naive B Cells and Memory B Cells (NBC_MBC)*__ and include DZ no proliferative cells from *GCBC*(to overcome proliferation influence in the umap obtention) cells to analyze the trajectory from a NBC. H

## Libraries 

```{r}
library(Seurat)
library(tidyverse)
library(harmony)
library(knitr)
```


## Parameters 

### Folder structure
```{r}
level <- "level_4"
cell_type <- "NBC_MBC"
base_folder <- here::here(paste0(level,"/",cell_type,"/"))
  data_folder <- paste0(base_folder, "data/")
  results_folder <- paste0(base_folder, "results/")
  figures_folder <- paste0(results_folder, "figures/")
  markers_folder <- paste0(results_folder, "objects/markers/")

level_5_folder = here::here(paste0("level_5/",cell_type,"/data/"))
```

### Palette and pt.size
```{r}
palette<-c('#e6194b', '#3cb44b', '#ffd8b1', '#4363d8', '#f58231', '#911eb4', '#46f0f0', '#f032e6', '#bcf60c', '#fabebe', '#008080', '#e6beff', '#9a6324', '#fffac8', '#800000', '#aaffc3', '#808000', '#ffe119', '#000075', '#808080', '#ffffff', '#000000')

pt_sizes <- c(
  NBC_MBC = 0.6,
  GCBC = 0.8,
  PC = 1.5
)
pt_size <- pt_sizes[cell_type]
```

## Load data 
```{r}
seurat_object_l4 <- readRDS(paste0(data_folder, cell_type, "_clustered_level_4.rds"))
```

## Functions

### Visualization exploratory analysis of a given object 

#### Data groups visualization
```{r}
data_groups_visualization <- function(seurat_object, group_variable_1, group_variable_2 = NULL, pt_size) {
        print(Seurat::DimPlot(seurat_object, 
                reduction = "umap",
                pt.size = pt_size,  
                cols= palette, 
                group.by = group_variable_1, 
                raster = FALSE) 
        )
        
        print(Seurat::DimPlot(seurat_object, 
                reduction = "umap",
                pt.size = pt_size,  
                cols= palette, 
                group.by = group_variable_1, 
                raster = FALSE, 
                label = TRUE)
        )
        
        if (!is.null(group_variable_2)) {
            print(Seurat::DimPlot(seurat_object, 
                reduction = "umap",
                pt.size = pt_size,  
                cols= palette, 
                group.by = group_variable_2, 
                raster = FALSE) 
            )
        }
        

        
        #### COMBINED EXPLORATORY DATA GROUPS ANALYSIS
        p1 <- Seurat::DimPlot(seurat_object, 
                reduction = "umap",
                pt.size = pt_size/3,  
                cols= palette, 
                group.by = group_variable_1, 
                raster = FALSE) 
        
        p2 <- Seurat::DimPlot(seurat_object, 
                reduction = "umap", 
                pt.size = pt_size/3,  
                cols= c("tan", "royalblue1", "red"), 
                group.by = "assay", 
                raster = FALSE) 
        
        p3 <- Seurat::DimPlot(seurat_object, 
                reduction = "umap", 
                pt.size = pt_size/3,  
                cols= palette, 
                group.by = "age", 
                raster = FALSE) 
        
        p4 <- Seurat::DimPlot(seurat_object, 
                reduction = "umap", 
                pt.size = pt_size/3,  
                cols= c("tan", "royalblue1", "red"), 
                group.by = "hospital", 
                raster = FALSE) 
        
        
        print(patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2) )
}
```

#### Data features Visualization
```{r}
data_features_visualization <- function(seurat_object, group_variable_1, pt_size) {

    p1 <- Seurat::FeaturePlot(seurat_object, 
                        features = "nCount_RNA", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        raster = FALSE) 
    
    p2 <- Seurat::FeaturePlot(seurat_object, 
                        features = "nFeature_RNA", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        raster = FALSE) 
    
    p3 <- Seurat::FeaturePlot(seurat_object, 
                        features = "pct_mt", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        raster = FALSE) 
    
    p4 <- Seurat::FeaturePlot(seurat_object, 
                        features = "pct_ribosomal", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        raster = FALSE) 
    
    
      print(patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2) )

    p1 <- Seurat::DimPlot(seurat_object, 
            reduction = "umap",
            pt.size = pt_size/3,  
            cols= palette, 
            group.by = group_variable_1, 
            raster = FALSE) 
    
    p2 <- Seurat::FeaturePlot(seurat_object, 
                        features = "pct_mt", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        raster = FALSE) 
    
    p3 <- Seurat::FeaturePlot(seurat_object, 
                        features = "nCount_RNA", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        raster = FALSE) 
    
    p4 <- Seurat::FeaturePlot(seurat_object, 
                        features = "nFeature_RNA", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        raster = FALSE) 
    
     print(patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2) )
    
    
    p1 <- Seurat::FeaturePlot(seurat_object, 
                        features = "nCount_RNA", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        min.cutoff = 'q10',
                        max.cutoff = 'q90', 
                        raster = FALSE) 
                        
    
    p2 <- Seurat::FeaturePlot(seurat_object, 
                        features = "nFeature_RNA", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        min.cutoff = 'q10',
                        max.cutoff = 'q90', 
                        raster = FALSE) 
                        
    
      print(patchwork::wrap_plots(p1, p2, ncol = 2) )
      
    g1 <- ggplot(seurat_object@meta.data, aes_string(x=group_variable_1, y="nFeature_RNA", fill=group_variable_1)) + 
        geom_boxplot() + 
        theme(legend.position = "none") + 
        scale_fill_manual(values=palette) + 
        theme(axis.text.x = element_text(angle = 90))
    g2 <- ggplot(seurat_object@meta.data, aes_string(x=group_variable_1, y="nCount_RNA", fill=group_variable_1)) + 
        geom_boxplot() + 
        scale_fill_manual(values=palette) + 
        theme(axis.text.x = element_text(angle = 90))
    
      print(patchwork::wrap_plots(g1, g2, ncol = 2) )

}
```

## Exploratory analysis of level 4 

### Explore the UMAP and different grouping variables 
```{r level4_exploratory_analysis_group_variables, fig.height = 12, fig.width = 16}
pdf(paste(figures_folder, "exploratory_analysis/level4_exploratory_analysis_data_groups.pdf", sep=""),width=20,height=10,paper='special') 
data_groups_visualization(seurat_object = seurat_object_l4, 
                          group_variable_1 = "annotation_level_3",
                          group_variable_2 = "naive_memory", 
                          pt_size = pt_size) 
dev.off()
```


### Explore the UMAP and different data variables 

```{r level4_exploratory_analysis_data_variables, fig.height = 12, fig.width = 16}
pdf(paste(figures_folder, "exploratory_analysis/level4_exploratory_analysis_data_variables.pdf", sep=""),width=20,height=10,paper='special') 
data_features_visualization(seurat_object = seurat_object_l4, 
                            group_variable_1 = "annotation_level_3",  
                            pt_size = pt_size) 
dev.off()
```

### Explore the doublet scores 
```{r level4_exploratory_analysis_data_variables, fig.height = 12, fig.width = 16}
pdf(paste(figures_folder, "exploratory_analysis/level4_exploratory_analysis_doublets.pdf", sep=""),width=20,height=10,paper='special')

p1 <- Seurat::DimPlot(seurat_object_l4, 
        reduction = "umap",
        pt.size = pt_size/3,  
        cols= palette, 
        group.by = "annotation_level_3", 
        raster = FALSE) 
  
p2 <- Seurat::FeaturePlot(seurat_object_l4, 
                    features = "scrublet_doublet_scores", 
                    reduction="umap", 
                    pt.size = pt_size/3, 
                    raster = FALSE) 
                    

p3 <- Seurat::FeaturePlot(seurat_object_l4, 
                    features = "scrublet_doublet_scores_scaled", 
                    reduction="umap", 
                    pt.size = pt_size/3, 
                    raster = FALSE) 

p4 <- Seurat::DimPlot(seurat_object_l4, 
        reduction = "umap",
        pt.size = pt_size/3,  
        cols= palette, 
        group.by = "scrublet_predicted_doublet", 
        raster = FALSE) 

patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)

cluster_name <- "unclear_doublets"
cluster_cells <- Cells(seurat_object_l4)[which(seurat_object_l4$annotation_level_3 == cluster_name)]
p1 <- Seurat::DimPlot(seurat_object_l4, 
        reduction = "umap",
        pt.size = pt_size/3,  
        cols= palette, 
        cells=cluster_cells, 
        group.by = "annotation_level_3", 
        raster = FALSE) 
  
p2 <- Seurat::FeaturePlot(seurat_object_l4, 
                    features = "scrublet_doublet_scores", 
                    reduction="umap", 
                    pt.size = pt_size/3, 
                    cells=cluster_cells, 
                    raster = FALSE) 
                    

p3 <- Seurat::FeaturePlot(seurat_object_l4, 
                    features = "scrublet_doublet_scores_scaled", 
                    reduction="umap", 
                    pt.size = pt_size/3, 
                    cells=cluster_cells, 
                    raster = FALSE) 

p4 <- Seurat::DimPlot(seurat_object_l4, 
        reduction = "umap",
        pt.size = pt_size/3,  
        cols= palette, 
        cells=cluster_cells, 
        group.by = "scrublet_predicted_doublet", 
        raster = FALSE) 

patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)

dev.off()
```


### Selection of the resolution 

#### Clusters 
```{r fig.height = 12, fig.width = 16}
pdf(paste(figures_folder, "exploratory_analysis/clusters_different_resolutions3.pdf", sep=""),width=20,height=10,paper='special') 
p1 <- Seurat::DimPlot(seurat_object_l4, 
        reduction = "umap",
        pt.size = pt_size/3,  
        cols= palette, 
        group.by = "annotation_level_3", 
        raster = FALSE) 
print(p1)

#resolution_vector <- c(0.025, 0.05, 0.1, 0.25, 0.35,  0.4, 0.46, 0.55, 0.6, 0.65, 0.75, 1)
resolution_vector <- c(0.35, 0.45, 0.46, 0.5, 0.55)

list_plots <- list()
for (index in 1:length(resolution_vector)) {
  resolution <- resolution_vector[index]
  resolution_wanted <- paste0("RNA_snn_res.", resolution)
  Idents(seurat_object_l4) <- resolution_wanted
  p <- Seurat::DimPlot(seurat_object_l4, reduction = "umap",  pt.size = pt_size/3, cols = c(palette, "black"), raster = FALSE, group.by = resolution_wanted)
  list_plots[[index]] <- p
}
print(patchwork::wrap_plots(list_plots))

dev.off()
```

```{r}
pdf(paste(figures_folder, "exploratory_analysis/selected_resolution.pdf", sep=""),width=20,height=10,paper='special') 
p1 <- Seurat::DimPlot(seurat_object_l4, 
        reduction = "umap",
        pt.size = pt_size/3,  
        cols= palette, 
        group.by = "annotation_level_3", 
        raster = FALSE) 
print(p1)

p2 <- Seurat::DimPlot(seurat_object_l4, 
        reduction = "umap",
        pt.size = pt_size/3,  
        cols= palette, 
        group.by = "RNA_snn_res.0.46", 
        raster = FALSE) 
print(p2)

print(patchwork::wrap_plots(p1, p2, ncol = 2))
dev.off()

```

```{r}
pdf(paste(figures_folder, "selected_resolution/quality_analysis.pdf", sep=""),width=20,height=10,paper='special') 
data_features_visualization(seurat_object = seurat_object_l4, 
                            group_variable_1 = "RNA_snn_res.0.46",  
                            pt_size = pt_size) 
dev.off()

```

```{r}
pdf(paste(figures_folder, "selected_resolution/selected_resolution_correspondence_level_3.pdf", sep=""),width=20,height=10,paper='special') 
p1 <- Seurat::DimPlot(seurat_object_l4, 
        reduction = "umap",
        pt.size = pt_size/3,  
        cols= palette, 
        group.by = "annotation_level_3", 
        raster = FALSE) 
print(p1)

p2 <- Seurat::DimPlot(seurat_object_l4, 
        reduction = "umap",
        pt.size = pt_size/3,  
        cols= palette, 
        group.by = "RNA_snn_res.0.46", 
        raster = FALSE) 
print(p2)

print(patchwork::wrap_plots(p1, p2, ncol = 2))

cell_type_pcts_df <- seurat_object_l4@meta.data %>%
  group_by(RNA_snn_res.0.46, annotation_level_3) %>%
  summarise(n_cells_cell_type = n()) %>% 
  group_by(RNA_snn_res.0.46) %>%
  mutate(
    total_cells = sum(n_cells_cell_type),
    pct_cells = n_cells_cell_type / total_cells * 100
  )

cell_type_pcts_gg <- cell_type_pcts_df %>%
  ggplot(aes(RNA_snn_res.0.46, pct_cells, fill = annotation_level_3)) +
    geom_col() +
    labs(x = "Clusters resolution 0.46", y = "Percentage of cells (%)", fill = "") +
    scale_fill_manual(values = palette) +
    theme_classic()

print(cell_type_pcts_gg)

dev.off()

pdf(paste(figures_folder, "selected_resolution/selected_resolution_quality_analysis.pdf", sep=""),width=20,height=10,paper='special') 
data_features_visualization(seurat_object = seurat_object_l4, 
                            group_variable_1 = "RNA_snn_res.0.46",  
                            pt_size = pt_size) 
dev.off()
```



## Find desired clusters

#### Cluster 1

```{r}
cluster_name <- "1"
subcluster_name <- paste0("cluster",cluster_name, "_subcluster")
```


##### Subclustering 

```{r}
### Find subcluster 
Idents(seurat_object_l4) <- "RNA_snn_res.0.46"
seurat_object_l4 <- FindSubCluster(
  seurat_object_l4,
  cluster = cluster_name,
  graph.name = "RNA_snn",
  subcluster.name = subcluster_name,
  resolution = 0.1
)
```

```{r}
cluster_cells <- Cells(seurat_object_l4)[which(seurat_object_l4$RNA_snn_res.0.46 == cluster_name)]
```

##### Visualization analysis
```{r}
pdf(paste(figures_folder, "subclustering/",cluster_name,"_subclustering.pdf", sep=""),width=20,height=10,paper='special') 
Seurat::DimPlot(seurat_object_l4, 
        reduction = "umap", 
        pt.size = pt_size,  
        cols= palette, 
        group.by = subcluster_name, 
        raster = FALSE) 

Seurat::DimPlot(seurat_object_l4, 
        reduction = "umap", 
        pt.size = pt_size,  
        cols= palette, 
        group.by = subcluster_name, 
        cells=cluster_cells)
dev.off()
```

#### Cluster 7

```{r}
cluster_name <- "7"
subcluster_name <- paste0("cluster",cluster_name, "_subcluster")
```


##### Subclustering 

```{r}
### Find subcluster 
Idents(seurat_object_l4) <- "cluster1_subcluster"
seurat_object_l4 <- FindSubCluster(
  seurat_object_l4,
  cluster = cluster_name,
  graph.name = "RNA_snn",
  subcluster.name = subcluster_name,
  resolution = 0.1
)
```

```{r}
cluster_cells <- Cells(seurat_object_l4)[which(seurat_object_l4$RNA_snn_res.0.46 == cluster_name)]
```

##### Visualization analysis
```{r}
pdf(paste(figures_folder, "subclustering/",cluster_name,"_subclustering.pdf", sep=""),width=20,height=10,paper='special') 
Seurat::DimPlot(seurat_object_l4, 
        reduction = "umap", 
        pt.size = pt_size,  
        cols= palette, 
        group.by = subcluster_name, 
        raster = FALSE) 

Seurat::DimPlot(seurat_object_l4, 
        reduction = "umap", 
        pt.size = pt_size,  
        cols= palette, 
        group.by = subcluster_name, 
        cells=cluster_cells)
dev.off()
```

#### Clusters 5-8-12 in UMAP from level 1

```{r}
cluster_cells <- Cells(seurat_object_l4)[which(seurat_object_l4$RNA_snn_res.0.46 %in% c(5,8,12))]
```

```{r}
umap <- seurat_object_l4@meta.data[,c("UMAP_1_level_1", "UMAP_2_level_1")]
seurat_object_l4[["level1_umap"]] <- CreateDimReducObject(embeddings = as.matrix(umap), key = "UMAP_", assay = "RNA")
colnames(x = seurat_object_l4[["level1_umap"]]@cell.embeddings) <- paste0("UMAP_", 1:2)

```


```{r}
pdf(paste(figures_folder, "subclustering/clusters5_8_12_umaplevel1.pdf", sep=""),width=20,height=10,paper='special') 
p1 <- Seurat::DimPlot(seurat_object_l4, 
        reduction = "level1_umap", 
        pt.size = pt_size,  
        cols= palette, 
        group.by = "cluster7_subcluster", 
        raster = FALSE) 
print(p1)
p2 <- Seurat::DimPlot(seurat_object_l4, 
        reduction = "level1_umap", 
        pt.size = pt_size,  
        cols= c("red", "lightblue", "gold"), 
        group.by = "cluster7_subcluster", 
        cells=cluster_cells)
print(p2)
patchwork::wrap_plots(p1, p2, ncol = 2)


dev.off()
```




#### Cluster 8

```{r}
cluster_name <- "8"
subcluster_name <- paste0("cluster",cluster_name, "_subcluster")
```


##### Subclustering 

```{r}
### Find subcluster 
Idents(seurat_object_l4) <- "cluster7_subcluster"
seurat_object_l4 <- FindSubCluster(
  seurat_object_l4,
  cluster = cluster_name,
  graph.name = "RNA_snn",
  subcluster.name = "Cluster8_subcluster_0.25res",
  resolution = 0.25
)
```

```{r}
cluster_cells <- Cells(seurat_object_l4)[which(seurat_object_l4$RNA_snn_res.0.46 == cluster_name)]
```

##### Visualization analysis
```{r}
pdf(paste(figures_folder, "subclustering/",cluster_name,"_subclustering.pdf", sep=""),width=20,height=10,paper='special') 
Seurat::DimPlot(seurat_object_l4, 
        reduction = "umap", 
        pt.size = pt_size,  
        cols= palette, 
        group.by = subcluster_name, 
        raster = FALSE) 

p2 <- Seurat::DimPlot(seurat_object_l4, 
        reduction = "umap", 
        pt.size = pt_size,  
        cols= palette, 
        group.by = subcluster_name, 
        cells=cluster_cells)
print(p2)

p3 <- Seurat::FeaturePlot(seurat_object_l4, 
                    features = "MEF2B", 
                    reduction="umap", 
                    pt.size = pt_size,  
                    cells=cluster_cells, 
                    raster = FALSE) 
print(p3)
patchwork::wrap_plots(p2, p3, ncol = 2)


dev.off()
```

#### Cluster 5

```{r}
cluster_name <- "5"
subcluster_name <- paste0("cluster",cluster_name, "_subcluster")
```


##### Subclustering 

```{r}
### Find subcluster 
Idents(seurat_object_l4) <- "cluster8_subcluster"
seurat_object_l4 <- FindSubCluster(
  seurat_object_l4,
  cluster = cluster_name,
  graph.name = "RNA_snn",
  subcluster.name = subcluster_name,
  resolution = 0.2
)
```

```{r}
cluster_cells <- Cells(seurat_object_l4)[which(seurat_object_l4$RNA_snn_res.0.46 == cluster_name)]
```

##### Visualization analysis
```{r}
pdf(paste(figures_folder, "subclustering/",cluster_name,"_subclustering.pdf", sep=""),width=20,height=10,paper='special') 
Seurat::DimPlot(seurat_object_l4, 
        reduction = "umap", 
        pt.size = pt_size,  
        cols= palette, 
        group.by = subcluster_name, 
        raster = FALSE) 

p2 <- Seurat::DimPlot(seurat_object_l4, 
        reduction = "umap", 
        pt.size = pt_size,  
        cols= palette, 
        group.by = subcluster_name, 
        cells=cluster_cells)
print(p2)

p3 <- Seurat::FeaturePlot(seurat_object_l4, 
                    features = c("MIR155HG", "MYC", "CCL3", "CCL4"), 
                    reduction="umap", 
                    pt.size = pt_size,  
                    cells=cluster_cells, 
                    raster = FALSE) 
print(p3)
patchwork::wrap_plots(p2, p3, ncol = 2)


dev.off()
```

#### Cluster 3

```{r}
cluster_name <- "3"
subcluster_name <- paste0("cluster",cluster_name, "_subcluster")
```


##### Subclustering 

```{r}
### Find subcluster 
Idents(seurat_object_l4) <- "cluster5_subcluster"
seurat_object_l4 <- FindSubCluster(
  seurat_object_l4,
  cluster = cluster_name,
  graph.name = "RNA_snn",
  subcluster.name = subcluster_name,
  resolution = 0.2
)
```

```{r}
cluster_cells <- Cells(seurat_object_l4)[which(seurat_object_l4$RNA_snn_res.0.46 == cluster_name)]
```

##### Visualization analysis
```{r}
pdf(paste(figures_folder, "subclustering/",cluster_name,"_subclustering.pdf", sep=""),width=20,height=10,paper='special') 
Seurat::DimPlot(seurat_object_l4, 
        reduction = "umap", 
        pt.size = pt_size,  
        cols= palette, 
        group.by = subcluster_name, 
        raster = FALSE) 

p2 <- Seurat::DimPlot(seurat_object_l4, 
        reduction = "umap", 
        pt.size = pt_size,  
        cols= palette, 
        group.by = subcluster_name, 
        cells=cluster_cells)
print(p2)

p3 <- Seurat::FeaturePlot(seurat_object_l4, 
                    features = c("IGHD", "IGHM", "IGHA1", "IGHA2", "IGHG1", "IGHG2"), 
                    reduction="umap", 
                    cells=cluster_cells, 
                    raster = FALSE) 
print(p3)
patchwork::wrap_plots(p2, p3, ncol = 2)
dev.off()
```

`

## Find final clusters after discussion with annotation team 

```{r}
seurat_object_l4$annotation_level_4_to_clean <- seurat_object_l4$cluster3_subcluster
NBC_MBC_seurat_object_level_4 <- seurat_object_l4
```

### Subcluster cluster 6  (HIPK2)
```{r}
cluster_name <- "6"
subcluster_name <- paste0("cluster",cluster_name, "_subcluster")
```

```{r}
Idents(NBC_MBC_seurat_object_level_4) <- "annotation_level_4_to_clean"
NBC_MBC_seurat_object_l4_to_l5 <- FindSubCluster(
  NBC_MBC_seurat_object_level_4,
  cluster = cluster_name,
  graph.name = "RNA_snn",
  subcluster.name = subcluster_name,
  resolution = 0.5
)
cluster_cells <- Cells(NBC_MBC_seurat_object_l4_to_l5)[which(NBC_MBC_seurat_object_l4_to_l5$annotation_level_4_to_clean == cluster_name)]

```

```{r}
pdf(paste(figures_folder, "subclustering_for_l5/",cluster_name,"_subclustering.pdf", sep=""),width=20,height=10,paper='special') 
Seurat::DimPlot(NBC_MBC_seurat_object_l4_to_l5, 
        reduction = "umap", 
        pt.size = pt_size,  
        cols= c(palette, palette),
        group.by = subcluster_name, 
        raster = FALSE) 

p2 <- Seurat::DimPlot(NBC_MBC_seurat_object_l4_to_l5, 
        reduction = "umap", 
        pt.size = pt_size,  
        cols= palette, 
        group.by = subcluster_name, 
        cells=cluster_cells)
print(p2)

p3 <- Seurat::FeaturePlot(NBC_MBC_seurat_object_l4_to_l5, 
                    features = c("HIPK2"), 
                    reduction="umap", 
                    cells=cluster_cells, 
                    raster = FALSE) 
print(p3)
patchwork::wrap_plots(p2, p3, ncol = 2)


dev.off()
```


### Merge cluster 4 with 6 - and 7_0 with 1_0
```{r}
NBC_MBC_seurat_object_l4_to_l5$annotation_level_4_to_clean <- ifelse(NBC_MBC_seurat_object_l4_to_l5$annotation_level_4_to_clean %in% c("6", "4"), "4_and_6", NBC_MBC_seurat_object_l4_to_l5$annotation_level_4_to_clean)

NBC_MBC_seurat_object_l4_to_l5$annotation_level_4_to_clean <- ifelse(NBC_MBC_seurat_object_l4_to_l5$annotation_level_4_to_clean %in% c("7_0", "1_0"), "1_0_and_7_0", NBC_MBC_seurat_object_l4_to_l5$annotation_level_4_to_clean)

NBC_MBC_seurat_object_l4_to_l5$annotation_level_4_to_clean <- ifelse(NBC_MBC_seurat_object_l4_to_l5$annotation_level_4_to_clean %in% c("3_0", "3_1", "3_2"), "3", NBC_MBC_seurat_object_l4_to_l5$annotation_level_4_to_clean)
```

### Cluster8 with three clusters
```{r}
cluster8_cells <- Cells(NBC_MBC_seurat_object_l4_to_l5)[which(NBC_MBC_seurat_object_l4_to_l5$annotation_level_4_to_clean %in% c("8_0", "8_1"))]
NBC_MBC_seurat_object_l4_to_l5@meta.data[cluster8_cells,"annotation_level_4_to_clean"] <- NBC_MBC_seurat_object_l4_to_l5@meta.data[cluster8_cells,"Cluster8_subcluster_0.2res"]
```

### Remove cluster 13
```{r}
Idents(NBC_MBC_seurat_object_l4_to_l5) <- "annotation_level_4_to_clean"
NBC_MBC_seurat_object_l4_to_l5 <- subset(
  NBC_MBC_seurat_object_l4_to_l5,
  cells = colnames(NBC_MBC_seurat_object_l4_to_l5)[Idents(NBC_MBC_seurat_object_l4_to_l5) != "13"]
)
```

```{r}
pdf(paste(figures_folder, "subclustering_for_l5/level4_final_clusters.pdf", sep=""),width=20,height=10,paper='special') 
Seurat::DimPlot(NBC_MBC_seurat_object_l4_to_l5, 
        reduction = "umap", 
        pt.size = pt_size,  
        label = TRUE,
        cols= c(palette, palette),
        group.by = "annotation_level_4_to_clean", 
        raster = FALSE) 
dev.off()
```

## Annotate clusters 

```{r}
NBC_MBC_seurat_object_l4_to_l5@meta.data <- NBC_MBC_seurat_object_l4_to_l5@meta.data %>% mutate(annotation_level_4 =
                                                  case_when(annotation_level_4_to_clean == "9" ~ "NBC_S100A+",
                                                            annotation_level_4_to_clean == "0" ~ "NBC main cluster",
                                                            annotation_level_4_to_clean == "1_0_and_7_0" ~ "NBC activated, first step",
                                                            annotation_level_4_to_clean == "1_1" ~ "NBC activated, first step (OAspecific",
                                                            annotation_level_4_to_clean == "5_0" ~ "GC-commited MIR155HG-",
                                                            annotation_level_4_to_clean == "5_1" ~ "Early GC-commited MYC+",     
                                                            annotation_level_4_to_clean == "5_2" ~ "GC-commited MIR155HG+",
                                                            annotation_level_4_to_clean == "10" ~ "NBC_CD229+",
                                                            annotation_level_4_to_clean == "7_1" ~ "NBC IFN-activated",
                                                            annotation_level_4_to_clean == "8_0" ~ "NBC preGC_2",
                                                            annotation_level_4_to_clean == "8_1" ~ "NBC preGC_1",
                                                            annotation_level_4_to_clean == "8_2" ~ "GC-related NBC",
                                                            annotation_level_4_to_clean == "12" ~ "Proliferative NBC, GC-related?",
                                                            annotation_level_4_to_clean == "2" ~ "ncsMBC",
                                                            annotation_level_4_to_clean == "3" ~ "MBC FCRL4+",
                                                            annotation_level_4_to_clean == "11" ~ "FCRL4-csMBC transition cells",        
                                                            annotation_level_4_to_clean == "4_and_6" ~ "csMBC"
                                                            ))
NBC_MBC_seurat_object_l4_to_l5$level4_clusters_clean <- NBC_MBC_seurat_object_l4_to_l5$annotation_level_4_to_clean
```

```{r}
pdf(paste(figures_folder, "subclustering_for_l5/level4_final_clusters_annotated.pdf", sep=""),width=20,height=10,paper='special') 
Seurat::DimPlot(NBC_MBC_seurat_object_l4_to_l5, 
        reduction = "umap", 
        pt.size = pt_size,  
        label = TRUE,
        cols= c(palette, palette),
        group.by = "annotation_level_4", 
        raster = FALSE) 
dev.off()
```

```{r}
saveRDS(NBC_MBC_seurat_object_l4_to_l5, paste0(data_folder, cell_type, "_clustered_level_4_to_clean_annotated.rds"))
```

## Include GCBC cells 

### Load GCBC cells
```{r}
GCBC_seurat_object_level_5 <- readRDS("/scratch/devel/saguilar/PhD/PROJECTS/tonsil_atlas/tonsil_atlas_annotation/level_5/GCBC/data/GCBC_seu_obj_level_5_eta.rds")
```

### DZ no proliferative from DZ non proliferative cluster
```{r}
set.seed(123)
DZ_noproli <- Cells(GCBC_seurat_object_level_5)[which(GCBC_seurat_object_level_5$names_level_5_clusters_eta == "DZ non proliferative")]
GC_cells_to_include1 <- sample(DZ_noproli,
                                 3000,
                                 replace=FALSE)
```

```{r}
GC_cells_to_include <- GC_cells_to_include1
```


### Visualization of included cells 
```{r}
pdf(paste(figures_folder, "level4_to_level5/DZ_noproli/DZ_NOproli_and_LZNOproli_selected_cells.pdf", sep=""),width=20,height=10,paper='special') 
Seurat::DimPlot(GCBC_seurat_object_level_5, 
        reduction = "umap", 
        pt.size = pt_size,  
        cols= c("lightgrey"), 
        group.by = "names_level_5_clusters_eta", 
        cells.highlight= list("GCBC_include_NBC" = GC_cells_to_include),
        cols.highlight = c("red"),
        raster=FALSE)
dev.off()
```


### Include GCBC in NBC object
```{r}
Idents(GCBC_seurat_object_level_5) <- "names_level_5_clusters_eta"
GC_interesting_seu_obj <- subset(
  GCBC_seurat_object_level_5,
  cells = GC_cells_to_include
)


GC_interesting_seu_obj@meta.data <- GC_interesting_seu_obj@meta.data %>% mutate(annotation_level_4 =
                                                  case_when(names_level_5_clusters_eta == "DZ non proliferative" ~ "GC_DZ_no_proli",
                                                            names_level_5_clusters_eta == "LZ" ~ "GC_LZ_NOproli"))   
table(GC_interesting_seu_obj$annotation_level_4)
```


```{r}
NBC_MBC_seurat_object_l4_to_l5_withGCcells <- merge(NBC_MBC_seurat_object_l4_to_l5, y = GC_interesting_seu_obj, merge.data = TRUE)
```

```{r}
name_project <- "3K_GC_DZ_NOproli"
```


### Re_processing

```{r}
seurat_list <- SplitObject(NBC_MBC_seurat_object_l4_to_l5_withGCcells, split.by = "assay")
seurat_list <- seurat_list[c("3P", "multiome")]
seurat_list <- purrr::map(
  seurat_list,
  FindVariableFeatures,
  nfeatures = 5000
)
hvg <- purrr::map(seurat_list, VariableFeatures)
shared_hvg <- intersect(hvg$`3P`, hvg$multiome)
# ElbowPlot(NBC_MBC_seurat_object_l4_to_l5_withGCcells, reduction = "harmony")
NBC_MBC_seurat_object_l4_to_l5_withGCcells <- NBC_MBC_seurat_object_l4_to_l5_withGCcells %>%
  ScaleData(features = shared_hvg) %>%
  RunPCA(features = shared_hvg) %>%
  RunHarmony(group.by.vars = "assay", reduction ="pca", dims = 1:30) %>%
  RunUMAP(reduction = "harmony", dims = 1:30)

seurat_object_l5 <- NBC_MBC_seurat_object_l4_to_l5_withGCcells
```

### New clustering

```{r}
resolution_vector <- c(0.01, 0.05, 0.1, 0.25, 0.35, 0.4, 0.45, 0.5, 0.55, 0,6, 0.75, 0.8, 0.9, 1, 1.1, 1.25, 1.5)
seurat_object_l5 <- seurat_object_l5 %>%
  FindNeighbors(dims = 1:30, reduction = "harmony")  %>%
  FindClusters(resolution = resolution_vector)
```

### Visualization of clusters 
```{r}
pdf(paste(level5_figures, "Level5_clusters/DZ_noproli/clusters_summary_resolutions.pdf", sep=""),width=20,height=10,paper='special')

list_plots <- list()
resolution_vector <- c(0.5, 0.75, 0.9, 1, 1.25, 1.5)

for (index in 1:length(resolution_vector)) {
  resolution <- resolution_vector[index]
  resolution_wanted <- paste0("RNA_snn_res.", resolution)
  Idents(seurat_object_l5) <- resolution_wanted
  p <- Seurat::DimPlot(seurat_object_l5, 
                       reduction = "umap",  
                       pt.size = pt_size/3, 
                       cols = c(palette, palette),
                       group.by = resolution_wanted, 
                       raster = FALSE ) + NoLegend()
  list_plots[[index]] <- p
}
print(patchwork::wrap_plots(list_plots), ncol =3)
dev.off()
```



## ETA annotation
Define clusters after discussion with annotation team

### Visualization of selected resolution
```{r}
p <- Seurat::DimPlot(NBC_MBC_seurat_object_l5, 
                       reduction = "umap",  
                       pt.size = pt_size,
                       label = TRUE,
                       cols = c(palette, palette),
                       group.by = "RNA_snn_res.0.9", 
                       raster = FALSE ) 
pdf(paste(figures_folder, "Level5_clusters/DZ_noproli/selected_resolution.pdf", sep=""),width=20,height=10,paper='special')
print(p)
dev.off()
```

### Subclusters of interesting clusters 
#### Subclustering of cluster 12
```{r}
Idents(NBC_MBC_seurat_object_l5) <- "RNA_snn_res.0.9"
### Find subcluster 
cluster_name <- "12"
subcluster_name <- paste0("cluster",cluster_name, "_subcluster")

NBC_MBC_seurat_object_l5 <- FindSubCluster(
  NBC_MBC_seurat_object_l5,
  cluster = cluster_name,
  graph.name = "RNA_snn",
  subcluster.name = subcluster_name,
  resolution = 0.1
)
table(NBC_MBC_seurat_object_l5@meta.data[,subcluster_name])

pdf(paste(figures_folder, "Level5_clusters/DZ_noproli/subclustering/",subcluster_name,".pdf", sep=""),width=20,height=10,paper='special') 
Seurat::DimPlot(NBC_MBC_seurat_object_l5, 
        reduction = "umap", 
        pt.size = pt_size,  
        cols= c(palette, palette), 
        group.by = subcluster_name, 
        raster = FALSE)
dev.off()
```


#### Subclustering of cluster 4
```{r}
Idents(NBC_MBC_seurat_object_l5) <- "cluster12_subcluster"
### Find subcluster 
cluster_name <- "4"
subcluster_name <- paste0("cluster",cluster_name, "_subcluster")

NBC_MBC_seurat_object_l5 <- FindSubCluster(
  NBC_MBC_seurat_object_l5,
  cluster = cluster_name,
  graph.name = "RNA_snn",
  subcluster.name = subcluster_name,
  resolution = 0.2
)
table(NBC_MBC_seurat_object_l5@meta.data[,subcluster_name])

pdf(paste(figures_folder, "Level5_clusters/DZ_noproli/subclustering/",subcluster_name,".pdf", sep=""),width=20,height=10,paper='special') 
Seurat::DimPlot(NBC_MBC_seurat_object_l5, 
        reduction = "umap", 
        pt.size = pt_size,  
        label = TRUE,
        cols= c(palette, palette), 
        group.by = subcluster_name, 
        raster = FALSE)
dev.off()
```

```{r}
NBC_MBC_seurat_object_l5$clusters_to_annotate <- NBC_MBC_seurat_object_l5$cluster4_subcluster
```

### Re-annotation 
```{r}
NBC_MBC_seurat_object_l5@meta.data <- NBC_MBC_seurat_object_l5@meta.data %>% mutate(names_level_5_clusters_eta =
  case_when(clusters_to_annotate  %in% c("1", "0", "9", "12_0") ~ "NBC",
  clusters_to_annotate  %in% c("2") ~ "NBC early activation",

  clusters_to_annotate  == "11" ~ "Early GC-commited NBC", 
  clusters_to_annotate  %in% c("7") ~ "GC-commited NBC",
  
  clusters_to_annotate  == "16" ~ "NBC IFN-activated", 
  clusters_to_annotate  == "17" ~ "NBC CD229+",
    
  clusters_to_annotate  %in% c("13", "20") ~ "preGC",

  clusters_to_annotate  == "15" ~ "Proliferative NBC",
  clusters_to_annotate  == "10" ~ "GC DZ Noproli",
    
    
  clusters_to_annotate  %in% c("6", "8", "18") ~ "ncsMBC",
  clusters_to_annotate  %in% c("3", "5", "19") ~ "csMBC",
  clusters_to_annotate  == "14" ~ "MBC FCRL5+",
  clusters_to_annotate  == "12_1" ~ "Early MBC",
  clusters_to_annotate  == "4_1" ~ "ncsMBC FCRL4+/FCRL5+",
  clusters_to_annotate  %in% c("4_0", "4_2", "4_3") ~ "csMBC FCRL4+/FCRL5+"
))
```



### Visualization of final clusters
```{r}
p <- Seurat::DimPlot(NBC_MBC_seurat_object_l5, 
                       reduction = "umap",  
                       pt.size = pt_size,
                       label = TRUE,
                       cols = c(palette, palette),
                       group.by = "names_level_5_clusters_eta", 
                       raster = FALSE ) 
pdf(paste(figures_folder, "Level5_clusters/DZ_noproli/Annotated_clusters.pdf", sep=""),width=20,height=10,paper='special')
print(p)
dev.off()
```

## Save object
```{r}
saveRDS(NBC_MBC_seurat_object_l5, paste0(data_folder, cell_type, "_seu_obj_level_5_eta_DZnoproli.rds"))
```



