---
title: "Separate microenvironment and tumoral cells (102)"
author: "Ramon Massoni-Badosa"
date: "`r format(Sys.time(), '%B %Y')`"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```


# Introduction

Here, we will separate the microenvironment from tumoral cells on the basis of CD79 expression. In addition, we will assess the success of the microenvironment selection (CD19-).


# Pre-processing

## Load packages

```{r}
library(Seurat)
library(Signac)
library(EnsDb.Hsapiens.v86)
library(ggpubr)
library(tidyverse)
set.seed(173)
```


## Define paths and parameters

```{r}
path_to_102 <- here::here("MCL/results/R_objects/5.seurat_merged_rna_processed_atac_processed_102.rds")
path_to_save_microenv_102 <- here::here("MCL/results/R_objects/6.seurat_microenvironment_102.rds")
path_to_save_tumoral_102 <- here::here("MCL/results/R_objects/6.seurat_tumoral_102.rds")
```


## Read data

```{r}
seurat <- readRDS(path_to_102)
```


# Cluster

We will cluster cells at a very low resolution, so that we can identify the microenvironment and tumoral compartments.

```{r}
# RNA
DefaultAssay(seurat) <- "RNA"
seurat <- FindNeighbors(seurat, reduction = "pca", dims = 1:30)
seurat <- FindClusters(seurat, resolution = 0.1)
DimPlot(seurat)
DimPlot(seurat, reduction = "atacUMAP")
FeaturePlot(seurat, features = c("CD79A", "CD3D", "NKG7", "LYZ"))
FeaturePlot(seurat, features = c("MS4A1"))
```


# Markers

```{r}
markers <- FindAllMarkers(
  seurat,
  only.pos = TRUE,
  logfc.threshold = 0.5
)
DT::datatable(markers)
```

Based on the markers, we conclude that cluster 0, 1 and 3 consist of tumoral cells; while cluster 2 and 5 consist of microenvironment cells (T cells, NK, myeloid). Although cluster 4 has a solid signature of T cells, it also expresses CD79. For now we will label it as microenvironment, and in downstream analysis we will recheck if they are singlets or doublets.


```{r}
seurat$is_tumoral <- ifelse(
  seurat$seurat_clusters %in% c("0", "1", "3"),
  "tumoral",
  "microenvironment"
)
```


# Subset

```{r}
tumoral <- subset(seurat, is_tumoral == "tumoral")
microenv <- subset(seurat, is_tumoral == "microenvironment")
```


# Save

```{r}
saveRDS(tumoral, path_to_save_tumoral_102)
saveRDS(microenv, path_to_save_microenv_102)
```


# Session information

```{r}
sessionInfo()
```


