---
title: "Separate microenvironment and tumoral cells (413)"
author: "Ramon Massoni-Badosa"
date: "`r format(Sys.time(), '%B %Y')`"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```


# Introduction

Here, we will separate the microenvironment from tumoral cells on the basis of CD79 expression. In addition, we will assess the success of the microenvironment selection (CD19-).


# Pre-processing

## Load packages

```{r}
library(Seurat)
library(Signac)
library(EnsDb.Hsapiens.v86)
library(ggpubr)
library(tidyverse)
set.seed(173)
```


## Define paths and parameters

```{r}
path_to_413 <- here::here("MCL/results/R_objects/5.seurat_merged_rna_processed_atac_processed_413.rds")
path_to_save_microenv_413 <- here::here("MCL/results/R_objects/6.seurat_microenvironment_413.rds")
path_to_save_tumoral_413 <- here::here("MCL/results/R_objects/6.seurat_tumoral_413.rds")
```


## Read data

```{r}
seurat <- readRDS(path_to_413)
```


# Cluster

We will cluster cells at a very low resolution, so that we can identify the microenvironment and tumoral compartments.

```{r}
# RNA
DefaultAssay(seurat) <- "RNA"
seurat <- FindNeighbors(seurat, reduction = "pca", dims = 1:30)
seurat <- FindClusters(seurat, resolution = 0.1)
DimPlot(seurat)
DimPlot(seurat, reduction = "atacUMAP")
FeaturePlot(seurat, features = c("CD79A", "CD3D", "NKG7", "LYZ"))
FeaturePlot(seurat, features = c("MS4A1"))
```


# Markers

```{r}
markers <- FindAllMarkers(
  seurat,
  only.pos = TRUE,
  logfc.threshold = 0.5
)
DT::datatable(markers)
```

Based on the markers, we conclude that cluster 0, 1, 5 and 6  consist of tumoral cells; while cluster 2, 4, 7, 8 and 9 consist of microenvironment cells (T cells, NK, myeloid). Cluster 3 seems to be of poor-quality cells, since it does not have any specific marker with a log2 fold-change > 0.5. Let us reduce this threshold to capture more subtle markers:

```{r}
markers_3 <- FindMarkers(seurat, ident.1 = "3", only.pos = TRUE)
DT::datatable(markers_3)
```


It expresses some markers of T cells, like ICOS or TOX2. Thus, we will consider microenvironment for now:

```{r}
seurat$is_tumoral <- ifelse(
  seurat$seurat_clusters %in% c("0", "1", "5", "6"),
  "tumoral",
  "microenvironment"
)
```


# Subset

```{r}
tumoral <- subset(seurat, is_tumoral == "tumoral")
microenv <- subset(seurat, is_tumoral == "microenvironment")
DimPlot(tumoral)
DimPlot(microenv)
```


# Save

```{r}
saveRDS(tumoral, path_to_save_tumoral_413)
saveRDS(microenv, path_to_save_microenv_413)
```


# Session information

```{r}
sessionInfo()
```


