---
title: "Cluster tumoral cells (102)"
author: "Ramon Massoni-Badosa"
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```


# Introduction

Here, we will cluster the tumoral cells for case 102.


# Pre-processing

## Load packages

```{r}
library(Seurat)
library(Signac)
library(SLOcatoR)
library(EnsDb.Hsapiens.v86)
library(ggpubr)
library(tidyverse)
library(pals)
library(openxlsx)
library(here)
#library(presto)
library(DT)
set.seed(173)
```


## Define paths and parameters

```{r}
path_to_obj <- here("MCL/results/R_objects/6.seurat_tumoral_102.rds")
path_to_save_xlsx <- here("MCL/3-clustering/tmp/7.seurat_tumoral_102_clustered_markers.xlsx")
path_to_save_obj <- here("MCL/results/R_objects/7.seurat_tumoral_102_clustered.rds")
path_to_save_shiny_metadata <- here("MCL/results/R_objects/7.seurat_tumoral_102_clustered_shiny_metadata.rds")
path_to_save_shiny_expression <- here("MCL/results/R_objects/7.seurat_tumoral_102_clustered_shiny_expression.rds")


# Colors
color_palette <-  c("#E6194B", "#3CB44B", "#FFD8B1", "#4363D8", "#F58231",
                    "#911EB4", "#46F0F0", "#F032E6", "#BCF60C", "#FABEBE",
                    "#008080", "#E6BEFF", "#9A6324", "#FFFAC8", "#800000",
                    "#AAFFC3", "#808000", "#FFE119", "#000075", "#808080",
                    "#000000", "tan", "darkgrey")


# Source functions
source(here::here("scRNA-seq/bin/utils.R"))


# Thresholds
chrY_cutoff <- -0.32
```


## Read data

```{r}
seurat <- readRDS(path_to_obj)
# seurat[["ATAC"]] <- NULL
```


# Dimensionality reduction

Let us score each cell for cell cycling singatures:

```{r}
seurat <- CellCycleScoring(
  seurat,
  s.features = cc.genes$s.genes,
  g2m.features = cc.genes$g2m.genes
)
FeaturePlot(seurat, c("G2M.Score", "S.Score")) &
  scale_color_viridis_c(option = "magma")
```


Since we subsetted the CD79A+ cells, let us rerun the general pipeline for dimensionality reduction:

```{r}
seurat <- seurat %>%
  FindVariableFeatures() %>%
  ScaleData() %>%
  RunPCA() %>%
  RunUMAP(dims = 1:20, reduction = "pca")
DimPlot(seurat, cols = color_palette, reduction = "umap")
```


# Classify chry +/-

Exploring our data, we have observed that the two main "blobs" might be two subclones, one chrY+ and the other chrY. Loss of chromosome Y is a [known feature of MCL](https://academic.oup.com/ajcp/article/116/6/886/1758329?login=false):

```{r}
FeaturePlot(
  seurat,
  features = c("UTY", "KDM5D", "DDX3Y", "USP9Y", "ZFY", "EIF1AY"),
  reduction = "umap"
) &
  scale_color_viridis_c(option = "magma")
```


Let us convert the expression of all genes located in chrY in a single score:

```{r}
# annotations <- GetGRangesFromEnsDb(ensdb = EnsDb.Hsapiens.v86)
# seqlevelsStyle(annotations) <- "UCSC"
# annotations_chrY <- annotations[annotations@seqnames == "chrY", "gene_name"]
# goi_chrY <- unique(annotations_chrY$gene_name)
goi_chrY <- c("UTY", "KDM5D", "DDX3Y", "USP9Y", "ZFY", "EIF1AY")
seurat <- AddModuleScore(seurat, features = list(goi_chrY), name = "chrY_score")
FeaturePlot(seurat, "chrY_score1", reduction = "umap")
(density_gg <- seurat@meta.data %>%
  ggplot(aes(chrY_score1)) +
  geom_density() +
  geom_vline(xintercept = chrY_cutoff, linetype = "dashed", color = "darkblue") +
  theme_classic())
seurat$has_loss_chrY <- ifelse(
  seurat$chrY_score1 > chrY_cutoff,
  "chrY+",
  "chrY-"
)
DimPlot(seurat, group.by = "has_loss_chrY", reduction = "umap")
```


# Cluster

```{r}
seurat <- FindNeighbors(seurat, dims = 1:20, reduction = "pca")
seurat <- FindClusters(seurat, resolution = 0.15)
DimPlot(seurat, cols = color_palette, reduction = "umap")
FeaturePlot(seurat, c("CD8A", "CD3D"), reduction = "umap", order = TRUE)
```


Let us subcluster and eliminate residual T cells:

```{r}
seurat <- FindSubCluster(
  seurat,
  cluster = "3",
  graph.name = "RNA_snn",
  subcluster.name = "T_cells",
  resolution = 0.25
)
DimPlot(seurat, group.by = "T_cells", reduction = "umap")
Idents(seurat) <- "T_cells"
markers_t <- FindMarkers(seurat, ident.1 = "3_3", ident.2 = c("3_0", "3_1", "3_2"))
head(markers_t, 20)
seurat <- subset(seurat, T_cells != "3_3")
DimPlot(seurat, reduction = "umap")
```


From our exploratory analysis we know that cluster 4 contains mostly poor-quality cells:

```{r}
markers_4 <- FindMarkers(seurat, ident.1 = "4", only.pos = TRUE, logfc.threshold = 0.65)
head(markers_4, 20)
seurat <- subset(seurat, seurat_clusters != "4")
```


We observe remarkable heterogeneity. Thus, let us increase the clustering resolution:

```{r}
seurat <- FindNeighbors(seurat, dims = 1:20, reduction = "pca")
seurat <- FindClusters(seurat, resolution = 0.7)
DimPlot(seurat, reduction = "umap")
```


# Markers

```{r}
#markers <- wilcoxauc(
#  seurat,
#  group_by = "seurat_clusters",
#  seurat_assay = "RNA",
#  y = levels(seurat$seurat_clusters)
#)
#markers <- markers[markers$padj < 0.001 & markers$logFC > 0.5, ]
#DT::datatable(markers)


# Dot plot
goi <- c("MT2A", "MT1E", "IGF2BP3", "RGS12", "RUBCNL",
         "RGS7", "LRRFIP1", "BANK1", "ZNF804A", "PLCL2",
         "RGS1", "TFEC", "CD69", "BTG2", "SIK3", "ECE1",
         "CD58", "KIF26B", "IRF4", "HDAC9", "TRAF1", "MIR155HG",
         "PCNA", "TOP2A", "STMN1", "CCDC88A", "NCL", "NFKB1",
         "CCL3", "CCL4", "CD55", "GALNT2", "IRAK2", "AFF3", 
         "CXCR4")
DotPlot(seurat, features = rev(goi)) +
  coord_flip()
```


Exploring our data, we have observed that the two main "blobs" might be two subclones, one chrY+ and the other chrY. Loss of chromosome Y is a [known feature of MCL](https://academic.oup.com/ajcp/article/116/6/886/1758329?login=false):

```{r}
FeaturePlot(
  seurat,
  features = c("UTY", "KDM5D", "DDX3Y", "USP9Y", "ZFY", "EIF1AY"),
  reduction = "umap"
) &
  scale_color_viridis_c(option = "magma")
```

We will explore this later with [infercnv](https://github.com/broadinstitute/infercnv) but, for now, we will try to find "mirror" clusters in both subclones:

```{r}
seurat <- FindClusters(seurat, resolution = 0.1)
DimPlot(seurat, reduction = "umap")

# chrY-
Idents(seurat) <- "seurat_clusters"
seurat <- FindSubCluster(
  seurat,
  cluster = "1",
  graph.name = "RNA_snn",
  subcluster.name = "chrYneg",
  resolution = 0.2
)
DimPlot(seurat, group.by = "chrYneg", reduction = "umap")
Idents(seurat) <- "chrYneg"
markers_1 <- FindAllMarkers(seurat, only.pos = TRUE, logfc.threshold = 0.8)
markers_1_1 <- FindMarkers(
  seurat,
  ident.1 = "1_1",
  ident.2 = c("1_2", "1_0"),
  only.pos = TRUE,
  logfc.threshold = 0.75
)
markers_1_0 <- FindMarkers(
  seurat,
  ident.1 = "1_0",
  ident.2 = c("1_2", "1_1"),
  only.pos = TRUE,
  logfc.threshold = 0.75
)
markers_1_2 <- FindMarkers(
  seurat,
  ident.1 = "1_2",
  ident.2 = c("1_0", "1_1"),
  only.pos = TRUE,
  logfc.threshold = 0.75
)
DT::datatable(markers_1_0, options = list(scrollX = TRUE))
DT::datatable(markers_1_1, options = list(scrollX = TRUE))
DT::datatable(markers_1_2, options = list(scrollX = TRUE))
seurat_chrYneg <- subset(seurat, idents = c("1_0", "1_1", "1_2"))
DotPlot(
  seurat_chrYneg,
  features = rev(c("MT2A", "MT1X", "MT1G", "INPP5A", "GAB2",
               "FOS", "JUN", "CD69", "MARCH1", "BANK1",
               "MIR155HG", "TRAF1", "NFKB1", "NFAT5", "MYC", "IL2RA"))) +
  coord_flip()


# chrY+
Idents(seurat) <- "seurat_clusters"
seurat <- FindSubCluster(
  seurat,
  cluster = "0",
  graph.name = "RNA_snn",
  subcluster.name = "chrYpos",
  resolution = 0.15
)
DimPlot(seurat, group.by = "chrYpos", reduction = "umap")
Idents(seurat) <- "chrYpos"
markers_0_0 <- FindMarkers(
  seurat,
  ident.1 = "0_0",
  ident.2 = c("0_1", "0_2", "0_3"),
  only.pos = TRUE,
  logfc.threshold = 0.75
)
markers_0_1 <- FindMarkers(
  seurat,
  ident.1 = "0_1",
  ident.2 = c("0_0", "0_2", "0_3"),
  only.pos = TRUE,
  logfc.threshold = 0.4
)
markers_0_2 <- FindMarkers(
  seurat,
  ident.1 = "0_2",
  ident.2 = c("0_0", "0_1", "0_3"),
  only.pos = TRUE,
  logfc.threshold = 0.6
)
markers_0_3 <- FindMarkers(
  seurat,
  ident.1 = "0_3",
  ident.2 = c("0_0", "0_1", "0_2"),
  only.pos = TRUE,
  logfc.threshold = 0.6
)
DT::datatable(markers_0_0, options = list(scrollX = TRUE))
DT::datatable(markers_0_1, options = list(scrollX = TRUE))
DT::datatable(markers_0_2, options = list(scrollX = TRUE))
DT::datatable(markers_0_3, options = list(scrollX = TRUE))
seurat_chrYpos <- subset(seurat, idents = c("0_0", "0_1", "0_2", "0_3"))
seurat_chrYpos$chrYpos <- factor(seurat_chrYpos$chrYpos, levels = c("0_0", "0_1", "0_2", "0_3"))

DotPlot(
  seurat_chrYpos,
  group.by = "chrYpos",
  features = rev(c("MT1X", "MT2A", "MT1G",
                   "GALNTL6", "BANK1", "RGS7",
                   "JUN", "FOS", "CD69", "RGS1",
                   "MIR155HG", "NFKB1", "NFKBIZ", "TRAF1", "IL2RA", "CD44"))) +
  coord_flip()
```


# Preliminary annotation

```{r}
seurat$annotation_20220518 <- case_when(
  seurat$chrYneg == "1_0" ~ "chrY- metallothionein",
  seurat$chrYneg == "1_1" ~ "chrY- CD69+AP1+",
  seurat$chrYneg == "1_2" ~ "chrY- MIR155HG+NFKB1+MYC+",
  seurat$chrYneg == "1_3" ~ "non-tumoral B-cells",
  seurat$chrYneg == "2" ~ "Cycling tumoral",
  seurat$chrYpos == "0_0" ~ "chrY+ metallothionein",
  seurat$chrYpos == "0_1" ~ "chrY+ RGS7+BANK1+",
  seurat$chrYpos == "0_2" ~ "chrY+ CD69+AP1+",
  seurat$chrYpos == "0_3" ~ "chrY+ MIR155HG+NFKB1+MYC+"
)
colors <- pals::polychrome()
names(colors) <- NULL
DimPlot(
  seurat,
  group.by = "annotation_20220518",
  cols = colors,
  reduction = "umap"
)
```



[Prognostic significance of metallothionein in B-cell lymphomas](https://ashpublications.org/blood/article/108/10/3514/22788/Prognostic-significance-of-metallothionein-in-B)
# Save

```{r}
#openxlsx::write.xlsx(list(markers = markers), file = path_to_save_xlsx)
saveRDS(seurat, path_to_save_obj)
```


Save input to shiny app:

```{r}
# input_shiny <- seurat2shiny(
#   seurat,
#   assay = "RNA",
#   slot = "data",
#   reduction = "umap"
# )
# saveRDS(input_shiny$metadata, path_to_save_shiny_metadata)
# saveRDS(input_shiny$expression, path_to_save_shiny_expression)
```



# Session information

```{r}
sessionInfo()
```
