---
title: "Harmonize CITE-seq object CD8 T"
author: "Ramon Massoni-Badosa"
date: "`r format(Sys.time(), '%B %Y')`"
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center',
                      message=FALSE, warning = FALSE)
options(width = 1200)
```


# Introduction

To be consistent with the other data modalities, we will harmonize the CITE-seq dataset for the CD4 T cells, including the variable names in the metadata, and the names of the ADT.


# Pre-processing

## Load packages

```{r}
library(Seurat)
library(harmony)
library(caret)
library(class)
library(tidyverse)
library(here)
```


## Parameters

```{r}
# Source functions
source(here("scRNA-seq/bin/utils_final_clusters.R"))
source(here("scRNA-seq/bin/SLOcatoR_functions.R"))


# Paths
path_to_donor_metadata <- here("data/tonsil_atlas_donor_metadata.csv")
path_to_sequencing_cite_metadata <- here("CITE-seq/1-cellranger_mapping/data/tonsil_atlas_metadata.csv")
```


## Read data

```{r}
donor_metadata <- read_csv(path_to_donor_metadata)
sequencing_cite_metadata <- read_csv(path_to_sequencing_cite_metadata)
cite_seq <- readRDS(path_to_save_tonsil_cite)
cd8 <- readRDS(path_to_save_cd8)
ilc_nk <- readRDS(path_to_save_ilc_nk)


# Merge cd8 and ilc_nk into a single reference
ilc_nk$UMAP_1_20220215 <- ilc_nk$UMAP_1_20220215 + 10
ilc_nk$UMAP_2_20220215 <- ilc_nk$UMAP_2_20220215 + 10
reference <- merge(x = cd8, y = ilc_nk)
rm(cd8, ilc_nk)


# Subset CITE-seq to cytotoxic clusters
DimPlot(cite_seq, cols = color_palette)
FeaturePlot(cite_seq, c("NKG7", "GNLY", "CD8", "GZMK"))
query <- subset(cite_seq, idents = c("6", "7", "8"))
```


# Remove doublets

```{r}
query <- query %>%
  FindVariableFeatures() %>%
  ScaleData() %>%
  RunPCA() %>%
  RunHarmony(dims = 1:20, reduction = "pca", group.by.vars = "subproject") %>%
  RunUMAP(dims = 1:20, reduction = "harmony")
DimPlot(query, reduction = "umap")
FeaturePlot(query, c("CD19", "CD4"), reduction = "umap", order = TRUE)
query <- FindNeighbors(query, reduction = "harmony", dims = 1:20)
query <- FindClusters(query, resolution = 0.3)
DimPlot(query, reduction = "umap")
markers_3 <- FindMarkers(
  query,
  ident.1 = "3",
  only.pos = TRUE,
  logfc.threshold = 0.8
) 
DT::datatable(markers_3)
markers_2 <- FindMarkers(
  query,
  ident.1 = "2",
  only.pos = TRUE,
  logfc.threshold = 0.8
) 
DT::datatable(markers_2)
```


* Cluster 3 is composed of doublets (T-B cells)
* Cluster 2 is composed of cycling cells. Cycling cells introduce a lot of variance, so we focused mostly on non-cycling cells.

```{r}
query <- subset(
  query,
  cells = colnames(query)[!(query$seurat_clusters %in% c("2", "3"))]
)
DimPlot(query, reduction = "umap")
```


# Merge reference and query and reintegrate

```{r}
reference$type <- "reference"
query$type <- "query"
merged <- merge(x = reference, y = query)
shared_hvg <- find_assay_specific_features(
  merged,
  assay_var = "assay",
  n_features = 5500
)
merged <- integrate_assays(merged, assay_var = "assay", shared_hvg = shared_hvg)
merged <- RunUMAP(merged, dims = 1:30, reduction = "harmony")
DimPlot(merged, split.by = "assay", group.by = "assay")
```


## Transfer labels

```{r}
# Define training and test sets
data_sets <- split_training_and_test_sets(
  merged,
  split_var = "type",
  referece_label = "reference",
  query_label = "query",
  reduction = "harmony",
  n_dims = 30
)


# Transfer label
annotation_query_df <- transfer_label(
  seurat_obj = merged,
  training_set = data_sets$training_set,
  test_set = data_sets$test_set,
  k = 10,
  response_var = "annotation_20220215"
)
```


## Transfer UMAP coordinates

```{r}
umap_test_df <- transfer_umap_coords(
  seurat_obj = merged,
  training_set = data_sets$training_set,
  test_set = data_sets$test_set,
  umap1_var = "UMAP_1_20220215",
  umap2_var = "UMAP_2_20220215",
  k = 10
)
```


# Plot

```{r}
rownames(annotation_query_df) <- annotation_query_df$query_cells
rownames(umap_test_df) <- umap_test_df$query_cells
query$UMAP_1_RNA_based <- umap_test_df[colnames(query), "UMAP1"]
query$UMAP_2_RNA_based <- umap_test_df[colnames(query), "UMAP2"]
query$annotation_20220215 <- annotation_query_df[colnames(query), "annotation"]
query$annotation_prob <- annotation_query_df[colnames(query), "annotation_prob"]
FeatureScatter(
  query,
  "UMAP_1_RNA_based",
  "UMAP_2_RNA_based",
  group.by = "annotation_20220215"
) 
```

# Save

```{r}
saveRDS(query, path_to_save_cite_cytotoxic)
```


# Session Info

```{r}
sessionInfo()
```

