---
title: "Transfer label microenvironment cells (102, CD4 T)"
author: "Ramon Massoni-Badosa"
date: "`r format(Sys.time(), '%B %Y')`"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```


# Introduction

In this notebook we will transfer the label and UMAP coordinates from the CD4 T cells of the tonsil atlas to the CD4 T cells of the MCL sample (case 102) using SLOcatoR.


# Pre-processing

## Load packages

```{r}
library(Seurat)
library(Signac)
library(SLOcatoR)
library(EnsDb.Hsapiens.v86)
library(ggpubr)
library(tidyverse)
library(pals)
library(here)
set.seed(173)
```


## Define paths and parameters

```{r}
# Paths
path_to_query <-  here("MCL/results/R_objects/6.seurat_microenvironment_102.rds")
path_to_save1 <- here("MCL/results/R_objects/7.merged_CD4_T_reference_query_102.rds")
path_to_save2 <- here("MCL/results/R_objects/7.seurat_CD4_T_102_annotated.rds")


# Misc
label_var <- "annotation_paper"
batch_var <- "assay"


# Source key functions and variables
source(here("scRNA-seq/bin/utils_final_clusters.R"))
source(here("scRNA-seq/bin/utils_figure2.R"))
```


## Read data

```{r}
reference <- readRDS(path_to_save_cd4)
reference$annotation_paper <- factor(
  reference$annotation_20220215,
  levels = names(colors_rna)
)
query <- readRDS(path_to_query)
```

Since here we will focus on the RNA, let us remove the ATAC assay:

```{r}
# query[["ATAC"]] <- NULL
```


# Transfer labels and UMAP coords

## Subset query to keep only CD4 T cells

```{r}
DimPlot(query)
FeaturePlot(query, c("CD3D", "NKG7", "LYZ"))
```

Cluster 5 is cleary myeloid and other non-lymphoid cells (CD3-LYZ+). Thus, let us subset to keep only clusters 2 and 4 and recluster:

```{r}
query <- subset(query, idents = c("2", "4"))
query <- query %>%
  FindVariableFeatures() %>%
  ScaleData() %>%
  RunPCA() %>%
  RunUMAP(dims = 1:20, reduction = "pca")
DimPlot(query, reduction = "umap")
```

Cluster with low resolution:

```{r}
query <- FindNeighbors(query, reduction = "pca", dims = 1:20)
query <- FindClusters(query, resolution = 0.15)
DimPlot(query, reduction = "umap")


# Find markers
markers_all <- FindAllMarkers(query, logfc.threshold = 0.5, only.pos = TRUE)
DT::datatable(markers_all)
```


* Cluster 0: cytotoxic cells (CD8 T and/or NK and/or ILC)
* Cluster 1: CD4 T cells
* Cluster 2: CD4 T cells 2 (probably a mix of effector and Treg, since they express PRDM1)
* Cluster 3: doublets with B cells
* Cluster 4: probably NK (GNLY, NCAM1)

Thus, we will conserve clusters 1 and 2

```{r}
query <- subset(query, idents = c("1", "2"))
```


Nevertheless, we know from other analysis that naive CD8 T cells cluster with naive CD4 T cells:

```{r}
FeaturePlot(query, c("CD8A", "NOSIP"), order = TRUE)
```

Thus, we subcluster cluster 1:

```{r}
query <- FindNeighbors(query, reduction = "pca", dims = 1:20)
query <- FindSubCluster(
  query,
  cluster = "1",
  subcluster.name = "naive_CD8",
  graph.name = "RNA_snn",
  resolution = 0.5
)
DimPlot(query, group.by = "naive_CD8")
Idents(query) <- "naive_CD8"
markers_1_3 <- FindMarkers(
  query,
  ident.1 = "1_3",
  only.pos = TRUE,
  logfc.threshold = 0.75
)
DT::datatable(markers_1_3)
```

We can see how subcluster 1_3 belongs to cytotoxic cells. Let us exclude it:

```{r}
query <- subset(query, cells = colnames(query)[query$naive_CD8 != "1_3"])
```



## Merge

We will merge both objects, keeping only the variables in the metadata that are common for both:

```{r}
# Define new variables for label transfer
reference$type <- "reference"
query$type <- "query"
reference$batch <- reference@meta.data[[batch_var]]
reference$label <- reference@meta.data[[label_var]]
reference$annotation_probability <- NA
reference$UMAP1 <- Embeddings(reference, "umap")[, "UMAP_1"]
reference$UMAP2 <- Embeddings(reference, "umap")[, "UMAP_2"]
query$label <- NA
query$annotation_probability <- NA
query$batch <- "query"
query$UMAP1 <- NA
query$UMAP2 <- NA


# Find common columns and subset
common_cols <- intersect(
  colnames(reference@meta.data),
  colnames(query@meta.data)
)
reference@meta.data <- reference@meta.data[, common_cols]
query@meta.data <- query@meta.data[, common_cols]
```


Now we can merge:

```{r}
merged <- merge(x = reference, y = query)
```


## Integrate


Let us start by finding the batch-specific highly variable features (HVG):

```{r}
shared_features <- find_assay_specific_features(merged, assay_var = "batch")
```

Integrate with Harmony:

```{r}
# Integrate
merged <- integrate_assays(
  merged,
  assay_specific = TRUE,
  assay_var = "batch",
  shared_hvg = shared_features,
  n_dim = 30
)


# Visualize integration
merged <- RunUMAP(merged, dims = 1:30, reduction = "harmony")
DimPlot(merged, group.by = "batch", split.by = "batch")
```

## Transfer labels

```{r}
# Define training and test sets
data_sets <- SLOcatoR::split_training_and_test_sets(
  merged,
  split_var = "type",
  referece_label = "reference",
  query_label = "query",
  reduction = "harmony",
  n_dims = 30
)


# Transfer label
annotation_query_df <- transfer_label(
  seurat_obj = merged,
  training_set = data_sets$training_set,
  test_set = data_sets$test_set,
  k = 10,
  response_var = "label"
)
```


## Transfer UMAP coordinates

```{r}
umap_test_df <- transfer_umap_coords(
  seurat_obj = merged,
  training_set = data_sets$training_set,
  test_set = data_sets$test_set,
  umap1_var = "UMAP1",
  umap2_var = "UMAP2",
  k = 10
)
```


# Plot

```{r}
merged$UMAP1[umap_test_df$query_cells] <- umap_test_df$UMAP1
merged$UMAP2[umap_test_df$query_cells] <- umap_test_df$UMAP2
merged$label[annotation_query_df$query_cells] <- annotation_query_df$annotation
merged$annotation_probability[annotation_query_df$query_cells] <- annotation_query_df$annotation_prob
(umap_gg <- merged@meta.data %>%
  mutate(type = factor(type, levels = c("reference", "query"))) %>%
  ggplot(aes(UMAP1, UMAP2, color = label)) +
    geom_point(size = 0.5) +
    facet_wrap(~type) +
    scale_color_manual(values = pals::glasbey()) +
    theme_classic() +
    theme(legend.title = element_blank()) +
    guides(color = guide_legend(override.aes = list(size = 2))))
```


Include annotations in query object:

```{r}
if (all(colnames(query) == annotation_query_df$query_cells)) {
  query$annotation_SLOcatoR <- annotation_query_df$annotation
  query$annotation_probability <- annotation_query_df$annotation_prob
}
if (all(colnames(query) == umap_test_df$query_cells)) {
  query$UMAP_1_SLOcatoR <- umap_test_df$UMAP1
  query$UMAP_2_SLOcatoR <- umap_test_df$UMAP2
}
query$label <- NULL
```


# Save

```{r}
saveRDS(merged, path_to_save1)
saveRDS(query, path_to_save2)
# saveRDS(merged@meta.data, path_to_save_df)
```


# Session Information

```{r}
sessionInfo()
```

