---
title: "Cell-of-origin prediction: Feature Selection (413)"
author: "Ramon Massoni-Badosa"
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```


# Introduction

In this and the following notebooks, we aim to predict the cell-of-origin of mantle cell lymphoma (MCL). The first step will be to define the features that we use for integration.


# Pre-processing

## Load packages

```{r}
library(Seurat)
library(Signac)
library(SLOcatoR)
library(EnsDb.Hsapiens.v86)
library(ggpubr)
library(tidyverse)
library(pals)
library(openxlsx)
set.seed(173)
```


## Define paths and parameters

```{r}
# Paths
path_to_reference <- here::here("scRNA-seq/results/R_objects/final_clusters/NBC_MBC_seu_obj_level_5_delta.rds")
path_to_query <- here::here("MCL/results/R_objects/6.seurat_tumoral_413.rds")
path_to_save <- here::here("MCL/results/R_objects/7.merged_tumoral_reference_query_413.rds")
path_to_save_query <- here::here("MCL/results/R_objects/M413_scRNAseq_seurat_object.rds")
path_to_save_df <- here::here("MCL/results/R_objects/dataframe_CD4_T_label_transfer_413.rds")
path_to_save_shiny_metadata <- here::here("MCL/results/R_objects/M413_scRNAseq_metadata_shiny.rds")
path_to_save_shiny_expression <- here::here("MCL/results/R_objects/M413_scRNAseq_expression_shiny.rds")
# path_to_genes1 <- here::here("MCL/results/tables/NBCpath_markers_genes.txt")
# path_to_genes2 <- here::here("MCL/results/tables/Additiona_markers-MBC_NBC.txt")


# Colors
color_palette <-  c("#E6194B", "#3CB44B", "#FFD8B1", "#4363D8", "#F58231",
                    "#911EB4", "#46F0F0", "#F032E6", "#BCF60C", "#FABEBE",
                    "#008080", "#E6BEFF", "#9A6324", "#FFFAC8", "#800000",
                    "#AAFFC3", "#808000", "#FFE119", "#000075", "#808080",
                    "#000000", "tan", "darkgrey")


# Misc
label_var <- "names_level_5_clusters_delta"
batch_var <- "assay"


# Source functions
source(here::here("scRNA-seq/bin/utils.R"))
```


## Read data

```{r}
reference <- readRDS(path_to_reference)
Idents(reference) <- "names_level_5_clusters_delta"
query <- readRDS(path_to_query)


DimPlot(reference, cols = color_palette)
DimPlot(query, cols = color_palette)
FeaturePlot(query, "CD79A")


# Read genes for curated feature selection
# genes1 <- read_tsv(path_to_genes1, col_names = "gene")
# genes2 <- read_tsv(path_to_genes2, col_names = "gene")
```


Since here we will focus on the RNA, let us remove the ATAC assay:

```{r}
query[["ATAC"]] <- NULL
```


# Reprocess query

Since we subsetted the CD79A+ cells, let us rerun the general pipeline for dimensionality reduction:

```{r}
query <- query %>%
  FindVariableFeatures() %>%
  ScaleData() %>%
  RunPCA() %>%
  RunUMAP(dims = 1:20, reduction = "pca")
DimPlot(query, cols = color_palette, reduction = "umap")
```


Cluster

```{r}
query <- FindNeighbors(query, dims = 1:20, reduction = "pca")
query <- FindClusters(query, resolution = 0.15)
query2 <- query
DimPlot(query, cols = color_palette, reduction = "umap")
```


Markers

```{r}
markers_all <- FindAllMarkers(query, only.pos = TRUE)
markers_list <- purrr::map(levels(markers_all$cluster), function(x) {
  df <- markers_all <- markers_all[markers_all$cluster == x, ]
  df
})
names(markers_list) <- levels(markers_all$cluster)
openxlsx::write.xlsx(
  markers_list,
  file = here::here("MCL/results/tables/markers_clusters_tumoral_cells_M413.xlsx")
)
```


# Assay-specific feature selection

```{r}
# hvg <- SLOcatoR::find_assay_specific_features(
#   reference,
#   assay_var = "assay",
#   n_features = 5000
# )
# # rm(reference)


# Intersect with the curated list of genes
# curated_genes <- c(genes1$gene, genes2$gene)
# curated_genes <- curated_genes[curated_genes %in% hvg]
# 
# 
# # Subset objects
# reference <- subset(reference, features = curated_genes)
# query <- subset(query, features = curated_genes)
# curated_genes <- hvg

# Define new variables for label transfer
reference$type <- "reference"
query$type <- "query"
reference$batch <- reference@meta.data[[batch_var]]
reference$label <- reference@meta.data[[label_var]]
reference$annotation_probability <- NA
reference$UMAP1 <- Embeddings(reference, "umap")[, "UMAP_1"]
reference$UMAP2 <- Embeddings(reference, "umap")[, "UMAP_2"]
query$label <- NA
query$annotation_probability <- NA
query$batch <- "query"
query$UMAP1 <- NA
query$UMAP2 <- NA


# Find common columns and subset
common_cols <- intersect(
  colnames(reference@meta.data),
  colnames(query@meta.data)
)
reference@meta.data <- reference@meta.data[, common_cols]
query@meta.data <- query@meta.data[, common_cols]
```


Can we still discriminate the clusters in the reference after manually selecting the genes?

```{r}
# reference <- integrate_assays(
#   reference,
#   assay_specific = TRUE,
#   shared_hvg = curated_genes,
#   assay_var = "batch"
# )
# reference <- RunUMAP(reference, dims = 1:15, reduction = "harmony")
```


# Transfer label and coordinates

```{r}
# Merge
merged <- merge(x = reference, y = query)


# HVG
hvg <- SLOcatoR::find_assay_specific_features(
  merged,
  assay_var = "batch",
  n_features = 5000
)

# rm(reference)

# Integrate
merged <- integrate_assays(
  merged,
  assay_specific = TRUE,
  assay_var = "batch",
  shared_hvg = hvg,
  n_dim = 30
)


# Visualize integration
merged <- RunUMAP(merged, dims = 1:15, reduction = "harmony")
DimPlot(merged, group.by = "batch", split.by = "batch")
# rm(reference)


# Define training and test sets
data_sets <- SLOcatoR::split_training_and_test_sets(
  merged,
  split_var = "type",
  referece_label = "reference",
  query_label = "query",
  reduction = "harmony",
  n_dims = 15
)


# Transfer label
annotation_query_df <- transfer_label(
  seurat_obj = merged,
  training_set = data_sets$training_set,
  test_set = data_sets$test_set,
  k = 10,
  response_var = "label"
)


# Transfer coordinates
umap_test_df <- transfer_umap_coords(
  seurat_obj = merged,
  training_set = data_sets$training_set,
  test_set = data_sets$test_set,
  umap1_var = "UMAP1",
  umap2_var = "UMAP2",
  k = 10
)


# Plot
merged$UMAP1[umap_test_df$query_cells] <- umap_test_df$UMAP1
merged$UMAP2[umap_test_df$query_cells] <- umap_test_df$UMAP2
merged$label[annotation_query_df$query_cells] <- annotation_query_df$annotation
merged$annotation_probability[annotation_query_df$query_cells] <- annotation_query_df$annotation_prob
(umap_gg <- merged@meta.data %>%
  mutate(type = factor(type, levels = c("reference", "query"))) %>%
  ggplot(aes(UMAP1, UMAP2, color = label)) +
    geom_point(size = 0.5) +
    facet_wrap(~type) +
    scale_color_manual(values = pals::glasbey()) +
    theme_classic() +
    theme(legend.title = element_blank()) +
    guides(color = guide_legend(override.aes = list(size = 2))))
```


Distribution by clusters

```{r}
query2$annotation_predicted <- merged$label[colnames(query2)]
proportions_df <- query2@meta.data %>%
  group_by(seurat_clusters, annotation_predicted) %>%
  summarise(n_cells = n()) %>%
  mutate(
      n_cells_total = sum(n_cells),
      percentage_cells = round(n_cells / n_cells_total * 100, 3)
  )
(proportions_gg <- proportions_df %>%
  ggplot(aes(seurat_clusters, percentage_cells, fill = annotation_predicted)) +
    geom_col() +
    scale_fill_manual(values = color_palette) +
    labs(x = "Clusters", y = "% of cells", fill = "") +
    theme_bw())
```



# Save

```{r}
saveRDS(merged, path_to_save)
saveRDS(query2, path_to_save_query)
```


Save input to shiny app:

```{r}
input_shiny <- seurat2shiny(
  query2,
  assay = "RNA",
  slot = "data",
  reduction = "umap"
)
saveRDS(input_shiny$metadata, path_to_save_shiny_metadata)
saveRDS(input_shiny$expression, path_to_save_shiny_expression)
```


# Session Information

```{r}
sessionInfo()
```

