---
title: "Obtantion of loom file from Seurat Object"
author: "Sergio Aguilar"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
In this Rmarkdown document we are going to convert a Seurat Objet to a loom file using the __SCopeLoomR package__. This script was initially thought to be used for obtaining the file needeed for running  __pySCENIC__ and compute the GRN; however, it could also be use to store data and load it into python in an easy way. Furthermore, we will also obtain a metadata.csv file to store the metadata.

## Libraries
```{r warning = FALSE, message = FALSE}
library(dplyr)
library(Seurat)
library(SCopeLoomR)
```



## Parameters
```{r}
#name of the project
project_name = "name_of_the_project"

#Folder structure
base_folder <- paste0("{path_to_project_folder}", project_name)
data_folder <- paste0(base_folder, "/data/")
```

## Load object
First, we load the initial seurat object from which we want to obtain the regulons using pySCENIC. 
```{r}
seurat_object <- readRDS(paste0(data_folder, project_name, "_seu_obj.rds"))
```

## Save metadata object
Secondly, we save the metadata for obtaining the regulon specificity score (RSS) from one of the columns
```{r}
subset <- seurat_object@meta.data

library(data.table)
fwrite(x = subset, row.names = TRUE, file = paste(data_folder, project_name, "_metadata.csv", sep=""))
```

## Obtain loom file
Finally, we obtain the loom file using SCopeLoomR package
```{r}
counts <- seurat_object@assays$RNA@counts
Emb_expr <- seurat_object[["umap"]]@cell.embeddings
default.emb <- Emb_expr
default.emb.name <- "Umap"

file.name <- paste0(data_folder, project_name, ".loom")
build_loom(
  file.name=file.name,
  dgem=counts,
  title=paste0("loom file from ", project_name),
  genome="Human", # Just for user information, not used internally
  default.embedding=default.emb,
  default.embedding.name=default.emb.name
)
```
