---
title: "CITEseq marker Differential"
author: "Sonal Rashmi"
date: "`r format(Sys.time(), '%B %Y')`"
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center',
                      message=FALSE, warning = FALSE)
options(width = 1200)
```


# Pre-processing

## Load packages

```{r libraries}
library(Seurat)
library(Signac)
library(flexclust)
library(tidyverse)
library(plyr)
library(harmony)
library(class)
library(ggplot2)
library(reshape2)
```

## Parameters

```{r parameters}
path_to_save_annotated_citeseq <- "/home/srashmi/Documents/tonsil_atlas_citeseq_vdj_20210505/10-sub_cluster_annotation/PC_differential/objects/PC_tonsil_cite_seq_annotated.rds"
color <- c("black", "gray", "red", "yellow", "violet", "green4",
                   "blue", "mediumorchid2", "coral2", "blueviolet",
                   "indianred4", "deepskyblue1", "dimgray", "deeppink1",
                   "greenyellow", "lightgray", "hotpink1", "chocolate", "aquamarine", "aliceblue", "burlywood", "blueviolet", "indianred", "gold","coral", "orangered")
```

## Variables

## Load data

```{r}
seurat_citeseq <- readRDS(path_to_save_annotated_citeseq)
DefaultAssay(seurat_citeseq) <- "ADT"
Idents(seurat_citeseq) <- seurat_citeseq@meta.data$PC_annotation_level_5
```


## GZ-LZ vs PC commited

```{r}
markers <- FindMarkers(object = seurat_citeseq, ident.1 = "GC-LZ", ident.2 = "PC committed GC-LZ", min.pct = 0.25)
DT::datatable(markers)
```


```{r}
DoHeatmap(seurat_citeseq, features = unique(rownames(markers)), label = F)
```
```{r , echo=FALSE}
seurat_citeseq@reductions$wnn.umap@cell.embeddings[,1] <- seurat_citeseq$PC_UMAP1
seurat_citeseq@reductions$wnn.umap@cell.embeddings[,2] <- seurat_citeseq$PC_UMAP2
defined_markers <- rownames(markers)
chunk <- function(x, n) split(x, sort(rank(x) %% n))
num<-round(length(defined_markers)/4)
defined_marker_list<-chunk(defined_markers,num)
purrr::map(defined_marker_list, function(l) {
  FeaturePlot(seurat_citeseq, features = l, pt.size = 0.1, order=TRUE) 
})
```


## GZ-LZ vs IgG+ PC precursor

```{r}
markers <- FindMarkers(object = seurat_citeseq, ident.1 = "GC-LZ", ident.2 = "IgG+ PC precursor", min.pct = 0.25)
DT::datatable(markers)
```


```{r}
DoHeatmap(seurat_citeseq, features = unique(rownames(markers)), label = F) 
```
```{r , echo=FALSE}
seurat_citeseq@reductions$wnn.umap@cell.embeddings[,1] <- seurat_citeseq$PC_UMAP1
seurat_citeseq@reductions$wnn.umap@cell.embeddings[,2] <- seurat_citeseq$PC_UMAP2
defined_markers <- rownames(markers)
chunk <- function(x, n) split(x, sort(rank(x) %% n))
num<-round(length(defined_markers)/4)
defined_marker_list<-chunk(defined_markers,num)
purrr::map(defined_marker_list, function(l) {
  FeaturePlot(seurat_citeseq, features = l, pt.size = 0.1, order=TRUE) 
})
```

## GZ-LZ vs Pre-Mature IgG+ PC

```{r}
markers <- FindMarkers(object = seurat_citeseq, ident.1 = "GC-LZ", ident.2 = "Pre-Mature IgG+ PC", min.pct = 0.25)
DT::datatable(markers)
```


```{r}
DoHeatmap(seurat_citeseq, features = unique(rownames(markers)), label = F) 
```


```{r , echo=FALSE}
seurat_citeseq@reductions$wnn.umap@cell.embeddings[,1] <- seurat_citeseq$PC_UMAP1
seurat_citeseq@reductions$wnn.umap@cell.embeddings[,2] <- seurat_citeseq$PC_UMAP2
defined_markers <- rownames(markers)
chunk <- function(x, n) split(x, sort(rank(x) %% n))
num<-round(length(defined_markers)/4)
defined_marker_list<-chunk(defined_markers,num)
purrr::map(defined_marker_list, function(l) {
  FeaturePlot(seurat_citeseq, features = l, pt.size = 0.1, order=TRUE) 
})
```

## GZ-LZ vs Mature PC ("MBC-derived IgG+ PC", "MBC-derived IgA+ PC", "Mature IgM+ PC", "Mature IgG+ PC", "Mature IgA+ PC")

```{r}
markers <- FindMarkers(object = seurat_citeseq, ident.1 = "GC-LZ", ident.2 = c("MBC-derived IgG+ PC", "MBC-derived IgA+ PC", "Mature IgM+ PC", "Mature IgG+ PC", "Mature IgA+ PC"), min.pct = 0.25)
DT::datatable(markers)
```


```{r}
DoHeatmap(seurat_citeseq, features = unique(rownames(markers)), label = F) 
```


```{r , echo=FALSE}
seurat_citeseq@reductions$wnn.umap@cell.embeddings[,1] <- seurat_citeseq$PC_UMAP1
seurat_citeseq@reductions$wnn.umap@cell.embeddings[,2] <- seurat_citeseq$PC_UMAP2
defined_markers <- rownames(markers)
chunk <- function(x, n) split(x, sort(rank(x) %% n))
num<-round(length(defined_markers)/4)
defined_marker_list<-chunk(defined_markers,num)
purrr::map(defined_marker_list, function(l) {
  FeaturePlot(seurat_citeseq, features = l, pt.size = 0.1, order=TRUE) 
})
```

## PC committed GC-LZ vs IgG+ PC precursor

```{r}
markers <- FindMarkers(object = seurat_citeseq, ident.1 = "PC committed GC-LZ", ident.2 = "IgG+ PC precursor", min.pct = 0.25)
DT::datatable(markers)
```


```{r}
DoHeatmap(seurat_citeseq, features = unique(rownames(markers)), label = F) 
```


```{r , echo=FALSE}
seurat_citeseq@reductions$wnn.umap@cell.embeddings[,1] <- seurat_citeseq$PC_UMAP1
seurat_citeseq@reductions$wnn.umap@cell.embeddings[,2] <- seurat_citeseq$PC_UMAP2
defined_markers <- rownames(markers)
chunk <- function(x, n) split(x, sort(rank(x) %% n))
num<-round(length(defined_markers)/4)
defined_marker_list<-chunk(defined_markers,num)
purrr::map(defined_marker_list, function(l) {
  FeaturePlot(seurat_citeseq, features = l, pt.size = 0.1, order=TRUE) 
})
```


## PC committed GC-LZ vs Pre-Mature IgG+ PC

```{r}
markers <- FindMarkers(object = seurat_citeseq, ident.1 = "PC committed GC-LZ", ident.2 = "Pre-Mature IgG+ PC", min.pct = 0.25)
DT::datatable(markers)
```


```{r}
DoHeatmap(seurat_citeseq, features = unique(rownames(markers)), label = F) 
```


```{r , echo=FALSE}
seurat_citeseq@reductions$wnn.umap@cell.embeddings[,1] <- seurat_citeseq$PC_UMAP1
seurat_citeseq@reductions$wnn.umap@cell.embeddings[,2] <- seurat_citeseq$PC_UMAP2
defined_markers <- rownames(markers)
chunk <- function(x, n) split(x, sort(rank(x) %% n))
num<-round(length(defined_markers)/4)
defined_marker_list<-chunk(defined_markers,num)
purrr::map(defined_marker_list, function(l) {
  FeaturePlot(seurat_citeseq, features = l, pt.size = 0.1, order=TRUE) 
})
```


## PC committed GC-LZ vs Mature PC ("MBC-derived IgG+ PC", "MBC-derived IgA+ PC", "Mature IgM+ PC", "Mature IgG+ PC", "Mature IgA+ PC")

```{r}
markers <- FindMarkers(object = seurat_citeseq, ident.1 = "PC committed GC-LZ", ident.2 = c("MBC-derived IgG+ PC", "MBC-derived IgA+ PC", "Mature IgM+ PC", "Mature IgG+ PC", "Mature IgA+ PC"), min.pct = 0.25)
DT::datatable(markers)
```


```{r}
DoHeatmap(seurat_citeseq, features = unique(rownames(markers)), label = F) 
```


```{r , echo=FALSE}
seurat_citeseq@reductions$wnn.umap@cell.embeddings[,1] <- seurat_citeseq$PC_UMAP1
seurat_citeseq@reductions$wnn.umap@cell.embeddings[,2] <- seurat_citeseq$PC_UMAP2
defined_markers <- rownames(markers)
chunk <- function(x, n) split(x, sort(rank(x) %% n))
num<-round(length(defined_markers)/4)
defined_marker_list<-chunk(defined_markers,num)
purrr::map(defined_marker_list, function(l) {
  FeaturePlot(seurat_citeseq, features = l, pt.size = 0.1, order=TRUE) 
})
```


## IgG+ PC precursor vs Pre-Mature IgG+ PC

```{r}
markers <- FindMarkers(object = seurat_citeseq, ident.1 = "IgG+ PC precursor", ident.2 = "Pre-Mature IgG+ PC", min.pct = 0.25)
DT::datatable(markers)
```


```{r}
DoHeatmap(seurat_citeseq, features = unique(rownames(markers)), label = F)  
```


```{r , echo=FALSE}
seurat_citeseq@reductions$wnn.umap@cell.embeddings[,1] <- seurat_citeseq$PC_UMAP1
seurat_citeseq@reductions$wnn.umap@cell.embeddings[,2] <- seurat_citeseq$PC_UMAP2
defined_markers <- rownames(markers)
chunk <- function(x, n) split(x, sort(rank(x) %% n))
num<-round(length(defined_markers)/4)
if(num==0){
  defined_marker_list<-chunk(defined_markers,1)
} else {
  defined_marker_list<-chunk(defined_markers,num)
}
purrr::map(defined_marker_list, function(l) {
  FeaturePlot(seurat_citeseq, features = l, pt.size = 0.1, order=TRUE) 
})
```


## IgG+ PC precursor vs Mature PC ("MBC-derived IgG+ PC", "MBC-derived IgA+ PC", "Mature IgM+ PC", "Mature IgG+ PC", "Mature IgA+ PC")

```{r}
markers <- FindMarkers(object = seurat_citeseq, ident.1 = "IgG+ PC precursor", ident.2 = c("MBC-derived IgG+ PC", "MBC-derived IgA+ PC", "Mature IgM+ PC", "Mature IgG+ PC", "Mature IgA+ PC"), min.pct = 0.25)
DT::datatable(markers)
```


```{r}
DoHeatmap(seurat_citeseq, features = unique(rownames(markers)), label = F)
```


```{r, echo=FALSE }
seurat_citeseq@reductions$wnn.umap@cell.embeddings[,1] <- seurat_citeseq$PC_UMAP1
seurat_citeseq@reductions$wnn.umap@cell.embeddings[,2] <- seurat_citeseq$PC_UMAP2
defined_markers <- rownames(markers)
chunk <- function(x, n) split(x, sort(rank(x) %% n))
num<-round(length(defined_markers)/4)
defined_marker_list<-chunk(defined_markers,num)
purrr::map(defined_marker_list, function(l) {
  FeaturePlot(seurat_citeseq, features = l, pt.size = 0.1, order=TRUE) 
})
```


## Pre-Mature IgG+ PC vs Mature PC ("MBC-derived IgG+ PC", "MBC-derived IgA+ PC", "Mature IgM+ PC", "Mature IgG+ PC", "Mature IgA+ PC")

```{r}
markers <- FindMarkers(object = seurat_citeseq, ident.1 = "Pre-Mature IgG+ PC", ident.2 = c("MBC-derived IgG+ PC", "MBC-derived IgA+ PC", "Mature IgM+ PC", "Mature IgG+ PC", "Mature IgA+ PC"), min.pct = 0.25)
DT::datatable(markers)
```


```{r}
DoHeatmap(seurat_citeseq, features = unique(rownames(markers)), label = F) 
```


```{r, echo=FALSE}
seurat_citeseq@reductions$wnn.umap@cell.embeddings[,1] <- seurat_citeseq$PC_UMAP1
seurat_citeseq@reductions$wnn.umap@cell.embeddings[,2] <- seurat_citeseq$PC_UMAP2
defined_markers <- rownames(markers)
chunk <- function(x, n) split(x, sort(rank(x) %% n))
num<-round(length(defined_markers)/4)
defined_marker_list<-chunk(defined_markers,num)
purrr::map(defined_marker_list, function(l) {
  FeaturePlot(seurat_citeseq, features = l, pt.size = 0.1, order=TRUE) 
})
```


# Session Information

```{r session}
sessionInfo()
```