---
title: "Germinal Center B cells level 3 to level 4"
author: "Sergio Aguilar Fernández"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  BiocStyle::html_document:
    df_print: paged
    toc: true
    toc_float: true
    number_sections: false
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      out.width = "100%", 
                      fig.align='center',
                      fig.path="figures_compiled_html/",
                      message=FALSE, 
                      warning = FALSE)
```

## Introduction

In this Rmd we are going to clean the level 3 of the __*Germinal Center B Cells (GCBC)*__. Here, we will focus on removing potential doublets and low quality cells to provide a curated level 4 for the *GCBC*. 

## Libraries 

```{r}
library(Seurat)
library(tidyverse)
library(harmony)
```


## Parameters 

### Folder structure
```{r}
level <- "level_3"
cell_type <- "GCBC"
base_folder <- here::here(paste0(level,"/",cell_type,"/"))
  data_folder <- paste0(base_folder, "data/")
  results_folder <- paste0(base_folder, "results/")
  figures_folder <- paste0(results_folder, "figures/")
  markers_folder <- paste0(results_folder, "objects/markers/")

level_4_folder = here::here(paste0("level_4/",cell_type,"/data/"))
```

### Palette and pt.size
```{r}
palette<-c('#e6194b', '#3cb44b', '#ffd8b1', '#4363d8', '#f58231', '#911eb4', '#46f0f0', '#f032e6', '#bcf60c', '#fabebe', '#008080', '#e6beff', '#9a6324', '#fffac8', '#800000', '#aaffc3', '#808000', '#ffe119', '#000075', '#808080', '#ffffff', '#000000')

pt_sizes <- c(
  NBC_MBC = 0.6,
  GCBC = 0.8,
  PC = 1.5
)
pt_size <- pt_sizes[cell_type]
```

## Functions

### Visualization exploratory analysis of a given object 

#### Data groups visualization
```{r}
data_groups_visualization <- function(seurat_object, group_variable_1, group_variable_2 = NULL, pt_size) {
        print(Seurat::DimPlot(seurat_object, 
                reduction = "umap",
                pt.size = pt_size,  
                cols= palette, 
                group.by = group_variable_1, 
                raster = FALSE) 
        )
        
        print(Seurat::DimPlot(seurat_object, 
                reduction = "umap",
                pt.size = pt_size,  
                cols= palette, 
                group.by = group_variable_1, 
                raster = FALSE, 
                label = TRUE)
        )
        
        if (!is.null(group_variable_2)) {
            print(Seurat::DimPlot(seurat_object, 
                reduction = "umap",
                pt.size = pt_size,  
                cols= palette, 
                group.by = group_variable_2, 
                raster = FALSE) 
            )
        }
        

        
        #### COMBINED EXPLORATORY DATA GROUPS ANALYSIS
        p1 <- Seurat::DimPlot(seurat_object, 
                reduction = "umap",
                pt.size = pt_size/3,  
                cols= palette, 
                group.by = group_variable_1, 
                raster = FALSE) 
        
        p2 <- Seurat::DimPlot(seurat_object, 
                reduction = "umap", 
                pt.size = pt_size/3,  
                cols= c("tan", "royalblue1", "red"), 
                group.by = "assay", 
                raster = FALSE) 
        
        p3 <- Seurat::DimPlot(seurat_object, 
                reduction = "umap", 
                pt.size = pt_size/3,  
                cols= palette, 
                group.by = "age_group", 
                raster = FALSE) 
        
        p4 <- Seurat::DimPlot(seurat_object, 
                reduction = "umap", 
                pt.size = pt_size/3,  
                cols= c("tan", "royalblue1", "red"), 
                group.by = "hospital", 
                raster = FALSE) 
        
        
        print(patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2) )
}
```

#### Data features Visualization
```{r}
data_features_visualization <- function(seurat_object, group_variable_1, pt_size) {

    p1 <- Seurat::FeaturePlot(seurat_object, 
                        features = "nCount_RNA", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        raster = FALSE) 
    
    p2 <- Seurat::FeaturePlot(seurat_object, 
                        features = "nFeature_RNA", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        raster = FALSE) 
    
    p3 <- Seurat::FeaturePlot(seurat_object, 
                        features = "pct_mt", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        raster = FALSE) 
    
    p4 <- Seurat::FeaturePlot(seurat_object, 
                        features = "pct_ribosomal", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        raster = FALSE) 
    
    
      print(patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2) )

    p1 <- Seurat::DimPlot(seurat_object, 
            reduction = "umap",
            pt.size = pt_size/3,  
            cols= palette, 
            group.by = group_variable_1, 
            raster = FALSE) 
    
    p2 <- Seurat::FeaturePlot(seurat_object, 
                        features = "pct_mt", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        raster = FALSE) 
    
    p3 <- Seurat::FeaturePlot(seurat_object, 
                        features = "nCount_RNA", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        raster = FALSE) 
    
    p4 <- Seurat::FeaturePlot(seurat_object, 
                        features = "nFeature_RNA", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        raster = FALSE) 
    
     print(patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2) )
    
    
    p1 <- Seurat::FeaturePlot(seurat_object, 
                        features = "nCount_RNA", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        min.cutoff = 'q10',
                        max.cutoff = 'q90', 
                        raster = FALSE) 
                        
    
    p2 <- Seurat::FeaturePlot(seurat_object, 
                        features = "nFeature_RNA", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        min.cutoff = 'q10',
                        max.cutoff = 'q90', 
                        raster = FALSE) 
                        
    
      print(patchwork::wrap_plots(p1, p2, ncol = 2) )
      
    g1 <- ggplot(seurat_object@meta.data, aes_string(x=group_variable_1, y="nFeature_RNA", fill=group_variable_1)) + 
        geom_boxplot() + 
        theme(legend.position = "none") + 
        scale_fill_manual(values=palette) + 
        theme(axis.text.x = element_text(angle = 90))
    g2 <- ggplot(seurat_object@meta.data, aes_string(x=group_variable_1, y="nCount_RNA", fill=group_variable_1)) + 
        geom_boxplot() + 
        scale_fill_manual(values=palette) + 
        theme(axis.text.x = element_text(angle = 90))
    
      print(patchwork::wrap_plots(g1, g2, ncol = 2) )

}
```

### Quality Control visualizarion analysis of a given cluster 
```{r}
cluster_qc_analysis_visualization <- function(seurat_object, group_variable_1, cluster_name, subcluster_name, pt_size) {
  
  cluster_cells <- Cells(seurat_object_l3_noRoyalLondon)[which(seurat_object_l3_noRoyalLondon@meta.data[,group_variable_1] == cluster_name)]
  
     ## INSPECT SUBCLUSTERS
    print(Seurat::DimPlot(seurat_object, 
            reduction = "umap", 
            pt.size = pt_size/2,  
            cols= c(palette, "black"),
            group.by = subcluster_name, 
            raster = FALSE) 
    ) 
    
    print(Seurat::DimPlot(seurat_object, 
            reduction = "umap", 
            pt.size = pt_size,  
            cols= c(palette, "black"),
            group.by = subcluster_name, 
            cells=cluster_cells)
    ) 
    
    ## INSPECT GENERAL VARIABLES
    p1 <- Seurat::DimPlot(seurat_object, 
            reduction = "umap",
            pt.size = pt_size/3,  
            cols= c(palette, "black"),
            group.by = subcluster_name, 
            raster = FALSE, 
            cells=cluster_cells)
    
    
    p2 <- Seurat::DimPlot(seurat_object, 
            reduction = "umap", 
            pt.size = pt_size/3,  
            cols=palette, 
            group.by = "assay", 
            raster = FALSE, 
            cells=cluster_cells)
    
    p3 <- Seurat::DimPlot(seurat_object, 
            reduction = "umap", 
            pt.size = pt_size/3,  
            cols= palette, 
            group.by = "age_group", 
            raster = FALSE, 
            cells=cluster_cells)
    
    p4 <- Seurat::DimPlot(seurat_object, 
            reduction = "umap", 
            pt.size = pt_size/3,  
            cols= palette, 
            group.by = "donor_id", 
            raster = FALSE, 
            cells=cluster_cells)
    
      print(patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)  ) 
    
    
    ## INSPECT GENERAL QUALITY 
    p1 <- Seurat::FeaturePlot(seurat_object, 
                        features = "nCount_RNA", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        cells=cluster_cells ) 
    
    
    p2 <- Seurat::FeaturePlot(seurat_object, 
                        features = "nFeature_RNA", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        cells=cluster_cells ) 
    
    
    p3 <- Seurat::FeaturePlot(seurat_object, 
                        features = "pct_mt", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        cells=cluster_cells ) 
    
    p4 <- Seurat::FeaturePlot(seurat_object, 
                        features = "pct_ribosomal", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        cells=cluster_cells ) 
    
    
      print(patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)  ) 
    
    
    ## INSPECT nCOUNTS and nFEATURES removing the extreme values 
    p1 <- Seurat::FeaturePlot(seurat_object, 
                        features = "nCount_RNA", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        min.cutoff = 'q10',
                        max.cutoff = 'q90', 
                        cells=cluster_cells ) 
                        
    
    p2 <- Seurat::FeaturePlot(seurat_object, 
                        features = "nFeature_RNA", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        min.cutoff = 'q10',
                        max.cutoff = 'q90', 
                        cells=cluster_cells ) 
    
      print(patchwork::wrap_plots(p1, p2, ncol = 2)  ) 
    
    ## INSPECT THE RELATIONSHIP BETWEEN QUALITY AND SUBCLUSTERS
    
    ### Combined plot
    p1 <- Seurat::DimPlot(seurat_object, 
            reduction = "umap", 
            pt.size = pt_size/3,  
            cols= c(palette, "black"),
            group.by = subcluster_name, 
            cells=cluster_cells)
    
    p2 <- Seurat::FeaturePlot(seurat_object, 
                        features = "pct_mt", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        cells=cluster_cells ) 
    
    p3 <- Seurat::FeaturePlot(seurat_object, 
                        features = "nCount_RNA", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        min.cutoff = 'q10',
                        max.cutoff = 'q90', 
                        cells=cluster_cells ) 
                        
    
    p4 <- Seurat::FeaturePlot(seurat_object, 
                        features = "nFeature_RNA", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        min.cutoff = 'q10',
                        max.cutoff = 'q90', 
                        cells=cluster_cells ) 
    
      print(patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)  ) 
    
    ### Boxplot
    df <- seurat_object@meta.data[cluster_cells,]
    
    g1 <- ggplot(df, aes_string(x=subcluster_name, y="nFeature_RNA", fill=subcluster_name)) + 
        geom_boxplot() + 
        theme(legend.position = "none") + 
        scale_fill_manual(values=palette) + 
        theme(axis.text.x = element_text(angle = 90))
    g2 <- ggplot(df, aes_string(x=subcluster_name, y="nCount_RNA", fill=subcluster_name)) + 
        geom_boxplot() + 
        scale_fill_manual(values=palette) + 
        theme(axis.text.x = element_text(angle = 90))
    
      print(patchwork::wrap_plots(g1, g2, ncol = 2) )      
}
```

## Load data 
```{r}
seurat_object_l3 <- readRDS(paste0(data_folder, cell_type, "_clustered_level_3_with_pre_freeze.rds"))
```


## Addition of proposed annotation by the annotation team
```{r}
seurat_object_l3$RNA_snn_res.0.45 <- as.factor(seurat_object_l3$RNA_snn_res.0.45)
seurat_object_l3$RNA_snn_res.0.45 <- factor(seurat_object_l3$RNA_snn_res.0.45, levels=c("0", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12"))

Idents(seurat_object_l3) <- "RNA_snn_res.0.45"


new.cluster.ids <- c("DZ/LZ", # merge with 6 
                     "Proliferative DZ1", 
                     "AID-non proliferative", # subcluster to find the intermediate
                     "AID-proliferative",
                     "B cell activation, pro-survival", # subcluster into two to find proliferative
                     "Proliferative DZ2", 
                     "DZ/LZ", # merge with 0
                     "Proliferative LZ", 
                     "Multiome old-adult specific", 
                     "Precursor Plasmablasts", # subcluster into two CD9_0, CD9_1
                     "Proliferation-technical cluster",
                     "RoyalLondon-specific",
                     "RoyalLondon-specific"
                     )
names(new.cluster.ids) <- levels(seurat_object_l3)
seurat_object_l3 <- RenameIdents(seurat_object_l3, new.cluster.ids)
seurat_object_l3$annotation_level_3 <- seurat_object_l3@active.ident
seurat_object_l3$annotation_level_3 <- as.character(seurat_object_l3$annotation_level_3)
```

## Exploratory analysis of level 3

### Explore the UMAP and different grouping variables 
```{r level3_exploratory_analysis_group_variables, fig.height = 12, fig.width = 16}
data_groups_visualization(seurat_object = seurat_object_l3, 
                          group_variable_1 = "annotation_level_3",
                          pt_size = pt_size) 
```


### Explore the UMAP and different data variables 

```{r level3_exploratory_analysis_data_variables, fig.height = 12, fig.width = 16}
data_features_visualization(seurat_object = seurat_object_l3, 
                            group_variable_1 = "annotation_level_3",  
                            pt_size = pt_size) 
```

### Explore the hospital procedence of the cells
```{r}
table(seurat_object_l3$hospital)
```

### Explore the relationship between the clusters and the assay and age groups
```{r}
table(seurat_object_l3$RNA_snn_res.0.45, seurat_object_l3$assay, seurat_object_l3$age_group)
table(seurat_object_l3$annotation_level_3, seurat_object_l3$assay,  seurat_object_l3$age_group)

```


## Removal of Royal London cells (Hamish King)
```{r}
Idents(seurat_object_l3) <- "hospital"
seurat_object_l3_clean <- subset(
  seurat_object_l3,
  cells = colnames(seurat_object_l3)[Idents(seurat_object_l3) != "Royal London"]
)
```


### Re-processing without Royal London cells

```{r}
seurat_list <- SplitObject(seurat_object_l3_clean, split.by = "assay")
seurat_list <- seurat_list[c("3P", "multiome")]
seurat_list <- purrr::map(
  seurat_list,
  FindVariableFeatures,
  nfeatures = 5000
)
hvg <- purrr::map(seurat_list, VariableFeatures)
shared_hvg <- intersect(hvg$`3P`, hvg$multiome)
seurat_object_l3_clean <- seurat_object_l3_clean %>%
  ScaleData(features = shared_hvg) %>%
  RunPCA(features = shared_hvg) %>%
  RunHarmony(group.by.vars = "assay", reduction ="pca", dims = 1:30) %>%
  RunUMAP(reduction = "harmony", dims = 1:30)

resolution_vector <- c(0.01, 0.05, 0.1, 0.25, 0.35, 0.5, 0.75, 0.9, 1)
seurat_object_l3_clean <- seurat_object_l3_clean %>%
  FindNeighbors(dims = 1:30, reduction = "harmony")  %>%
  FindClusters(resolution = resolution_vector)
```


```{r}
seurat_object_l3_noRoyalLondon <- seurat_object_l3_clean
```

### Exploratory analysis without Royal London Cells 

#### Data groups
```{r level3_without_RL_group_variables, fig.height = 12, fig.width = 16}
data_groups_visualization(seurat_object = seurat_object_l3_noRoyalLondon, 
                          group_variable_1 = "annotation_level_3", 
                          pt_size = pt_size) 
```

```{r Level3_without_RL_resolutions, fig.height = 12, fig.width = 16}
list_plots <- list()
for (index in 1:length(resolution_vector)) {
  resolution <- resolution_vector[index]
  resolution_wanted <- paste0("RNA_snn_res.", resolution)
  Idents(seurat_object_l3_noRoyalLondon) <- resolution_wanted
  p <- Seurat::DimPlot(seurat_object_l3_noRoyalLondon, reduction = "umap",  pt.size = pt_size/3, raster = FALSE, group.by = resolution_wanted)
  list_plots[[index]] <- p
}
print(patchwork::wrap_plots(list_plots))

#### COMBINED EXPLORATORY DATA GROUPS ANALYSIS
p1 <- Seurat::DimPlot(seurat_object_l3_noRoyalLondon, 
        reduction = "umap",
        pt.size = pt_size/3,  
        cols= palette, 
        group.by = "RNA_snn_res.0.35", 
        raster = FALSE)  

p2 <- Seurat::DimPlot(seurat_object_l3_noRoyalLondon, 
        reduction = "umap", 
        pt.size = pt_size/3,  
        cols= c("tan", "royalblue1"), 
        group.by = "assay", 
        raster = FALSE) 

p3 <- Seurat::DimPlot(seurat_object_l3_noRoyalLondon, 
        reduction = "umap", 
        pt.size = pt_size/3,  
        cols= palette, 
        group.by = "age_group", 
        raster = FALSE) 

p4 <- Seurat::DimPlot(seurat_object_l3_noRoyalLondon, 
        reduction = "umap", 
        pt.size = pt_size/3,  
        cols= c("tan", "royalblue1", "red"), 
        group.by = "hospital", 
        raster = FALSE) 

patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)
```

#### Data features

```{r level3_without_RL_data_variables, fig.height = 12, fig.width = 16}
data_features_visualization(seurat_object = seurat_object_l3_noRoyalLondon,
                            group_variable_1 = "annotation_level_3",  
                            pt_size = pt_size) 
```

```{r}
table(seurat_object_l3_noRoyalLondon$annotation_level_3, seurat_object_l3_noRoyalLondon$assay,  seurat_object_l3_noRoyalLondon$age_group)
```

The __Proliferation−technical__ cluster and __Multiome old−adult specific__ cluster seem to be the ones with the lowest number of counts and features. Additionally, those clusters have a higher mt% and hence, are candidates for removal. However, we first analyse them in more depth 

## Analysis of low quality clusters 

### Multiome old−adult specific cluster 

```{r}
cluster_name <- "Multiome old-adult specific"
subcluster_name <- "Multiome_oldadult_subcluster"
```


#### Subclustering 

```{r}
### Find subcluster 
Idents(seurat_object_l3_noRoyalLondon) <- "annotation_level_3"
seurat_object_l3_noRoyalLondon <- FindSubCluster(
  seurat_object_l3_noRoyalLondon,
  cluster = cluster_name,
  graph.name = "RNA_snn",
  subcluster.name = subcluster_name,
  resolution = 0.25
)
```

#### Visualization analysis
```{r Multiome_old_adult_qc_analysis, fig.height = 12, fig.width = 16}
cluster_qc_analysis_visualization(seurat_object = seurat_object_l3_noRoyalLondon, 
                                  group_variable_1 = "annotation_level_3", 
                                  cluster_name = cluster_name, 
                                  subcluster_name = subcluster_name, 
                                  pt_size = pt_size)
```



### Proliferation−technical cluster 

```{r}
cluster_name <- "Proliferation-technical cluster"
subcluster_name <- "Proliferation_technical_subcluster"
```


#### Subclustering 
```{r}
### Find subcluster 
Idents(seurat_object_l3_noRoyalLondon) <- "annotation_level_3"
seurat_object_l3_noRoyalLondon <- FindSubCluster(
  seurat_object_l3_noRoyalLondon,
  cluster = cluster_name,
  graph.name = "RNA_snn",
  subcluster.name = subcluster_name,
  resolution = 0.5
)
```

#### Visualization analysis
```{r Proliferation-technical_qc_analysis, fig.height = 12, fig.width = 16}
cluster_qc_analysis_visualization(seurat_object = seurat_object_l3_noRoyalLondon, 
                                  group_variable_1 = "annotation_level_3", 
                                  cluster_name = cluster_name, 
                                  subcluster_name = subcluster_name, 
                                  pt_size = pt_size)
```




### DZ / LZ cluster (top right region)

#### Subclustering 
```{r}
cluster_name <- "DZ/LZ"
subcluster_name <- "DZ_LZ_subcluster"
```

```{r}
### Find subcluster 
Idents(seurat_object_l3_noRoyalLondon) <- "annotation_level_3"
seurat_object_l3_noRoyalLondon <- FindSubCluster(
  seurat_object_l3_noRoyalLondon,
  cluster = cluster_name,
  graph.name = "RNA_snn",
  subcluster.name = subcluster_name,
  resolution = 0.25
)
```

#### Visualization analysis
```{r DZ_LZ_qc_analysis, fig.height = 12, fig.width = 16}
cluster_qc_analysis_visualization(seurat_object = seurat_object_l3_noRoyalLondon, 
                                  group_variable_1 = "annotation_level_3", 
                                  cluster_name = cluster_name, 
                                  subcluster_name = subcluster_name, 
                                  pt_size = pt_size)
```

## Removal of Low quality clusters 

```{r}
### Remove the Multiome old-adult specific_0 subcluster (1344 cells)
Idents(seurat_object_l3_noRoyalLondon) <- "Multiome_oldadult_subcluster"
seurat_object_l3_clean <- subset(
  seurat_object_l3_noRoyalLondon,
  cells = colnames(seurat_object_l3_noRoyalLondon)[Idents(seurat_object_l3_noRoyalLondon) != "Multiome old-adult specific_0"]
)

### Remove completely the Proliferation−technical cluster  (1111 cells)
Idents(seurat_object_l3_clean) <- "annotation_level_3"
seurat_object_l3_clean <- subset(
  seurat_object_l3_clean,
  cells = colnames(seurat_object_l3_clean)[Idents(seurat_object_l3_clean) != "Proliferation-technical cluster"]
)

### Remove the DZ/LZ_3 (2337 cells)
Idents(seurat_object_l3_clean) <- "DZ_LZ_subcluster"
seurat_object_l3_clean <- subset(
  seurat_object_l3_clean,
  cells = colnames(seurat_object_l3_clean)[Idents(seurat_object_l3_clean) != "DZ/LZ_3"]
)
```


### Analysis of removed cells
```{r}
initial_cells <- ncol(seurat_object_l3)
final_cells <- ncol(seurat_object_l3_clean)
royal_london_cells_removed <- initial_cells - ncol(seurat_object_l3_noRoyalLondon) 
low_quality_cells_removed <- initial_cells - royal_london_cells_removed - final_cells
total_cells_removed <- royal_london_cells_removed + low_quality_cells_removed

```

```{r}
df <- as.data.frame(c(initial_cells, royal_london_cells_removed, low_quality_cells_removed, total_cells_removed, final_cells)) 
colnames(df) <- c("Number of cells")
rownames(df) <- c("Initial number of cells at level 3", "Royal London cells", "Low quality Cells", "Total number of cells removed", "Final number of cells")
```

```{r}
df 
```



## Reprocessing with the cleaned level 3 + cluster 3 from level 2 

### Addition of cluster 3 from Level 2 
```{r}
seurat_object_l2 <- readRDS(paste0(data_folder, cell_type, "_clustered_level_2.rds"))
```

After analysing the level 3, the annotation team realized that a cluster at level 2 annotated as doublets could have been pre-memory B cells. Consequently, we introduce those cells again for a better determination of their identity  

```{r}
seurat_object_l2$original_level <- "level_2"
seurat_object_l3_clean$original_level <- "level_3"
```

```{r}
Idents(seurat_object_l2) <- "RNA_snn_res.0.1"
clus3_level2 <- subset(
  seurat_object_l2,
  cells = colnames(seurat_object_l2)[Idents(seurat_object_l2) == 3]
)
```

```{r}
table(clus3_level2$hospital)
Idents(clus3_level2) <- "hospital"
clus3_level2 <- subset(
  clus3_level2,
  cells = colnames(clus3_level2)[Idents(clus3_level2) != "Royal London"]
)
table(clus3_level2$hospital)

```


```{r}
seurat_object_l3_clean_clus2 <- merge(seurat_object_l3_clean, y = clus3_level2, merge.data = TRUE)
```

### Reprocessing 
```{r}
seurat_object_l3_clean_clus2 <- merge(seurat_object_l3_clean, y = clus3_level2, merge.data = TRUE)

seurat_list <- SplitObject(seurat_object_l3_clean_clus2, split.by = "assay")
seurat_list <- seurat_list[c("3P", "multiome")]
seurat_list <- purrr::map(
  seurat_list,
  FindVariableFeatures,
  nfeatures = 5000
)
hvg <- purrr::map(seurat_list, VariableFeatures)
shared_hvg <- intersect(hvg$`3P`, hvg$multiome)
seurat_object_l3_clean_clus2 <- seurat_object_l3_clean_clus2 %>%
  ScaleData(features = shared_hvg) %>%
  RunPCA(features = shared_hvg) %>%
  RunHarmony(group.by.vars = "assay", reduction ="pca", dims = 1:30) %>%
  RunUMAP(reduction = "harmony", dims = 1:30)

resolution_vector <- c(0.01, 0.05, 0.1, 0.25, 0.35, 0.4, 0.45, 0.5, 0.75, 0.9, 1)
seurat_object_l3_clean_clus2 <- seurat_object_l3_clean_clus2 %>%
  FindNeighbors(dims = 1:30, reduction = "harmony")  %>%
  FindClusters(resolution = resolution_vector)
```

## Save object for Level 4 annotation
```{r save_level4_object}
seurat_object_l4 <- seurat_object_l3_clean_clus2
#saveRDS(seurat_object_l4, paste0(level_4_folder, cell_type, "_clustered_level_4.rds"))
```










