---
author: "Marc Elosua-Bayes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    toc: yes
    toc_float: yes
    df_print: paged
title: "Figure 1 - Spatial Figure-1`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message = FALSE, warning = FALSE)
options(width = 1200)
```

## Introduction
In this script we are going to make the spatial panels for Figure-1

## Libraries

```{r}
library(Seurat)
library(ggpubr)
library(cowplot)
library(dplyr)
library(ggplot2)
library(stringr)
library(readr)
library(Matrix)
library(SPOTlight)
```

## Setting parameters
Loading necessary paths and parameters
```{r}
set.seed(123)
source(here::here("misc/paths.R"))
source(here::here("utils/bin.R"))

"{fig1}/{plt_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(path = .,
             showWarnings = FALSE,
             recursive = TRUE)

"{fig1}/{robj_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(path = .,
             showWarnings = FALSE,
             recursive = TRUE)

clust_vr <- "annotation_figure_1"
```

```{r}
cols_fig1 <- c(
  "NBC" = "#dcf0f4",
  "Activated NBC" = "#7bc6d6", 
  "GCBC" = "#398D9F",
  "MBC" = "#025566", 
  "PC" = "#032872", 
  "CD4 T" = "#93331b", 
  "Naive CD4 T" = "#be8476", 
  "Naive CD8 T" = "#e5b7c6", 
  "CD8 T" = "#cb708e", 
  "NK" = "#67253a", 
  "DN" = "#d2a027", 
  "ILC" = "#88657f",
  "cycling T"  = "#cc0000",
  "preB/T" = "#4363d8", 
  "DC" = "#b6dec3", 
  "Mono/Macro" = "#6ac087", 
  "Mast" = "#3a8b55", 
  "Granulocytes" = "#32d732", 
  "cycling myeloid" = "#2b6840",
  "FDC" = "#d5ff00", 
  "cycling FDC" = "#bfe600", 
  "epithelial" = "#ffdc9f", 
  "PDC" = "#f032e6",
  "Cycling" = "lightgrey")

col_df <- data.frame(
  ct = names(cols_fig1),
  ct_mod = str_replace_all(names(cols_fig1), "[[:punct:]]|[[:space:]]", "."),
  color = cols_fig1)
```

## Load data
We have 8 different datasets that we have integrated in previous scripts - *03-clustering/03-clustering_integration.Rmds*.
We are going to analyze the integrated dataset all together since the regions are shared across all

```{r}
# Tonsils integrated
sp_obj <- "misc/{robj_dir}/20220215_tonsil_atlas_spatial_seurat_obj.rds" %>%
  glue::glue() %>%
  here::here() %>%
  readRDS(file = .)

decon_mtrx_ls <- "{fig1}/{robj_dir}/decon_mtrx_{clust_vr}.rds" %>%
  glue::glue() %>%
  here::here() %>%
  readRDS(file = .)
```

## Analysis
Change old names to ones unmodified without dots
```{r}
decon_mtrx <- decon_mtrx_ls[[2]]
decon_mtrx <- decon_mtrx[, colnames(decon_mtrx) != "res_ss"]
decon_mtrx <- decon_mtrx[, !is.na(colnames(decon_mtrx))]

# Set as 0 cell types predicted to be under 3 % of the spot
decon_mtrx[decon_mtrx < 0.03] <- 0

# new_colnames <- data.frame(ct_name = colnames(decon_mtrx),
#                            stringsAsFactors = FALSE) %>%
#   dplyr::left_join(col_df, by = c("ct_name" = "ct_mod")) %>%
#   dplyr::pull(ct)

# colnames(decon_mtrx) <- new_colnames
```

We are going to add the deconvolution to the Seurat object.
```{r}
sp_obj@meta.data <- cbind(sp_obj@meta.data, decon_mtrx)
```

Save the seurat object with the predicted cell-type proportions
```{r eval=FALSE}
"{fig1}/{robj_dir}/se_deconv_{clust_vr}.rds" %>%
  glue::glue() %>%
  here::here() %>%
  saveRDS(
    object = sp_obj,
    file = .)
```

### Clustering
Subset slide of interest
```{r}
i <- "esvq52_nluss5"
print(i)
# Subset sp object
keep <- sp_obj$gem_id == i
sp_sub <- sp_obj[, names(keep[keep])]
```

### Visualization

Before even looking at the decomposed spots we can gain insight on how well the model performed by looking at the topic profiles for the cell types.

#### Cell type location
Look at the location of each cell type in each slice separately
```{r fig.height=25, fig.width=25}
ct <- colnames(decon_mtrx)

# https://stackoverflow.com/questions/38722202/how-do-i-change-the-number-of-decimal-places-on-axis-labels-in-ggplot2
# Our transformation function to ensure 2 decimals in the legend
scaleFUN <- function(x) sprintf("%.2f", x)

# Iterate over imagesample_id
# lapply(id_sp_df$gem_id, function(i) {
# i <- "esvq52_nluss5"
# print(i)
# # Subset sp object
# sp_sub <- subset(sp_obj, subset = gem_id == i)

# Set filename
donor <- id_sp_df %>% 
dplyr::filter(gem_id == i) %>%
  dplyr::pull(donor_id)

fn <- "{fig1}/{plt_dir}/{i}_{clust_vr}_celltype.pdf" %>%
glue::glue() %>%
here::here()

# Iterate over cell types
lapply(ct, function(ii) {
  tmp_plt <- Seurat::SpatialFeaturePlot(
    object = sp_sub, 
    features = ii,
    alpha = c(0.1, 1),
    images = i)
  
if (sum(sp_sub@meta.data[, ii]) == 0) {
  tmp_plt <- tmp_plt + ggplot2::scale_alpha(range = c(0,0))
} else {
  tmp_plt <- tmp_plt + ggplot2::scale_alpha(range = c(0, 1))
}

  tmp_plt <- tmp_plt +
    ggplot2::theme(legend.title = ggplot2::element_blank()) +
    ggplot2::labs(title = ii) +
    ggplot2::scale_fill_gradientn(
      colors = grDevices::heat.colors(10, rev = TRUE),
      # Same number of breaks for all plots
      breaks = seq(min(sp_sub@meta.data[, ii]),
                   max(sp_sub@meta.data[, ii]),
                   length.out = 4),
      limits = c(0, max(sp_sub@meta.data[, ii])),
      # 2 decimals in the legend
      labels = scaleFUN
      # limits = c(0, 1)
    )
  tmp_plt
}) %>%
  cowplot::plot_grid(
    plotlist = .,
    align = "hv",
    axis = "trbl") %>%
  cowplot::save_plot(
    filename = fn,
    plot = .,
    base_height = 25,
    base_width = 25)
# })
```

#### Spatial scatterpie
```{r eval=TRUE, fig.height=10, fig.width=8}
# lapply(id_sp_df$gem_id, function(i) {
# print(i)
# Subset sp object
# sp_sub <- subset(sp_obj, subset = gem_id == i)

sct_plt <- SPOTlight::scatterpie_plot(
  se_obj = sp_sub,
  cell_types_all = ct,
  pie_scale = 0.425,
  slice = i) +
  ggplot2::scale_fill_manual(
    values = col_df[col_df$ct_mod %in% ct, "color"],
    breaks = col_df[col_df$ct_mod %in% ct, "ct_mod"],
    labels = stringr::str_wrap(string = col_df[col_df$ct_mod %in% ct, "ct"],
                               width = 25)) +
  ggplot2::coord_fixed(ratio = 1) +
  ggplot2::theme(
    legend.position = "top",
    legend.text = ggplot2::element_text(size = 12),
    legend.title = ggplot2::element_blank()) +
  ggplot2::guides(fill = ggplot2::guide_legend(nrow = 3, byrow = TRUE))

# sct_plt

# Set filename
"{fig1}/{plt_dir}/{i}_{clust_vr}_scatterpie_allct.pdf" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    filename = .,
    plot = sct_plt,
    base_width = 9,
    base_height = 9)

"{fig1}/{plt_dir}/{i}_{clust_vr}_scatterpie_allct.rds" %>%
  glue::glue() %>%
  here::here() %>%
  saveRDS(
    object = sct_plt,
    file = .)
```

We can also take a look at the spatial scatterpie by looking at cell types which are not present throughout the entire tissue.
```{r eval=TRUE, echo=FALSE, fig.height=10, fig.width=8}
lapply(id_sp_df$gem_id, function(i) {
  print(i)
  # Subset sp object
  sp_sub <- subset(sp_obj, subset = gem_id == i)
  # Subset cell types from metadata
  metadata_subset <- sp_sub@meta.data[, ct]
  
  # Create masks
  keep_0.7 <- colSums(metadata_subset > 0) < 0.9 * ncol(sp_sub)
  keep_g0 <- colSums(metadata_subset > 0) > 0
  
  # Select cell types fullfiling the conditions
  ct_var <- colnames(metadata_subset)[keep_0.7 & keep_g0]
  
  sct_int_plt <- SPOTlight::scatterpie_plot(
    se_obj = sp_sub,
    cell_types_all = ct_var,
    pie_scale = 0.4,
    slice = i) +
  ggplot2::scale_fill_manual(
    values = col_df[col_df$ct_mod %in% ct, "color"],
    breaks = col_df[col_df$ct_mod %in% ct, "ct_mod"],
    labels = stringr::str_wrap(string = col_df[col_df$ct_mod %in% ct, "ct"],
                               width = 25)) +
            ggplot2::coord_fixed(ratio = 1) +
            ggplot2::theme(
              legend.position = "top",
              legend.text = ggplot2::element_text(size = 12),
              legend.title = ggplot2::element_blank()) +
            ggplot2::guides(fill = ggplot2::guide_legend(nrow = 3, byrow = TRUE))
  
  # sct_int_plt
  
  # Set filename
  donor <- id_sp_df %>% 
  dplyr::filter(gem_id == i) %>%
    dplyr::pull(donor_id)
  
  "{fig1}/{plt_dir}/{i}_{clust_vr}_scatterpie_subset.pdf" %>%
    glue::glue() %>%
    here::here() %>%
    cowplot::save_plot(
      filename = .,
      plot = sct_int_plt,
      base_width = 9,
      base_height = 9)
  })
```

#### Cell-Cell correlation plot
By looking at correlations we want to assess if there are cell types which are correlating in space. So when one is present is another one also present?

We will be using `ggcorrplot` to show these correlation plots a bit nicer:
http://www.sthda.com/english/wiki/ggcorrplot-visualization-of-a-correlation-matrix-using-ggplot2

Get correlation matrix between all cell types
```{r echo=FALSE, fig.height=12, fig.width=12}
# Extract metadata features
mtrx <- sp_sub@meta.data[, ct]

# Remove those cell types that are all 0s
mtrx <- mtrx[, colSums(mtrx) > 0]

# Compute correlation matrix
mtrx_cor <- cor(as.matrix(mtrx))

# Correlation P value test
p.mat <- corrplot::cor.mtest(mat = mtrx, conf_int = 0.95, 
    method = "pearson")

# Modify column and row names
colnames(mtrx_cor) <- stringr::str_wrap(string = colnames(mtrx_cor), 
    width = 30)
rownames(mtrx_cor) <- stringr::str_wrap(string = rownames(mtrx_cor), 
    width = 30)

# Remove diagonal
diag(mtrx_cor) <- 0

colors = c("#6D9EC1", "white", "#E46726")
# Plot correlation heatmap
cor_plt <- ggcorrplot::ggcorrplot(
  corr = mtrx_cor,
  p.mat = p.mat[[1]], 
  hc.order = TRUE,
  type = "full",
  insig = "blank",
  lab = FALSE, 
  outline.col = "lightgrey",
  method = "square",
  colors = c("#6D9EC1", "white", "#E46726"),
  legend.title = glue::glue("Correlation\n(Pearson)")) +
  ggplot2::scale_fill_gradient2(
    low = colors[1],
    high = colors[3], 
    mid = colors[2],
    midpoint = 0,
    space = "Lab", 
    # name = legend.title,
    limits = c(min(mtrx_cor) - 0.01, max(mtrx_cor) + 0.01)) +
  ggplot2::theme(
    plot.title = ggplot2::element_text(
      size = 22,
      hjust = 0.5,
      face = "bold"),
    legend.text = ggplot2::element_text(size = 12), 
    legend.title = ggplot2::element_text(size = 15),
    axis.text.x = ggplot2::element_text(angle = 90,
                                        vjust = 0.5),
    axis.text = ggplot2::element_text(size = 18,
                                      vjust = 0.5))

cor_plt

"{fig1}/{plt_dir}/correlation_plot_{i}_{clust_vr}.pdf" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    plot = cor_plt,
    filename = .,
    base_width = 10,
    base_height = 8)
```

## Session Info
```{r}
sessionInfo()
```

