---
author: "Marc Elosua-Bayes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    toc: yes
    toc_float: yes
    df_print: paged
title: "Figure 1 - Spatial Panels"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message = FALSE, warning = FALSE)
options(width = 1200)
```

## Introduction
In this Rmarkdown document we are going to create the spatial pannels necessary for Figure 1.

## Libraries

```{r}
library(Seurat)
library(ggpubr)
library(cowplot)
library(dplyr)
library(ggplot2)
```

## Setting parameters
Loading necessary paths and parameters
```{r}
set.seed(123)
source(here::here("misc/paths.R"))
source(here::here("utils/bin.R"))

"{fig1}/{plt_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(path = .,
             showWarnings = FALSE,
             recursive = TRUE)

"{fig1}/{robj_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(path = .,
             showWarnings = FALSE,
             recursive = TRUE)

```

## Load data
The data used in this Rmarkdown document comes from **03-clustering_integration.Rmd** where the data was integrated.
```{r}
merged_se <- "misc/{robj_dir}/20220215_tonsil_atlas_spatial_seurat_obj.rds" %>%
  glue::glue() %>%
  here::here() %>%
  readRDS(file = .)
```

## Panels

### Panel 1
Update labels
```{r}
merged_se@meta.data <- merged_se@meta.data %>%
  mutate(
    annot_fig1 = case_when(
      annotation_20220215 %in% c("T:B Border", "Inter-Follicular Zone", "T cell zone") ~ "Inter-Follicular Zone",
      annotation_20220215 %in% c("Germinal Center", "Proliferating Follicle") ~ "Follicle",
      TRUE ~ annotation_20220215)
  )
```


Look at the spatial compartments
```{r fig.width=12, fig.height=9}
se_sub <- subset(merged_se, subset = gem_id == "esvq52_nluss5")
se_sub
se_sub@images <- se_sub@images[Seurat::Images(se_sub) == "esvq52_nluss5"]

# library(colorBlindness)
# colorBlindness::displayAvailablePalette()
# colorBlindness::paletteMartin
palette <- c(
  "Inter-Follicular Zone" = "#ff6db6",
  "Epithelial-1" = "#db6d00",
  "Epithelial-2" = "#924900",
  "Follicle" = "#004949", 
  "Muscle" = "#920000",
  "Subepithelial" = "#490092")

# (sp_plt <- SpatialDimPlot(
#   object = se_sub,
#   group.by = "annotation_20220215",
#   cols = palette,
#   image.alpha = 0,
#   stroke = 0
#   ) +
#         theme(
#         legend.box.background = element_blank(),
#         legend.key=element_blank()) + 
#         guides(fill = guide_legend(override.aes = list(size = 4))))

(sp_plt <- ggplot() +
    geom_point(aes(
        x = se_sub@images$esvq52_nluss5@coordinates$imagecol,
        y = se_sub@images$esvq52_nluss5@coordinates$imagerow,
        color = se_sub$annot_fig1),
        size = 1.25) +
    labs(color = "") +
    scale_y_reverse() +
    scale_color_manual(values = palette) +
    coord_fixed() +
    theme_void() +
    guides(color = guide_legend(override.aes = list(size = 4))))

"{fig1}/{plt_dir}/spatial_stratification_plt.pdf" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    filename = .,
    plot = sp_plt,
    base_height = 4,
    base_width = 7)

"{fig1}/{plt_dir}/spatial_stratification_plt.rds" %>%
  glue::glue() %>%
  here::here() %>%
  saveRDS(
    object = sp_plt,
    file = .)
```

## Panel 2
```{r}
img <- "{fig1}/{plt_dir}/visium_stacked.png" %>%
  glue::glue() %>%
  here::here() %>%
  png::readPNG(source = .)
library(grid)
x <- rasterGrob(img,
            interpolate = FALSE,
            width = unit(1, "npc"),
            height = unit(1, "npc"))

(img_gg <- ggplot() +
  annotation_custom(
    grob = x,
    xmin = 0,
    xmax = ncol(x$raster),
    ymin = 0,
    ymax = nrow(x$raster)
    ) +
  coord_fixed(
    # ratio = 1
    xlim = c(0, ncol(x$raster)),
    ylim = c(0, nrow(x$raster))
    ) +
  theme_void())

"{fig1}/{plt_dir}/visium_img_plt.pdf" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    filename = .,
    plot = img_gg,
    base_height = 9,
    base_width = 15)

"{fig1}/{plt_dir}/visium_img_plt.rds" %>%
  glue::glue() %>%
  here::here() %>%
  saveRDS(
    object = img_gg,
    file = .)
```

## Session Info
```{r}
sessionInfo()
```

