---
title: "Normalization and Batch Effect Correction (without doublets)"
author: "Ramon Massoni-Badosa"
date: "2021/01/29"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```


# Introduction

In previous notebooks and scripts, we excluded doublets and merged our Seurat object with the scRNA-seq dataset of tonsillar cells from [Hamish et al.](https://immunology.sciencemag.org/content/6/56/eabe6291). Here, we will correct for batch effects with [Harmony](https://www.nature.com/articles/s41592-019-0619-0), and visualize the intermixing of different potential confounders pre- and post-integration.

In addition, we will save the dimensionality reduction (PCA) matrices before and after integration to further quantify the effect of the aforementioned confounders in high dimensional space. 


# Pre-processing

## Load packages

```{r}
library(Seurat)
library(SeuratWrappers)
library(harmony)
library(tidyverse)
```


## Parameters

```{r}
# Paths
path_to_obj <- here::here("scRNA-seq/results/R_objects/seurat_merged_with_king_et_al.rds")
path_to_save_obj <- here::here("scRNA-seq/results/R_objects/seurat_merged_with_king_et_al_integrated.rds")
path_tmp_dir <- here::here("scRNA-seq/2-QC/5-batch_effect_correction/2-data_integration_king_et_al/tmp/")
path_to_save_dimred_uncorrect <- str_c(path_tmp_dir, "batch_uncorrected_pca.rds", sep = "")
path_to_save_dimred_correct <- str_c(path_tmp_dir, "batch_corrected_pca.rds", sep = "")
path_to_save_confounders_df <- str_c(path_tmp_dir, "confounders_df.rds", sep = "") 
```


## Load data

```{r}
tonsil <- readRDS(path_to_obj)
tonsil
```


# Visualize UMAP without batch effect correction

```{r}
# Process Seurat object
tonsil <- tonsil %>%
  NormalizeData(normalization.method = "LogNormalize", scale.factor = 1e4) %>% 
  FindVariableFeatures(nfeatures = 3000) %>%
  ScaleData() %>% 
  RunPCA() %>%
  RunUMAP(reduction = "pca", dims = 1:30)


# Visualize UMAP
confounders <- c("library_name", "sex", "age_group", "is_hashed",
                 "hospital", "assay")
umaps_before_integration <- purrr::map(confounders, function(x) {
  p <- DimPlot(tonsil, group.by = x, pt.size = 0.1)
  p
})
names(umaps_before_integration) <- confounders
print("UMAP colored by GEM:")
umaps_before_integration$library_name + NoLegend()
print("UMAP colored by sex, age group, cell hashing status, sampling center and assay:")
umaps_before_integration[2:length(umaps_before_integration)]
```


# Run and visualize Harmony's integration

```{r}
tonsil <- tonsil %>%
  RunHarmony(reduction = "pca", dims = 1:30, group.by.vars = "gem_id") %>%
  RunUMAP(reduction = "harmony", dims = 1:30)
umaps_after_integration <- purrr::map(confounders, function(x) {
  p <- DimPlot(tonsil, group.by = x, pt.size = 0.1)
  p
})
names(umaps_after_integration) <- confounders
print("UMAP colored by GEM:")
umaps_after_integration$library_name + NoLegend()
print("UMAP colored by sex, age group, cell hashing status, sampling center and assay:")
umaps_after_integration[2:length(umaps_before_integration)]
```


# Save

```{r}
# If it doesn't exist create temporal directory
dir.create(path_tmp_dir, showWarnings = FALSE) 


# Save integrated Seurat object
saveRDS(tonsil, path_to_save_obj)


# Save PCA matrices to compute the Local Inverse Simpson Index (LISI)
confounders_df <- tonsil@meta.data[, confounders]
saveRDS(confounders_df, path_to_save_confounders_df)
saveRDS(
  tonsil@reductions$pca@cell.embeddings[, 1:30],
  path_to_save_dimred_uncorrect
)
saveRDS(
  tonsil@reductions$harmony@cell.embeddings[, 1:30],
  path_to_save_dimred_correct
)
```


# Session Information

```{r}
sessionInfo()
```
