---
author: "Marc Elosua-Bayes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
title: "Plasma deconvolution assessment"
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message = FALSE, warning = FALSE, cache = FALSE,
                      cache.lazy = FALSE)
options(width = 1200)
```

## Introduction
In this script we are going to map Plasma cell onto the Visium slides.

## Libraries
```{r}
library(Seurat)
library(ggpubr)
library(cowplot)
library(dplyr)
library(ggplot2)
library(stringr)
library(readr)
library(SPOTlight)
library(SPATA2)
```

## Setting parameters
Loading necessary paths and parameters
```{r}
set.seed(123)
source(here::here("misc/paths.R"))
source(here::here("utils/bin.R"))

"{plasma}/{plt_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(path = ,
             showWarnings = FALSE,
             recursive = TRUE)

"{plasma}/{robj_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(path = ,
             showWarnings = FALSE,
             recursive = TRUE)
```

## Load data
We have 8 different datasets that we are going to analyze separately.
The spatial data comes from the script *03-clustering/03-clustering_integration.Rmd* while the sc data can be found in Ramon's scRNAseq analysis: **/scratch/devel/rmassoni/tonsil_atlas_private/2-DOWNSTREAM_PROCESSING/results/R_objects/processed_seurat_objects/processed_seurat_objects/tonsil_integrated_with_harmony_scrublet_annotated.rds**. \
```{r}
sp_obj <- "{anot}/{robj_dir}/integrated_spatial_annot.rds" %>%
  glue::glue() %>%
  here::here() %>%
  readRDS(file = .)

# Load SPOTlight data
spotlight_ls <- "{plasma}/{robj_dir}/spotlight_ls_plasma.rds" %>%
  glue::glue() %>% 
  here::here() %>%
  readRDS(file = .)
```

Set color dataframe for the cell types
```{r}
nm_df <- "{plasma}/{robj_dir}/plasma_nm_df.rds" %>%
  glue::glue() %>% 
  here::here() %>%
  readRDS(file = .)
```

Add colors to cell types
```{r}
library(RColorBrewer)
# Define the number of colors you want
nb.cols <- nrow(nm_df)
mycolors <- colorRampPalette(RColorBrewer::brewer.pal(8, "Set2"))(nb.cols)
# https://sashamaps.net/docs/resources/20-colors/
# color <- c('#e6194B', '#3cb44b', '#ffe119', '#4363d8', '#f58231', '#911eb4',
#            '#42d4f4', '#f032e6', '#bfef45', '#fabed4', '#469990', '#dcbeff',
#            '#9A6324', '#fffac8', '#800000', '#aaffc3', '#808000', '#ffd8b1',
#            '#000075', '#a9a9a9', '#ffffff', '#000000', '#b00b69')

col_df <- dplyr::bind_cols(nm_df, color = mycolors[1:nrow(nm_df)])
t <- data.frame(mod_nm = "Cycling", plt_nm = "Cycling", color = "black")
col_df <- rbind(col_df, t)
```

## Analysis

### Preprocess data
```{r}
decon_mtrx <- spotlight_ls[[2]]
decon_mtrx <- decon_mtrx[, colnames(decon_mtrx) != "res_ss"]

# Set as 0 cell types predicted to be under 3 % of the spot
decon_mtrx[decon_mtrx < 0.03] <- 0
```

Change column names
```{r eval=FALSE}
new_cn <- data.frame(mod_nm = colnames(decon_mtrx)) %>%
  dplyr::left_join(col_df, by = "mod_nm") %>%
  # Central.Mem.PASK. fives some trouble because it only changes between + an -
  # negative goes first and distinct solves it automatically
  dplyr::distinct() %>%
  dplyr::pull(plt_nm)

colnames(decon_mtrx) <- new_cn
```

We are going to add the deconvolution to the Seurat object.
```{r}
sp_obj@meta.data <- cbind(sp_obj@meta.data, decon_mtrx)
```

### Look at SPOTlight results
Check Topic profiles
```{r fig.height=8, fig.width=8}
nmf_mod_ls <- spotlight_ls[[1]]
nmf_mod <- nmf_mod_ls[[1]]

h <- NMF::coef(nmf_mod)
rownames(h) <- paste("Topic", 1:nrow(h), sep = "_")
topic_profile_plts <- SPOTlight::dot_plot_profiles_fun(
  h = h,
  train_cell_clust = nmf_mod_ls[[2]])

topic_profile_plts[[2]] +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90), 
                 axis.text = ggplot2::element_text(size = 12))
```

Look at all cells profiles
```{r fig.height=20, fig.width=20}
topic_profile_plts[[1]] +
  ggplot2::theme(
    axis.text.y = ggplot2::element_blank(),
    axis.text.x = ggplot2::element_blank(),
    axis.title = ggplot2::element_blank())
```

Look at cells topic profile
```{r}
basis_spotlight <- data.frame(NMF::basis(spotlight_ls[[1]][[1]]))

train_labs <- spotlight_ls[[1]][[2]]
colnames(basis_spotlight) <- unique(stringr::str_wrap(train_labs, width = 30))

basis_spotlight[basis_spotlight < 0.0000001] <- 0

DT::datatable(basis_spotlight, filter = "top")
```

#### Cell type location
Look at the location of each cell type in each slice separately
```{r fig.height=25, fig.width=25}
# Iterate over cell types
ct <- colnames(decon_mtrx)

# Iterate over images
lapply(Seurat::Images(sp_obj), function(nm) {
  print(nm)
  nm_donor <- id_sp_df %>% dplyr::filter(gem_id == nm) %>% dplyr::pull(donor_id)
  # Iterate over cell types
  ct_plt_ls <- lapply(ct, function(i) {
    tmp_plt <- Seurat::SpatialFeaturePlot(
      object = sp_obj,
      features = i,
      alpha = c(0, 1),
      images = nm) +
      ggplot2::scale_fill_gradientn(
        colors = heat.colors(10, rev = TRUE)) +
      ggplot2::scale_alpha(range = c(0, 1)) +
      ggplot2::labs(title = stringr::str_wrap(string = i,
                                     width = 25),
           fill = "") +
      ggplot2::theme(
        plot.title = ggplot2::element_text(
          hjust = 0.5,
          size = 20,
          face = "bold"))
    
    return(tmp_plt)
  })
  
  (plt_arr <- cowplot::plot_grid(
    plotlist = ct_plt_ls,
    axis = "trbl",
    align = "hv",
    ncol = 6))
  
  "{plasma}/{plt_dir}/cell_type_location_plasma_{nm_donor}.pdf" %>%
    glue::glue() %>%
    here::here() %>%
    cowplot::save_plot(
      filename = .,
      plot = plt_arr,
      base_height = 25,
      base_width = 25)
  })
```


### Boxplots

Make boxplots for each slice
```{r fig.width=30, fig.height=20}
# Iterate over images
lapply(Seurat::Images(sp_obj), function(nm) {
  nm_donor <- id_sp_df %>% dplyr::filter(gem_id == nm) %>% dplyr::pull(donor_id)
  # Prepare data for boxplots
  metadata_long <- sp_obj@meta.data %>% 
    # tidyr::pivot_longer(cols = c("annotation"),
    #                     names_to = "stratification_id",
    #                     values_to = "stratification_val") %>%
    tidyr::pivot_longer(cols = dplyr::all_of(ct),
                        names_to = "ct_key",
                        values_to = "ct_val") %>%
    # dplyr::left_join(col_df, by = c("ct_key" = "ct_name")) %>%
    dplyr::mutate(ct_val = dplyr::if_else(ct_val > 0.001, ct_val, 0)) %>%
    dplyr::filter(gem_id == nm)
  
  # Box plot of cell type proportion between stratified regions showing the unadjusted ANOVA Pvalue
  
  keep_ct <- metadata_long %>%
    dplyr::group_by(ct_key) %>%
    dplyr::summarise(prop_sum = sum(ct_val)) %>% 
    dplyr::filter(prop_sum > 0) %>%
    dplyr::pull(ct_key)
  
  (bplt <- metadata_long %>%
    dplyr::filter(ct_key %in% keep_ct) %>%
    dplyr::mutate(
      ct_key = stringr::str_wrap(string = ct_key,
                                     width = 30)) %>%
    # dplyr::mutate(stratification_val = factor(stratification_val, 
    #                                           levels = c("Fibrotic", "HER2+/ESR1+",
    #                                                      "HER2+/ESR1-", "HER2-/ESR1+",
    #                                                      "HER2-/ESR1-"))) %>%
    # dplyr::filter(! plt_name %in% c("CD8 EM-like", "CD4 naive-like")) %>%
    ggpubr::ggboxplot(data = .,
                      x = "annotation.general",
                      y = "ct_val",
                      facet.by = "ct_key",
                      color = "annotation.general",
                      fill = "annotation.general",
                      add = "jitter",
                      scales = "free",
                      repel = TRUE,
                      outlier.shape = NA,
                      alpha = 0.6,
                      palette = "Set1",
                      ncol = 5) +
    ggplot2::theme(
      strip.text = ggplot2::element_text(size = 18, face = "bold"),
      axis.text.y = ggplot2::element_text(size = 16),
      axis.title.y = ggplot2::element_text(size = 20),
      # axis.text.x = element_text(size = 12, angle = 90, vjust = 0.5, hjust = 0.5),
      axis.text.x = ggplot2::element_blank(),
      legend.text = ggplot2::element_text(size = 18),
      legend.title = ggplot2::element_blank(),
      strip.background = ggplot2::element_blank()) +
    ggplot2::labs(
      y = "Proportion",
      color = "Regions",
      fill = "Regions"))
  # bplt <- bplt +
  #   ggpubr::stat_compare_means(method = "anova", size = 6) +
  #   ggplot2::scale_y_continuous(
  #     expand = expansion(mult = c(0, 0.1)),
  #     labels = function(x) sprintf("%.2f", x))
  
  "{plasma}/{plt_dir}/strat_bplot_{nm_donor}.pdf" %>%
    glue::glue() %>%
    here::here() %>%
    cowplot::save_plot(
      filename = .,
      plot = bplt,
      base_height = 20,
      base_width = 25)

})
```

### Integrated UMAP
```{r fig.width=12, fig.height=5}
Seurat::DimPlot(
  object = sp_obj,
  group.by = c("annotation", "sample_id"))
```

Next we want to look at the piecharts in the UMAP
```{r fig.width=25, fig.height=20, eval = FALSE}
# Loading libraries
suppressMessages(require(ggplot2))
suppressMessages(require(dplyr))
suppressMessages(require(tibble))
suppressMessages(require(grid))


metadata_ds <- data.frame(sp_obj@meta.data)

colnames(metadata_ds) <- colnames(sp_obj@meta.data)
cell_types_all <- ct
cell_types_interest <- cell_types_all

# If not all cell types are in the cell types of interest we only want to keep those spots which have at least one of the cell types of interest
if (!all(cell_types_all %in% cell_types_interest)) {

  metadata_ds <- metadata_ds %>%
    tibble::rownames_to_column("barcodeID") %>%
    dplyr::mutate(rsum = base::rowSums(.[, cell_types_interest,
                                         drop = FALSE])) %>%
    dplyr::filter(rsum != 0) %>%
    dplyr::select("barcodeID") %>%
    dplyr::left_join(metadata_ds %>% tibble::rownames_to_column("barcodeID"),
                     by = "barcodeID") %>%
    tibble::column_to_rownames("barcodeID")
}

## Preprocess data
umap_coord <- data.frame(sp_obj@reductions$umap@cell.embeddings[, c("UMAP_1", "UMAP_2")]) %>%
  tibble::rownames_to_column("barcodeID") %>%
  dplyr::inner_join(metadata_ds %>% tibble::rownames_to_column("barcodeID"),
                    by = "barcodeID")

## Plot spatial scatterpie plot
(umap_pie_plt <- ggplot2::ggplot() +
    scatterpie::geom_scatterpie(
      data = umap_coord,
      ggplot2::aes(
        x = UMAP_1,
        y = UMAP_2),
      # Fill the pie chart
      cols = cell_types_all,
      pie_scale = 0.15,
      # Remove black outline
      color = NA) +
    ggplot2::scale_fill_manual(
      breaks = col_df$plt_nm,
      values = col_df$color))

"{plasma}/{plt_dir}/UMAP_piehcart_integrated.pdf" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    filename = .,
    plot = umap_pie_plt,
    base_height = 20,
    base_width = 25)
```

### Cell type correlation matrix

We look at the cell-type correlation for all the slices together
```{r}
# se_sub <- subset(merged_se, subset = gem_id == "esvq52_nluss5")
# se_sub
# se_sub@images <- se_sub@images[Seurat::Images(se_sub) == "esvq52_nluss5"]

(cor_mtrx_ct <- SCrafty::correlation_heatmap( 
  se = sp_obj,
  feats = ct,
  assay = "Spatial",
  slot = "data") +
   ggplot2::labs(
     title = "Integrated cell-type correlation matrix"))

"{plasma}/{plt_dir}/cor-mtrx_cell-type_integrated.pdf" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    filename = .,
    plot = cor_mtrx_ct,
    base_height = 9,
    base_width = 10)
```

We look at the cell-type correlation for each slices together
```{r}
# Iterate over images
lapply(Seurat::Images(sp_obj), function(nm) {

  se_sub <- subset(sp_obj, subset = gem_id == nm)
  se_sub@images <- se_sub@images[Seurat::Images(se_sub) == nm]
  
  (cor_mtrx_ct <- SCrafty::correlation_heatmap( 
    se = se_sub,
    feats = ct,
    assay = "Spatial",
    slot = "data") +
     ggplot2::labs(
       title = "{nm} - Integrated cell-type correlation matrix"))
  
  "{plasma}/{plt_dir}/cor-mtrx_cell-type_integrated_{nm}.pdf" %>%
    glue::glue() %>%
    here::here() %>%
    cowplot::save_plot(
      filename = .,
      plot = cor_mtrx_ct,
      base_height = 9,
      base_width = 10)
})
```

## Session Info
```{r}
sessionInfo()
```
