---
author: "Marc Elosua-Bayes"
title: "04-DE analysis integrated dataset"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message = FALSE, warning = FALSE, cache = TRUE)
options(width = 1200)
```

## Introduction
In this script we are going to compute marker genes for the samples at varying resolutions.

## Libraries
```{r}
library(Seurat)
library(ggpubr)
library(cowplot)
library(dplyr)
library(ggplot2)
library(stringr)
library(readr)
```

## Setting parameters
Loading necessary paths and parameters
```{r}
source(here::here("misc/paths.R"))
source(here::here("utils/bin.R"))

set.seed(123)
source(here::here("misc/paths.R"))
source(here::here("utils/bin.R"))

"{anot}/{plt_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(path = .,
             showWarnings = FALSE,
             recursive = TRUE)

"{anot}/{robj_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(path = .,
             showWarnings = FALSE,
             recursive = TRUE)
```

## Load data
The data used in this Rmarkdown document comes from **03-clustering_integration.Rmd** where the data was integrated.
```{r}
merged_se <- "{clust}/{robj_dir}/integrated_spatial.rds" %>%
  glue::glue() %>%
  here::here() %>%
  readRDS(file = .)
```

## Analysis

First we are going to look at the cluster resolution we are going to check
```{r fig.width=5, fig.height=4}
Seurat::DimPlot(merged_se, group.by = "Spatial_snn_res.0.3")
```

```{r fig.width=18, fig.height=8}
res_vec <- c("Spatial_snn_res.0.01", "Spatial_snn_res.0.05", 
             "Spatial_snn_res.0.1", "Spatial_snn_res.0.3",
             "Spatial_snn_res.0.8", "Spatial_snn_res.1",
             "Spatial_snn_res.1.25", "Spatial_snn_res.1.5")
# Iterate over resolutions
lapply(res_vec, function(i) {
  # Iterate over slices
  lapply(id_sp_df$gem_id, function(id) {
    Seurat::SpatialPlot(
      merged_se,
      group.by = i,
      images = id,
      image.alpha = 0,
      crop = FALSE,
      pt.size.factor = 1.25)
  }) %>% 
    patchwork::wrap_plots(., nrow = 2, guides = "collect")
})
```

We are going to use the function `FindAllMarkers` to get the marker genes for all the clusters at the varying resolutions:
```{r eval=FALSE}
# Assign Identities
Seurat::Idents(merged_se) <- merged_se@meta.data[, "Spatial_snn_res.0.3"]

# Compute markers
markers03 <- Seurat::FindAllMarkers(
  object = merged_se,
  assay = "Spatial",
  slot = "data",
  only.pos = TRUE)

DT::datatable(
  data = markers03,
  filter = "top")
```

### Save markers for resolution 0.3
```{r eval=FALSE}
# Assign Identities
Seurat::Idents(merged_se) <- merged_se@meta.data[, "Spatial_snn_res.0.3"]

# Compute markers
markers03 <- Seurat::FindAllMarkers(
  object = merged_se,
  assay = "Spatial",
  slot = "data",
  only.pos = TRUE)

DT::datatable(
  data = markers03,
  filter = "top")
```

```{r eval=FALSE}
out_file03 <- "{anot}/{robj_dir}/markers_Spatial_snn_res.0.3.xlsx" %>%
  glue::glue() %>%
  here::here()

# Remove file to write it from scratch
file.remove(out_file03)

lapply(unique(as.character(markers03$cluster)), function(clust) {
  
  markers03 %>%
    dplyr::filter(cluster == clust) %>%
    dplyr::arrange(dplyr::desc(avg_log2FC)) %>%
    xlsx::write.xlsx(
    .,
    file = out_file03,
    row.names = FALSE,
    sheetName = glue::glue("Cluster-{clust}"),
    append = file.exists(out_file03))
  })
```

### Save markers for resolution 0.8
```{r eval=TRUE}
# Assign Identities
Seurat::Idents(merged_se) <- merged_se@meta.data[, "Spatial_snn_res.0.8"]

# Compute markers
markers08 <- Seurat::FindAllMarkers(
  object = merged_se,
  assay = "Spatial",
  slot = "data",
  only.pos = TRUE)

DT::datatable(
  data = markers08,
  filter = "top")
```

```{r eval=TRUE}
out_file08 <- "{anot}/{robj_dir}/markers_Spatial_snn_res.0.8.xlsx" %>%
  glue::glue() %>%
  here::here()

# Remove file to write it from scratch
file.remove(out_file08)

lapply(unique(as.character(markers08$cluster)), function(clust) {
  
  markers08 %>%
    dplyr::filter(cluster == clust) %>%
    dplyr::arrange(dplyr::desc(avg_log2FC)) %>%
    xlsx::write.xlsx(
    .,
    file = out_file08,
    row.names = FALSE,
    sheetName = glue::glue("Cluster-{clust}"),
    append = file.exists(out_file08))
  })
```

## Compute markers Resolution 1.5
We are going to use the function `FindAllMarkers` to get the marker genes for all the clusters at the varying resolutions:
```{r eval=FALSE}
# Assign Identities
Seurat::Idents(merged_se) <- merged_se@meta.data[, "Spatial_snn_res.1.5"]

# Compute markers
markers_15 <- Seurat::FindAllMarkers(
  object = merged_se,
  assay = "Spatial",
  slot = "data",
  only.pos = TRUE)

DT::datatable(
  data = markers_15,
  filter = "top")
```

### Save markers for resolution 1.5
```{r eval=FALSE}
out_file_15 <- "{anot}/{robj_dir}/markers_Spatial_snn_res.1.5.xlsx" %>%
  glue::glue() %>%
  here::here()

# Remove file to write it from scratch
file.remove(out_file_15)

lapply(unique(as.character(markers_15$cluster)), function(clust) {
  markers_15 %>%
    dplyr::filter(cluster == clust) %>%
    dplyr::arrange(dplyr::desc(avg_log2FC)) %>%
    xlsx::write.xlsx(
    .,
    file = out_file_15,
    row.names = FALSE,
    sheetName = glue::glue("Cluster-{clust}"),
    append = file.exists(out_file_15))
  })
```

### Clsutering Annotation
```{r}
merged_se[["annotation-general"]] <- dplyr::case_when(
  merged_se@meta.data$Spatial_snn_res.0.3 == 0 ~ "T cell zone",
  merged_se@meta.data$Spatial_snn_res.0.3 == 1 ~ "T:B Border",
  merged_se@meta.data$Spatial_snn_res.0.3 == 2 ~ "Epithelial-1",
  merged_se@meta.data$Spatial_snn_res.0.3 == 3 ~ "Inter-Follicular Zone",
  merged_se@meta.data$Spatial_snn_res.0.3 == 4 ~ "Proliferating Follicle",
  merged_se@meta.data$Spatial_snn_res.0.3 == 5 ~ "Germinal Center",
  merged_se@meta.data$Spatial_snn_res.0.3 == 6 ~ "Subepithelial",
  merged_se@meta.data$Spatial_snn_res.0.3 == 7 ~ "Epithelial-2",
  merged_se@meta.data$Spatial_snn_res.0.3 == 8 ~ "Muscle"
)
```

### Histological Annotation

Furthermore, we annotate these slides following a pathologists histological annotation.

#### HA Marco
```{r}
# load annotations
ann_dir <- "{anot}/{robj_dir}/annotations-histological" %>%
  glue::glue() %>%
  here::here()

fn_ls <- list.files(ann_dir, recursive = TRUE, full.names = TRUE)
# Remove Gerard's annotation file
fn_ls <- fn_ls[str_detect(fn_ls, "gerard", negate = TRUE)]

# Read in data
hist_df_marco <- lapply(fn_ls, function(n) {
  n_mt <- str_split(n, "/", simplify = TRUE)
  slide_dir <- n_mt[1, ncol(n_mt) - 2]
  id <- substr(slide_dir, 1, 13)  # Extract first 13 characters
  region <- n_mt[1, ncol(n_mt) - 1]
  
  
  bc <- readr::read_csv(file = n) %>% dplyr::pull("barcode")
    
  data.frame(
    "barcodes" = bc,
    "histological_annotation_marco" = region,
    "gem_id" = id)
  }) %>% dplyr::bind_rows()

# Make sure there are no duplicate rows
hist_df_marco <- dplyr::distinct(hist_df_marco)

# Make sure one spot is not assigned to 2 regions
hist_df_marco %>%
  dplyr::count(barcodes) %>%
  dplyr::filter(n > 1)
```

Quick reprocessing of the names
```{r}
hist_df_marco <- hist_df_marco %>%
    mutate(
        histological_annotation_marco = case_when(
            histological_annotation_marco == "mesenchymal areas" ~ "Mesenchymal Areas",
            histological_annotation_marco == "mantle zones" ~ "Mantle Zone",
            histological_annotation_marco == "vascular spaces" ~ "Vascular Spaces",
            TRUE ~ histological_annotation_marco
        )
    )

unique(hist_df_marco$histological_annotation_marco)
```

#### HA Gerard
```{r}
library(openxlsx)
an_g <- "{anot}/{robj_dir}/annotations-histological/Tonsil_Annotation_gerard.xlsx" %>%
    glue::glue() %>%
    here::here() %>%
    openxlsx::read.xlsx(xlsxFile = ., sheet = 1)

cnames <- colnames(an_g)
rr <- which(cnames %in% c("LIGHT.ZONE", "DARK.ZONE", "INTERFOLLICULAR.T-CELL.ZONE"))
hist_df_gerard <- lapply(cnames[-rr], function(i) {
    print(i)
    bc_g <- an_g[, i, drop = FALSE] %>%
        tidyr::drop_na() %>%
        pull(1) %>%
        gsub(pattern = '"', replacement = '', x = .) %>%
        str_split(., ",") %>%
        purrr::map(~ .[5]) %>%
        unlist()
    
    data.frame(
    "barcodes" = bc_g[-1],
    "histological_annotation_gerard" = i,
    "gem_id" = str_sub(bc_g[-1], 1, 13))
    }) %>%
    bind_rows()

# Make sure there are no duplicate rows
hist_df_gerard <- dplyr::distinct(hist_df_gerard)

# Make sure one spot is not assigned to 2 regions
hist_df_gerard %>%
  dplyr::count(barcodes) %>%
  dplyr::filter(n > 1)

# There are some duplicated spots
bc_dup <- hist_df_gerard %>%
    dplyr::count(barcodes) %>%
    dplyr::filter(n > 1) %>%
    dplyr::pull(barcodes)

# We are going to manually annotate them
dup_ls <- hist_df_gerard[hist_df_gerard$barcodes %in% bc_dup, ] %>%
    arrange(barcodes) %>%
    group_split(barcodes)

lapply(seq_len(length(dup_ls)), function(i) {
    tmp <- dup_ls[[i]]
    if ("MANTLE.ZONE" %in% tmp$histological_annotation_gerard) {
        tmp %>%
            filter(histological_annotation_gerard == "MANTLE.ZONE")
    } else if ("FOLDED.AREA.(NOT.EVALUABLE" %in% tmp$histological_annotation_gerard) {
        tmp %>%
            filter(histological_annotation_gerard == "FOLDED.AREA.(NOT.EVALUABLE")
    }
})

dup_bc_vec <- dup_ls %>% bind_rows() %>% pull(barcodes) %>% unique()
hist_df_gerard <- hist_df_gerard %>%
    mutate(
        keep = case_when(
            # Unice case for barcode esvq52_nluss5_ATGGAACCTTTGCACA-1
            barcodes == "esvq52_nluss5_ATGGAACCTTTGCACA-1" & histological_annotation_gerard == "FOLDED.AREA.(NOT.EVALUABLE" ~ TRUE,
            barcodes == "esvq52_nluss5_ATGGAACCTTTGCACA-1" & histological_annotation_gerard == "GERMINAL.CENTER" ~ FALSE,
            ! barcodes %in% dup_bc_vec ~ TRUE,
            barcodes %in% dup_bc_vec & histological_annotation_gerard == "FOLDED.AREA.(NOT.EVALUABLE" ~ TRUE,
            barcodes %in% dup_bc_vec & histological_annotation_gerard == "GERMINAL.CENTER" ~ TRUE,
            TRUE ~ FALSE)
    ) %>%
    dplyr::filter(keep)

# Make sure one spot is not assigned to 2 regions
hist_df_gerard %>%
  dplyr::count(barcodes) %>%
  dplyr::filter(n > 1)

# Check unique  case
hist_df_gerard %>% filter(barcodes == "esvq52_nluss5_ATGGAACCTTTGCACA-1")
```

Quick reprocessing of the names
```{r}
hist_df_gerard <- hist_df_gerard %>%
    mutate(
        histological_annotation_gerard = case_when(
            histological_annotation_gerard == "SURFACE.EPITHELIUM" ~ "Surface Epithelium",
            histological_annotation_gerard == "CRYPT.EPITHELIUM" ~ "Crypt Epithelium",
            histological_annotation_gerard == "FOLDED.AREA.(NOT.EVALUABLE" ~ "Folded",
            histological_annotation_gerard == "GERMINAL.CENTER" ~ "Germinal Center",
            histological_annotation_gerard == "MANTLE.ZONE" ~ "Mantle Zone",
            histological_annotation_gerard == "MESENCHIMA./.STROMA" ~ "Mesenchymal/Stromal",
            histological_annotation_gerard == "VASCULAR" ~ "Vascular",
            TRUE ~ "other"
        )
    )
```

#### Add histological annotation
Add histological annotation to seurat object
```{r}
merged_se@meta.data <- merged_se@meta.data %>%
    tibble::rownames_to_column("barcodes") %>%
    dplyr::left_join(hist_df_marco, by = c("barcodes", "gem_id")) %>%
    dplyr::left_join(hist_df_gerard, by = c("barcodes", "gem_id")) %>%
    dplyr::mutate(
    histological_annotation_marco = dplyr::case_when(
      !is.na(histological_annotation_marco) ~ histological_annotation_marco,
      Spatial_snn_res.0.1 == 0 ~ "Inter-Follicular Zone",
      Spatial_snn_res.0.1 == 1 ~ "Germinal Center",
      Spatial_snn_res.0.1 == 2 ~ "Epithelium",
      TRUE ~ "unannotated"),
    histological_annotation_gerard = dplyr::case_when(
      str_detect(histological_annotation_gerard, "Epithelium") ~ "Epithelium",
      !is.na(histological_annotation_gerard) ~ histological_annotation_gerard,
      Spatial_snn_res.0.1 == 0 ~ "Inter-Follicular Zone",
      TRUE ~ "unannotated")
        ) %>%
    tibble::column_to_rownames("barcodes") %>%
    dplyr::select(-keep)


table(merged_se@meta.data$histological_annotation_gerard)
table(merged_se@meta.data$histological_annotation_marco)
```

Visualize the annotation
```{r fig.width=15, fig.height=7}
library(colorBlindness)

palette_marco<- palette <- c(
  "Inter-Follicular Zone" = "#ff6db6",
  "Epithelium" = "#924900",
  "Mesenchymal Areas" = "#db6d00",
  "Germinal Center" = "#004949",
  "Mantle Zone" = "#00e8e8",
  "Vascular Spaces" = "#490092",
  "unannotated" = "lightgrey"
  )

palette_gerard <- c(
  "Inter-Follicular Zone" = "#ff6db6",
  "Epithelium" = "#924900",
  "Mesenchymal/Stromal" = "#db6d00",
  "Germinal Center" = "#004949",
  "Mantle Zone" = "#00e8e8",
  "Vascular" = "#490092",
  "unannotated" = "lightgrey",
  "Folded" = "#1a1a1a"
  )

pn <- SpatialPlot(
    object = merged_se,
    group.by = "histological_annotation_marco",
    images = "esvq52_nluss5",
    crop = FALSE,
    pt.size.factor = 1.25) &
    scale_fill_manual(values = palette_marco) &
    theme(legend.position = "top") &
    guides(fill= guide_legend(override.aes = list(size = 5)))

pg <- SpatialPlot(
    object = merged_se,
    group.by = "histological_annotation_gerard",
    images = "esvq52_nluss5",
    crop = FALSE,
    pt.size.factor = 1.25) &
    scale_fill_manual(values = palette_gerard) &
    theme(legend.position = "top") &
    guides(fill = guide_legend(override.aes = list(size = 5)))

pn | pg

SpatialPlot(
    object = merged_se,
    group.by = "Spatial_snn_res.1.5",
    images = c("esvq52_nluss5"),
    crop = FALSE,
    pt.size.factor = 1.25)
```

### Check how Marco & Gerard correspond
```{r fig.width=6, fig.height=6}
merged_se@meta.data %>%
    dplyr::filter(gem_id == "esvq52_nluss5") %>% 
    data.frame() %>%
    dplyr::count(histological_annotation_marco, histological_annotation_gerard) %>%
    group_by(histological_annotation_marco) %>%
    mutate(freq = n / sum(n)) %>%
    ungroup() %>%
    # Only show those populations that have more than 5 cells
    filter(n >= 5) %>%
    ggplot() +
        geom_point(aes(
            x = histological_annotation_marco,
            y = histological_annotation_gerard,
            color = n,
            size = freq)) +
        scale_color_viridis_c() +
        labs(
            title = "Slide esvq52_nluss5 annotation overlap",
            subtitle = "Proportion of Gerard's histological annotation by within Marco's annotation") +
        theme_minimal() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


## Save object
```{r}
"{anot}/{robj_dir}/integrated_spatial_annot.rds" %>%
  glue::glue() %>%
  here::here() %>%
  saveRDS(object = merged_se, file = .)
```

## Session Info
```{r}
sessionInfo()
```

