---
title: 'Exclude Doublets'
author: "Ramon Massoni-Badosa"
date: "1/28/2021"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```


# Introduction

The objectives of this notebook are:

1. Include the scores computed in the previous notebook to the metadata of the Seurat object.
2. Project such scores and annotations in the UMAP to get a sense of where our predicted doublets are located.
3. Exclude the doublets and save the object.


# Pre-processing

## Load packages

```{r}
library(AnnotationDbi)
library(org.Hs.eg.db)
library(Seurat)
library(Matrix)
library(tidyverse)
```


## Parameters

```{r}
# Paths
path_to_obj <- here::here("scRNA-seq/results/R_objects/seurat_integrated_with_doublets.rds")
path_to_doubl_annot <- here::here("scRNA-seq/2-QC/5-batch_effect_correction/1-doublet_exclusion/tmp/doublet_final_annotations_and_pDNN.rds")
path_to_save <- here::here("scRNA-seq/results/R_objects/seurat_without_doublets_missing_reintegration.rds")


# Functions
source(here::here("scRNA-seq/bin/utils.R"))
```


## Load data

```{r}
tonsil <- readRDS(path_to_obj)
doublet_annot <- readRDS(path_to_doubl_annot)
```


# Add metadata

```{r}
if (all(rownames(doublet_annot) == colnames(tonsil))) {
  tonsil$pDNN_hashing <- doublet_annot$pDNN_hashing
  tonsil$pDNN_scrublet <- doublet_annot$pDNN_scrublet
  tonsil$pDNN_union <- doublet_annot$pDNN_union
} else {
  warning("barcodes are not equal")
}
```


# Visualize

## Proportion of doublet nearest neighbors (pDNN)

```{r fig.wide=TRUE}
pDNN_vars <- c("pDNN_hashing", "pDNN_scrublet", "pDNN_union")
pDNN_ggs <- purrr::map(pDNN_vars, function(x) {
  p <- feature_plot_doublets(seurat_obj = tonsil, feature = x)
  p
})
names(pDNN_ggs) <- pDNN_vars
pDNN_ggs
```


## Doublet annotations

```{r fig.wide=TRUE}
doublet_vars <- c("HTO_classification.global", "scrublet_predicted_doublet",
                  "has_high_lib_size")
doublet_ggs <- purrr::map(doublet_vars, function(x) {
  p <- DimPlot(tonsil, group.by = x, pt.size = 0.3) +
    ggtitle(x) +
    theme(
      plot.title = element_text(size = 14, hjust = 0.5),
      axis.title = element_text(size = 13),
      axis.text = element_text(size = 11),
      legend.text = element_text(size = 11)
    )
  p
})
names(doublet_ggs) <- doublet_vars
doublet_ggs


# Zoom in for cell hashing
umap_hashing <- tonsil@reductions$umap@cell.embeddings %>%
  as.data.frame() %>%
  mutate(HTO_classification.global = tonsil$HTO_classification.global) %>%
  ggplot(aes(UMAP_1, UMAP_2, color = HTO_classification.global)) +
    geom_point(size = 0.01, alpha = 0.5) +
    facet_wrap(~HTO_classification.global) +
    theme_classic()
umap_hashing
```


## Canonical markers

```{r}
canonical_markers <- c("CD79A", "CD79B", "CD3D", "CD3E", "NKG7", "LYZ", "FDCSP")
canonical_markers_gg <- purrr::map(canonical_markers, function(x) {
  p <- feature_plot_doublets(seurat_obj = tonsil, feature = x)
  p
})
names(canonical_markers_gg) <- canonical_markers
canonical_markers_gg
```


## pre-Plasmablast and tingible-body macrophages

We want to be particularly careful with two cell types: pre-plasmablasts and [tingible-body macrophages](https://en.wikipedia.org/wiki/Tingible_body_macrophage), since they might be confused by doublets (specially using scrublet). Thus, we will visualize a [prePB](https://immunology.sciencemag.org/content/6/56/eabe6291) and apoptosis signature:

```{r fig.wide=TRUE}
# Define signatures
apoptosis_signature <- AnnotationDbi::select(
  org.Hs.eg.db,
  keys = "GO:0006915",
  keytype = "GO",
  columns = "SYMBOL"
)$SYMBOL
apoptosis_signature <- unique(apoptosis_signature)
prePB_signature <- c("RASSF6", "FRZB", "HOPX", "BRNL9", "FGFR1")
signatures_list <- list(
  prePB_signature = prePB_signature,
  apoptosis_signature = apoptosis_signature
)


# Compute score
tonsil <- AddModuleScore(
  tonsil,
  features = signatures_list,
  name = c("prePB_signature", "apoptosis_signature")
)


# Visualize
prePB_gg <- feature_plot_doublets(
  seurat_obj = tonsil,
  feature = "prePB_signature1"
)
apopotosis_gg <- feature_plot_doublets(
  seurat_obj = tonsil,
  feature = "apoptosis_signature2"
)
prePB_gg
apopotosis_gg
```


# Exclude doublets

```{r}
tonsil
table(tonsil$HTO_classification.global)
tonsil <- subset(tonsil, subset = HTO_classification.global != "Doublet")
tonsil
```


# Save

```{r}
saveRDS(tonsil, path_to_save)
```


# Session Information

```{r}
sessionInfo()
```
