---
title: "Motif analysis JASPAR"
author: "Paula Soler-Vila"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, out.width = "100%", fig.align = "center",
  message = FALSE, warning = FALSE
)
options(width = 1200)
```

# Introduction
The transcription factors (TF) and DNA recognitions depend on multiples levels of interactions. The first level involves chromatin accessibility, where nuclesosome-depleted regions are highly associated with TFs binding compared to the closed chromatin, which is often inaccesible to most TFs. The second represents the existence of the consensus binding motif in the DNA sequence, a point that we are going to identify in this notebook in each cell type using Signac and ChromVar.

# Pre-processing

## Load packages

```{r libraries} 
library(Seurat)
library(Signac)
library(reshape)
library(ggplot2)
library(JASPAR2020)
library(TFBSTools)
library(BSgenome.Hsapiens.UCSC.hg38)
library(patchwork)
library(chromVAR)
library(motifmatchr)
library(ggpubr)
library(data.table)
library(chromVARmotifs)
library(dplyr)
library(purrr)
library(readxl)
library(writexl)
library(pheatmap)
library(factoextra)
library(corrplot)

set.seed(1234)
```

## Parameters
```{r parameters}
cell_type = "GCBC"

# Paths
path_to_obj <- paste0(
  here::here("scATAC-seq/results/R_objects/level_4/"),
  cell_type,
  "/",
  cell_type,
  "_integration_peak_calling_level_4.rds",
  sep = ""
)


color_palette <-  c("#1CFFCE", "#90AD1C", "#C075A6", "#85660D", 
                    "#5A5156", "#AA0DFE", "#F8A19F", "#F7E1A0",
                    "#1C8356", "#FEAF16", "#822E1C", "#C4451C", 
                    "#1CBE4F", "#325A9B", "#F6222E", "#FE00FA",
                    "#FBE426", "#16FF32",  "black",  "#3283FE",
                    "#B00068", "#DEA0FD", "#B10DA1", "#E4E1E3", 
                    "#90AD1C", "#FE00FA", "#85660D", "#3B00FB", 
                    "#822E1C", "coral2",  "#1CFFCE", "#1CBE4F", 
                    "#3283FE", "#FBE426", "#F7E1A0", "#325A9B", 
                    "#2ED9FF", "#B5EFB5", "#5A5156", "#DEA0FD",
                    "#FEAF16", "#683B79", "#B10DA1", "#1C7F93", 
                    "#F8A19F", "dark orange", "#FEAF16", 
                    "#FBE426", "Brown")

path_to_save <- paste0(
  here::here("scATAC-seq/results/R_objects/level_4/"),
  cell_type,
  "/",
  cell_type,
  "_chromVar_JASPAR_level_4.rds",
  sep = ""
)

path_to_save_TF_motifs <- paste0(
  here::here("scATAC-seq/results/files/"),
  cell_type,
  "/",
  cell_type,
  "_chromVar_JASPAR_level_4.xlsx",
  sep = ""
)
```

## Functions
```{r}
remove_correlated_helper <- function(mat, val, cutoff = 0.9) {
  stopifnot(nrow(mat) == length(val))
  cormat <- cor(t(mat), use = "pairwise.complete.obs")
  diag(cormat) <- NA
  keep <- seq_len(nrow(mat))
  for (i in order(val, decreasing = TRUE)) {
    if (i %in% keep) {
      toremove <- which(cormat[keep, i] >= cutoff)
      if (length(toremove) > 0) 
        keep <- keep[-toremove]
    }
  }
  return(keep)
}
```

# Load data
```{r}
seurat <- readRDS(path_to_obj)
seurat

DimPlot(
  seurat, 
  group.by = "level_5",
  cols = color_palette,
  pt.size = 0.1
)

DimPlot(
  seurat, 
  split.by = "level_5",
  cols = color_palette,
  pt.size = 0.1,
  ncol = 5
) + NoLegend()
```

# Adding motif information to the Seurat object

## Retrieving matrices from JASPAR2020 database
JASPAR is a collection of transcription factor DNA-binding preferences, modeled as matrices. To have a detaill explanation of it, visit the following link, http://jaspar.genereg.net/about/.

```{r}
opts <- list()
opts[["tax_group"]] <- "vertebrates"
pfm <- getMatrixSet(JASPAR2020, opts)
length(pfm)
```

## Retrieving matrices from cisBP database
Curated collection of human motifs from cisBP database

```{r}
data("human_pwms_v1")
length(human_pwms_v1)

human_pwms_v1[grep(pattern = "FOXO",x = name(human_pwms_v1))]
```


## Add motif information

```{r}
seurat <- AddMotifs(
  object = seurat,
  genome = BSgenome.Hsapiens.UCSC.hg38,
  pfm = pfm
)

seurat[["peaks_redefined"]]
```


# Computing motif activities
The TF motif enrichments (that help us to predict potential specific cell-type regulators) previously computed are not calculated per-cell and they do not take into account the insertion sequence bias of the Tn5 transposase. To account for these issues we can use 
chromVAR that computes for each motif annotation and each cell a bias corrected "desviation" in accessibility from a expected accessibility based on the average of all the cells. This allows us to visualize motif activities per cell, and also provides an alternative method of identifying differentially-active motifs between cell types. 

```{r}
# The RunChromVAR function retrieved the deviationScores, the Z-scores for each bias corrected deviations.
seurat <- RunChromVAR(
  object = seurat,
  genome = BSgenome.Hsapiens.UCSC.hg38
)

saveRDS(seurat, path_to_save)
```

## Averaging the Z-Score by all the cell that conform a cell-type group.
```{r}
avgexpr_mat <- AverageExpression(
  seurat,
  assays = "chromvar",
  return.seurat = F,
  group.by = "ident",
  slot = "data")
```

```{r}
res.pca <- prcomp(t(avgexpr_mat$chromvar),scale. = T)

options(repr.plot.width=6, repr.plot.height=8)
fviz_pca_ind(res.pca,
             repel = TRUE)
```

```{r}
pheatmap(avgexpr_mat$chromvar, scale = "row",
         border_color = "black",
         cluster_rows = T,
         cluster_cols = T,
         fontsize_row= 3,
         clustering_distance_rows = "euclidean",
         clustering_distance_cols = "euclidean", 
         clustering_method = "ward.D2",
         cutree_rows = NA, 
         cutree_cols = 2)
```

## Averaging the Z-Score & Correction proposed by chromVar
We compute the standar desviation for each annotated motif. Then, we select motifs that have a standar desviation higher than a specific theshold and use them to perform a correlation test and a principal component analysis. Note, that we are going to use the function "remove_correlated_helper", to eliminate the variables that present a correlation greater than 0.9.

```{r}
seurat_average <- AverageExpression(
  seurat,
  assays = "chromvar",
  return.seurat = T,
  group.by = "ident")


threshold = 1
matrix <- seurat_average[["chromvar"]]@data
vars <- matrixStats::rowSds(matrix, na_rm = TRUE)
boxplot(vars)
abline(h=1)

ix <- which(vars >= threshold)
ix2 <- ix[remove_correlated_helper(matrix[ix, , drop = FALSE], 
                                   vars[ix])]

cormat <- cor(matrix[ix2,], 
              use = "pairwise.complete.obs")

corrplot(cormat, type="upper", order="hclust", tl.col="black", tl.srt=45)

pc_res <- prcomp(t(matrix[ix2, ]))
fviz_pca_ind(pc_res, repel = TRUE)
```

# Subclustering of the re-entry compartment using scATAC data.

```{r}
DefaultAssay(seurat) <- 'peaks_redefined'

seurat <- FindNeighbors(object = seurat, reduction = 'lsi', dims = 2:30)
seurat <- FindClusters(object = seurat, 
                       verbose = FALSE,
                       algorithm = 3, 
                       resolution = 0.3)

p1 <- DimPlot(object = seurat, label = F)

seurat <- FindSubCluster(object = seurat, 
                         graph.name = "peaks_redefined_snn",
                         subcluster.name = "re_entry_7",
                         cluster = "7",
                         resolution = 0.15)

p2 <- DimPlot(object = seurat, label = F, group.by = "re_entry_7")


p1 | p2

```


## Performing differential motif activity analysis. 

```{r}
Idents(seurat) <- seurat$re_entry_7
DefaultAssay(seurat) <- 'chromvar'

da_regions_7_0 <- FindMarkers(
 ident.1 = "7_0",
 object = seurat,
 only.pos = TRUE,
 min.pct = 0.1,
 test.use = 'LR',
 latent.vars = 'nCount_peaks_redefined'
)

da_regions_7_1 <- FindMarkers(
 ident.1 = "7_1",
 object = seurat,
 only.pos = TRUE,
 min.pct = 0.1,
 test.use = 'LR',
 latent.vars = 'nCount_peaks_redefined'
)

da_regions_7_0_1 <- FindMarkers(
 ident.1 = "7_0",
 ident.2 = "7_1",
 object = seurat,
 only.pos = TRUE,
 min.pct = 0.1,
 test.use = 'LR',
 latent.vars = 'nCount_peaks_redefined'
)

da_regions = da_regions_7_0
  
motif_name <- sapply(row.names(da_regions), function(x) {name(getMatrixByID(JASPAR2020, ID = x))})
da_regions$motif_name <- motif_name

families <- sapply(row.names(da_regions), function(x) {tags(getMatrixByID(JASPAR2020, ID = x))$family})
da_regions$family <- families
da_regions$family <- gsub('\\s+', '',da_regions$family)

DT::datatable(da_regions)

da_regions_selected <- (da_regions[da_regions$p_val_adj < 0.005 & 
                                     da_regions$avg_log2FC > 0.5, ])
```


```{r}
da_regions_selected_sorted <- da_regions_selected[order(-da_regions_selected$avg_log2FC),]
```


### Plotting UMAP TF motif activity
```{r}
FeaturePlot(
  object = seurat,
  features = row.names(da_regions_selected_sorted)[1:4],
  min.cutoff = 'q5',
  max.cutoff = 'q95',
  pt.size = 0.1,
  ncol = 2
)
```


```{r}
FeaturePlot(
  object = seurat,
  features = c("MA1954.1"),
  min.cutoff = 'q5',
  max.cutoff = 'q95',
  pt.size = 0.1,
  ncol = 2
)

```


# Session info
```{r}
sessionInfo()
```

