---
author: "Marc Elosua-Bayes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    toc: yes
    toc_float: yes
    df_print: paged
title: "Figure 3 - Spatial Panels"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message = FALSE, warning = FALSE)
options(width = 1200)
```

## Introduction
In this Rmarkdown document we are going to create the spatial pannels necessary for Figure 3.
This Figure contains data regarding Epithelial and non-lymphoid tissue.

## Libraries

```{r}
library(Seurat)
library(ggpubr)
library(cowplot)
library(dplyr)
library(ggplot2)
library(SPATA2)
```

## Setting parameters
Loading necessary paths and parameters
```{r}
set.seed(123)
source(here::here("misc/paths.R"))
source(here::here("utils/bin.R"))

"{fig3}/{plt_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(path = .,
             showWarnings = FALSE,
             recursive = TRUE)

"{fig3}/{robj_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(path = .,
             showWarnings = FALSE,
             recursive = TRUE)

clust_vr <- "annotation_figure_1"
```

## Load data
The data used in this Rmarkdown document comes from **04-annotation/04-DE_markers_integrated** where the data was integrated.
```{r}
sp <- "{anot}/{robj_dir}/integrated_spatial_annot.rds" %>%
  glue::glue() %>%
  here::here() %>%
  readRDS(file = .)
```

Load deconvolution from **spatial_transcriptomics/epithelium_integration/epithelium-deconvolution.Rmd**
```{r}
# Load SPOTlight data
spotlight_ls <- "{epithelium}/{robj_dir}/spotlight_ls_epithelium.rds" %>%
  glue::glue() %>% 
  here::here() %>%
  readRDS(file = .)
```

Set correspondance dataframe for the cell types
```{r}
nm_df <- "{epithelium}/{robj_dir}/epithelium_nm_df.rds" %>%
  glue::glue() %>% 
  here::here() %>%
  readRDS(file = .)
```

Add colors to cell types
```{r}
library(RColorBrewer)
# Define the number of colors you want
nb.cols <- nrow(nm_df)
mycolors <- colorRampPalette(RColorBrewer::brewer.pal(8, "Set2"))(nb.cols)
# https://sashamaps.net/docs/resources/20-colors/
color <- c('#e6194B', '#3cb44b', '#ffe119', '#4363d8', '#f58231', '#911eb4',
           '#42d4f4', '#f032e6', '#bfef45', '#fabed4', '#469990', '#dcbeff',
           '#9A6324', '#fffac8', '#800000', '#aaffc3', '#808000', '#ffd8b1',
           '#000075', '#a9a9a9', '#ffffff', '#000000', '#b00b69')

col_df <- dplyr::bind_cols(nm_df, color = color[1:nrow(nm_df)])
```

## Analysis

### Preprocess data
```{r}
decon_mtrx <- spotlight_ls[[2]]
decon_mtrx <- decon_mtrx[, colnames(decon_mtrx) != "res_ss"]

# Set as 0 cell types predicted to be under 3 % of the spot
decon_mtrx[decon_mtrx < 0.03] <- 0
```

Change column names
```{r}
new_cn <- data.frame(mod_nm = colnames(decon_mtrx)) %>%
  dplyr::left_join(nm_df, by = "mod_nm") %>%
  # Central.Mem.PASK. fives some trouble because it only changes between + an -
  # negative goes first and distinct solves it automatically
  dplyr::distinct() %>%
  dplyr::pull(plt_nm)

colnames(decon_mtrx) <- new_cn
```

We are going to add the deconvolution to the Seurat object.
```{r}
sp@meta.data <- cbind(sp@meta.data, decon_mtrx)
```

### Panel
Location of cell types of interest
```{r}
ct_int <- c("Surface epithelium", "Basal cells", "Crypt")

# Decimal selectino for legend
scaleFUN <- function(x) sprintf("%.2f", x)

nm <- "tarwe1_xott6q"

nm_donor <- id_sp_df %>%
  dplyr::filter(gem_id == nm) %>%
  dplyr::pull(donor_id)
  
# Subset seurat object
se_sub <- subset(sp, subset = gem_id == nm)
se_sub@images <- se_sub@images[Seurat::Images(se_sub) == nm]

"{fig3}/{plt_dir}/Spatial-Epithelial-deconv-object.rds" %>%
  glue::glue() %>%
  here::here() %>%
  saveRDS(
    object = se_sub,
    file = .)

# First make a plot with just the image
tmp_plt <- Seurat::SpatialPlot(
    object = se_sub,
    features = "nCount_Spatial",
    alpha = c(0, 0),
    images = nm,
    image.alpha = 1) +
  Seurat::NoLegend()
  
# Iterate over cell types
ct_plt_ls <- lapply(ct_int, function(i) {
  tmp_plt <- Seurat::SpatialPlot(
    object = se_sub,
    features = i,
    alpha = c(0, 1),
    images = nm,
    image.alpha = 0) +
    ggplot2::scale_fill_gradientn(
      colors = grDevices::heat.colors(10, rev = TRUE),
      # Same number of breaks for all plots
      breaks = seq(min(se_sub@meta.data[, i]),
                   max(se_sub@meta.data[, i]),
                   length.out = 4),
      # 2 decimals in the legend
      labels = scaleFUN) +
    ggplot2::scale_alpha(range = c(0, 1)) +
    ggplot2::labs(
      title = stringr::str_wrap(string = i,
                                width = 25),
      fill = "") +
    ggplot2::theme(
      plot.title = ggplot2::element_text(
        hjust = 0.5,
        size = 20,
        face = "bold"))
  
  return(tmp_plt)
})

ct_plt_ls[[4]] <- tmp_plt

(plt_arr <- cowplot::plot_grid(
  plotlist = ct_plt_ls,
  axis = "trbl",
  align = "hv",
  ncol = 2))

"{fig3}/{plt_dir}/Spatial-Epithelial-deconv.pdf" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    filename = .,
    plot = plt_arr,
    base_height = 12,
    base_width = 10)

"{fig3}/{plt_dir}/Spatial-Epithelial-deconv.rds" %>%
  glue::glue() %>%
  here::here() %>%
  saveRDS(
    object = plt_arr,
    file = .)
```

## Session Info
```{r}
sessionInfo()
```
