---
title: 'NBC/MBC: Label and coordinate transfer (validation cohort)'
author: "Ramon Massoni-Badosa"
date: "`r Sys.Date()`"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```


# Introduction

Here, we will transfer the label and UMAP coordinates from the discovery cohort to the validation cohort. We will do it one compartment at a time (e.g. nbc_mbc), so we can validate that the transfer is robust

Specifically, we will follow these steps for every compartment:

1. Include or update the annotation label for this object.
2. Plot the integrated object split by type (quey/reference) to visualize if discovery and validation cohorts were integrated properly.
3. Transfer label.
4. Transfer UMAP coordinates + visualize
5. Exclude cells with a low annotation probability and/or a high doublet score
6. Visualize heatmap of specific markers to validate that the cluster and annotation holds for the new cohort.

# Pre-processing

## Load packages

```{r}
library(Seurat)
library(scCustomize)
library(tidyverse)
library(caret)
library(class)
library(harmony)
library(here)
library(glue)
library(DT)
library(pheatmap)
```


## Source functions

```{r}
source(here("scRNA-seq/bin/SLOcatoR_functions.R"))
source(here("scRNA-seq/bin/utils.R"))
source(here("scRNA-seq/bin/utils_final_clusters.R"))
color_palette <-  c("#1CFFCE", "#90AD1C", "#C075A6", "#85660D", "#5A5156", "#AA0DFE",
                    "#F8A19F", "#F7E1A0", "#1C8356", "#FEAF16", "#822E1C", "#C4451C",
                    "#1CBE4F", "#325A9B", "#F6222E", "#FE00FA", "#FBE426", "#16FF32",
                    "black",   "#3283FE", "#B00068", "#DEA0FD", "#B10DA1", "#E4E1E3",
                    "#90AD1C", "#FE00FA", "#85660D", "#3B00FB", "#822E1C", "coral2",
                    "#1CFFCE", "#1CBE4F", "#3283FE", "#FBE426", "#F7E1A0", "#325A9B",
                    "#2ED9FF", "#B5EFB5", "#5A5156", "#DEA0FD", "#FEAF16", "#683B79",
                    "#B10DA1", "#1C7F93", "#F8A19F", "dark orange", "#FEAF16", "#FBE426",
                    "Brown")
```


# NBC/MBC

```{r}
# Read data, clean and include annotation
nbc_mbc <- readRDS(here("scRNA-seq/results/R_objects/seurat_objects_revision/4.10-NBC_MBC_seurat_object_discovery_validation_cohorts.rds"))
load(here("scRNA-seq/7-analysis_additional_samples/tmp/NBC_MBC_annotation_df.rda"))
```


We proceed to update the annotation:

```{r}
nbc_mbc$annotation_20220619 <- NA
nbc_mbc$annotation_20220619[NBC_MBC_annotation_df$barcode] <- NBC_MBC_annotation_df$annotation_20220619
nbc_mbc$annotation_20230508 <- nbc_mbc$annotation_20220619
nbc_mbc$annotation_20230508[is.na(nbc_mbc$annotation_20230508)] <- "unannotated"
nbc_mbc$annotation_20230508 <- factor(
  nbc_mbc$annotation_20230508,
  levels = c(names(colors_20230508[["NBC_MBC"]]), "unannotated")
)
Idents(nbc_mbc) <- "annotation_20230508"
nbc_mbc$is_reference <- ifelse(
  nbc_mbc$cohort_type == "discovery",
  "reference",
  "query"
)
DimPlot(
  nbc_mbc,
  group.by = "annotation_20230508",
  split.by = "cohort_type",
  cols = colors_20230508[["NBC_MBC"]]
)
```

Let us explore if there are lingering doublets:

```{r}
Idents(nbc_mbc) <- "cohort_type"
FeaturePlot(nbc_mbc, c("CD3D", "LYZ")) &
  scale_color_viridis_c(option = "magma")
FeaturePlot(nbc_mbc[, nbc_mbc$cohort_type == "validation"], "doublet_score_scDblFinder") &
  scale_color_viridis_c(option = "magma")
```

Indeed, one cluster is composed of doublets between B and T cells. Let us find it and remove it:

```{r}
nbc_mbc <- FindNeighbors(nbc_mbc, dims = 1:25, reduction = "harmony")
nbc_mbc <- FindClusters(nbc_mbc, resolution = 0.1)
DimPlot(nbc_mbc, cols = rev(color_palette))
markers6 <- FindMarkers(nbc_mbc, ident.1 = "6", only.pos = TRUE, logfc.threshold = 0.5)
head(markers6, 30)
nbc_mbc <- nbc_mbc[, !(nbc_mbc$seurat_clusters == "6" & nbc_mbc$cohort_type == "validation")]
DimPlot(nbc_mbc, cols = rev(color_palette), split.by = "cohort_type")
```

Reintegrate:

```{r}
nbc_mbc$type <- str_c(nbc_mbc$cohort_type, nbc_mbc$assay, sep = "_")
shared_hvg <- find_assay_specific_features(nbc_mbc, assay_var = "type")
nbc_mbc <- integrate_assays(
  nbc_mbc,
  assay_var = "type",
  shared_hvg = shared_hvg
)
nbc_mbc <- RunUMAP(nbc_mbc, dims = 1:25, reduction = "harmony")
DimPlot(
  nbc_mbc,
  group.by = "annotation_20230508",
  split.by = "cohort_type",
  cols = c(colors_20230508[["NBC_MBC"]], "unannotated" = "gray")
)  &
  theme(legend.text = element_text(size = 6))
```


Transfer label:

```{r}
# Split training and test sets
data_sets <- split_training_and_test_sets(
  nbc_mbc,
  split_var = "cohort_type",
  referece_label = "discovery",
  query_label = "validation",
  reduction = "harmony",
  n_dims = 25
)


# Transfer label
annotation_query_df <- transfer_label(
  seurat_obj = nbc_mbc,
  training_set = data_sets$training_set,
  test_set = data_sets$test_set,
  k = 30,
  response_var = "annotation_20230508"
)

# Include annotation and UMAP coords in Seurat object
nbc_mbc$annotation_20230508[annotation_query_df$query_cells] <- annotation_query_df$annotation
Idents(nbc_mbc) <- "annotation_20230508"
nbc_mbc$annotation_20230508_probability <- NA
nbc_mbc$annotation_20230508_probability[annotation_query_df$query_cells] <- annotation_query_df$annotation_prob
p1 <- FeaturePlot(
  nbc_mbc[, annotation_query_df$query_cells],
  "annotation_20230508_probability"
) + scale_color_viridis_c(option = "magma")
p2 <- FeaturePlot(
  nbc_mbc[, annotation_query_df$query_cells],
  "doublet_score_scDblFinder"
) + scale_color_viridis_c(option = "magma")
p1 | p2
DimPlot(
  nbc_mbc,
  group.by = "annotation_20230508",
  split.by = "cohort_type",
  cols = colors_20230508[["NBC_MBC"]]
)
table(nbc_mbc$annotation_20230508_probability)
```


Plot dot plot

```{r}
goi_rna <- c("TCL1A", "FCER2", "CD27", "TNFRSF13B", "IGHD", "IGHM", "IGHA1",
             "IGHG1", "CD69", "EGR2", "CCL3", "MYC", "BATF", "CCND2", "NFKB1",
             "NFKB2", "RELB", "CD40", "BCL2A1", "MIR155HG", "EBI3", "TRAF1",
             "MEF2B", "RGS13", "MKI67", "TOP2A", "LY9", "LILRA4", "IFIT1", 
             "IFIT3", "FCRL4", "FCRL5")
avgexpr_mat <- AverageExpression(
  nbc_mbc[, nbc_mbc$cohort_type == "validation"],
  features = goi_rna,
  assays = "RNA",
  return.seurat = FALSE,
  group.by = "annotation_20230508",
  slot = "data"
)$RNA


# Define annotation column
cell_types <- levels(nbc_mbc$annotation_20230508)
mycolors <- list(cell_type = c(colors_20230508$NBC_MBC))
annotation_col <- data.frame(cell_type = cell_types) 
rownames(annotation_col) <- cell_types


# Scale per row between 0 and 1 
input_mat <- t(apply(avgexpr_mat, 1, function(x) (x - min(x)) / diff(range(x))))


# Plot heatmap
pheatmap(
  t(input_mat),
  annotation_row = annotation_col,
  annotation_colors = mycolors,
  annotation_names_row = FALSE,
  annotation_legend = FALSE,
  show_rownames = FALSE,
  show_colnames = TRUE, 
  border_color = NA,
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  gaps_col = c(1, 3, 8),
  fontsize_row = 6
)
```


Plot annotation probability:

```{r}
nbc_mbc@meta.data %>%
  filter(cohort_type == "validation") %>%
  ggplot(aes(annotation_20230508, annotation_20230508_probability, fill = annotation_20230508)) +
    geom_boxplot() +
    scale_fill_manual(values = colors_20230508$NBC_MBC) +
    theme_classic() +
    coord_flip()
```



# Save

```{r}
saveRDS(nbc_mbc, here("scRNA-seq/results/R_objects/seurat_objects_revision/5.10-NBC_MBC_seurat_object_discovery_validation_cohorts_annotation.rds"))
```


# Session Information

```{r}
sessionInfo()
```

