---
title: 'GCBC: Label and coordinate transfer (validation cohort)'
author: "Ramon Massoni-Badosa"
date: "`r Sys.Date()`"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```


# Introduction

Here, we will transfer the label and UMAP coordinates from the discovery cohort to the validation cohort. We will do it one compartment at a time (e.g. gcbc), so we can validate that the transfer is robust

Specifically, we will follow these steps for every compartment:

1. Include or update the annotation label for this object.
2. Plot the integrated object split by type (quey/reference) to visualize if discovery and validation cohorts were integrated properly.
3. Transfer label.
4. Transfer UMAP coordinates + visualize
5. Exclude cells with a low annotation probability and/or a high doublet score
6. Visualize heatmap of specific markers to validate that the cluster and annotation holds for the new cohort.

# Pre-processing

## Load packages

```{r}
library(Seurat)
library(scCustomize)
library(tidyverse)
library(caret)
library(class)
library(harmony)
library(here)
library(glue)
library(DT)
library(pheatmap)
```


## Source functions

```{r}
source(here("scRNA-seq/bin/SLOcatoR_functions.R"))
source(here("scRNA-seq/bin/utils.R"))
source(here("scRNA-seq/bin/utils_final_clusters.R"))
color_palette <-  c("#1CFFCE", "#90AD1C", "#C075A6", "#85660D", "#5A5156", "#AA0DFE",
                    "#F8A19F", "#F7E1A0", "#1C8356", "#FEAF16", "#822E1C", "#C4451C",
                    "#1CBE4F", "#325A9B", "#F6222E", "#FE00FA", "#FBE426", "#16FF32",
                    "black",   "#3283FE", "#B00068", "#DEA0FD", "#B10DA1", "#E4E1E3",
                    "#90AD1C", "#FE00FA", "#85660D", "#3B00FB", "#822E1C", "coral2",
                    "#1CFFCE", "#1CBE4F", "#3283FE", "#FBE426", "#F7E1A0", "#325A9B",
                    "#2ED9FF", "#B5EFB5", "#5A5156", "#DEA0FD", "#FEAF16", "#683B79",
                    "#B10DA1", "#1C7F93", "#F8A19F", "dark orange", "#FEAF16", "#FBE426",
                    "Brown")
```


# GCBC

```{r}
# Read data, clean and include annotation
gcbc <- readRDS(here("scRNA-seq/results/R_objects/seurat_objects_revision/4.11-GCBC_seurat_object_discovery_validation_cohorts.rds"))
```


We proceed to update the annotation:

```{r}
gcbc <- updateAnnotation(gcbc)
gcbc$annotation_20230508 <- gcbc$annotation_20220619
gcbc$annotation_20230508[is.na(gcbc$annotation_20230508)] <- "unannotated"
gcbc$annotation_20230508 <- factor(
  gcbc$annotation_20230508,
  levels = c(names(colors_20230508[["GCBC"]]), "unannotated")
)
Idents(gcbc) <- "annotation_20230508"
gcbc$is_reference <- ifelse(
  gcbc$cohort_type == "discovery",
  "reference",
  "query"
)
DimPlot(
  gcbc,
  group.by = "annotation_20230508",
  split.by = "cohort_type",
  cols = colors_20230508[["GCBC"]]
)
```

Let us explore if there are lingering doublets:

```{r}
Idents(gcbc) <- "cohort_type"
FeaturePlot(gcbc, c("CD3D", "LYZ")) &
  scale_color_viridis_c(option = "magma")
FeaturePlot(gcbc[, gcbc$cohort_type == "validation"], "doublet_score_scDblFinder") &
  scale_color_viridis_c(option = "magma")
```

Indeed, one cluster is composed of doublets between B and T cells. Let us find it and remove it:

```{r}
gcbc <- FindNeighbors(gcbc, dims = 1:25, reduction = "harmony")
gcbc <- FindClusters(gcbc, resolution = 1)
DimPlot(gcbc, cols = rev(color_palette))
markers16 <- FindMarkers(gcbc, ident.1 = "16", only.pos = TRUE, logfc.threshold = 0.5)
head(markers16, 30)
gcbc <- gcbc[, !(gcbc$seurat_clusters == "16" & gcbc$cohort_type == "validation")]
DimPlot(gcbc, cols = rev(color_palette), split.by = "cohort_type")
```

Reintegrate:

```{r}
gcbc$type <- str_c(gcbc$cohort_type, gcbc$assay, sep = "_")
shared_hvg <- find_assay_specific_features(gcbc, assay_var = "type")
gcbc <- integrate_assays(
  gcbc,
  assay_var = "type",
  shared_hvg = shared_hvg
)
gcbc <- RunUMAP(gcbc, dims = 1:25, reduction = "harmony")
DimPlot(
  gcbc,
  group.by = "annotation_20230508",
  split.by = "cohort_type",
  cols = c(colors_20230508[["GCBC"]], "unannotated" = "gray")
)  &
  theme(legend.text = element_text(size = 6))
```


Transfer label:

```{r}
# Split training and test sets
data_sets <- split_training_and_test_sets(
  gcbc,
  split_var = "cohort_type",
  referece_label = "discovery",
  query_label = "validation",
  reduction = "harmony",
  n_dims = 25
)


# Transfer label
annotation_query_df <- transfer_label(
  seurat_obj = gcbc,
  training_set = data_sets$training_set,
  test_set = data_sets$test_set,
  k = 30,
  response_var = "annotation_20230508"
)

# Include annotation and UMAP coords in Seurat object
gcbc$annotation_20230508[annotation_query_df$query_cells] <- annotation_query_df$annotation
Idents(gcbc) <- "annotation_20230508"
gcbc$annotation_20230508_probability <- NA
gcbc$annotation_20230508_probability[annotation_query_df$query_cells] <- annotation_query_df$annotation_prob
p1 <- FeaturePlot(
  gcbc[, annotation_query_df$query_cells],
  "annotation_20230508_probability"
) + scale_color_viridis_c(option = "magma")
p2 <- FeaturePlot(
  gcbc[, annotation_query_df$query_cells],
  "doublet_score_scDblFinder"
) + scale_color_viridis_c(option = "magma")
p1 | p2
DimPlot(
  gcbc,
  group.by = "annotation_20230508",
  split.by = "cohort_type",
  cols = colors_20230508[["GCBC"]]
)
table(gcbc$annotation_20230508_probability)
```


Plot dot plot

```{r}
goi_rna <- c("CXCR4", "AICDA", "FOXP1", "MME", "CD83", "LMO2", "POLA1", "HIST1H4C",
             "MKI67", "TOP2A", "CDC20", "CCNB1", "STMN1", "HMGB2", "CD40", "BCL2A1",
             "MIR155HG", "EBI3", "TRAF1", "NFKB1", "NFKB2", "RELB", "MYC", "BATF",
             "CCR6", "CELF2", "BANK1", "PRDM1", "XBP1")
avgexpr_mat <- AverageExpression(
  gcbc[, gcbc$cohort_type == "validation"],
  features = goi_rna,
  assays = "RNA",
  return.seurat = FALSE,
  group.by = "annotation_20230508",
  slot = "data"
)$RNA


# Define annotation column
cell_types <- levels(gcbc$annotation_20230508)
mycolors <- list(cell_type = c(colors_20230508$GCBC))
annotation_col <- data.frame(cell_type = cell_types) 
rownames(annotation_col) <- cell_types


# Scale per row between 0 and 1 
input_mat <- t(apply(avgexpr_mat, 1, function(x) (x - min(x)) / diff(range(x))))


# Plot heatmap
pheatmap(
  t(input_mat),
  annotation_row = annotation_col,
  annotation_colors = mycolors,
  annotation_names_row = FALSE,
  annotation_legend = FALSE,
  show_rownames = FALSE,
  show_colnames = TRUE, 
  border_color = NA,
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  gaps_col = c(1, 3, 8),
  fontsize_row = 6
)
```


# Save

```{r}
saveRDS(gcbc, here("scRNA-seq/results/R_objects/seurat_objects_revision/5.11-GCBC_seurat_object_discovery_validation_cohorts_annotation.rds"))
```


# Session Information

```{r}
sessionInfo()
```

