---
title: 'PDC: Label and coordinate transfer (validation cohort)'
author: "Ramon Massoni-Badosa"
date: "`r Sys.Date()`"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```


# Introduction

Here, we will transfer the label and UMAP coordinates from the discovery cohort to the validation cohort. We will do it one compartment at a time (here, follicular dendritic cells, pdc), so we can validate that the transfer is robust.

Specifically, we will follow these steps for every compartment:

1. Include or update the annotation label for this object.
2. Plot the integrated object split by type (quey/reference) to visualize if discovery and validation cohorts were integrated properly.
3. Transfer label.
4. Transfer UMAP coordinates + visualize
5. Exclude cells with a low annotation probability and/or a high doublet score
6. Visualize heatmap of specific markers to validate that the cluster and annotation holds for the new cohort.

# Pre-processing

## Load packages

```{r}
library(Seurat)
library(tidyverse)
library(caret)
library(class)
library(harmony)
library(here)
library(glue)
library(DT)
```


## Source functions

```{r}
source(here("scRNA-seq/bin/SLOcatoR_functions.R"))
source(here("scRNA-seq/bin/utils.R"))
source(here("scRNA-seq/bin/utils_final_clusters.R"))
color_palette <-  c("#1CFFCE", "#90AD1C", "#C075A6", "#85660D", "#5A5156", "#AA0DFE",
                    "#F8A19F", "#F7E1A0", "#1C8356", "#FEAF16", "#822E1C", "#C4451C",
                    "#1CBE4F", "#325A9B", "#F6222E", "#FE00FA", "#FBE426", "#16FF32",
                    "black",   "#3283FE", "#B00068", "#DEA0FD", "#B10DA1", "#E4E1E3",
                    "#90AD1C", "#FE00FA", "#85660D", "#3B00FB", "#822E1C", "coral2",
                    "#1CFFCE", "#1CBE4F", "#3283FE", "#FBE426", "#F7E1A0", "#325A9B",
                    "#2ED9FF", "#B5EFB5", "#5A5156", "#DEA0FD", "#FEAF16", "#683B79",
                    "#B10DA1", "#1C7F93", "#F8A19F", "dark orange", "#FEAF16", "#FBE426",
                    "Brown")
```


# PDC

```{r}
# Read data, clean and include annotation
pdc <- readRDS(here("scRNA-seq/results/R_objects/seurat_objects_revision/4.3-PDC_seurat_object_discovery_validation_cohorts.rds"))
pdc <- updateAnnotation(pdc)
table(pdc$assay)
pdc <- pdc[, pdc$assay == "3P"]
pdc$annotation_20230508 <- pdc$annotation_20220619
pdc$annotation_20230508[is.na(pdc$annotation_20230508)] <- "unannotated"
pdc$annotation_20230508 <- factor(
  pdc$annotation_20230508,
  levels = c(names(colors_20230508[["PDC"]]), "unannotated")
)
Idents(pdc) <- "annotation_20230508"
pdc$is_reference <- ifelse(
  pdc$cohort_type == "discovery",
  "reference",
  "query"
)
DimPlot(
  pdc,
  group.by = "annotation_20230508",
  split.by = "is_reference",
  cols = colors_20230508[["PDC"]]
)
DimPlot(
  pdc,
  group.by = "assay",
  split.by = "is_reference"
)
```


We see that we have too few multiome cells for a proper label transfer. Thus, we will label them as "pdc", and focus on transering the label from the discovery scRNA-seq dataset to validation scRNA-seq. Let us integrate again this dataset:

```{r}
Idents(pdc) <- "gem_id"
shared_hvg <- find_assay_specific_features(pdc, assay_var = "cohort_type")
print(glue("The number of shared HVG is: {length(shared_hvg)}"))
table(pdc$cohort_type)
pdc <- integrate_assays(
  pdc,
  assay_var = "cohort_type",
  shared_hvg = shared_hvg
)
pdc <- RunUMAP(pdc, dims = 1:25, reduction = "harmony")
DimPlot(
  pdc,
  group.by = "annotation_20230508",
  split.by = "cohort_type",
  cols = colors_20230508[["PDC"]]
)
DimPlot(
  pdc,
  group.by = "cohort_type"
)
FeaturePlot(pdc, c("CD79B", "CD3G")) &
  scale_color_viridis_c(option = "magma")
pdc <- FindNeighbors(pdc, reduction = "harmony", dims = 1:25, k.param = 10)
pdc <- FindClusters(pdc, resolution = 0.8)
DimPlot(pdc, cols = color_palette)
```

We see that there are two clusters of doublets with B and T cells (5, 10). Let's exclude them and reintegrate:

```{r}
markers_5 <- FindMarkers(pdc, ident.1 = "5", only.pos = TRUE, logfc.threshold = 0.7)
DT::datatable(markers_5)
markers_10 <- FindMarkers(pdc, ident.1 = "10", only.pos = TRUE, logfc.threshold = 0.7)
DT::datatable(markers_10)
pdc <- pdc[, !(pdc$seurat_clusters %in% c("5", "10") & pdc$cohort_type == "validation")]
DimPlot(pdc)
table(pdc$cohort_type)
shared_hvg <- find_assay_specific_features(pdc, assay_var = "cohort_type")
print(glue("The number of shared HVG is: {length(shared_hvg)}"))
table(pdc$cohort_type)
pdc <- integrate_assays(
  pdc,
  assay_var = "cohort_type",
  shared_hvg = shared_hvg
)
pdc <- RunUMAP(pdc, dims = 1:25, reduction = "harmony")
DimPlot(
  pdc,
  group.by = "annotation_20230508",
  split.by = "cohort_type",
  cols = colors_20230508[["PDC"]]
)
```


Now we can transfer the label and UMAP coords

```{r}
# Split training and test sets
data_sets <- split_training_and_test_sets(
  pdc,
  split_var = "cohort_type",
  referece_label = "discovery",
  query_label = "validation",
  reduction = "harmony",
  n_dims = 25
)

# Transfer label
annotation_query_df <- transfer_label(
  seurat_obj = pdc,
  training_set = data_sets$training_set,
  test_set = data_sets$test_set,
  k = 8,
  response_var = "annotation_20230508"
)

# Transfer UMAP coords
# umap_test_df <- transfer_umap_coords(
#   seurat_obj = pdc,
#   training_set = data_sets$training_set,
#   test_set = data_sets$test_set,
#   umap1_var = "UMAP_1_20220215",
#   umap2_var = "UMAP_2_20220215",
#   k = 15
# )


# Include annotation and UMAP coords in Seurat object
pdc$annotation_20230508[annotation_query_df$query_cells] <- annotation_query_df$annotation
Idents(pdc) <- "annotation_20230508"
pdc$annotation_20230508_probability <- NA
pdc$annotation_20230508_probability[annotation_query_df$query_cells] <- annotation_query_df$annotation_prob
# pdc$UMAP_1_20220215[umap_test_df$query_cells] <- umap_test_df$UMAP1
# pdc$UMAP_2_20220215[umap_test_df$query_cells] <- umap_test_df$UMAP2
p1 <- FeaturePlot(
  pdc[, annotation_query_df$query_cells],
  "annotation_20230508_probability"
) + scale_color_viridis_c(option = "magma")
p2 <- FeaturePlot(
  pdc[, annotation_query_df$query_cells],
  "doublet_score_scDblFinder"
) + scale_color_viridis_c(option = "magma")
p1 | p2
DimPlot(
  pdc,
  group.by = "annotation_20230508",
  split.by = "cohort_type",
  cols = colors_20230508[["PDC"]]
)
table(pdc$annotation_20230508_probability)
FeaturePlot(pdc, c("nFeature_RNA", "pct_mt"))
```

We see that there is a cluster with hallmarks of poor-quality cells: high % of mitochondrial expression and low number of detected features. We will exclude it:

```{r}
pdc <- FindNeighbors(pdc, k.param = 15, dims = 1:20, reduction = "harmony")
pdc <- FindClusters(pdc, resolution = 0.2)
DimPlot(pdc)
markers_2 <- FindMarkers(pdc, ident.1 = "2", only.pos = TRUE)
DT::datatable(markers_2)
pdc <- pdc[, !(pdc$cohort_type == "validation" & pdc$seurat_clusters == "2")]
DimPlot(pdc, cols = color_palette)
goi <- c("IRF8", "IL3RA", "LILRA4", "IFI44", "IFI6")
FeaturePlot(pdc, goi) &
  scale_color_viridis_c(option = "magma")
DimPlot(pdc, group.by = "annotation_20230508", cols = colors_20230508$PDC)
```


# Save

```{r}
saveRDS(pdc, here("scRNA-seq/results/R_objects/seurat_objects_revision/5.3-PDC_seurat_object_discovery_validation_cohorts_annotation.rds"))
```


# Session Information

```{r}
sessionInfo()
```

