---
title: "Harmonize scATAC-seq object Cytotoxic"
author: "Ramon Massoni-Badosa"
date: "`r format(Sys.time(), '%B %Y')`"
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center',
                      message=FALSE, warning = FALSE)
options(width = 1200)
```


# Introduction

To be consistent with the other data modalities, we will harmonize the scATAC-seq dataset for cytotoxic (CD8 T, NK, ILC) cells, including the variable names in the metadata. We will also transfer the annotation and UMAP coordinates from RNA to ATAC-seq.

# Pre-processing

## Load packages

```{r}
library(Seurat)
library(Signac)
library(harmony)
library(tidyverse)
library(caret)
library(class)
library(here)
library(glue)
```


## Parameters

```{r}
# Source functions
source(here("scRNA-seq/bin/utils_final_clusters.R"))
source(here("scRNA-seq/bin/SLOcatoR_functions.R"))


# Misc
optimal_k <- 5
```



## Read data

```{r}
atac <- readRDS(path_to_atac_cytotoxic_not_harmonized)
rna_cd8 <- readRDS(path_to_save_cd8)
rna_ilc_nk <- readRDS(path_to_save_ilc_nk)
```


# Merge cytotoxic RNA and reprocess

```{r}
# Merge
rna_ilc_nk$UMAP_1_20220215 <- rna_ilc_nk$UMAP_1_20220215 + 10
rna_ilc_nk$UMAP_2_20220215 <- rna_ilc_nk$UMAP_2_20220215 + 10
rna <- merge(rna_cd8, rna_ilc_nk)


# Reprocess
shared_hvg <- find_assay_specific_features(rna)
rna <- integrate_assays(rna, shared_hvg = shared_hvg, n_dim = 30)
rna <- RunUMAP(rna, dims = 1:30, reduction = "harmony")
DimPlot(rna, split.by = "assay")
```


# Harmonize scATAC-seq dataset

```{r}
selected_cols_atac <- selected_cols_atac[which(selected_cols_atac != "annotation_figure_1")]
atac$barcode <- colnames(atac)
atac$is_facs_sorted <- atac$technique
atac$UMAP_1_20220215 <- Embeddings(atac, "umap")[, "UMAP_1"]
atac$UMAP_2_20220215 <- Embeddings(atac, "umap")[, "UMAP_2"]
if (all(selected_cols_atac %in% colnames(atac@meta.data))) {
  atac@meta.data <- atac@meta.data[, selected_cols_atac]
} else {
  stop("Some columns are not present!")
}
```


# Transfer UMAP coordinates

```{r}
# Define variables to predict
ref_cells <- colnames(rna)[colnames(rna) %in% colnames(atac)]
atac$UMAP_1_RNA_based <- NA
atac$UMAP_2_RNA_based <- NA
atac$UMAP_1_RNA_based[ref_cells] <- rna$UMAP_1_20220215[ref_cells]
atac$UMAP_2_RNA_based[ref_cells] <- rna$UMAP_2_20220215[ref_cells]
atac$type <- ifelse(
  colnames(atac) %in% ref_cells,
  "reference",
  "query"
)


# Create training and test sets
data_sets <- split_training_and_test_sets(
  atac,
  split_var = "type",
  referece_label = "reference",
  query_label = "query",
  reduction = "harmony",
  n_dims = 30
)
data_sets$training_set <- data_sets$training_set[, 2:30]
data_sets$test_set <- data_sets$test_set[, 2:30]


# Predict UMAP coords
umap_coords <- transfer_umap_coords(
  seurat_obj = atac,
  training_set = data_sets$training_set,
  test_set = data_sets$test_set,
  k = optimal_k,
  umap1_var = "UMAP_1_RNA_based",
  umap2_var = "UMAP_2_RNA_based"
)
atac$UMAP_1_RNA_based[umap_coords$query_cells] <- umap_coords$UMAP1
atac$UMAP_2_RNA_based[umap_coords$query_cells] <- umap_coords$UMAP2
```


# Transfer label

```{r}
# Predict UMAP coords
atac$annotation_20220215 <- NA
atac$annotation_20220215[ref_cells] <- rna$annotation_20220215[ref_cells]
annotation_data <- transfer_label(
  seurat_obj = atac,
  training_set = data_sets$training_set,
  test_set = data_sets$test_set,
  k = optimal_k,response_var = "annotation_20220215"
)
atac$annotation_20220215[annotation_data$query_cells] <- annotation_data$annotation
atac$annotation_prob <- NA
atac$annotation_prob[annotation_data$query_cells] <- annotation_data$annotation_prob
FeatureScatter(
  atac,
  "UMAP_1_RNA_based",
  "UMAP_2_RNA_based",
  group.by = "annotation_20220215",
  cols = color_palette,
  pt.size = 0.4
) + ggtitle("")
```


# Recreate fragment objects

As we changed the names of the cells and the path to the fragment files, let us recreate the Fragment objects:

```{r}
# Create new fragment files
gem_ids <- unique(atac$gem_id[atac$assay == "multiome"])
fragments_list <- purrr::map(gem_ids, function(x) {
  fragments <- CreateFragmentObject(
    path = here(glue("data/raw_data_figures/scATAC_CD4_T/fragments/{x}_atac_fragments_renamed_cells.tsv.gz")),
    cells = colnames(atac)[atac$gem_id == x],
    validate.fragments = TRUE
  )
  fragments
})
names(fragments_list) <- gem_ids
fragments_atac <- CreateFragmentObject(
  path = here("data/raw_data_figures/scATAC_CD4_T/fragments/scATAC_fragments.tsv.gz"),
  cells = colnames(atac)[atac$assay == "scATAC"],
  validate.fragments = TRUE
)
fragments_list <- c(
  "scATAC-seq" = fragments_atac,
  fragments_list
)
atac[["peaks_redefined"]]@fragments <- fragments_list
```


# Save

```{r}
saveRDS(atac, path_to_save_atac_cytotoxic)
```


# Session Info

```{r}
sessionInfo()
```

