---
title: "Harmonize scATAC-seq object B cells"
author: "Ramon Massoni-Badosa"
date: "`r format(Sys.time(), '%B %Y')`"
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center',
                      message=FALSE, warning = FALSE)
options(width = 1200)
```


# Introduction

To be consistent with the other data modalities, we will harmonize the scATAC-seq dataset for the B cells (NBC/MBC, GCBC, PC), including the variable names in the metadata. We will also transfer the UMAP coordinates from RNA to ATAC-seq.

# Pre-processing

## Load packages

```{r}
library(Seurat)
library(Signac)
library(tidyverse)
library(caret)
library(class)
library(here)
library(glue)
```


## Parameters

```{r}
# Source functions
source(here("scRNA-seq/bin/utils_final_clusters.R"))
source(here("scRNA-seq/bin/SLOcatoR_functions.R"))


# Misc
optimal_k <- 5
```


## Read data

```{r}
atac_nbc_mbc <- readRDS(path_to_atac_nbc_mbc_not_harmonized)
atac_gcbc <- readRDS(path_to_atac_gcbc_not_harmonized)
atac_pc<- readRDS(path_to_atac_pc_not_harmonized)
# rna_nbc_mbc <- readRDS(path_to_save_nbc_mbc)
# rna_gcbc <- readRDS(path_to_save_gcbc)
# rna_pc <- readRDS(path_to_save_pc)
```


# Harmonize scATAC-seq dataset

```{r}
unwanted_cols <- c("annotation_figure_1", "nCount_peaks_macs",
                   "nFeature_peaks_macs")
selected_cols_atac <- selected_cols_atac[which(!(selected_cols_atac %in% unwanted_cols))]


# NBC/MBC
atac_nbc_mbc <- RenameAssays(atac_nbc_mbc, peaks_macs = "ATAC")
atac_nbc_mbc$barcode <- colnames(atac_nbc_mbc)
atac_nbc_mbc$UMAP_1_20220215 <- Embeddings(atac_nbc_mbc, "umap")[, "UMAP_1"]
atac_nbc_mbc$UMAP_2_20220215 <- Embeddings(atac_nbc_mbc, "umap")[, "UMAP_2"]
atac_nbc_mbc$is_facs_sorted <- atac_nbc_mbc$technique
atac_nbc_mbc$nCount_ATAC <- atac_nbc_mbc$nCount_peaks_macs
atac_nbc_mbc$nFeature_ATAC <- atac_nbc_mbc$nFeature_peaks_macs
if (all(selected_cols_atac %in% colnames(atac_nbc_mbc@meta.data))) {
  atac_nbc_mbc@meta.data <- atac_nbc_mbc@meta.data[, selected_cols_atac]
} else {
  stop("Some columns are not present!")
}


# GCBC
atac_gcbc <- RenameAssays(atac_gcbc, peaks_redefined = "ATAC")
atac_gcbc$barcode <- colnames(atac_gcbc)
atac_gcbc$UMAP_1_20220215 <- Embeddings(atac_gcbc, "umap")[, "UMAP_1"]
atac_gcbc$UMAP_2_20220215 <- Embeddings(atac_gcbc, "umap")[, "UMAP_2"]
atac_gcbc$is_facs_sorted <- atac_gcbc$technique
atac_gcbc$nCount_ATAC <- atac_gcbc$nCount_peaks_redefined
atac_gcbc$nFeature_ATAC <- atac_gcbc$nFeature_peaks_redefined
atac_gcbc$annotation_20220215 <- atac_gcbc$level_5
if (all(selected_cols_atac %in% colnames(atac_gcbc@meta.data))) {
  atac_gcbc@meta.data <- atac_gcbc@meta.data[, c(selected_cols_atac, "annotation_20220215")]
} else {
  stop("Some columns are not present!")
}


# PC
atac_pc <- RenameAssays(atac_pc, peaks_level_5 = "ATAC")
atac_pc$barcode <- colnames(atac_pc)
atac_pc$UMAP_1_20220215 <- Embeddings(atac_pc, "umap")[, "UMAP_1"]
atac_pc$UMAP_2_20220215 <- Embeddings(atac_pc, "umap")[, "UMAP_2"]
atac_pc$is_facs_sorted <- atac_pc$technique
atac_pc$nCount_ATAC <- atac_pc$nCount_peaks_level_5
atac_pc$nFeature_ATAC <- atac_pc$nFeature_peaks_level_5
atac_pc$annotation_20220215 <- atac_pc$annotation_level_5
unwanted_cols2 <- c("UMAP_1_level_1", "UMAP_2_level_1")
selected_cols_atac_pc <- selected_cols_atac[which(!(selected_cols_atac %in% unwanted_cols2))]
if (all(selected_cols_atac_pc %in% colnames(atac_pc@meta.data))) {
  atac_pc@meta.data <- atac_pc@meta.data[, selected_cols_atac_pc]
} else {
  stop("Some columns are not present!")
}
```


# Transfer UMAP coordinates

```{r}
paths_rna <- c(
  NBC_MBC = path_to_save_nbc_mbc,
  GCBC = path_to_save_gcbc,
  PC = path_to_save_pc
)
atac_list <- list(
  NBC_MBC = atac_nbc_mbc,
  GCBC = atac_gcbc,
  PC = atac_pc
)
rm(atac_nbc_mbc, atac_gcbc, atac_pc)
gc(verbose = TRUE)
for (x in names(atac_list)) {
  print(x)
  rna <- readRDS(paths_rna[x])
  
  # Define variables to predict
  ref_cells <- colnames(rna)[colnames(rna) %in% colnames(atac_list[[x]])]
  atac_list[[x]]$UMAP_1_RNA_based <- NA
  atac_list[[x]]$UMAP_2_RNA_based <- NA
  atac_list[[x]]$UMAP_1_RNA_based[ref_cells] <- rna$UMAP_1_20220215[ref_cells]
  atac_list[[x]]$UMAP_2_RNA_based[ref_cells] <- rna$UMAP_2_20220215[ref_cells]
  atac_list[[x]]$type <- ifelse(
    colnames(atac_list[[x]]) %in% ref_cells,
    "reference",
    "query"
  )
  
  # Create training and test sets
  data_sets <- split_training_and_test_sets(
    atac_list[[x]],
    split_var = "type",
    referece_label = "reference",
    query_label = "query",
    reduction = "harmony",
    n_dims = 30
  )
  data_sets$training_set <- data_sets$training_set[, 2:30]
  data_sets$test_set <- data_sets$test_set[, 2:30]
  
  # Predict UMAP coords
  umap_coords <- transfer_umap_coords(
    seurat_obj = atac_list[[x]],
    training_set = data_sets$training_set,
    test_set = data_sets$test_set,
    k = optimal_k,
    umap1_var = "UMAP_1_RNA_based",
    umap2_var = "UMAP_2_RNA_based"
  )
  atac_list[[x]]$UMAP_1_RNA_based[umap_coords$query_cells] <- umap_coords$UMAP1
  atac_list[[x]]$UMAP_2_RNA_based[umap_coords$query_cells] <- umap_coords$UMAP2
  
  # Clean
  rm(rna)
  gc(verbose = TRUE)
}
```


# Transfer label

```{r}
# Predict label
rna <- readRDS(paths_rna["NBC_MBC"])
atac_list$NBC_MBC$annotation_20220215 <- NA
ref_cells <- colnames(rna)[colnames(rna) %in% colnames(atac_list$NBC_MBC)]
atac_list$NBC_MBC$annotation_20220215[ref_cells] <- rna$annotation_20220215[ref_cells]
data_sets <- split_training_and_test_sets(
  atac_list$NBC_MBC,
  split_var = "type",
  referece_label = "reference",
  query_label = "query",
  reduction = "harmony",
  n_dims = 30
)
data_sets$training_set <- data_sets$training_set[, 2:30]
data_sets$test_set <- data_sets$test_set[, 2:30]
annotation_data <- transfer_label(
  seurat_obj = atac_list$NBC_MBC,
  training_set = data_sets$training_set,
  test_set = data_sets$test_set,
  k = optimal_k,
  response_var = "annotation_20220215"
)
atac_list$NBC_MBC$annotation_20220215[annotation_data$query_cells] <- annotation_data$annotation
atac_list$NBC_MBC$annotation_20220215_prob <- NA
atac_list$NBC_MBC$annotation_20220215_prob[annotation_data$query_cells] <- annotation_data$annotation_prob
FeatureScatter(
  atac_list$NBC_MBC,
  "UMAP_1_RNA_based",
  "UMAP_2_RNA_based",
  group.by = "annotation_20220215",
  cols = color_palette,
  pt.size = 0.4
) + ggtitle("")
```


# Recreate fragment objects

As we changed the names of the cells and the path to the fragment files, let us recreate the Fragment objects:

```{r}
# Create new fragment files
atac_list <- map(atac_list, function(atac) {
  gem_ids <- unique(atac$gem_id[atac$assay == "multiome"])
  fragments_list <- purrr::map(gem_ids, function(x) {
    fragments <- CreateFragmentObject(
      path = here(glue("data/raw_data_figures/scATAC_CD4_T/fragments/{x}_atac_fragments_renamed_cells.tsv.gz")),
      cells = colnames(atac)[atac$gem_id == x],
      validate.fragments = TRUE
    )
    fragments
  })
  names(fragments_list) <- gem_ids
  fragments_atac <- CreateFragmentObject(
    path = here("data/raw_data_figures/scATAC_CD4_T/fragments/scATAC_fragments.tsv.gz"),
    cells = colnames(atac)[atac$assay == "scATAC"],
    validate.fragments = TRUE
  )
  fragments_list <- c(
    "scATAC-seq" = fragments_atac,
    fragments_list
  )
  atac[["ATAC"]]@fragments <- fragments_list
  atac
})
```


# Save

```{r}
saveRDS(atac_list$NBC_MBC, path_to_save_atac_nbc_mbc)
saveRDS(atac_list$GCBC, path_to_save_atac_gcbc)
saveRDS(atac_list$PC, path_to_save_atac_pc)
```


# Session Info

```{r}
sessionInfo()
```

