---
title: "Merge and integrate all B cells"
author: "Ramon Massoni-Badosa & Paula Soler-Vila"
date: "2021/05/13"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```


# Introduction

The purpose of this notebook is to integrate all B cells into a single Seurat object. 

# Pre-processing

## Load packages

```{r}
library(Seurat)
library(SeuratWrappers)
library(harmony)
library(tidyverse)
```


## Parameters

```{r}
path_to_obj_general <- here::here("scATAC-seq/results/R_objects/8.3.tonsil_peakcalling_annotation_level1_signature.rds")

path_to_NBC_MBC <- here::here("scATAC-seq/results/R_objects/level_3/NBC_MBC/NBC_MBC_integrated_level_3.rds")
path_to_GCBC <- here::here("scATAC-seq/results/R_objects/level_3/GCBC/GCBC_integrated_level_3.rds")
path_to_PC <- here::here("scATAC-seq/results/R_objects/level_3/PC/PC_integrated_level_3.rds")

path_to_NBC_MBC_RNA <- here::here("scRNA-seq/3-clustering/3-level_3/NBC_MBC_clustered_level_3_with_pre_freeze.rds")
path_to_GCBC_RNA <- here::here("scRNA-seq/3-clustering/3-level_3/GCBC_clustered_level_3_with_pre_freeze.rds")
path_to_PC_RNA <- here::here("scRNA-seq/3-clustering/3-level_3/PC_clustered_level_3_with_pre_freeze.rds")

path_to_save <- here::here("scATAC-seq/results/R_objects/B_cells_integrated.rds")
```

## Load data
```{r}
NBC_MBC_RNA <- readRDS(path_to_NBC_MBC_RNA)
NBC_MBC_RNA

GCBC_RNA <- readRDS(path_to_GCBC_RNA)
GCBC_RNA

PC_RNA <- readRDS(path_to_PC_RNA)
PC_RNA

tonsil <- readRDS(path_to_obj_general)
tonsil

B_cell_lineage <- subset(tonsil, annotation_level_1 == "NBC_MBC" |
                          annotation_level_1 == "GCBC" |
                          annotation_level_1 == "PC" )
B_cell_lineage

DimPlot(
  B_cell_lineage,
  pt.size = 0.2,
  group.by = "annotation_level_1"
)

nbc_mbc <- readRDS(path_to_NBC_MBC)
gcbc <- readRDS(path_to_GCBC)
pc <- readRDS(path_to_PC)
```

## Clean the data taking into account the level 3
We exclude the clusters of doublets or poor-quality cells detected at level 3.

```{r}
selected_cells <- c(colnames(nbc_mbc),colnames(gcbc),colnames(pc))

B_cell_lineage_clean <- subset(B_cell_lineage, 
                               cells = selected_cells)
B_cell_lineage_clean

DimPlot(
  B_cell_lineage_clean,
  pt.size = 0.2,
  group.by = "annotation_level_1"
)
```

# Integration of the B-cell lineage data.

## Normalization, dimensionality reduction 
```{r}
B_cell_lineage_clean <-  RunTFIDF(B_cell_lineage_clean) %>% 
FindTopFeatures(min.cutoff = 10) %>% RunSVD()  

DepthCor(B_cell_lineage_clean)

B_cell_lineage_clean <- RunUMAP(object = B_cell_lineage_clean, 
                                reduction = 'lsi', 
                                dims = 2:40)
DimPlot(B_cell_lineage_clean, 
        group.by ="assay")
```

## Batch correction
```{r}
B_cell_lineage_clean <- RunHarmony(
  object = B_cell_lineage_clean,
  dims = 2:40,
  group.by.vars = 'gem_id',
  reduction = 'lsi',
  assay.use = 'peaks_macs',
  project.dim = FALSE,
  max.iter.harmony = 20
)

# Non-linear dimension reduction and clustering
B_cell_lineage_clean <- RunUMAP(B_cell_lineage_clean, 
                    dims = 2:40, 
                    reduction = 'harmony') 

B_cell_lineage_clean <- FindNeighbors(object = B_cell_lineage_clean, 
                                      reduction = 'harmony', 
                                      dims = 2:40)
B_cell_lineage_clean <- FindClusters(object = B_cell_lineage_clean, 
                                     verbose = FALSE,
                                     resolution = 0.18)
DimPlot(object = B_cell_lineage_clean, label = TRUE) + NoLegend()
```


# Annotation
We used the clusters defined at level 3 in the scRNA-seq data. 
```{r}
cluster_Naive  <- colnames(NBC_MBC_RNA)[NBC_MBC_RNA$seurat_clusters == 0 | NBC_MBC_RNA$seurat_clusters == 2]

DimPlot(B_cell_lineage_clean, 
  cols.highlight = "darkred", cols= "grey",
  cells.highlight= c(cluster_Naive),
  pt.size = 0.1
)
```

```{r}
cluster_Memory_cs  <- colnames(NBC_MBC_RNA)[NBC_MBC_RNA$seurat_clusters == 1]
cluster_Memory_ncs  <- colnames(NBC_MBC_RNA)[NBC_MBC_RNA$seurat_clusters == 3]

DimPlot(B_cell_lineage_clean, 
  cols.highlight = "brown1", cols= "grey",
  cells.highlight= c(cluster_Memory_cs),
  pt.size = 0.1
)

DimPlot(B_cell_lineage_clean, 
  cols.highlight = "brown4", cols= "grey",
  cells.highlight= c(cluster_Memory_ncs),
  pt.size = 0.1
)
```

```{r}
cluster_Plasma <- colnames(PC_RNA)

DimPlot(B_cell_lineage_clean, 
  cols.highlight = "cyan4", cols= "grey",
  cells.highlight= c(cluster_Plasma),
  pt.size = 0.1
)
```

```{r}
cluster_Germinal <- colnames(GCBC_RNA)

cluster_Germinal_DZ <- colnames(GCBC_RNA)[GCBC_RNA$seurat_clusters == 5 | 
                                                    GCBC_RNA$seurat_clusters == 1 | 
                                                    GCBC_RNA$seurat_clusters == 3]

cluster_Germinal_LZ  <- colnames(GCBC_RNA)[GCBC_RNA$seurat_clusters == 4 |
                                             GCBC_RNA$seurat_clusters == 0 |
                                             GCBC_RNA$seurat_clusters == 6]

DimPlot(B_cell_lineage_clean, 
  cols.highlight = "green4", cols= "grey",
  cells.highlight= c(cluster_Germinal_DZ),
  pt.size = 0.1
)

DimPlot(B_cell_lineage_clean,
  cols.highlight = "green3", cols= "grey",
  cells.highlight= c(cluster_Germinal_LZ),
  pt.size = 0.1
)

```

# Save

```{r}
saveRDS(B_cell_lineage_clean, path_to_save)
```


# Session Info

```{r}
sessionInfo()
```
