---
title: "Germinal Center B cells level 4 to level 5"
author: "Sergio Aguilar Fernández"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  BiocStyle::html_document:
    df_print: paged
    toc: true
    toc_float: true
    number_sections: false
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      out.width = "100%", 
                      fig.align='center',
                      fig.path="figures_compiled_html/",
                      message=FALSE, 
                      warning = FALSE)
```

## Introduction

In this Rmd we are going to clean the level 4 of the __*Germinal Center B Cells (GCBC)*__. Here, we will focus on removing potential doublets and low quality cells to provide a curated level 5 for the *GCBC*. 

## Libraries 

```{r}
library(Seurat)
library(tidyverse)
library(harmony)

```

## Parameters 

### Folder structure
```{r}
level <- "level_4"
cell_type <- "GCBC"
base_folder <- here::here(paste0(level,"/",cell_type,"/"))
  data_folder <- paste0(base_folder, "data/")
  results_folder <- paste0(base_folder, "results/")
  figures_folder <- paste0(results_folder, "figures/")
  markers_folder <- paste0(results_folder, "objects/markers/")

level_5_folder = here::here(paste0("level_5/",cell_type,"/data/"))
```

### Palette and pt.size
```{r}
palette<-c('#e6194b', '#3cb44b', '#ffd8b1', '#4363d8', '#f58231', '#911eb4', '#46f0f0', '#f032e6', '#bcf60c', '#fabebe', '#008080', '#e6beff', '#9a6324', '#fffac8', '#800000', '#aaffc3', '#808000', '#ffe119', '#000075', '#808080', '#ffffff', '#000000')

pt_sizes <- c(
  NBC_MBC = 0.6,
  GCBC = 0.8,
  PC = 1.5
)
pt_size <- pt_sizes[cell_type]
```


## Load data 
```{r}
seurat_object_l4 <- readRDS(paste0(data_folder, cell_type, "_clustered_level_4.rds"))
```
## Functions

## Visualization exploratory analysis of a given object 

#### Data groups visualization
```{r}
data_groups_visualization <- function(seurat_object, group_variable_1, group_variable_2 = NULL, pt_size) {
        print(Seurat::DimPlot(seurat_object, 
                reduction = "umap",
                pt.size = pt_size,  
                cols= palette, 
                group.by = group_variable_1, 
                raster = FALSE) 
        )
        
        print(Seurat::DimPlot(seurat_object, 
                reduction = "umap",
                pt.size = pt_size,  
                cols= palette, 
                group.by = group_variable_1, 
                raster = FALSE, 
                label = TRUE)
        )
        
        if (!is.null(group_variable_2)) {
            print(Seurat::DimPlot(seurat_object, 
                reduction = "umap",
                pt.size = pt_size,  
                cols= palette, 
                group.by = group_variable_2, 
                raster = FALSE) 
            )
        }
        

        
        #### COMBINED EXPLORATORY DATA GROUPS ANALYSIS
        p1 <- Seurat::DimPlot(seurat_object, 
                reduction = "umap",
                pt.size = pt_size/3,  
                cols= palette, 
                group.by = group_variable_1, 
                raster = FALSE) 
        
        p2 <- Seurat::DimPlot(seurat_object, 
                reduction = "umap", 
                pt.size = pt_size/3,  
                cols= c("tan", "royalblue1", "red"), 
                group.by = "assay", 
                raster = FALSE) 
        
        p3 <- Seurat::DimPlot(seurat_object, 
                reduction = "umap", 
                pt.size = pt_size/3,  
                cols= palette, 
                group.by = "age", 
                raster = FALSE) 
        
        p4 <- Seurat::DimPlot(seurat_object, 
                reduction = "umap", 
                pt.size = pt_size/3,  
                cols= c("tan", "royalblue1", "red"), 
                group.by = "hospital", 
                raster = FALSE) 
        
        
        print(patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2) )
}
```

#### Data features Visualization
```{r}
data_features_visualization <- function(seurat_object, group_variable_1, pt_size) {

    p1 <- Seurat::FeaturePlot(seurat_object, 
                        features = "nCount_RNA", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        raster = FALSE) 
    
    p2 <- Seurat::FeaturePlot(seurat_object, 
                        features = "nFeature_RNA", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        raster = FALSE) 
    
    p3 <- Seurat::FeaturePlot(seurat_object, 
                        features = "pct_mt", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        raster = FALSE) 
    
    p4 <- Seurat::FeaturePlot(seurat_object, 
                        features = "pct_ribosomal", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        raster = FALSE) 
    
    
      print(patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2) )

    p1 <- Seurat::DimPlot(seurat_object, 
            reduction = "umap",
            pt.size = pt_size/3,  
            cols= palette, 
            group.by = group_variable_1, 
            raster = FALSE) 
    
    p2 <- Seurat::FeaturePlot(seurat_object, 
                        features = "pct_mt", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        raster = FALSE) 
    
    p3 <- Seurat::FeaturePlot(seurat_object, 
                        features = "nCount_RNA", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        raster = FALSE) 
    
    p4 <- Seurat::FeaturePlot(seurat_object, 
                        features = "nFeature_RNA", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        raster = FALSE) 
    
     print(patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2) )
    
    
    p1 <- Seurat::FeaturePlot(seurat_object, 
                        features = "nCount_RNA", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        min.cutoff = 'q10',
                        max.cutoff = 'q90', 
                        raster = FALSE) 
                        
    
    p2 <- Seurat::FeaturePlot(seurat_object, 
                        features = "nFeature_RNA", 
                        reduction="umap", 
                        pt.size = pt_size/3, 
                        min.cutoff = 'q10',
                        max.cutoff = 'q90', 
                        raster = FALSE) 
                        
    
      print(patchwork::wrap_plots(p1, p2, ncol = 2) )
      
    g1 <- ggplot(seurat_object@meta.data, aes_string(x=group_variable_1, y="nFeature_RNA", fill=group_variable_1)) + 
        geom_boxplot() + 
        theme(legend.position = "none") + 
        scale_fill_manual(values=palette) + 
        theme(axis.text.x = element_text(angle = 90))
    g2 <- ggplot(seurat_object@meta.data, aes_string(x=group_variable_1, y="nCount_RNA", fill=group_variable_1)) + 
        geom_boxplot() + 
        scale_fill_manual(values=palette) + 
        theme(axis.text.x = element_text(angle = 90))
    
      print(patchwork::wrap_plots(g1, g2, ncol = 2) )

}
```

## Exploratory analysis of level 4 

### Explore the UMAP and different grouping variables 
```{r level4_exploratory_analysis_group_variables, fig.height = 12, fig.width = 16}
pdf(paste(figures_folder, "exploratory_analysis/level4_exploratory_analysis_data_groups.pdf", sep=""),width=20,height=10,paper='special') 
data_groups_visualization(seurat_object = seurat_object_l4, 
                          group_variable_1 = "annotation_level_3",
                          pt_size = pt_size) 
dev.off()
```


### Explore the UMAP and different data variables 

```{r level4_exploratory_analysis_data_variables, fig.height = 12, fig.width = 16}
pdf(paste(figures_folder, "exploratory_analysis/level4_exploratory_analysis_data_variables.pdf", sep=""),width=20,height=10,paper='special') 
data_features_visualization(seurat_object = seurat_object_l4, 
                            group_variable_1 = "annotation_level_3",  
                            pt_size = pt_size) 
dev.off()
```


### Selection of the resolution 

#### Clusters 
```{r fig.height = 12, fig.width = 16}
pdf(paste(figures_folder, "exploratory_analysis/clusters_different_resolutions_all.pdf", sep=""),width=20,height=10,paper='special') 
p1 <- Seurat::DimPlot(seurat_object_l4, 
        reduction = "umap",
        pt.size = pt_size/3,  
        cols= palette, 
        group.by = "annotation_level_3", 
        raster = FALSE) 
print(p1)

resolution_vector <- c(0.01, 0.05, 0.1, 0.25, 0.35, 0.4, 0.45, 0.5, 0.75, 0.9, 1)
#resolution_vector <- c(0.35, 0.4, 0.45, 0.5)

list_plots <- list()
for (index in 1:length(resolution_vector)) {
  resolution <- resolution_vector[index]
  resolution_wanted <- paste0("RNA_snn_res.", resolution)
  Idents(seurat_object_l4) <- resolution_wanted
  p <- Seurat::DimPlot(seurat_object_l4, reduction = "umap",  pt.size = pt_size/3, cols = c(palette, "black"), raster = FALSE, group.by = resolution_wanted)
  list_plots[[index]] <- p
}
print(patchwork::wrap_plots(list_plots))

dev.off()
```

```{r}
pdf(paste(figures_folder, "selected_resolution/selected_resolution.pdf", sep=""),width=20,height=10,paper='special') 
p1 <- Seurat::DimPlot(seurat_object_l4, 
        reduction = "umap",
        pt.size = pt_size/3,  
        cols= palette, 
        group.by = "annotation_level_3", 
        raster = FALSE) 
print(p1)

p2 <- Seurat::DimPlot(seurat_object_l4, 
        reduction = "umap",
        pt.size = pt_size/3,  
        cols= palette, 
        group.by = "RNA_snn_res.0.45", 
        raster = FALSE) 
print(p2)

print(patchwork::wrap_plots(p1, p2, ncol = 2))

dev.off()

pdf(paste(figures_folder, "selected_resolution/quality_analysis.pdf", sep=""),width=20,height=10,paper='special') 
data_features_visualization(seurat_object = seurat_object_l4, 
                            group_variable_1 = "RNA_snn_res.0.45",  
                            pt_size = pt_size) 
dev.off()

```

```{r}
pdf(paste(figures_folder, "selected_resolution/comparison_with_annotation_level_3.pdf", sep=""),width=20,height=10,paper='special') 
cell_type_pcts_df <- seurat_object_l4@meta.data %>%
  group_by(RNA_snn_res.0.45, annotation_level_3) %>%
  summarise(n_cells_cell_type = n()) %>% 
  group_by(RNA_snn_res.0.45) %>%
  mutate(
    total_cells = sum(n_cells_cell_type),
    pct_cells = n_cells_cell_type / total_cells * 100
  )

cell_type_pcts_gg <- cell_type_pcts_df %>%
  ggplot(aes(RNA_snn_res.0.45, pct_cells, fill = annotation_level_3)) +
    geom_col() +
    labs(x = "Cluster", y = "Percentage of cells (%)", fill = "") +
    scale_fill_manual(values = palette) +
    theme_classic()

print(cell_type_pcts_gg)
dev.off()
```


```{r}
pdf(paste(figures_folder, "selected_resolution/level4_exploratory_analysis_data_groups.pdf", sep=""),width=20,height=10,paper='special') 
data_groups_visualization(seurat_object = seurat_object_l4, 
                          group_variable_1 = "RNA_snn_res.0.45",
                          pt_size = pt_size) 
dev.off()

```


## Analysis of selected resolution clusters 
```{r}
seurat_object_l4$level4_clusters <- seurat_object_l4$RNA_snn_res.0.45
```

### Analysis of low_region of cluster 9 

#### Cluster 9 cells

```{r}
cluster_name <- "9"
subcluster_name <- paste0("cluster", cluster_name, "_subcluster")
```

##### Subclustering 

```{r}
### Find subcluster 
Idents(seurat_object_l4) <- "level4_clusters"
seurat_object_l4 <- FindSubCluster(
  seurat_object_l4,
  cluster = "9",
  graph.name = "RNA_snn",
  subcluster.name = subcluster_name,
  resolution = 0.25
)
```

```{r}
cluster_cells <- Cells(seurat_object_l4)[which(seurat_object_l4$level4_clusters == cluster_name)]
```

##### Visualization analysis

```{r}
pdf(paste(figures_folder, "subtype_analysis/cluster9_subclusters.pdf", sep=""),width=20,height=10,paper='special') 
Seurat::DimPlot(seurat_object_l4, 
        reduction = "umap", 
        pt.size = pt_size/2,  
        cols= palette, 
        group.by = subcluster_name, 
        raster = FALSE) 

Seurat::DimPlot(seurat_object_l4, 
        reduction = "umap", 
        pt.size = pt_size,  
        cols= palette, 
        group.by = subcluster_name, 
        cells=cluster_cells)
dev.off()
```


```{r}
pdf(paste(figures_folder, "subtype_analysis/cluster9_lowquality_region_selection.pdf", sep=""),width=20,height=10,paper='special') 
## INSPECT SUBCLUSTERS
Seurat::DimPlot(seurat_object_l4, 
        reduction = "umap", 
        pt.size = pt_size/2,  
        cols= palette, 
        group.by = subcluster_name, 
        raster = FALSE) 

Seurat::DimPlot(seurat_object_l4, 
        reduction = "umap", 
        pt.size = pt_size,  
        cols= palette, 
        group.by = subcluster_name, 
        cells=cluster_cells)

## INSPECT GENERAL QUALITY 
p1 <- Seurat::FeaturePlot(seurat_object_l4, 
                    features = "nCount_RNA", 
                    reduction="umap", 
                    min.cutoff = 'q10',
                    max.cutoff = 'q90',                    
                    pt.size = pt_size/3, 
                    cells=cluster_cells ) 


p2 <- Seurat::FeaturePlot(seurat_object_l4, 
                    features = "nFeature_RNA", 
                    reduction="umap", 
                    pt.size = pt_size/3, 
                    min.cutoff = 'q10',
                    max.cutoff = 'q90', 
                    cells=cluster_cells ) 


p3 <- Seurat::FeaturePlot(seurat_object_l4, 
                    features = "pct_mt", 
                    reduction="umap", 
                    pt.size = pt_size/3, 
                    cells=cluster_cells ) 

p4 <- Seurat::FeaturePlot(seurat_object_l4, 
                    features = "pct_ribosomal", 
                    reduction="umap", 
                    pt.size = pt_size/3, 
                    cells=cluster_cells ) 


patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)

## INSPECT THE RELATIONSHIP BETWEEN QUALITY AND SUBCLUSTERS

### Combined plot
p1 <- Seurat::DimPlot(seurat_object_l4, 
        reduction = "umap", 
        pt.size = pt_size/3,  
        cols= palette, 
        group.by = subcluster_name, 
        cells=cluster_cells)

p2 <- Seurat::FeaturePlot(seurat_object_l4, 
                    features = "pct_mt", 
                    reduction="umap", 
                    pt.size = pt_size/3, 
                    cells=cluster_cells ) 

p3 <- Seurat::FeaturePlot(seurat_object_l4, 
                    features = "nCount_RNA", 
                    reduction="umap", 
                    pt.size = pt_size/3, 
                    min.cutoff = 'q10',
                    max.cutoff = 'q90', 
                    cells=cluster_cells ) 
                    

p4 <- Seurat::FeaturePlot(seurat_object_l4, 
                    features = "nFeature_RNA", 
                    reduction="umap", 
                    pt.size = pt_size/3, 
                    min.cutoff = 'q10',
                    max.cutoff = 'q90', 
                    cells=cluster_cells ) 

patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)

### Boxplot
df <- seurat_object_l4@meta.data[cluster_cells,]

g1 <- ggplot(df, aes_string(x=subcluster_name, y="nFeature_RNA", fill=subcluster_name)) + 
    geom_boxplot() + 
    theme(legend.position = "none") + 
    scale_fill_manual(values=palette) + 
    theme(axis.text.x = element_text(angle = 90))
g2 <- ggplot(df, aes_string(x=subcluster_name, y="nCount_RNA", fill=subcluster_name)) + 
    geom_boxplot() + 
    scale_fill_manual(values=palette) + 
    theme(axis.text.x = element_text(angle = 90))

patchwork::wrap_plots(g1, g2, ncol = 2)

dev.off()
```

```{r}
Idents(seurat_object_l4) <- subcluster_name
markers_df <- FindMarkers(
  seurat_object_l4,
  ident.1 = "9_3",
  ident.2 = c("9_0","9_1","9_2"),
  test.use = "wilcox",
  only.pos = TRUE,
  verbose = TRUE
)
markers <- markers_df %>%
  dplyr::arrange(desc(abs(avg_log2FC))) %>% 
  dplyr::filter(p_val_adj < 0.001 & avg_log2FC > 0.7) %>% # changed avg_log2FC > 0.9 after discussion with Ramon Massoni
  tibble::rownames_to_column(var = "gene")

markers
```

We see FTH1 and FTL as key marker gennes which were described as the most upregulated genes in cellular stress in "paper cold shock"

##### Remove cluster 9_3: Low quality
```{r}
Idents(seurat_object_l4) <- subcluster_name
seurat_object_l4_clean <- subset(
  seurat_object_l4,
  cells = colnames(seurat_object_l4)[Idents(seurat_object_l4) != "9_3"]
)
```


### Remove cluster 10: multiome specific
```{r}
table(seurat_object_l4_clean$level4_clusters, seurat_object_l4_clean$assay)
```


```{r}
Idents(seurat_object_l4_clean) <- "level4_clusters"
seurat_object_l4_clean <- subset(
  seurat_object_l4_clean,
  cells = colnames(seurat_object_l4_clean)[Idents(seurat_object_l4_clean) != "10"]
)
```


### Remove selected cells from level 2

```{r}
table(duplicated(colnames(seurat_object_l4_clean)))
table(duplicated(seurat_object_l4_clean$barcode))
seurat_object_l4_clean$barcode <- colnames(seurat_object_l4_clean)
```





## Reprocessing without low quality cells 
```{r}
seurat_list <- SplitObject(seurat_object_l4_clean, split.by = "assay")
seurat_list <- seurat_list[c("3P", "multiome")]
seurat_list <- purrr::map(
  seurat_list,
  FindVariableFeatures,
  nfeatures = 5000
)
hvg <- purrr::map(seurat_list, VariableFeatures)
shared_hvg <- intersect(hvg$`3P`, hvg$multiome)
# ElbowPlot(seurat_object_l4_clean, reduction = "harmony")
seurat_object_l4_clean <- seurat_object_l4_clean %>%
  ScaleData(features = shared_hvg) %>%
  RunPCA(features = shared_hvg) %>%
  RunHarmony(group.by.vars = "assay", reduction ="pca", dims = 1:30) %>%
  RunUMAP(reduction = "harmony", dims = 1:30)

resolution_vector <- c(0.01, 0.05, 0.1, 0.25, 0.35, 0.4, 0.45, 0.5, 0.75, 0.9, 1)
seurat_object_l4_clean <- seurat_object_l4_clean %>%
  FindNeighbors(dims = 1:30, reduction = "harmony")  %>%
  FindClusters(resolution = resolution_vector)
```

## Exploratory analysis of level 4 without low quality cells 

```{r}
seurat_object_l4_clean <- readRDS(paste0(data_folder, cell_type, "_clustered_clean_level_4_new.rds"))
```


### Data groups and features visualization

#### Data groups

```{r fig.height = 12, fig.width = 16}
pdf(paste(figures_folder, "cleaned_level4/level4_cleaned_exploratory_analysis_data_groups.pdf", sep=""),width=20,height=10,paper='special') 
data_groups_visualization(seurat_object = seurat_object_l4_clean, 
                          group_variable_1 = "level4_clusters",
                          pt_size = pt_size) 
dev.off()
```


#### Data features

```{r fig.height = 12, fig.width = 16}
pdf(paste(figures_folder, "cleaned_level4/level4_cleaned_exploratory_analysis_data_variables.pdf", sep=""),width=20,height=10,paper='special') 
data_features_visualization(seurat_object = seurat_object_l4_clean, 
                            group_variable_1 = "level4_clusters",  
                            pt_size = pt_size) 
dev.off()
```

### Selection of the resolution 

#### Clusters 
```{r fig.height = 12, fig.width = 16}
pdf(paste(figures_folder, "cleaned_level4/clusters_different_resolutions_all.pdf", sep=""),width=20,height=10,paper='special') 
p1 <- Seurat::DimPlot(seurat_object_l4_clean, 
        reduction = "umap",
        pt.size = pt_size/3,  
        cols= palette, 
        group.by = "annotation_level_3") 
print(p1)

list_plots <- list()
for (index in 1:length(resolution_vector)) {
  resolution <- resolution_vector[index]
  resolution_wanted <- paste0("RNA_snn_res.", resolution)
  Idents(seurat_object_l4_clean) <- resolution_wanted
  p <- Seurat::DimPlot(seurat_object_l4_clean, reduction = "umap",  pt.size = pt_size/3, cols = palette, group.by = resolution_wanted)
  list_plots[[index]] <- p
}
list_plots[[(length(resolution_vector) + 1)]] <- p1


print(patchwork::wrap_plots(list_plots))
dev.off()

pdf(paste(figures_folder, "cleaned_level4/clusters_different_resolutions.pdf", sep=""),width=20,height=10,paper='special') 
p1 <- Seurat::DimPlot(seurat_object_l4_clean, 
        reduction = "umap",
        pt.size = pt_size/3,  
        cols= palette, 
        group.by = "annotation_level_3") 
print(p1)

resolution_vector <- c(0.4, 0.45, 0.5)
list_plots <- list()
for (index in 1:length(resolution_vector)) {
  resolution <- resolution_vector[index]
  resolution_wanted <- paste0("RNA_snn_res.", resolution)
  Idents(seurat_object_l4_clean) <- resolution_wanted
  p <- Seurat::DimPlot(seurat_object_l4_clean, reduction = "umap",  pt.size = pt_size/3, cols = palette, group.by = resolution_wanted)
  list_plots[[index]] <- p
}
list_plots[[(length(resolution_vector) + 1)]] <- p1


print(patchwork::wrap_plots(list_plots))
dev.off()

```

```{r}
pdf(paste(figures_folder, "cleaned_level4/selected_resolution_correspondence_level_3.pdf", sep=""),width=20,height=10,paper='special') 
p1 <- Seurat::DimPlot(seurat_object_l4_clean, 
        reduction = "umap",
        pt.size = pt_size/3,  
        cols= palette, 
        group.by = "annotation_level_3", 
        raster = FALSE) 
print(p1)

p2 <- Seurat::DimPlot(seurat_object_l4_clean, 
        reduction = "umap",
        pt.size = pt_size/3,  
        cols= palette, 
        group.by = "RNA_snn_res.0.45", 
        raster = FALSE) 
print(p2)

print(patchwork::wrap_plots(p1, p2, ncol = 2))

cell_type_pcts_df <- seurat_object_l4_clean@meta.data %>%
  group_by(RNA_snn_res.0.45, annotation_level_3) %>%
  summarise(n_cells_cell_type = n()) %>% 
  group_by(RNA_snn_res.0.45) %>%
  mutate(
    total_cells = sum(n_cells_cell_type),
    pct_cells = n_cells_cell_type / total_cells * 100
  )

cell_type_pcts_gg <- cell_type_pcts_df %>%
  ggplot(aes(RNA_snn_res.0.45, pct_cells, fill = annotation_level_3)) +
    geom_col() +
    labs(x = "Cluster", y = "Percentage of cells (%)", fill = "") +
    scale_fill_manual(values = palette) +
    theme_classic()

print(cell_type_pcts_gg)

dev.off()

pdf(paste(figures_folder, "cleaned_level4/selected_resolution_quality_analysis.pdf", sep=""),width=20,height=10,paper='special') 
data_features_visualization(seurat_object = seurat_object_l4_clean, 
                            group_variable_1 = "RNA_snn_res.0.45",  
                            pt_size = pt_size) 
dev.off()

```

```{r}
pdf(paste(figures_folder, "cleaned_level4/comparison_with_annotation_level_3.pdf", sep=""),width=20,height=10,paper='special') 
cell_type_pcts_df <- seurat_object_l4@meta.data %>%
  group_by(RNA_snn_res.0.45, annotation_level_3) %>%
  summarise(n_cells_cell_type = n()) %>% 
  group_by(RNA_snn_res.0.45) %>%
  mutate(
    total_cells = sum(n_cells_cell_type),
    pct_cells = n_cells_cell_type / total_cells * 100
  )

cell_type_pcts_gg <- cell_type_pcts_df %>%
  ggplot(aes(RNA_snn_res.0.45, pct_cells, fill = annotation_level_3)) +
    geom_col() +
    labs(x = "Cluster", y = "Percentage of cells (%)", fill = "") +
    scale_fill_manual(values = palette) +
    theme_classic()

print(cell_type_pcts_gg)
dev.off()
```

## Find desired clusters after discussion with annotation team

### Subcluster cluster 1  (CENPE, HELLS)
```{r}
cluster_name <- "1"
subcluster_name <- paste0("cluster",cluster_name, "_subcluster")
```

```{r}
Idents(GCBC_seurat_object_level_4) <- "RNA_snn_res.0.45"
GCBC_seurat_object_level_4_to_l5 <- FindSubCluster(
  GCBC_seurat_object_level_4,
  cluster = cluster_name,
  graph.name = "RNA_snn",
  subcluster.name = subcluster_name,
  resolution = 0.5
)
table(GCBC_seurat_object_level_4_to_l5@meta.data[,subcluster_name])
cluster_cells <- Cells(GCBC_seurat_object_level_4_to_l5)[which(GCBC_seurat_object_level_4_to_l5$RNA_snn_res.0.45 == cluster_name)]

```

```{r}
pdf(paste(figures_folder, "subclustering_for_l5/",cluster_name,"_subclustering.pdf", sep=""),width=20,height=10,paper='special') 
p1 <- Seurat::DimPlot(GCBC_seurat_object_level_4_to_l5, 
        reduction = "umap", 
        pt.size = pt_size/2,  
        cols= c(palette, palette),
        group.by = subcluster_name, 
        raster = FALSE) 
p2 <- Seurat::FeaturePlot(GCBC_seurat_object_level_4_to_l5, 
                    pt.size = pt_size/2,  
                    features = c("CENPE", "HELLS"), 
                    reduction="umap", 
                    ncol = 1,
                    #cells=cluster_cells, 
                    raster = FALSE) 
patchwork::wrap_plots(p1, p2, ncol = 2)

dev.off()
```




```{r}
GCBC_seurat_object_level_4_to_l5$cluster1_subcluster <- ifelse(GCBC_seurat_object_level_4_to_l5$cluster1_subcluster %in% c("1_0", "1_1", "1_3", "1_6"), "1_0", 
                                                               ifelse(GCBC_seurat_object_level_4_to_l5$cluster1_subcluster %in% c("1_2", "1_4", "1_5"), "1_1",
                                                                      GCBC_seurat_object_level_4_to_l5$cluster1_subcluster)
                                                               )

pdf(paste(figures_folder, "subclustering_for_l5/",cluster_name,"_subclustering_changed.pdf", sep=""),width=20,height=10,paper='special') 
p1 <- Seurat::DimPlot(GCBC_seurat_object_level_4_to_l5, 
        reduction = "umap", 
        pt.size = pt_size/2,  
        cols= c(palette, palette),
        group.by = subcluster_name, 
        raster = FALSE) 
p2 <- Seurat::FeaturePlot(GCBC_seurat_object_level_4_to_l5, 
                    pt.size = pt_size/2,  
                    features = c("CENPE", "HELLS"), 
                    reduction="umap", 
                    ncol = 1,
                    #cells=cluster_cells, 
                    raster = FALSE) 
patchwork::wrap_plots(p1, p2, ncol = 2)

dev.off()
```





### Subcluster cluster 2  (CCNB, STMN1, FOXP1)

```{r}
cluster_name <- "2"
subcluster_name <- paste0("cluster",cluster_name, "_subcluster")
```

```{r}
Idents(GCBC_seurat_object_level_4_to_l5) <- "cluster1_subcluster"
GCBC_seurat_object_level_4_to_l5 <- FindSubCluster(
  GCBC_seurat_object_level_4_to_l5,
  cluster = cluster_name,
  graph.name = "RNA_snn",
  subcluster.name = subcluster_name,
  resolution = 0.2
)
table(GCBC_seurat_object_level_4_to_l5@meta.data[,subcluster_name])
cluster_cells <- Cells(GCBC_seurat_object_level_4_to_l5)[which(GCBC_seurat_object_level_4_to_l5$RNA_snn_res.0.45 == cluster_name)]

```

```{r}
pdf(paste(figures_folder, "subclustering_for_l5/",cluster_name,"_subclustering.pdf", sep=""),width=20,height=10,paper='special') 
p1 <- Seurat::DimPlot(GCBC_seurat_object_level_4_to_l5, 
        reduction = "umap", 
        pt.size = pt_size/2,  
        cols= c(palette, palette),
        group.by = subcluster_name, 
        raster = FALSE) 
p2 <- Seurat::FeaturePlot(GCBC_seurat_object_level_4_to_l5, 
                    pt.size = pt_size/2,  
                    features = c("CCNB2", "STMN1", "FOXP1"), 
                    reduction="umap", 
                    ncol = 1,
                    #cells=cluster_cells, 
                    raster = FALSE) 
patchwork::wrap_plots(p1, p2, ncol = 2)

dev.off()
```


```{r}
GCBC_seurat_object_level_4_to_l5$cluster2_subcluster <- ifelse(GCBC_seurat_object_level_4_to_l5$cluster2_subcluster %in% c("2_0","2_1"), "2_0", 
                                                               ifelse(GCBC_seurat_object_level_4_to_l5$cluster2_subcluster %in% c("2_2"), "2_1",
                                                                      GCBC_seurat_object_level_4_to_l5$cluster2_subcluster)
                                                               )

pdf(paste(figures_folder, "subclustering_for_l5/",cluster_name,"_subclustering_changed.pdf", sep=""),width=20,height=10,paper='special') 
p1 <- Seurat::DimPlot(GCBC_seurat_object_level_4_to_l5, 
        reduction = "umap", 
        pt.size = pt_size/2,  
        cols= c(palette, palette),
        group.by = subcluster_name, 
        raster = FALSE) 
p2 <- Seurat::FeaturePlot(GCBC_seurat_object_level_4_to_l5, 
                    pt.size = pt_size/2,  
                    features = c("CENPE", "HELLS"), 
                    reduction="umap", 
                    ncol = 1,
                    #cells=cluster_cells, 
                    raster = FALSE) 
patchwork::wrap_plots(p1, p2, ncol = 2)

dev.off()
```


### Subcluster cluster 3  

```{r}
cluster_name <- "3"
subcluster_name <- paste0("cluster",cluster_name, "_subcluster")
```

```{r}
Idents(GCBC_seurat_object_level_4_to_l5) <- "cluster2_subcluster"
GCBC_seurat_object_level_4_to_l5 <- FindSubCluster(
  GCBC_seurat_object_level_4_to_l5,
  cluster = cluster_name,
  graph.name = "RNA_snn",
  subcluster.name = subcluster_name,
  resolution = 0.25
)
table(GCBC_seurat_object_level_4_to_l5@meta.data[,subcluster_name])
cluster_cells <- Cells(GCBC_seurat_object_level_4_to_l5)[which(GCBC_seurat_object_level_4_to_l5$RNA_snn_res.0.45 == cluster_name)]

```

```{r}
pdf(paste(figures_folder, "subclustering_for_l5/",cluster_name,"_subclustering.pdf", sep=""),width=20,height=10,paper='special') 
p1 <- Seurat::DimPlot(GCBC_seurat_object_level_4_to_l5, 
        reduction = "umap", 
        pt.size = pt_size/2,  
        cols= c(palette, palette),
        group.by = subcluster_name, 
        raster = FALSE) 
p2 <- Seurat::FeaturePlot(GCBC_seurat_object_level_4_to_l5, 
                    pt.size = pt_size/2,  
                    features = c("MIR155HG", "MYC", "BCL2A1"), 
                    reduction="umap", 
                    ncol = 1,
                    #cells=cluster_cells, 
                    raster = FALSE) 
patchwork::wrap_plots(p1, p2, ncol = 2)

dev.off()
```


### Subcluster cluster 5  

```{r}
cluster_name <- "5"
subcluster_name <- paste0("cluster",cluster_name, "_subcluster")
```

```{r}
Idents(GCBC_seurat_object_level_4_to_l5) <- "cluster3_subcluster"
GCBC_seurat_object_level_4_to_l5 <- FindSubCluster(
  GCBC_seurat_object_level_4_to_l5,
  cluster = cluster_name,
  graph.name = "RNA_snn",
  subcluster.name = subcluster_name,
  resolution = 0.2
)
table(GCBC_seurat_object_level_4_to_l5@meta.data[,subcluster_name])
cluster_cells <- Cells(GCBC_seurat_object_level_4_to_l5)[which(GCBC_seurat_object_level_4_to_l5$RNA_snn_res.0.45 == cluster_name)]

```

```{r}
pdf(paste(figures_folder, "subclustering_for_l5/",cluster_name,"_subclustering.pdf", sep=""),width=20,height=10,paper='special') 
p1 <- Seurat::DimPlot(GCBC_seurat_object_level_4_to_l5, 
        reduction = "umap", 
        pt.size = pt_size/2,  
        cols= c(palette, palette),
        group.by = subcluster_name, 
        raster = FALSE) 
p2 <- Seurat::FeaturePlot(GCBC_seurat_object_level_4_to_l5, 
                    pt.size = pt_size/2,  
                    features = c("BCL2A1"), 
                    reduction="umap", 
                    ncol = 1,
                    #cells=cluster_cells, 
                    raster = FALSE) 
patchwork::wrap_plots(p1, p2, ncol = 2)

dev.off()
```


```{r}
cluster_name <- "5_0"
subcluster_name <- paste0("cluster",cluster_name, "_subcluster")
```

```{r}
Idents(GCBC_seurat_object_level_4_to_l5) <- "cluster5_subcluster"
GCBC_seurat_object_level_4_to_l5 <- FindSubCluster(
  GCBC_seurat_object_level_4_to_l5,
  cluster = cluster_name,
  graph.name = "RNA_snn",
  subcluster.name = subcluster_name,
  resolution = 0.2
)
table(GCBC_seurat_object_level_4_to_l5@meta.data[,subcluster_name])
cluster_cells <- Cells(GCBC_seurat_object_level_4_to_l5)[which(GCBC_seurat_object_level_4_to_l5$RNA_snn_res.0.45 == cluster_name)]

```

```{r}
pdf(paste(figures_folder, "subclustering_for_l5/",cluster_name,"_subclusterin.pdf", sep=""),width=20,height=10,paper='special') 
p1 <- Seurat::DimPlot(GCBC_seurat_object_level_4_to_l5, 
        reduction = "umap", 
        pt.size = pt_size/2,  
        cols= c(palette, palette),
        group.by = subcluster_name, 
        raster = FALSE) 
p2 <- Seurat::FeaturePlot(GCBC_seurat_object_level_4_to_l5, 
                    pt.size = pt_size/2,  
                    features = c("BCL2A1"), 
                    reduction="umap", 
                    ncol = 1,
                    #cells=cluster_cells, 
                    raster = FALSE) 
patchwork::wrap_plots(p1, p2, ncol = 2)

dev.off()
```



### Subcluster cluster 2_1  
```{r}
cluster_name <- "2_1"
subcluster_name <- paste0("cluster",cluster_name, "_subcluster")
```

```{r}
Idents(GCBC_seurat_object_level_4) <- "cluster5_subcluster"
GCBC_seurat_object_level_4 <- FindSubCluster(
  GCBC_seurat_object_level_4,
  cluster = cluster_name,
  graph.name = "RNA_snn",
  subcluster.name = subcluster_name,
  resolution = 0.2
)
table(GCBC_seurat_object_level_4@meta.data[,subcluster_name])

```

```{r}
pdf(paste(figures_folder, "subclustering_for_l5/",cluster_name,"_subclustering.pdf", sep=""),width=20,height=10,paper='special') 
p1 <- Seurat::DimPlot(GCBC_seurat_object_level_4, 
        reduction = "umap", 
        pt.size = pt_size/2,  
        cols= c(palette, palette),
        group.by = subcluster_name, 
        raster = FALSE) 
print(p1)
dev.off()
```

### Subcluster cluster 7  
```{r}
cluster_name <- "7"
subcluster_name <- paste0("cluster",cluster_name, "_subcluster")
```

```{r}
Idents(GCBC_seurat_object_level_4) <- "cluster2_1_subcluster"
GCBC_seurat_object_level_4 <- FindSubCluster(
  GCBC_seurat_object_level_4,
  cluster = cluster_name,
  graph.name = "RNA_snn",
  subcluster.name = subcluster_name,
  resolution = 0.2
)
table(GCBC_seurat_object_level_4@meta.data[,subcluster_name])

```

```{r}
pdf(paste(figures_folder, "subclustering_for_l5/",cluster_name,"_subclustering.pdf", sep=""),width=20,height=10,paper='special') 
p1 <- Seurat::DimPlot(GCBC_seurat_object_level_4, 
        reduction = "umap", 
        pt.size = pt_size/2,  
        cols= c(palette, palette),
        group.by = subcluster_name, 
        raster = FALSE) 
print(p1)
dev.off()
```

### Subcluster cluster 8  
```{r}
cluster_name <- "8"
subcluster_name <- paste0("cluster",cluster_name, "_subcluster")
```

```{r}
Idents(GCBC_seurat_object_level_4) <- "cluster7_subcluster"
GCBC_seurat_object_level_4 <- FindSubCluster(
  GCBC_seurat_object_level_4,
  cluster = cluster_name,
  graph.name = "RNA_snn",
  subcluster.name = subcluster_name,
  resolution = 0.2
)
table(GCBC_seurat_object_level_4@meta.data[,subcluster_name])

```

```{r}
pdf(paste(figures_folder, "subclustering_for_l5/",cluster_name,"_subclustering.pdf", sep=""),width=20,height=10,paper='special') 
p1 <- Seurat::DimPlot(GCBC_seurat_object_level_4, 
        reduction = "umap", 
        pt.size = pt_size/2,  
        cols= c(palette, palette),
        group.by = subcluster_name, 
        raster = FALSE) 
print(p1)
dev.off()
```


```{r}
saveRDS(GCBC_seurat_object_level_4, paste0(data_folder, cell_type, "_clustered_clean_level_4_with_final_clusters.rds"))
```

## Annotate clusters: 

### BETA 
```{r}
GCBCseurat_object_l5 <- GCBC_seurat_object_level_4
```


#### Cluster names
```{r}
library(plyr)
GCBCseurat_object_l5$numbers_level_5_clusters_beta <- GCBCseurat_object_l5$cluster8_subcluster
GCBCseurat_object_l5$names_level_5_clusters_beta <- revalue(GCBCseurat_object_l5$numbers_level_5_clusters_beta,
                                      c("0"= "DZ/LZ",
                                        "1_0"="DZ_Sphase_HistoneHigh",
                                        "1_1"="DZ_G2M_HistoneHigh",
                                        "2_0"="DZ-nonproliferative_FOXP1hi",
                                        "2_1_0"="DZ-nonproliferative",
                                        "2_1_1"="DZ-cell cycle exit",
                                        "3_0"="LZ",
                                        "3_1"="LZ-DZ-re-entry early commitment",
                                        "3_2"="LZ-BCL2A1 neg",
                                        "4"="DZ_Sphase",
                                        "5_0"="LZ-proliferative_BCL2A1neg",
                                        "5_1"="LZ-proliferative_BCL2A1pos",
                                        "6"="DZ_G2M_CCNBHigh",
                                        "7_0"="MBC-like_nonproli",
                                        "7_1"="MBC-like_proli1",
                                        "8_0"="MBC-like_FCRL4+",
                                        "8_1"="MBC-like_proli2",
                                        "8_2"="MBC-like_proli3",
                                        "9"="PC-precursors"))
```

```{r}
pdf(paste(figures_folder, "cluster_visualization/cluster_names.pdf", sep=""),width=20,height=10,paper='special') 
Seurat::DimPlot(GCBCseurat_object_l5, 
        reduction = "umap", 
        pt.size = pt_size,  
        cols= c(palette, palette), 
        label = TRUE,
        group.by = "names_level_5_clusters_beta", 
        raster = FALSE) + NoLegend()

Seurat::DimPlot(GCBCseurat_object_l5, 
        reduction = "umap", 
        pt.size = pt_size,  
        cols= c(palette, palette), 
        group.by = "names_level_5_clusters_beta", 
        raster = FALSE) 
dev.off()
```

### ETA 
#### Cluster names
```{r}
GCBCseurat_object_l5@meta.data <- GCBCseurat_object_l5@meta.data %>% mutate(names_level_5_clusters_eta =
  case_when(names_level_5_clusters_beta  == "DZ_Sphase" ~ "DZ early Sphase",
  names_level_5_clusters_beta  == "DZ_Sphase_HistoneHigh" ~ "DZ late Sphase",
  names_level_5_clusters_beta  == "DZ_G2M_HistoneHigh" ~ "DZ early G2Mphase",
  names_level_5_clusters_beta  == "DZ_G2M_CCNBHigh"~ "DZ late G2Mphase",
  
  names_level_5_clusters_beta  == "DZ-cell cycle exit" ~ "DZ cell cycle exit",
  names_level_5_clusters_beta  == "DZ-nonproliferative" ~ "DZ cell cycle exit",
  names_level_5_clusters_beta  ==  "DZ-nonproliferative_FOXP1hi"~ "DZ non proliferative",
  
  names_level_5_clusters_beta  == "DZ/LZ" ~ "DZ_LZ transition", 
  
  names_level_5_clusters_beta  == "LZ" ~ "LZ", 
  names_level_5_clusters_beta  == "LZ-BCL2A1 neg"~ "LZ",
  names_level_5_clusters_beta  == "LZ-DZ-re-entry early commitment" ~ "LZ_DZ reentry commitment",
  names_level_5_clusters_beta  ==  "LZ-proliferative_BCL2A1pos" ~ "LZ proliferative",
  names_level_5_clusters_beta  == "LZ-proliferative_BCL2A1neg" ~ "LZ_DZ transition",
  
  names_level_5_clusters_beta  == "MBC-like_nonproli" ~ "Precursor MBCs",
  names_level_5_clusters_beta  == "MBC-like_proli1" ~ "Precursor MBCs",
  names_level_5_clusters_beta  == "MBC-like_proli2"~ "Reactivated proliferative MBCs",
  names_level_5_clusters_beta  == "MBC-like_proli3" ~ "Reactivated proliferative MBCs",
  names_level_5_clusters_beta  == "MBC-like_FCRL4+"~ "Reactivated proliferative MBCs",
  
  names_level_5_clusters_beta  == "PC-precursors" ~ "prePC"))
```

```{r}
GCBCseurat_object_l5@meta.data <- GCBCseurat_object_l5@meta.data %>% mutate(names_level_5_clusters_eta_atac = 
    case_when(names_level_5_clusters_eta  == "DZ early Sphase" ~ "DZ Sphase",
              names_level_5_clusters_eta  == "DZ late Sphase" ~ "DZ Sphase",
              names_level_5_clusters_eta  == "DZ early G2Mphase" ~ "DZ G2Mphase",
              names_level_5_clusters_eta  == "DZ late G2Mphase" ~ "DZ G2Mphase",
              names_level_5_clusters_eta  == "DZ cell cycle exit" ~ "DZ non proliferative",
              names_level_5_clusters_eta  == "DZ non proliferative" ~ "DZ non proliferative",
              names_level_5_clusters_eta  == "DZ_LZ transition" ~ "DZ_LZ transition",
              names_level_5_clusters_eta  == "LZ" ~ "LZ",
              names_level_5_clusters_eta  == "LZ_DZ reentry commitment" ~ "LZ-DZ-re-entry",
              names_level_5_clusters_eta  == "LZ_DZ transition" ~ "LZ-DZ-re-entry",
              names_level_5_clusters_eta  == "LZ proliferative" ~ "LZ-DZ-re-entry", 
              names_level_5_clusters_eta  == "Precursor MBC" ~ "MBC",
              names_level_5_clusters_eta  == "Reactivated proliferative MBCs" ~ "MBC",
              names_level_5_clusters_eta  == "prePC" ~ "prePC"
      ))
```

```{r}
#saveRDS(GCBCseurat_object_l5, paste0(data_folder, cell_type, "_seu_obj_level_5_eta.rds"))
```




