---
title: "Normalization and Batch Effect Correction: MCL"
author: "Ramon Massoni-Badosa"
date: "`r format(Sys.time(), '%B %Y')`"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```


# Objective

The objective of this notebook is to perform pre-processing and dimensional reduction on both assays independently, using standard approaches for RNA and ATAC-seq data. Then, we will follow the ["Joint RNA and ATAC analysis: 10x multiomic"](https://satijalab.org/signac/articles/pbmc_multiomic.html) vignette from Signac to obtain a joint visualization using both modalities.


# Pre-processing

## Load packages

```{r}
library(Signac)
library(Seurat)
library(GenomicRanges)
library(future)
library(SeuratWrappers)
library(harmony)
library(EnsDb.Hsapiens.v86)
library(stringr)
library(dplyr)
library(ggplot2)

set.seed(173)
```


## Parameters

```{r}
path_to_102 <- "~/Desktop/data_MCL/results/R_objects/seurat_102_filtered.rds"
path_to_save_102 <- "~/Desktop/data_MCL/results/R_objects/seurat_102_filtered_normalized.rds"
```


## Load Multiome filtered data

```{r}
seurat_102 <- readRDS(path_to_102)
seurat_102
```


# Without Harmony integration

## scATAC

### Normalization and linear dimensional reduction

```{r}
DefaultAssay(seurat_102) <- "ATAC"
seurat_102 <- RunTFIDF(seurat_102)
seurat_102 <- FindTopFeatures(seurat_102, min.cutoff = "q0")
seurat_102 <- RunSVD(seurat_102)
```

```{r fig.wide=TRUE}
DepthCor(seurat_102)
```


### UMAP representation

```{r fig.wide=TRUE}
seurat_102 <- RunUMAP(
  seurat_102,
  dims = 2:40,
  reduction = "lsi",
  reduction.name = "umap.atac",
  reduction.key = "atacUMAP_"
)
DimPlot(
  seurat_102,
  reduction = "umap.atac",
  group.by = "library_name",
  pt.size = 0.1
)
```


## scRNA

### Normalization and linear dimensional reduction

```{r}
DefaultAssay(seurat_102) <- "RNA"
seurat_102 <- NormalizeData(
  seurat_102,
  normalization.method = "LogNormalize",
  scale.factor = 1e4
)
seurat_102 <- seurat_102 %>%
  FindVariableFeatures(nfeatures = 3000) %>%
  ScaleData() %>% 
  RunPCA() 
```


### UMAP representation

```{r}
seurat_102 <- RunUMAP(
  seurat_102,
  dims = 1:30,
  reduction = "pca",
  reduction.name = "umap.rna",
  reduction.key = "rnaUMAP_"
)
DimPlot(
  seurat_102,
  reduction = "umap.rna",
  group.by = "library_name",
  pt.size = 0.1
)
```


## Joint

```{r}
seurat_102 <- FindMultiModalNeighbors(
  seurat_102,
  reduction.list = list("pca", "lsi"),
  dims.list = list(1:30, 2:40)
)
seurat_102 <- RunUMAP(
  seurat_102,
  nn.name = "weighted.nn",
  reduction.name = "wnn.umap",
  reduction.key = "wnnUMAP_"
)
DimPlot(
  seurat_102,
  reduction = "wnn.umap",
  group.by = "library_name",
  pt.size = 0.1
)
DimPlot(
  seurat_102,
  reduction = "wnn.umap",
  group.by = "library_name",
  split.by = "library_name",
  pt.size = 0.1
)
```


# With Harmony integration

## scATAC

```{r}
DefaultAssay(seurat_102) <- "ATAC"
seurat_102 <- RunHarmony(
  object = seurat_102,
  reduction = "lsi",
  dims = 2:40,
  group.by.vars = "gem_id",
  assay.use = "ATAC",
  project.dim = FALSE,
  reduction.save = "harmony_peaks"
)
```


### UMAP representation

```{r}
seurat_102 <- RunUMAP(
  seurat_102,
  dims = 2:40,
  reduction = "harmony_peaks",
  reduction.name = "umap.atac",
  reduction.key = "atacUMAP_"
)
DimPlot(
  seurat_102,
  reduction = "umap.atac",
  group.by = "library_name",
  pt.size = 0.1
)
```


## scRNA

```{r}
DefaultAssay(seurat_102) <- "RNA"
seurat_102 <- RunHarmony(
  object = seurat_102,
  reduction = "pca",
  dims = 1:30,
  group.by.vars = "gem_id",
  assay.use = "RNA",
  project.dim = FALSE,
  reduction.save = "harmony_RNA"
)
```


```{r}
seurat_102 <- RunUMAP(
  seurat_102,
  dims = 1:30,
  reduction = "harmony_RNA",
  reduction.name = "umap.rna",
  reduction.key = "rnaUMAP_"
)
DimPlot(
  seurat_102,
  reduction = "umap.rna",
  group.by = "library_name",
  pt.size = 0.1
)
```


## Joint

```{r}
seurat_102 <- FindMultiModalNeighbors(
  seurat_102,
  reduction.list = list("harmony_peaks", "harmony_RNA"),
  dims.list = list(2:40, 1:30)
)
seurat_102 <- RunUMAP(
  seurat_102,
  nn.name = "weighted.nn",
  reduction.name = "wnn.umap",
  reduction.key = "wnnUMAP_"
)
DimPlot(
  seurat_102,
  reduction = "wnn.umap",
  group.by = "library_name",
  pt.size = 0.1
)
```


# Save

We will save the resulting object and use it in the following notebook to exclude doublets:

```{r}
saveRDS(tonsil_filtered, path_to_save)
```


# Session Information

```{r}
sessionInfo()
```
