---
title: "Myeloid: Level 4"
author: "Ramon Massoni-Badosa"
date: "2021/05/13"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```

NOTE: from level 4/5 onwards I will execute most steps locally, as the datasets are not as large anymore.


# Introduction

Here, we will include the annotation of the myeloid cells (level 4).


## Load packages

```{r}
library(Seurat)
library(SeuratWrappers)
library(harmony)
library(clusterProfiler)
library(org.Hs.eg.db)
library(tidyverse)
```


## Parameters

```{r}
# Paths
path_to_obj <- "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/results/R_objects/level_3/myeloid/myeloid_clustered_level_3_with_pre_freeze.rds"
path_to_save <- "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/results/R_objects/level_4/myeloid/myeloid_annotated_level_4.rds"


# Colors
color_palette <-  c("#1CFFCE", "#90AD1C", "#C075A6", "#85660D", "#5A5156", "#AA0DFE",   
                    "#F8A19F", "#F7E1A0", "#1C8356", "#FEAF16", "#822E1C", "#C4451C",   
                    "#1CBE4F", "#325A9B", "#F6222E", "#FE00FA", "#FBE426", "#16FF32", 
                    "black",   "#3283FE", "#B00068", "#DEA0FD", "#B10DA1", "#E4E1E3",   
                    "#90AD1C", "#FE00FA", "#85660D", "#3B00FB", "#822E1C", "coral2", 
                    "#1CFFCE", "#1CBE4F", "#3283FE", "#FBE426", "#F7E1A0", "#325A9B",   
                    "#2ED9FF", "#B5EFB5", "#5A5156", "#DEA0FD", "#FEAF16", "#683B79",   
                    "#B10DA1", "#1C7F93", "#F8A19F", "dark orange", "#FEAF16", "#FBE426",  
                    "Brown")

# Functions
source("~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/bin/utils.R")
```


## Load data

```{r}
seurat <- readRDS(path_to_obj)
seurat
DimPlot(seurat, cols = color_palette)
```

# Subcluster


* Subcluster cluster 4 to distinguish between DC2 and DC4.
* Subcluster cluster 8 to distinguish proliferative myeloid from doublets.


```{r}
# Subcluster 4 (DC2 and DC4)
seurat <- FindSubCluster(
  seurat,
  cluster = "4",
  graph.name = "RNA_snn",
  subcluster.name = "DC2_DC4",
  resolution = 0.4
)
Idents(seurat) <- "DC2_DC4"
clusters_4 <- seurat$DC2_DC4 %>%
  unique() %>%
  str_subset("^4_") %>%
  sort()
markers_4 <- purrr::map(clusters_4, function(x) {
  group_1 <- clusters_4[which(clusters_4 == x)]
  group_2 <- clusters_4[which(clusters_4 != x)]
  df <- FindMarkers(
    seurat,
    ident.1 = group_1,
    ident.2 = group_2,
    only.pos = TRUE,
    logfc.threshold = 0.5,
    verbose = TRUE
  )
  df <- df %>%
    rownames_to_column(var = "gene") %>%
    arrange(desc(avg_log2FC))
  df
})
names(markers_4) <- clusters_4
DT::datatable(markers_4$`4_0`)
DT::datatable(markers_4$`4_1`)
DT::datatable(markers_4$`4_2`)
DT::datatable(markers_4$`4_3`)
DimPlot(seurat, group.by = "DC2_DC4", cols = color_palette)
```


* 4_0: DC2
* 4_1: doublets
* 4_2: DC4 
* 4_3: DC3


Cycling cells:

```{r}
seurat <- FindSubCluster(
  seurat,
  cluster = "8",
  graph.name = "RNA_snn",
  subcluster.name = "cycling",
  resolution = 0.1
)
DimPlot(seurat, group.by = "cycling", cols = color_palette)
FeaturePlot(seurat, "MS4A1")
Idents(seurat) <- "cycling"
```

* 8_0: doublets
* 8_1: Cycling


# Annotation

```{r}
seurat$annotation_level_3 <- case_when(
  seurat$cycling == "0" ~ "poor-quality",
  seurat$cycling == "1" ~ "SELENOP/APOC Macrophages",
  seurat$cycling == "2" ~ "poor-quality",
  seurat$cycling == "3" ~ "TBD_3_level_3",
  seurat$cycling == "4_0" ~ "DC2",
  seurat$cycling == "4_1" ~ "poor-quality",
  seurat$cycling == "4_2" ~ "DC4",
  seurat$cycling == "4_3" ~ "DC3",
  seurat$cycling == "5" ~ "TBD_5_level_3",
  seurat$cycling == "6" ~ "aDC",
  seurat$cycling == "7" ~ "TBD_7_level_3",
  seurat$cycling == "8_0" ~ "poor-quality",
  seurat$cycling == "8_1" ~ "Cycling",
  seurat$cycling == "9" ~ "Neutrophil Granulocytes",
  seurat$cycling == "10" ~ "TBD_10_level_3",
  seurat$cycling == "11" ~ "DC1 precursor",
  seurat$cycling == "12" ~ "DC5",
  seurat$cycling == "13" ~ "Monocytes",
  seurat$cycling == "14" ~ "DC1 mature",
  seurat$cycling == "15" ~ "M1 Macrophages",
  seurat$cycling == "16" ~ "poor-quality",
  seurat$cycling == "17" ~ "Mast cells",
  seurat$cycling == "18" ~ "TBD_18_level_3",
  seurat$cycling == "19" ~ "IL7R DC",
  seurat$cycling == "20" ~ "poor-quality",
  seurat$cycling == "21" ~ "poor-quality"
)
DimPlot(seurat, cols = color_palette, group.by = "annotation_level_3")
seurat <- subset(seurat, annotation_level_3 != "poor-quality")
seurat <- subset(seurat, donor_id != "BCP2")
DimPlot(seurat, cols = color_palette, group.by = "annotation_level_3")
Idents(seurat) <- "annotation_level_3"
```


# Dimensionality reduction

```{r}
seurat <- seurat %>%
  FindVariableFeatures(nfeatures = 3000) %>%
  ScaleData() %>%
  RunPCA() %>%
  RunHarmony(group.by.vars = "donor_id", reduction = "pca", dims = 1:30)
seurat <- RunUMAP(seurat, reduction = "harmony", dims = 1:30)
seurat <- FindNeighbors(seurat, reduction = "harmony", dims = 1:30)
DimPlot(seurat, cols = color_palette, group.by = "annotation_level_3")
```


# Recluster aDC

```{r}
Idents(seurat) <- "annotation_level_3"
seurat <- FindSubCluster(
  seurat,
  cluster = "aDC",
  graph.name = "RNA_snn",
  subcluster.name = "aDC",
  resolution = 0.3
)
DimPlot(seurat, group.by = "aDC", cols = color_palette)

Idents(seurat) <- "aDC"
clusters_aDC <- seurat$aDC %>%
  unique() %>%
  str_subset("^aDC_") %>%
  sort()
markers_aDC <- purrr::map(clusters_aDC, function(x) {
  group_1 <- clusters_aDC[which(clusters_aDC == x)]
  group_2 <- clusters_aDC[which(clusters_aDC != x)]
  df <- FindMarkers(
    seurat,
    ident.1 = group_1,
    ident.2 = group_2,
    only.pos = TRUE,
    logfc.threshold = 0.5,
    verbose = TRUE
  )
  df <- df %>%
    rownames_to_column(var = "gene") %>%
    arrange(desc(avg_log2FC))
  df
})
names(markers_aDC) <- clusters_aDC
DT::datatable(markers_aDC$aDC_0)
DT::datatable(markers_aDC$aDC_1)
DT::datatable(markers_aDC$aDC_2)
seurat$aDC[seurat$aDC == "aDC_0"]  <- "aDC3" #MT2A, CCL19
seurat$aDC[seurat$aDC == "aDC_1"]  <- "aDC1" #CD274 CD80 CD40
seurat$aDC[seurat$aDC == "aDC_2"]  <- "aDC2" #AIRE CCL22
```


# Subset and recluster macrophages

```{r}
seurat$annotation_level_3 <- seurat$aDC
Idents(seurat) <- "annotation_level_3"
DimPlot(seurat, cols = color_palette)
macrophages_levels <- c("TBD_3_level_3", "TBD_5_level_3", "TBD_7_level_3",
                        "TBD_10_level_3", "TBD_10_level_3", "SELENOP/APOC Macrophages")
macrophages <- subset(seurat, idents = macrophages_levels)
macrophages <- macrophages %>%
  FindVariableFeatures(nfeatures = 2500) %>%
  ScaleData() %>%
  RunPCA() %>%
  RunHarmony(group.by.vars = "donor_id", reduction = "pca", dims = 1:25)
macrophages <- RunUMAP(
  macrophages,
  reduction = "harmony",
  dims = 1:25,
  n.neighbors = 10
)
macrophages <- FindNeighbors(
  macrophages,
  reduction = "harmony",
  dims = 1:25,
  k.param = 20
)
DimPlot(macrophages, cols = color_palette)
macrophages <- FindClusters(macrophages, resolution = 0.4)
DimPlot(macrophages, cols = color_palette)
FeaturePlot(macrophages, "MS4A1")


# Markers
markers_macrophages <- FindAllMarkers(
  macrophages,
  only.pos = TRUE,
  logfc.threshold = 0.75
)
markers_macrophages <- markers_macrophages %>%
  group_by(cluster) %>%
  arrange(desc(avg_log2FC), .by_group = TRUE)
DT::datatable(markers_macrophages)
```


# Save

```{r}
# Macrophages
input_shiny_macrophages <- seurat2shiny(macrophages, slot = "data", reduction = "umap")
saveRDS(
  input_shiny_macrophages$expression,
  "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/results/R_objects/level_4/myeloid/macrophages_expression_to_shiny_app_level_4.rds"
)
saveRDS(
  input_shiny_macrophages$metadata,
  "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/results/R_objects/level_4/myeloid/macrophages_metadata_to_shiny_app_level_4.rds"
)
umap_level_4 <- DimPlot(macrophages, cols = color_palette, pt.size = 1)
ggsave(
  filename = "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/3-clustering/4-level_4/myeloid/tmp/macrophages_umap_level_4.png",
  plot = umap_level_4,
  width = 14,
  height = 12,
  units = "cm"
)
saveRDS(macrophages, "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/results/R_objects/level_4/myeloid/macrophages_clustered_level_4.rds")
markers_macrophages_list <- purrr::map(unique(markers_macrophages$cluster), function(x) {
  df <- markers_macrophages[markers_macrophages$cluster == x, ]
  df
})
names(markers_macrophages_list) <- unique(markers_macrophages$cluster)
openxlsx::write.xlsx(
  markers_macrophages_list,
  "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/3-clustering/4-level_4/myeloid/tmp/macrophages_markers.xlsx"
)


# umap_level_3 <- DimPlot(seurat, cols = color_palette, pt.size = 1)
ggsave(
  filename = "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/3-clustering/4-level_4/myeloid/tmp/myeloid_umap_level_4.png",
  plot = umap_level_4,
  width = 14,
  height = 12,
  units = "cm"
)
saveRDS(seurat, path_to_save)


# Shiny app
input_shiny <- seurat2shiny(seurat, slot = "data", reduction = "umap")
saveRDS(
  input_shiny$expression,
  "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/results/R_objects/level_4/myeloid/myeloid_expression_to_shiny_app_level_4.rds")
saveRDS(
  input_shiny$metadata,
  "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/results/R_objects/level_4/myeloid/myeloid_metadata_to_shiny_app_level_4.rds")
```


# Session Info

```{r}
sessionInfo()
```

