---
author: "Paula Soler-Vila & Ramon Massoni-Badosa"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
params:
  cell_type: PC
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
title <- paste("Visualize integration (level 2):", params$cell_type, sep = " ")
```

---
title: "`r title`"
---

# Introduction

The objectives of this notebook are:

- Visualize how well we could remove technical variability associated with the assay (scATAC or multiome). We should see a high degree of intermixing between the 2 assays.

- Plot several markers of potential doublets or problematic cells: 
       - scRNAseq doublets determined by 10X Multiome
       - Scrublet doublet predictions
       - nCount and nFeatures of scATAC
       - Chromatin Signature
       - Projection of the scRNA-seq clusters defined at level 3
       
The broader objective of level 2 is to eliminate most remaining doublets and poor-quality cells, as we will discuss in future notebooks. These visualizations will allow us to explore this. We will also plot their location in the UMAP obtained at level 1.


# Pre-processing

## Load packages

```{r}
library(Seurat)
library(tidyverse)
library(reshape2)
library(ggpubr)
```


## Parameters
```{r}
# Paths
path_to_obj <- str_c(
  here::here("scATAC-seq/results/R_objects/level_2/"),
  params$cell_type,
  "/",
  params$cell_type,
  "_integrated_level_2.rds",
  sep = ""
)

path_to_obj_RNA <- str_c(
  here::here("scRNA-seq/3-clustering/3-level_3/"),
  params$cell_type,
  "_clustered_level_3_with_pre_freeze.rds",
  sep = ""
)

path_to_doublets <- here::here("scRNA-seq/3-clustering/2-level_2/tmp/doublets_multiome_df_all.rds")

# Functions
source(here::here("scRNA-seq/bin/utils.R"))
```


## Load data

```{r}
# Seurat object
seurat <- readRDS(path_to_obj)
seurat

seurat_RNA <- readRDS(path_to_obj_RNA)
seurat_RNA

tonsil_RNA_annotation <- seurat_RNA@meta.data %>%
  rownames_to_column(var = "cell_barcode") %>%
  dplyr::filter(assay == "multiome") %>%
  dplyr::select("cell_barcode", "seurat_clusters")
```


### Visualization of the data
```{r fig.wide=TRUE}
p1 <- DimPlot(seurat,
  pt.size = 0.1)

p2 <- DimPlot(seurat_RNA,
  group.by = "seurat_clusters",
  pt.size = 0.1, label = T) + NoLegend()

p1 + p2
```

# Assess integration

## Technical variability |

```{r}
p_assay <- plot_split_umap(seurat, var = "assay")
p_assay
```

## Technical variability ||

```{r}
p_assay <- plot_split_umap(seurat, var = "hospital")
p_assay
```

## Biological variability | 

```{r}
p_assay <- plot_split_umap(seurat, var = "age_group")
p_assay
```

## Biological variability || 

```{r}
p_assay <- plot_split_umap(seurat, var = "sex")
p_assay
```

# Spot potential doublets

## Multiome scRNA-seq data

### Doublets detected in scRNA-seq data from Multiome.
Here, we can see the count of doublets detected by cell-type. Note that for the level1 annotation there are not doublets detected in the Naive and Memory B cell cluster. 

```{r fig.wide=TRUE}
multiome_doublets <- readRDS(path_to_doublets)

dfm = melt(table(multiome_doublets$cell_type))
dfm$value = as.numeric(as.character(dfm$value))
ggbarplot(dfm, x = "Var1", y = "value",
          fill = "Var1",   
          palette = "jco",
          sort.val = "desc",         
          sort.by.groups = FALSE,    
          x.text.angle = 90          
          )
```


```{r}
doublets_cells <- colnames(seurat)[which(colnames(seurat) %in% multiome_doublets$barcode)]
length(doublets_cells)
```

## Visualize the projection of the doublets cells onto the scATAC-seq UMAP

```{r fig.wide=TRUE}
DimPlot(
  seurat, reduction = "umap",
  cols.highlight = "darkred", cols= "grey",
  cells.highlight= doublets_cells,
  pt.size = 0.1
)

```
## Scrublet prediction

```{r}
# Scrublet
DimPlot(seurat, group.by = "scrublet_predicted_doublet_atac")
table(seurat$scrublet_predicted_doublet_atac)
```

## QC metrics

```{r}
qc_vars <- c(
  "nCount_peaks",
  "nFeature_peaks"
)
qc_gg <- purrr::map(qc_vars, function(x) {
  p <- FeaturePlot(seurat, features = x)
  p
})
qc_gg
```

## Annotation probability

```{r}
qc_vars <- c(
  "annotation_prob")
qc_gg <- purrr::map(qc_vars, function(x) {
  p <- FeaturePlot(seurat, features = x)
  p
})
qc_gg
```

## Chromatin Signature

```{r}
qc_vars <- c("NBC.MBC", "GCBC", "PC", "CD4.T", "Cytotoxic",
             "myeloid", "FDC", "PDC")
qc_gg <- purrr::map(qc_vars, function(x) {
  p <- FeaturePlot(seurat, feature = x, max.cutoff = 4, min.cutoff = -4) + scale_color_viridis_c(option = "magma")
  p
})
qc_gg
```

## Projection of the scRNA-seq clusters to scATAC-seq UMAP
Annotation level 3 for scATAC will be defined "a priori" as unannotated and the scRNA annotation will be transfered to the scATAC-multiome cells based on the same cell barcode.

```{r fig.wide=TRUE}
cell_testing <- c("NBC_MBC", "GCBC", "PC", "CD4_T","Cytotoxic")

if (unique(seurat$annotation_level_1) %in% cell_testing){
  seurat_df <- data.frame(cell_barcode = colnames(seurat)[seurat$assay == "scATAC"])
  seurat_df$seurat_clusters <- "unannotated"
  
  df_all <- rbind(tonsil_RNA_annotation,seurat_df)
  rownames(df_all) <- df_all$cell_barcode
  df_all <- df_all[colnames(seurat), ]
  
  seurat$seurat_clusters <- df_all$seurat_clusters
  seurat@meta.data$annotation_prob  <- 1
  melt(table(seurat$seurat_clusters))
  table(is.na(seurat$seurat_clusters))

  p3 <- DimPlot(seurat,
              group.by = "seurat_clusters",
              pt.size = 0.5) 
  p3
}

p2
```

## UMAP level 1

```{r}
umap_level_1 <- FeatureScatter(
  seurat,
  "UMAP_1_level_1",
  "UMAP_2_level_1",
  group.by = "annotation_level_1"
)
umap_level_1 <- umap_level_1 +
  theme(
    #legend.position = "none",
    plot.title = element_blank()
  )
umap_level_1
```


# Session Information

```{r}
sessionInfo()
```
