---
title: "Seurat harmony and WNN assess mixibility (ADT+GEX)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: Sonal 
output:
  html_document:
    toc: yes
    toc_float: yes
    number_sections: yes
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(Seurat)
library(dplyr)
library(ggplot2)
library(patchwork)
library(scales)
library(tidyverse)
library(harmony)
```

# Define Parameters

```{r}
# Paths
path_to_save_obj <- "../results"
path_to_saved_post_QC_obj <- str_c(
  path_to_save_obj,
  "seurat_object_cite_seq_seurat_doublets_excluded.rds",
  sep = "/"
)
path_to_save_seurat_wnn_obj <- str_c(
  path_to_save_obj,
  "seurat_object_cite_seq_seurat_wnn.rds",
  sep = "/"
)
saved_cell_cycle_obj <- str_c(
  "../data/cycle.rda",
  sep = "/"
)
```

## Read saved seurat object post QC

```{r}
filtered_bcll.combined <- readRDS(path_to_saved_post_QC_obj)
```

# ADT and RNA multimodal analysis

```{r}
DefaultAssay(filtered_bcll.combined) <- 'RNA'
filtered_bcll.combined <- NormalizeData(filtered_bcll.combined) %>% FindVariableFeatures() %>% ScaleData() %>% RunPCA()  %>%
  RunHarmony(reduction = "pca",dims = 1:30,group.by.vars = "gemid",assay.use = "RNA",reduction.save = "harmony_RNA")
```

# Check for Cell cycle stages variability between clusters

```{r}
load(saved_cell_cycle_obj)
filtered_bcll.combined <- CellCycleScoring(filtered_bcll.combined, 
                                 g2m.features = g2m_genes, 
                                 s.features = s_genes)
```


```{r}
DefaultAssay(filtered_bcll.combined) <- 'ADT'
# we will use all ADT features for dimensional reduction
# we set a dimensional reduction name to avoid overwriting the 
VariableFeatures(filtered_bcll.combined) <- rownames(filtered_bcll.combined[["ADT"]])
filtered_bcll.combined <- NormalizeData(filtered_bcll.combined, normalization.method = 'CLR', margin = 2) %>% 
  ScaleData() %>% RunPCA(reduction.name = 'apca') %>%
  RunHarmony(reduction = "apca",dims = 1:20, group.by.vars = "gemid", assay.use = "ADT", reduction.save = "harmony_protein")
```

# Multimodal Neighbor Identification

```{r}
# Identify multimodal neighbors. These will be stored in the neighbors slot, 
# and can be accessed using filtered_bcll.combined[['weighted.nn']]
# The WNN graph can be accessed at filtered_bcll.combined[["wknn"]], 
# and the SNN graph used for clustering at filtered_bcll.combined[["wsnn"]]
# Cell-specific modality weights can be accessed at filtered_bcll.combined$RNA.weight
filtered_bcll.combined <- FindMultiModalNeighbors(
  filtered_bcll.combined, reduction.list = list("harmony_RNA", "harmony_protein"),
  dims.list = list(1:30, 1:20), modality.weight.name = "RNA.weight"
)
```


# UMAP visualisation

```{r}
feat_clust <- c(0.1,0.5,1,1.5,2)
filtered_bcll.combined <- RunUMAP(filtered_bcll.combined, nn.name = "weighted.nn", reduction.name = "wnn.umap", reduction.key = "wnnUMAP_")
filtered_bcll.combined <- FindClusters(filtered_bcll.combined, graph.name = "wsnn", algorithm = 3, resolution = feat_clust, verbose = FALSE)
```

```{r}
Seurat::DimPlot(filtered_bcll.combined, reduction = 'wnn.umap', label = TRUE, repel = TRUE, label.size = 2.5, group.by = c("wsnn_res.0.1","wsnn_res.0.5","wsnn_res.1","wsnn_res.1.5","wsnn_res.2")) + NoLegend()  & Seurat::NoLegend()
```

```{r}
filtered_bcll.combined <- FindClusters(filtered_bcll.combined, graph.name = "wsnn", algorithm = 3, resolution = 0.1, verbose = FALSE)
filtered_bcll.combined <- RunUMAP(filtered_bcll.combined, reduction = 'pca', dims = 1:30, assay = 'RNA', 
              reduction.name = 'rna.umap', reduction.key = 'rnaUMAP_')
filtered_bcll.combined <- RunUMAP(filtered_bcll.combined, reduction = 'apca', dims = 1:18, assay = 'ADT', 
              reduction.name = 'adt.umap', reduction.key = 'adtUMAP_')

```


```{r}
p1 <- DimPlot(filtered_bcll.combined, reduction = 'rna.umap', label = TRUE, 
              repel = TRUE, label.size = 2.5) + NoLegend()
p2 <- DimPlot(filtered_bcll.combined, reduction = 'adt.umap', label = TRUE, 
              repel = TRUE, label.size = 2.5) + NoLegend()
```

```{r}
p1
```


```{r}
p2
```

```{r}
Seurat::DimPlot(filtered_bcll.combined, reduction = 'wnn.umap', label = TRUE, repel = TRUE, label.size = 2.5, group.by = "wsnn_res.0.1")
```

```{r}
metadata <- filtered_bcll.combined@meta.data
df = count(metadata,metadata$wsnn_res.0.1)
colnames(df) = c("Cluster","Number_of_cells")
library(knitr)
library(kableExtra)
kable(df) %>%
  kable_styling("striped", full_width = T)

```

# Check Batch effect 

## Check for subproject variability

```{r}
DimPlot(filtered_bcll.combined, reduction = "wnn.umap", group.by = 'subproject')
```

## Cell Cycle 

```{r}
# Plot the PCA colored by cell cycle phase
DimPlot(filtered_bcll.combined,
        reduction = "pca",
        group.by= "Phase",
        split.by = "Phase")
```

```{r}
DimPlot(filtered_bcll.combined, reduction = "wnn.umap", group.by = 'Phase')
```


## Scrublet Doublets

```{r}
DimPlot(filtered_bcll.combined, reduction = "wnn.umap", group.by = 'scrublet_predicted_doublet')
```


```{r}
FeaturePlot(filtered_bcll.combined, features = "scrublet_doublet_scores")
```

## Check vireo doublets

```{r}
DimPlot(filtered_bcll.combined, group.by = "genotype_based_doublet_flag", pt.size = 0.1)
```

## Unassigned to any donor 

```{r}
DimPlot(filtered_bcll.combined, group.by = "genotype_based_unassigned_flag", pt.size = 0.1)
```

## Plot UMAP with respect to TCR

```{r}
DimPlot(filtered_bcll.combined, group.by = "tcr_flag", pt.size = 0.1)
```

## Plot UMAP with respect to BCR

```{r}
DimPlot(filtered_bcll.combined, group.by = "bcr_flag", pt.size = 0.1)
```


## Plot MAIT cells

```{r}
DimPlot(filtered_bcll.combined, group.by = "mait_evidence", pt.size = 0.1)
```

## Plot iNKT cells

```{r}
DimPlot(filtered_bcll.combined, group.by = "inkt_evidence", pt.size = 0.1)
```

# RNA and ADT weight per cluster

## Violin plot to check the RNA weight

```{r}
Seurat::VlnPlot(filtered_bcll.combined, features = "RNA.weight", sort = TRUE, pt.size = 0.1) & Seurat::NoLegend()
```

## Violin plot to check the ADT weight

```{r}
Seurat::VlnPlot(filtered_bcll.combined, features = "ADT.weight", sort = TRUE, pt.size = 0.1) & Seurat::NoLegend()
```


## Save the seurat object 

```{r}
saveRDS(filtered_bcll.combined, file = path_to_save_seurat_wnn_obj)
```

# Session Info
```{r}
sessionInfo()
```
