---
title: 'CD4 T: scATAC annotation level 5'
author: "Paula Soler-Vila & Ramon Massoni-Badosa"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```

# Introduction

The objective of this notebook is to refine the clustering annotation done at level 4. This refinement is the result of a manual curation carried out by specialists to remove poor quality cells, misclassified cells or clusters with very few cells and redefine possible new clusters.

Taking into account the data from scRNAseq, we will proceed to:

* Remove of 1 cluster: mitochondrial+ T cells.
* Subcluster Follicular Th CXCL13+CBLB+ into 2 clusters

# Pre-processing

## Load packages

```{r}
library(Seurat)
library(SeuratWrappers)
library(Signac)
library(tidyverse)
library(reshape2)
library(ggpubr)
library(harmony)
```


## Parameters
```{r}
cell_type = "CD4_T"

# Paths
path_to_obj <- str_c(
  here::here("scATAC-seq/results/R_objects/level_4/"),
  cell_type,
  "/04.",
  cell_type,
  "_integration_peak_calling_level_4.rds",
  sep = ""
)

path_to_obj_RNA <- str_c(
  here::here("scRNA-seq/3-clustering/5-level_5/"),
  cell_type,
    "/",
  cell_type,
  "_subseted_integrated_level_5.rds",
  sep = ""
)

# Functions
source(here::here("scRNA-seq/bin/utils.R"))


# Colors
color_palette <-  c("#1CFFCE", "#90AD1C", "#C075A6", "#85660D", 
                    "#5A5156", "#AA0DFE", "#F8A19F", "#F7E1A0", 
                    "#1C8356", "#FEAF16", "#822E1C", "#C4451C", 
                    "#1CBE4F", "#325A9B", "#F6222E")


path_to_level_5 <- here::here("scATAC-seq/results/R_objects/level_5/CD4_T/")
path_to_save <- str_c(path_to_level_5, "01.CD4_T_integrated_level_5.rds")
```

## Load data
```{r}
# Seurat object
seurat <- readRDS(path_to_obj)
seurat

seurat_RNA <- readRDS(path_to_obj_RNA)
seurat_RNA
```

## Visualization of the data
```{r fig.wide=TRUE}
p1 <- DimPlot(seurat,
      group.by = "annotation_level_3",
      cols = color_palette,
      pt.size = 0.2)

p2 <- DimPlot(seurat_RNA,
  group.by = "annotation_level_5",
  pt.size = 0.1,cols = color_palette)

p1
p2
```

# Removing mitochondrial+ T cells.

```{r}
selected_cells_clusters <- colnames(seurat)[!(seurat$annotation_level_3 == "Mitochondrial+ T cells")]
seurat <- subset(seurat, cells = selected_cells_clusters)
seurat$annotation_level_3 <- droplevels(seurat$annotation_level_3)

table(seurat$annotation_level_3)
```

## Visualization after removing problematic cells

```{r fig.wide=TRUE}
DimPlot(seurat,
      group.by = "annotation_level_3",
      cols = color_palette,
      pt.size = 0.1)
```

# Integration

```{r fig.wide=TRUE}
seurat <- seurat %>%
  RunTFIDF() %>%
  FindTopFeatures(min.cutoff = 10) %>%
  RunSVD()

DepthCor(seurat)
seurat <- RunUMAP(object = seurat, reduction = 'lsi', dims = 2:40)

DimPlot(seurat,
      group.by = "annotation_level_3",
      cols = color_palette,
      pt.size = 0.2)

seurat <- RunHarmony(
  object = seurat,
  dims = 2:40,
  group.by.vars = 'assay',
  reduction = 'lsi',
  assay.use = 'peaks_redefined',
  project.dim = FALSE,
  max.iter.harmony = 20
)

seurat <- RunUMAP(seurat, reduction = "harmony", dims = 2:40)
seurat <- FindNeighbors(seurat, reduction = "harmony", dims = 2:40)

DimPlot(seurat,
      group.by = "annotation_level_3",
      cols = color_palette,
      pt.size = 0.2)
```

# Save
```{r}
saveRDS(seurat, path_to_save)
```

# Session Information

```{r}
sessionInfo()
```
