---
title: "FDC: Level 4"
author: "Ramon Massoni-Badosa"
date: "2021/08/18"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```

NOTE: from level 4/5 onwards I will execute most steps locally, as the datasets are not as large anymore.


# Introduction

Here, we will include the annotation of the FDC cells (level 4).


## Load packages

```{r}
library(Seurat)
library(SeuratWrappers)
library(harmony)
library(clusterProfiler)
library(org.Hs.eg.db)
library(tidyverse)
```


## Parameters

```{r}
# Paths
path_to_obj <- "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/results/R_objects/level_3/FDC/FDC_clustered_level_3_with_pre_freeze.rds"
path_to_save_proliferative <- "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/results/R_objects/level_4/FDC/FDC_proliferative_level_4.rds"
path_to_save_epithelial <- "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/results/R_objects/level_4/FDC/epithelial_level_4.rds"
path_to_save_fdc <- "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/results/R_objects/level_4/FDC/FDC_all_level_4.rds"


# Colors
color_palette <-  c("#1CFFCE", "#90AD1C", "#C075A6", "#85660D", "#5A5156", "#AA0DFE",   
                    "#F8A19F", "#F7E1A0", "#1C8356", "#FEAF16", "#822E1C", "#C4451C",   
                    "#1CBE4F", "#325A9B", "#F6222E", "#FE00FA", "#FBE426", "#16FF32", 
                    "black",   "#3283FE", "#B00068", "#DEA0FD", "#B10DA1", "#E4E1E3",   
                    "#90AD1C", "#FE00FA", "#85660D", "#3B00FB", "#822E1C", "coral2", 
                    "#1CFFCE", "#1CBE4F", "#3283FE", "#FBE426", "#F7E1A0", "#325A9B",   
                    "#2ED9FF", "#B5EFB5", "#5A5156", "#DEA0FD", "#FEAF16", "#683B79",   
                    "#B10DA1", "#1C7F93", "#F8A19F", "dark orange", "#FEAF16", "#FBE426",  
                    "Brown")

# Functions
source("~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/bin/utils.R")
```


## Load data

```{r}
seurat <- readRDS(path_to_obj)
seurat
DimPlot(seurat, cols = color_palette)
```

# Explore

```{r}
FeaturePlot(seurat, c("CD79A", "CD3D", "TOP2A"), pt.size = 0.5) &
  scale_color_viridis_c(option = "magma")
FeaturePlot(seurat, "MS4A1") +
  scale_color_viridis_c(option = "magma")
FeaturePlot(seurat, c("KRT15", "KRT19"), pt.size = 0.5) &
  scale_color_viridis_c(option = "magma")
```


# TODO

* Merge clusters 0 and 2. Cluster 2 has a poorly defined signature, and some markers of B cells. We will subcluster it to see if we can separate doublets and FDC.
* Remove cluster 4 (doublets with B cells).
* Remove cluster 6 (doublets: top half doublets with T cells, bottom half doublet with B cell).
* Divide cluster 3 in cycling (left) + doublets (right). Remove doublets
* Subset cluster 9 and merge it with epithelial cells (it expresses KRT19 and KRT15).


Important literature

* [Application of Tonsil-Derived Mesenchymal Stem Cells in Tissue Regeneration: Concise Review](https://stemcellsjournals.onlinelibrary.wiley.com/doi/10.1002/stem.3058)

# Remove

```{r}
seurat$annotation_level_3 <- as.character(seurat$seurat_clusters)
selected_cells <- colnames(seurat)[!(seurat$annotation_level_3 %in% c("4", "6"))]
seurat <- subset(seurat, cells = selected_cells)
DimPlot(seurat, group.by = "annotation_level_3", cols = color_palette)
```


# Subcluster

### Cluster 2

```{r}
Idents(seurat) <- "annotation_level_3"
seurat <- FindNeighbors(seurat, reduction = "harmony", dims = 1:20)
seurat <- FindSubCluster(seurat, "2", "RNA_snn", "sub2", resolution = 0.6)
DimPlot(seurat, group.by = "sub2", cols = color_palette)
Idents(seurat) <- "sub2"
markers_sub2 <- find_markers_subclusters(seurat, "sub2", "^2_")
DT::datatable(markers_sub2$`2_0`)
DT::datatable(markers_sub2$`2_1`)
DT::datatable(markers_sub2$`2_2`)
```


Clearly, cluster 2_2 expressed B cell markers, so we will remove it:

```{r}
seurat <- subset(seurat, sub2 != "2_2")
DimPlot(seurat, group.by = "sub2", cols = color_palette)
markers_2vs0 <- FindMarkers(seurat, ident.1 = c("2_0", "2_1"), ident.2 = "0", only.pos = TRUE)
DT::datatable(markers_2vs0)
```


New cluster 2 does not have distinctive markers when compared to cluster 0, so we will merge both:

```{r}
seurat$annotation_level_3[seurat$annotation_level_3 %in% c("0", "2")] <- "0_2"
Idents(seurat) <- "annotation_level_3"
DimPlot(seurat, group.by = "annotation_level_3")
```


### Cluster 3

```{r}
seurat <- FindNeighbors(seurat, reduction = "harmony", dims = 1:15)
seurat <- FindSubCluster(seurat, "3", "RNA_snn", "sub3", resolution = 0.3)
DimPlot(seurat, group.by = "sub3", cols = color_palette)
Idents(seurat) <- "sub3"
markers_sub3 <- FindMarkers(seurat, ident.1 = "3_1", ident.2 = c("3_0", "3_2"), only.pos = FALSE)
markers_sub3 <- markers_sub3 %>%
  rownames_to_column("gene") %>%
  arrange(desc(avg_log2FC))
DT::datatable(markers_sub3)
markers_sub3 <- find_markers_subclusters(seurat, "sub3", "^3_")
DT::datatable(markers_sub3$`3_0`)
DT::datatable(markers_sub3$`3_1`)
```

Since it is hard to separate, I propose to keep this cluster and label it "proliferative FDC/doublets". However, to gain more resolution in the annotation of the other clusters, we can subset it (similar to CD4 T cells):

```{r}
proliferative_doublets <- subset(seurat, annotation_level_3 == "3")
seurat <- subset(seurat, annotation_level_3 != "3")
DimPlot(seurat, group.by = "annotation_level_3")
```


# Get cluster 9 (epithelial)

```{r}
epithelial <- subset(seurat, annotation_level_3 == "9")
seurat <- subset(seurat, annotation_level_3 != "9")
DimPlot(epithelial, cols = color_palette)
Idents(seurat) <- "annotation_level_3"
DimPlot(seurat, cols = color_palette)
```

# Reintegrate

```{r}
seurat <- seurat %>%
  FindVariableFeatures() %>%
  ScaleData() %>%
  RunPCA() %>%
  RunHarmony(group.by.vars = "donor_id", reduction = "pca", dims = 1:20) %>%
  RunUMAP(reduction = "harmony", dims = 1:20, n.neighbors = 20)
DimPlot(seurat, group.by = "annotation_level_3", cols = color_palette)
```

# Recluster

```{r}
seurat <- FindNeighbors(
  seurat,
  dims = 1:20,
  reduction = "harmony",
  k.param = 20
)
seurat <- FindClusters(seurat, resolution = 0.7)
DimPlot(seurat, cols = color_palette)
```
Let us join clusters 1, 2 and 4 in a single cluster; and split cluster 3 in 3:

```{r}
# Merge
seurat$annotation_level_4 <- as.character(seurat$seurat_clusters)
seurat$annotation_level_4[seurat$annotation_level_4 %in% c("1", "2", "4")] <- "1"


# Split
Idents(seurat) <- "annotation_level_4"
seurat <- FindSubCluster(seurat, "3", "RNA_snn", "sub3_level_4", resolution = 0.7)
DimPlot(seurat, group.by = "sub3_level_4", cols = color_palette)

# Rename
seurat$annotation_level_4 <- case_when(
  seurat$sub3_level_4 == "0" ~ "0",
  seurat$sub3_level_4 == "1" ~ "1",
  seurat$sub3_level_4 == "3_0" ~ "2",
  seurat$sub3_level_4 == "3_1" ~ "3",
  seurat$sub3_level_4 == "3_2" ~ "4",
  seurat$sub3_level_4 == "5" ~ "5",
  seurat$sub3_level_4 == "6" ~ "6"
)
Idents(seurat) <- "annotation_level_4"
DimPlot(seurat, group.by = "annotation_level_4")
```


## Markers

```{r}
markers_new <- FindAllMarkers(seurat, only.pos = TRUE, logfc.threshold = 0.75)
```


# Generate outputs

```{r}
shiny_output <- seurat2shiny(seurat, assay = "RNA", reduction = "umap", slot = "data")
```


# Final annotation

* 0: COL27A1_DZ_FDC_or_FRC
* 1: LZ_FDC
* 2: precursor_FDC
* 3: KLF_FDC
* 4: T-MSC
* 5: doublets
* 6: CD14_CD55_FDC


# Save

```{r}
# Save Seurat objects
saveRDS(seurat, path_to_save_fdc)
saveRDS(epithelial, path_to_save_epithelial)
saveRDS(proliferative_doublets, path_to_save_proliferative)
```
```{r}
# Save shiny
saveRDS(
  shiny_output$metadata,
  "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/results/R_objects/level_4/FDC/FDC_metadata_to_shiny_app_level_4.rds"
)
saveRDS(shiny_output$expression, "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/results/R_objects/level_4/FDC/FDC_expression_to_shiny_app_level_4.rds")
```
```{r}
# Markers
markers_new_list <- purrr::map(unique(markers_new$cluster), function(x) {
  df <- markers_new[markers_new$cluster == x, ]
  df
})
names(markers_new_list) <- unique(markers_new$cluster)
markers_new_list <- markers_new_list[as.character(0:6)]
openxlsx::write.xlsx(x = markers_new_list, "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/3-clustering/4-level_4/FDC/tmp/markers_FDC_level_4.xlsx")
```
```{r}
# Save
ggsave(
  filename = "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/3-clustering/4-level_4/FDC/tmp/FDC_umap_level_4.png",
  plot = DimPlot(seurat, cols = color_palette),
  width = 14,
  height = 12,
  units = "cm"
)
```


# Session Information

```{r}
sessionInfo()
```

