---
title: "Explore doublets: Multiome"
author: "Paula Soler-Vila and Ramon Massoni-Badosa"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```


# Introduction

In this notebook we will plot the scrublet doublet scores on the UMAPs computed in the previous notebook.


# Pre-processing

## Load packages

```{r}
library(Seurat)
library(Signac)
library(tidyverse)
```


## Parameters

```{r}
# Paths
path_to_obj <- here::here("multiome/results/R_objects/5.tonsil_multiome_integrated_using_wnn.rds")


# Thresholds
max_doublet_score_rna <- 0.3
max_doublet_score_atac <- 0.3
```


## Load data

```{r}
tonsil <- readRDS(path_to_obj)
```


# Doublet score

## Scrublet doublet score (ATAC)

```{r fig.wide=TRUE}
hist_atac <- tonsil@meta.data %>%
  ggplot(aes(scrublet_doublet_scores_atac)) +
    geom_histogram(bins = 30) +
    geom_vline(
      xintercept = max_doublet_score_atac,
      linetype = "dashed",
      color = "red"
    ) +
    xlab("Doublet Score (ATAC)") +
    theme_bw() +
    theme(
      axis.title = element_text(size = 13),
      axis.text = element_text(size = 11)
    )
hist_atac
```


## Scrublet doublet score (RNA)

```{r fig.wide=TRUE}
hist_rna <- tonsil@meta.data %>%
  ggplot(aes(scrublet_doublet_scores)) +
    geom_histogram(bins = 30) +
    geom_vline(
      xintercept = max_doublet_score_rna,
      linetype = "dashed",
      color = "red"
    ) +
    xlab("Doublet Score (RNA)") +
    theme_bw() +
    theme(
      axis.title = element_text(size = 13),
      axis.text = element_text(size = 11)
    )
hist_rna
```


## Correlation

```{r fig.wide=TRUE}
scatter_plot <- tonsil@meta.data %>%
  ggplot(aes(scrublet_doublet_scores, scrublet_doublet_scores_atac)) +
    geom_point(size = 0.01, alpha = 0.75) +
    geom_vline(xintercept = max_doublet_score_rna, linetype = "dashed", color = "red") +
    geom_hline(yintercept = max_doublet_score_atac, linetype = "dashed", color = "red") +
    labs(x = "Doublet Score (RNA)", y = "Doublet Score (ATAC)") +
    theme_bw() +
    theme(
      axis.title = element_text(size = 13),
      axis.text = element_text(size = 11)
    )
scatter_plot
```


## Projection

### ATAC-driven UMAP

```{r}
# Define doublets
doublets_comparison <- 
  tonsil$scrublet_doublet_scores > max_doublet_score_rna |
  tonsil$scrublet_doublet_scores_atac > max_doublet_score_atac
tonsil$is_doublet <- doublets_comparison


# Plot
feat_plot1 <- FeaturePlot(
  tonsil,
  features = "scrublet_doublet_scores",
  reduction = "umap.atac",
  pt.size = 0.1
)
feat_plot2 <- FeaturePlot(
  tonsil,
  features = "scrublet_doublet_scores_atac",
  reduction = "umap.atac",
  pt.size = 0.1
)
dim_plot1 <- DimPlot(
  tonsil,
  group.by = "is_doublet",
  reduction = "umap.atac",
  pt.size = 0.1
)
```


```{r fig.wide=TRUE}
feat_plot1
```
```{r fig.wide=TRUE}
feat_plot2
```
```{r fig.wide=TRUE}
dim_plot1
```


### RNA-driven UMAP

```{r}
feat_plot3 <- FeaturePlot(
  tonsil,
  features = "scrublet_doublet_scores",
  reduction = "umap.rna",
  pt.size = 0.1
)
feat_plot4 <- FeaturePlot(
  tonsil,
  features = "scrublet_doublet_scores_atac",
  reduction = "umap.rna",
  pt.size = 0.1
)
dim_plot2 <- DimPlot(
  tonsil,
  group.by = "is_doublet",
  reduction = "umap.rna",
  pt.size = 0.1
)
```

```{r fig.wide=TRUE}
feat_plot3
```
```{r fig.wide=TRUE}
feat_plot4
```
```{r fig.wide=TRUE}
dim_plot2
```

Canonical markers:

```{r fig.wide=TRUE}
canonical_markers <- c("CD3D", "CD79B", "NKG7", "LYZ", "FDCSP")
canonical_markers_gg <- purrr::map(canonical_markers, function(x) {
  p <- FeaturePlot(
    tonsil,
    features = x,
    reduction = "umap.rna",
    pt.size = 0.1
  )
  p
})
canonical_markers_gg
```


As we can see, there are clearly two clusters of cells that are B-T cell clusters. In the next notebook we will overcluster to discard those.



# Session Information

```{r}
sessionInfo()
```
