---
title: "Doublet exclusion: Multiome"
author: "Paula Soler-Vila and Ramon Massoni-Badosa"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```


# Introduction

In this notebook we will exclude the doublets in the multiome using two approaches:

- Applying permissive thresholds to the scrublet doublet scores (ATAC and RNA).
- Overcluster to find and remove the clusters of B-T doublets.


# Pre-processing 

## Load packages

```{r}
library(Seurat)
library(Signac)
library(tidyverse)
```


## Parameters

```{r}
# Paths
path_to_obj <- here::here("multiome/results/R_objects/5.tonsil_multiome_integrated_using_wnn.rds")
path_to_save <- here::here("multiome/results/R_objects/6.tonsil_multiome_integrated_using_wnn_no_doublets.rds")


# Thresholds
max_doublet_score_rna <- 0.3
max_doublet_score_atac <- 0.3
```


## Load data

```{r}
tonsil <- readRDS(path_to_obj)
```


# Cluster


```{r}
DefaultAssay(tonsil) <- "RNA"
tonsil <- FindNeighbors(tonsil, reduction = "harmony_RNA", dims = 1:30)
tonsil <- FindClusters(tonsil, resolution = c(1.25, 1.5, 1.75, 2))
print(colnames(tonsil@meta.data))
```


## Visualize clusters

```{r fig.wide=TRUE}
vars <- str_subset(colnames(tonsil@meta.data), "^RNA_snn_res")
clusters_gg <- purrr::map(vars, function(x) {
  p <- DimPlot(
    tonsil,
    group.by = x,
    reduction = "umap.rna",
    pt.size = 0.1, label = T
  )
  p
})
clusters_gg
```


# Filter

We will exclude cluster 8 (resolution = 1.25), as we saw in the previous notebook that corresponds to B-T doublets and has a high scrublet score. We will also filter cells with extreme doublet score:

```{r}
tonsil$is_doublet <- 
  tonsil$RNA_snn_res.1.25 == "8" |
  tonsil$scrublet_doublet_scores > max_doublet_score_rna |
  tonsil$scrublet_doublet_scores_atac > max_doublet_score_atac
tonsil <- subset(tonsil, subset = is_doublet == FALSE)
```

Plot:

```{r fig.wide=TRUE}
p <- DimPlot(
  tonsil,
  group.by = "library_name",
  reduction = "umap.rna",
  pt.size = 0.1
)
p
```


# Save

```{r}
saveRDS(tonsil, path_to_save)
```


# Session Information

```{r}
sessionInfo()
```

