---
title: "Transfer labels RNA to ATAC: NBC/MBC"
author: "Ramon Massoni-Badosa"
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```


# Introduction

Here, we will transfer the label from RNA to ATAC for the naive and memory B cells (NBC/MBC) using a KNN classifier.


# Pre-processing

## Load packages

```{r}
library(Seurat)
library(Signac)
library(SLOcatoR)
library(EnsDb.Hsapiens.v86)
library(ggpubr)
library(tidyverse)
library(pals)
library(openxlsx)
set.seed(173)
```


## Define paths and parameters

```{r}
# Paths
path_to_rna <- here::here("scRNA-seq/results/R_objects/final_clusters/NBC_MBC_seu_obj_level_5_delta.rds")
path_to_atac <- here::here("scATAC-seq/results/R_objects/NBC_MBC_integrated_level_3.rds")
path_to_save <- here::here("scATAC-seq/results/R_objects/NBC_MBC_annotated.rds")


# Colors
color_palette <-  c("#E6194B", "#3CB44B", "#FFD8B1", "#4363D8", "#F58231",
                    "#911EB4", "#46F0F0", "#F032E6", "#BCF60C", "#FABEBE",
                    "#008080", "#E6BEFF", "#9A6324", "#FFFAC8", "#800000",
                    "#AAFFC3", "#808000", "#FFE119", "#000075", "#808080",
                    "#000000", "tan", "darkgrey")


# Misc
optimal_k <- 5
```


## Read data

```{r}
rna <- readRDS(path_to_rna)
rna$names_level_5_clusters_delta[rna$names_level_5_clusters_delta == "ncsMBC_FCRL4"] <- "ncsMBC FCRL4+"
Idents(rna) <- "names_level_5_clusters_delta"
atac <- readRDS(path_to_atac)
DimPlot(rna, cols = color_palette)
DimPlot(atac, cols = color_palette)
```


# Subset object

We noticed that the ATAC-seq fraction of the BCLL-2-T donor (old adult) profiled with multiome was of poor-quality, so we excluded it. Thus, to perform the label transfer we first need to find the intersection of cell barcodes for cells from the multiome experiments.

```{r}
# Intersection of cell barcodes
common_cells <- intersect(colnames(rna), colnames(atac))


# Subset
selected_cells <- c(
  colnames(atac)[atac$assay == "scATAC"],
  common_cells
)
atac <- subset(atac, cells = selected_cells)
atac$type <- ifelse(
  atac$assay == "scATAC",
  "query",
  "reference"
)
Idents(atac) <- "assay"
atac$annotation <- NA
atac$UMAP1_nbc_mbc <- NA
atac$UMAP2_nbc_mbc <- NA


# Define variables to predict
ref_cells <- colnames(atac)[atac$type == "reference"]
umap_df <- as.data.frame(Embeddings(rna, "umap"))
atac$annotation[ref_cells] <- rna$names_level_5_clusters_delta[ref_cells]
atac$UMAP1_nbc_mbc[ref_cells] <- umap_df[ref_cells, "UMAP_1"]
atac$UMAP2_nbc_mbc[ref_cells] <- umap_df[ref_cells, "UMAP_2"]
DimPlot(atac, group.by = "annotation", cols = color_palette, split.by = "type")
```


# Create training and test sets


```{r}
data_sets <- split_training_and_test_sets(
  atac,
  split_var = "type",
  referece_label = "reference",
  query_label = "query",
  reduction = "harmony",
  n_dims = 40
)
data_sets$training_set <- data_sets$training_set[, 2:40]
data_sets$test_set <- data_sets$test_set[, 2:40]
```


# Transfer label

```{r}
annotation_data  <- SLOcatoR::transfer_label(
  seurat_obj = atac,
  training_set = data_sets$training_set,
  test_set = data_sets$test_set,
  k = optimal_k,
  response_var = "annotation"
)
```


# Transfer UMAP coords

```{r}
umap_coords <- transfer_umap_coords(
  seurat_obj = atac,
  training_set = data_sets$training_set,
  test_set = data_sets$test_set,
  k = optimal_k,
  umap1_var = "UMAP1_nbc_mbc",
  umap2_var = "UMAP2_nbc_mbc"
)
annotation_atac_df <- left_join(
  annotation_data,
  umap_coords,
  by = "query_cells"
)
p1 <- ggplot(annotation_atac_df, aes(UMAP1, UMAP2, color = annotation)) +
  geom_point(size = 0.4) +
  scale_color_manual(values = color_palette) +
  theme_classic()
p2 <- ggplot(annotation_atac_df, aes(UMAP1, UMAP2, color = annotation_prob)) +
  geom_point(size = 0.4) +
  scale_color_viridis_c(option = "magma") +
  theme_classic()
p1 | p2
```


# Add info to Seurat object

```{r}
rownames(annotation_atac_df) <- annotation_atac_df$query_cells
query_cells <- colnames(atac)[!(colnames(atac) %in% ref_cells)]
atac$annotation[query_cells] <- annotation_atac_df[query_cells, "annotation"]
atac$annotation_prob[query_cells] <- annotation_atac_df[query_cells, "annotation_prob"]
atac$UMAP1_nbc_mbc[query_cells] <- annotation_atac_df[query_cells, "UMAP1"]
atac$UMAP2_nbc_mbc[query_cells] <- annotation_atac_df[query_cells, "UMAP2"]
```


# Save

```{r}
saveRDS(atac, path_to_save)
```


# Session Information

```{r}
sessionInfo()
```

