---
title: "Call accessibility matrix (query)"
author: "Ramon Massoni-Badosa"
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```


# Introduction

In the previous notebook we defined our reference as the NBC/MBC cells profiled with scATAC-seq, to which we transfered the label defined by scRNA-seq. Here, we will recompute the accessibility matrix for the query cells (M102), so that we can merge both datasets.


# Pre-processing

## Load packages

```{r}
library(Seurat)
library(Signac)
library(SLOcatoR)
library(EnsDb.Hsapiens.v86)
library(ggpubr)
library(tidyverse)
library(pals)
library(openxlsx)
set.seed(173)
```


## Define paths and parameters

```{r}
# Paths
path_to_reference <- here::here("scATAC-seq/results/R_objects/NBC_MBC_annotated.rds")
path_to_query <- here::here("MCL/results/R_objects/6.seurat_tumoral_102.rds")


# Colors
color_palette <-  c("#E6194B", "#3CB44B", "#FFD8B1", "#4363D8", "#F58231",
                    "#911EB4", "#46F0F0", "#F032E6", "#BCF60C", "#FABEBE",
                    "#008080", "#E6BEFF", "#9A6324", "#FFFAC8", "#800000",
                    "#AAFFC3", "#808000", "#FFE119", "#000075", "#808080",
                    "#000000", "tan", "darkgrey")
```


## Read data

```{r}
ref <- readRDS(path_to_reference)
Idents(ref) <- "annotation"
query <- readRDS(path_to_query)
DefaultAssay(query) <- "ATAC"
query@assays$RNA <- NULL
DimPlot(ref, cols = color_palette)
DimPlot(query, cols = color_palette)
```


# Call accessibility matrix

Let's correct the paths for the fragments files:
