---
title: "Normalization and Batch Effect Correction after the new Peak calling"
author: "Paula Soler-Vila & Sergio Aguilar-Fernández"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```


# Introduction

The new peak calling round applied on the previous notebook significantly increased the number of the features identified in our dataset. Therefore, we must need to repeat the standard downstream analysis, including data normalization, dimensionality reduction analysis and batch correction to account for this change in the number of features detected.


# Pre-processing

## Load packages

```{r}
library(Seurat)
library(SeuratWrappers)
library(Signac)
library(harmony)
library(tidyverse)
library(unikn) 

set.seed(1234)
```

## Parameters

```{r parameters}
cell_type = "PC"

# Paths
path_to_obj <- paste0(
  here::here("scATAC-seq/results/R_objects/level_5/"),
  cell_type,
  "/03.",
  cell_type,
  "_peak_calling_level_5.rds",
  sep = ""
)

path_to_save <- paste0(
  here::here("scATAC-seq/results/R_objects/level_5/"),
  cell_type,
  "/04.",
  cell_type,
  "_integration_peak_calling_level_5.rds",
  sep = ""
)
```

## Color palette
```{r palette}
Precursor <- c("PC committed Light Zone GCBC", "Early PC precursor")
IgM_pathway <- c("Short lived IgM+ PC","IgM+ early PC precursor", "IgM+ PC precursor", "preMature IgM+ PC", "Mature IgM+ PC")
IgG_pathway <- c("IgG+ PC precursor", "preMature IgG+ PC")
Proli <- c("PB committed early PC precursor", "Transitional PB", "PB")
Mature_IgG <- c( "Mature IgG+ PC", "MBC derived IgG+ PC")
Mature_IgA <- c( "Mature IgA+ PC", "MBC derived IgA+ PC")
IgD <- c("IgD PC precursor")
Memory <- c("class switch MBC", "MBC derived early PC precursor", "MBC derived PC precursor")
DZ <- c("Dark Zone GCBC", "Light Zone GCBC", "DZ migratory PC precursor")



my_pal_prec <- c(seecol("pal_peach", n=5)[c(2,5)])
names(my_pal_prec) <- Precursor
#my_pal_igm <- seecol("pal_seegruen", n=length(IgM_pathway))
my_pal_igm <- c("#99CC99", "#66B266", "#198C19", "#006600", "#003300")
names(my_pal_igm) <- IgM_pathway
my_pal_igg <- seecol("pal_seeblau", n=5)[c(2,5)]
names(my_pal_igg) <- IgG_pathway
my_pal_matureigg <- rev(seecol("pal_karpfenblau", n=5)[c(3,5)])
names(my_pal_matureigg) <- Mature_IgG
#my_pal_matureiga <- rev(seecol("pal_bordeaux", n=5)[c(2,5)])
my_pal_matureiga <- rev(c("#896B60", "#4A332D"))
names(my_pal_matureiga) <- Mature_IgA
my_pal_mbc <- rev(seecol("pal_petrol",  n=5)[c(1,3,5)])
names(my_pal_mbc) <- Memory
my_pal_proli <- rev(seecol("pal_pinky", n=length(Proli)))
names(my_pal_proli) <- Proli
my_pal_igd <- c('#FFD8B1')
names(my_pal_igd) <- IgD
my_pal_dz <- c(seecol("pal_grau", n=2), "#D94701") #E56C72
names(my_pal_dz) <- DZ


palette <- c(my_pal_prec, my_pal_igm, my_pal_igg, my_pal_matureigg, my_pal_matureiga, my_pal_mbc, my_pal_proli, my_pal_igd, my_pal_dz)
```

## Load data
```{r}
seurat <- readRDS(path_to_obj)
seurat
```

# Visualize UMAP without batch effect correction
```{r}
# Normalization, dimensionality reduction 
seurat <- seurat %>%
  RunTFIDF() %>% 
  FindTopFeatures(min.cutoff = 5) %>%
  RunSVD() %>%
  RunUMAP(reduction = "lsi", dims = 2:40)

DepthCor(seurat)
```

```{r}
DimPlot(seurat, 
        pt.size = 0.1)
```

```{r}
# Visualize UMAP's confounders
confounders <- c("library_name", "sex", "age_group", "hospital", "assay")
umaps_before_integration <- purrr::map(confounders, function(x) {
  p <- DimPlot(seurat, group.by = x, pt.size = 0.1)
  p
})
names(umaps_before_integration) <- confounders
print("UMAP colored by GEM:")
umaps_before_integration$library_name + NoLegend()
print("UMAP colored by sex, age group, cell hashing status, sampling center and assay:")
umaps_before_integration[2:length(umaps_before_integration)]
```

# Run and visualize Harmony’s integration
```{r}
seurat <- RunHarmony(
  object = seurat,
  dims = 2:40,
  group.by.vars = 'gem_id',
  reduction = 'lsi',
  assay.use = 'peaks_level_5',
  project.dim = FALSE,
  max.iter.harmony = 20)
    
# Non-linear dimension reduction and clustering
seurat <- RunUMAP(seurat, dims = 2:11, reduction = 'harmony')

p1 <- DimPlot(seurat, order = F,
  cols = c("#73787E", "#B8BCC1",
           "#FECFC7" ,"#FF8E7B",
           "#a13b53",
           "#A6E1F4", "#586BA4", 
           "#323734","#035F72",
           "#9CC6CF" ,"#198C19",
           "#006600","#FFD8B1"),
  pt.size = 0.6) + NoLegend()

p2 <- FeaturePlot(seurat,"annotation_prob")

p1 

p2

pdf(file = here::here("scATAC-seq/results/plots/PC/UMAP_final_annotation.pdf"), 
    width = 6, height = 6)
p1 
dev.off()
```

```{r}
# Visualize UMAP's confounders
umaps_after_integration <- purrr::map(confounders, function(x) {
  p <- DimPlot(seurat, group.by = x, pt.size = 0.1)
  p
})
names(umaps_after_integration) <- confounders
print("UMAP colored by GEM:")
umaps_after_integration$library_name + NoLegend()
print("UMAP colored by sex, age group, cell hashing status, sampling center and assay:")
umaps_after_integration[2:length(umaps_before_integration)]
```

# Save

```{r}
# Save integrated Seurat object
saveRDS(seurat, path_to_save)
```

# Session Information

```{r}
sessionInfo()
```





