---
author: "Ramon Massoni-Badosa"
date: "2021/03/15"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
title <- paste(cell_type, ": Cluster level 3 (optimal resolution)", sep = "")
```

---
title: "`r title`"
---


# Introduction

In this notebook, we will cluster the `r cell_type` cells of the tonsil atlas (level 3) using the optimal resolution we predicted previously. We will save it to annotate it in the future.


# Pre-processing

## Load packages

```{r}
library(Seurat)
library(tidyverse)
```


## Parameters

```{r}
# Paths
path_to_level_3 <- here::here("scRNA-seq/results/R_objects/level_3/")
path_to_level_3_cell_type <- str_c(path_to_level_3, cell_type, sep = "")
path_to_obj <- str_c(
  path_to_level_3_cell_type,
  "/",
  cell_type,
  "_integrated_level_3.rds",
  sep = ""
)
path_to_save <- str_c(
  path_to_level_3_cell_type,
  "/",
  cell_type,
  "_clustered_level_3.rds",
  sep = ""
)
path_to_tmp <- str_c(
  here::here("scRNA-seq/3-clustering/3-level_3/tmp/"),
  cell_type,
  sep = ""
)
dir.create(path_to_tmp, showWarnings = FALSE)
path <- str_c(
  path_to_tmp,
  "/",
  cell_type,
  sep = ""
)
path_to_save_shiny_metadata <- str_c(path, "_metadata_to_shiny_app.rds")
path_to_save_shiny_expression <- str_c(path, "_expression_to_shiny_app.rds")
path_to_save_umap <- str_c(path, "_umap_clusters_level_3.png")
path_to_save_umap_king <- str_c(path, "_umap_annotation_king_et_al_level_3.png")
path_to_save_umap_level_1 <- str_c(path, "_umap_clusters_level_1.png")
path_to_save_barplot_king <- str_c(path, "_barplot_annotation_king_et_al_level_3.png")


# Optimal resolution
optimal_resolutions <- c(
  NBC_MBC = 0.5,
  GCBC = 0.45,
  CD4_T = 0.6,
  Cytotoxic = 0.8,
  PC = 0.7,
  myeloid = 1.1,
  epithelial = 0.5,
  PDC = 0.75,
  FDC = 0.8
)
optimal_resolution <- optimal_resolutions[cell_type]


# Point sizes
pt_sizes <- c(
  NBC_MBC = 0.15,
  GCBC = 0.15,
  CD4_T = 0.25,
  Cytotoxic = 0.4,
  PC = 0.5,
  myeloid = 0.6,
  FDC = 0.6,
  PDC = 1,
  epithelial = 1
)
pt_size <- pt_sizes[cell_type]


# Colors
color_palette <- c("black", "gray", "red", "yellow", "plum4", "green4",
                   "blue", "mediumorchid2", "coral2", "blueviolet",
                   "indianred4", "deepskyblue1", "dimgray", "deeppink1",
                   "green", "lightgray", "hotpink1", "gold", "brown",
                   "mediumvioletred", "mediumaquamarine")


# Source functions
source(here::here("scRNA-seq/bin/utils.R"))
```


## Load data

```{r}
seurat <- readRDS(path_to_obj)
seurat
```


# Cluster

```{r}
seurat <- FindClusters(seurat, resolution = optimal_resolution)
```


# Visualize UMAP

```{r}
p <- DimPlot(
  seurat,
  group.by = "seurat_clusters",
  pt.size = pt_size,
  cols = color_palette
)
p
```

Let us plot the annotation of [King et al.](https://immunology.sciencemag.org/content/6/56/eabe6291):

```{r}
umap_king <- plot_annotation_king(
  seurat,
  pt_size = pt_size,
  color_palette = color_palette
)
umap_king <- umap_king +
  guides(colour = guide_legend(override.aes = list(size = 2)))
umap_king


# Stacked bar plot
cell_type_pcts_df <- seurat@meta.data %>%
  filter(assay == "5P") %>%
  group_by(seurat_clusters, cell_type) %>%
  summarise(n_cells_cell_type = n()) %>% 
  group_by(seurat_clusters) %>%
  mutate(
    total_cells = sum(n_cells_cell_type),
    pct_cells = n_cells_cell_type / total_cells * 100
  )
cell_type_pcts_gg <- cell_type_pcts_df %>%
  ggplot(aes(seurat_clusters, pct_cells, fill = cell_type)) +
    geom_col() +
    labs(x = "Cluster", y = "Percentage of cells (%)", fill = "") +
    scale_fill_manual(values = color_palette) +
    theme_classic()
cell_type_pcts_gg
```


To aid in the annotation even more, let us plot the cells in the UMAP of the level 1, which we saved in the metadata:

```{r}
umap_level_1 <- seurat@meta.data %>%
  ggplot(aes(UMAP_1_level_1, UMAP_2_level_1, color = seurat_clusters)) +
    geom_point(size = 0.5) +
    scale_color_manual(values = color_palette) +
    labs(x = "UMAP_1", y = "UMAP_2") +
    theme_classic() +
    theme(legend.title = element_blank()) +
    guides(colour = guide_legend(override.aes = list(size = 2)))
umap_level_1
```



# Save

```{r}
saveRDS(seurat, path_to_save)
```


Save input to shiny app:

```{r}
input_shiny <- seurat2shiny(
  seurat,
  assay = "RNA",
  slot = "data",
  reduction = "umap"
)
saveRDS(input_shiny$metadata, path_to_save_shiny_metadata)
saveRDS(input_shiny$expression, path_to_save_shiny_expression)
```


Save plots:

```{r}
ggsave(
  filename = path_to_save_umap,
  plot = p,
  width = 14,
  height = 12,
  units = "cm"
)
ggsave(
  filename = path_to_save_umap_king,
  plot = umap_king,
  width = 14,
  height = 12,
  units = "cm"
)
ggsave(
  filename = path_to_save_umap_level_1,
  plot = umap_level_1,
  width = 14,
  height = 12,
  units = "cm"
)
ggsave(
  filename = path_to_save_barplot_king,
  plot = cell_type_pcts_gg,
  width = 16,
  height = 11,
  units = "cm"
)
```


# Session Information

```{r}
sessionInfo()
```

