---
title: "Myeloid Signature Projection"
author: "Marc Elosua-Bayes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    toc: yes
    toc_float: yes
    number_sections: yes
    df_print: paged
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message = FALSE, warning = FALSE, cache = FALSE)
options(width = 1200)
```

## Introduction
In this Rmarkdown document we are going to project the myeloid gene signatures onto the Tonsil Visium slices. \

## Libraries
```{r}
library(ggpubr)
library(cowplot)
library(dplyr)
library(dtplyr)
library(ggplot2)
library(stringr)
library(readr)
library(data.table)
library(Matrix)
```

## Setting parameters
Loading necessary paths and parameters
```{r}
source(here::here("misc/paths.R"))
source(here::here("utils/bin.R"))
source(here::here("utils/stereoseq_bin.R"))

"{myeloid}/{plt_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(path = .,
             showWarnings = FALSE,
             recursive = TRUE)

"{myeloid}/{robj_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(path = .,
             showWarnings = FALSE,
             recursive = TRUE)
```

## Load data
Data loaded here is generated in the previous script *myeloid_signatures.Rmd* and *03-clustering/03-clustering_integration.Rmds*
```{r}
# Markers for signatures
markers_subset <- "{myeloid}/{robj_dir}/myeloid_markers_signature.rds" %>%
  glue::glue() %>%
  here::here() %>%
  readRDS(file = .)

# Tonsils integrated
se_obj <- "{clust}/{robj_dir}/integrated_spatial.rds" %>%
  glue::glue() %>%
  here::here() %>%
  readRDS(file = .)
```

## Analysis

### Signature scores
We start by computing the gene signatures for each cell type:
```{r}
ct_vec <- as.character(unique(markers_subset$cluster))

for (i in ct_vec) {
  print(i)
  tmp_df <- markers_subset %>%
    dplyr::filter(cluster == i) %>%
    dplyr::arrange(dplyr::desc(avg_log2FC)) %>%
    dplyr::top_n(n = 15, wt = avg_log2FC)
  
  se_obj <- Seurat::AddModuleScore(
    object = se_obj,
    features = list(tmp_df$gene),
    assay = "Spatial",
    name = glue::glue("{i}-Signature"))
}
head(se_obj@meta.data)
```

### Visualize Signatures
```{r}
sign_vec <- colnames(se_obj@meta.data)[stringr::str_detect(
  string = colnames(se_obj@meta.data),
  pattern = "Signature1$")]
```

Iterate over signatures and slides
```{r fig.width=12, fig.height=8}
# Iterate over signature
plt_all <- lapply(sign_vec, function(sign) {
  # Iterate over slides
  plt_ls <- lapply(Seurat::Images(se_obj), function(img) {
    Seurat::SpatialPlot(
      object = se_obj,
      features = sign,
      images = img,
      alpha = c(0, 1),
      crop = FALSE,
      pt.size.factor = 1) +
      ggplot2::scale_fill_gradient(
        low = "yellow", high = "red",
        limits = c(min(se_obj@meta.data[, sign]),
                   max(se_obj@meta.data[, sign]))) +
      # Remove title
      ggplot2::theme(legend.title = ggplot2::element_blank())
  })
  # Combine all the plots
  cowplot::plot_grid(plotlist = plt_ls, align = "hv", axis = "trbl", ncol = 4) +
    # Add title
    patchwork::plot_annotation(
      # Remove the trailin1 added by module score
      title = stringr::str_sub(string = sign, start = 1, end = nchar(sign)-1))
      # title = sign)
})

plt_all
```

Add empty plot as the 1st element to have the reference files
```{r fig.width=12, fig.height=6}
plt_hist <- lapply(Seurat::Images(se_obj), function(img) {
    Seurat::SpatialPlot(
      object = se_obj,
      features = "CD3D",
      images = img,
      alpha = c(0, 0),
      crop = FALSE,
      pt.size.factor = 1) +
    Seurat::NoLegend() +
    # Remove title
    ggplot2::theme(legend.title = ggplot2::element_blank())
  }) %>%
  # Combine all the plots
  cowplot::plot_grid(plotlist = ., align = "hv", axis = "trbl", ncol = 4)
```

## Save Plots

```{r}
"{myeloid}/{plt_dir}/myeloid_signature_plots.pdf" %>%
  glue::glue() %>%
  here::here() %>%
  ggpubr::ggexport(
    plotlist = plt_all,
    filename = .,
    ncol = 1,
    nrow = 1,
    width = 20,
    height = 12)

"{myeloid}/{plt_dir}/tissue_plots.pdf" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    plot = plt_hist,
    filename = .,
    base_width = 20,
    base_height = 10)
```

## Session Info
```{r}
sessionInfo()
```
