---
author: "Marc Elosua-Bayes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
title: "Myeloid deconvolution assessment"
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message = FALSE, warning = FALSE, cache = FALSE,
                      cache.lazy = FALSE)
options(width = 1200)
```

## Introduction
In this script we are going to map Myeloid subtypes onto the Visium slides.

## Libraries
```{r}
library(Seurat)
library(ggpubr)
library(cowplot)
library(dplyr)
library(ggplot2)
library(stringr)
library(readr)
library(SPOTlight)
library(SPATA2)
```

## Setting parameters
Loading necessary paths and parameters
```{r}
source(here::here("misc/paths.R"))
source(here::here("utils/bin.R"))

"{myeloid}/{plt_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(path = ,
             showWarnings = FALSE,
             recursive = TRUE)

"{myeloid}/{robj_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(path = ,
             showWarnings = FALSE,
             recursive = TRUE)

set.seed(123)
```

Extract sample id and get Donor ID
```{r}
# sample_id <- params$sample_id
sample_id <- "esvq52_nluss5"
donor_id <- id_sp_df[id_sp_df$gem_id == sample_id, ] %>% dplyr::pull(donor_id)
```

## Load data
We have 8 different datasets that we are going to analyze separately.
The spatial data comes from the script *03-clustering/03-clustering_integration.Rmd* while the sc data can be found in Ramon's scRNAseq analysis: **/scratch/devel/rmassoni/tonsil_atlas_private/2-DOWNSTREAM_PROCESSING/results/R_objects/processed_seurat_objects/processed_seurat_objects/tonsil_integrated_with_harmony_scrublet_annotated.rds**. \
```{r}
sp_obj <- "misc/{robj_dir}/20220215_tonsil_atlas_spatial_seurat_obj.rds" %>%
  glue::glue() %>%
  here::here() %>%
  readRDS(file = .)

# Load SPOTlight data
spotlight_ls <- "{myeloid}/{robj_dir}/spotlight_ls_myeloid.rds" %>%
  glue::glue() %>% 
  here::here() %>%
  readRDS(file = .)

# Load names df data
nm_df <- "{myeloid}/{robj_dir}/myeloid_nm_df.rds" %>%
  glue::glue() %>% 
  here::here() %>%
  readRDS(file = .)
```

Add colors to cell types
```{r}
# (nm_df <- data.frame(col_vec))
# nm_df$plt_nm <- rownames(nm_df)
# nm_df$mod_nm <- stringr::str_replace_all(
#   string = rownames(nm_df),
#   pattern = "[[:punct:]]|[[:space:]]",
#   replacement = ".")
```

## Analysis

### Preprocess data
```{r}
decon_mtrx <- spotlight_ls[[2]]
decon_mtrx <- decon_mtrx[, colnames(decon_mtrx) != "res_ss"]

# Set as 0 cell types predicted to be under 3 % of the spot
decon_mtrx[decon_mtrx < 0.03] <- 0
```

Change column names
```{r}
new_cn <- data.frame(mod_nm = colnames(decon_mtrx)) %>%
  dplyr::left_join(nm_df, by = "mod_nm") %>%
  dplyr::mutate(plt_nm = dplyr::if_else(is.na(plt_nm), mod_nm, plt_nm)) %>%
  dplyr::distinct() %>%
  dplyr::pull(plt_nm)

colnames(decon_mtrx) <- new_cn
```

We are going to add the deconvolution to the Seurat object.
```{r}
sp_obj@meta.data <- cbind(sp_obj@meta.data, decon_mtrx)
```

Subset sample of interest
```{r}
sp_sub <- sp_obj[, sp_obj@meta.data$gem_id == sample_id]
sp_sub@images <- sp_sub@images[Seurat::Images(sp_sub) == sample_id]
```

### Look at SPOTlight results
Check Topic profiles
```{r fig.height=8, fig.width=8}
nmf_mod_ls <- spotlight_ls[[1]]
nmf_mod <- nmf_mod_ls[[1]]

h <- NMF::coef(nmf_mod)
rownames(h) <- paste("Topic", 1:nrow(h), sep = "_")
topic_profile_plts <- SPOTlight::dot_plot_profiles_fun(
  h = h,
  train_cell_clust = nmf_mod_ls[[2]])

topic_profile_plts[[2]] +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90), 
                 axis.text = ggplot2::element_text(size = 12))
```

Look at all cells profiles
```{r fig.height=20, fig.width=20}
topic_profile_plts[[1]] +
  ggplot2::theme(
    axis.text.y = ggplot2::element_blank(),
    axis.text.x = ggplot2::element_blank(),
    axis.title = ggplot2::element_blank())
```

Look at cells topic profile
```{r}
basis_spotlight <- data.frame(NMF::basis(spotlight_ls[[1]][[1]]))

train_labs <- spotlight_ls[[1]][[2]]
colnames(basis_spotlight) <- unique(stringr::str_wrap(train_labs, width = 30))

basis_spotlight[basis_spotlight < 0.0000001] <- 0

DT::datatable(basis_spotlight, filter = "top")
```

#### Cell type location
Look at the location of each cell type in each slice separately
```{r fig.height=30, fig.width=25}
# Iterate over cell types
ct <- colnames(decon_mtrx)

# Iterate over images
lapply(names(sp_obj@images), function(nm) {
  print(nm)
  nm_donor <- id_sp_df %>% dplyr::filter(gem_id == nm) %>% dplyr::pull(donor_id)
  # Iterate over cell types
  ct_plt_ls <- lapply(ct, function(i) {
    tmp_plt <- Seurat::SpatialFeaturePlot(
      object = sp_obj,
      features = i,
      alpha = c(0, 1),
      images = nm) +
      ggplot2::scale_fill_gradientn(
        colors = heat.colors(10, rev = TRUE)) +
      ggplot2::scale_alpha(range = c(0, 1)) +
      ggplot2::labs(title = stringr::str_wrap(string = i,
                                     width = 25),
           fill = "") +
      ggplot2::theme(
        plot.title = ggplot2::element_text(
          hjust = 0.5,
          size = 20,
          face = "bold"))
    
    if (sum(sp_sub@meta.data[, i]) == 0) {
      tmp_plt <- suppressMessages(tmp_plt + ggplot2::scale_alpha(range = c(0,0)))
    }
    
    return(tmp_plt)
  })
  
  (plt_arr <- cowplot::plot_grid(
    plotlist = ct_plt_ls,
    axis = "trbl",
    align = "hv",
    nrow = 6,
    ncol = 5))
  
  "{myeloid}/{plt_dir}/cell_type_location_myeloid_{nm_donor}_new.pdf" %>%
    glue::glue() %>%
    here::here() %>%
    cowplot::save_plot(
      filename = .,
      plot = plt_arr,
      base_height = 30,
      base_width = 25)
  })
```

### Boxplots
Prepare data for boxplots
```{r}
metadata_long <- sp_sub@meta.data %>% 
  # tidyr::pivot_longer(cols = c("annotation"),
  #                     names_to = "stratification_id",
  #                     values_to = "stratification_val") %>%
  tidyr::pivot_longer(cols = all_of(ct), names_to = "ct_key", values_to = "ct_val") %>%
  # dplyr::left_join(col_df, by = c("ct_key" = "ct_name")) %>%
  dplyr::mutate(ct_val = dplyr::if_else(ct_val > 0.001, ct_val, 0))
```

Box plot of cell type proportion between stratified regions showing the unadjusted ANOVA Pvalue

```{r fig.width=30, fig.height=20}
keep_ct <- metadata_long %>%
  dplyr::group_by(ct_key) %>%
  dplyr::summarise(prop_sum = sum(ct_val)) %>% 
  dplyr::filter(prop_sum > 0) %>%
  dplyr::pull(ct_key)

(bplt <- metadata_long %>%
  dplyr::filter(ct_key %in% keep_ct) %>%
  dplyr::mutate(
    ct_key = stringr::str_wrap(string = ct_key,
                                   width = 30)) %>%
  ggpubr::ggboxplot(data = .,
                    x = "annotation_20220215",
                    y = "ct_val",
                    facet.by = "ct_key",
                    color = "annotation_20220215",
                    fill = "annotation_20220215",
                    add = "jitter",
                    scales = "free",
                    repel = TRUE,
                    outlier.shape = NA,
                    alpha = 0.6,
                    palette = "Set1",
                    ncol = 6) +
  ggplot2::theme(
    strip.text = ggplot2::element_text(size = 18, face = "bold"),
    axis.text.y = ggplot2::element_text(size = 16),
    axis.title.y = ggplot2::element_text(size = 20),
    # axis.text.x = element_text(size = 12, angle = 90, vjust = 0.5, hjust = 0.5),
    axis.text.x = ggplot2::element_blank(),
    legend.text = ggplot2::element_text(size = 18),
    legend.title = ggplot2::element_blank(),
    strip.background = ggplot2::element_blank()) +
  ggplot2::labs(
    y = "Proportion",
    color = "Regions",
    fill = "Regions"))
# bplt <- bplt +
#   ggpubr::stat_compare_means(method = "anova", size = 6) +
#   ggplot2::scale_y_continuous(
#     expand = expansion(mult = c(0, 0.1)),
#     labels = function(x) sprintf("%.2f", x))

"{myeloid}/{plt_dir}/strat_bplot_{donor_id}.pdf" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    filename = .,
    plot = bplt,
    base_height = 20,
    base_width = 25)
```

### Cell type correlation matrix

Slide-specific
```{r}

lapply(Images(sp_obj), function(i) {
    
    se_sub <- subset(sp_obj, subset = gem_id == i)
    # se_sub
    se_sub@images <- se_sub@images[Seurat::Images(se_sub) == i]
    donor_id <- id_sp_df[id_sp_df$gem_id == sample_id, ] %>%
        dplyr::pull(donor_id)
    
    (cor_mtrx_ct <- SCrafty::correlation_heatmap( 
      se = sp_obj,
      feats = ct,
      assay = "Spatial",
      slot = "data") +
       ggplot2::labs(
         title = glue::glue("{donor_id}: Cell-type correlation matrix")))
    
    "{myeloid}/{plt_dir}/cor-mtrx_cell-type_{donor_id}.pdf" %>%
      glue::glue() %>%
      here::here() %>%
      cowplot::save_plot(
        filename = .,
        plot = cor_mtrx_ct,
        base_height = 9,
        base_width = 10)
})
```

Integrated
```{r}
# se_sub <- subset(merged_se, subset = gem_id == "esvq52_nluss5")
# se_sub
# se_sub@images <- se_sub@images[Seurat::Images(se_sub) == "esvq52_nluss5"]

(cor_mtrx_ct <- SCrafty::correlation_heatmap( 
  se = sp_obj,
  feats = ct,
  assay = "Spatial",
  slot = "data") +
   ggplot2::labs(
     title = "Integrated cell-type correlation matrix"))

"{myeloid}/{plt_dir}/cor-mtrx_cell-type_integrated.pdf" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    filename = .,
    plot = cor_mtrx_ct,
    base_height = 9,
    base_width = 10)
```

## Session Info
```{r}
sessionInfo()
```
