---
title: "Make Myeloid signatures"
author: "Marc Elosua-Bayes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    toc: yes
    toc_float: yes
    number_sections: yes
    df_print: paged
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message = FALSE, warning = FALSE, cache = FALSE)
options(width = 1200)
```

## Introduction
In this Rmarkdown document we are going to load the level-5 annotation of the myeloid cells and get gene signatures for each cell type. \

## Libraries
```{r}
library(ggpubr)
library(cowplot)
library(dplyr)
library(dtplyr)
library(ggplot2)
library(stringr)
library(readr)
library(data.table)
library(Matrix)
```

## Setting parameters
Loading necessary paths and parameters
```{r}
source(here::here("misc/paths.R"))
source(here::here("utils/bin.R"))
source(here::here("utils/stereoseq_bin.R"))

"{myeloid}/{plt_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(path = .,
             showWarnings = FALSE,
             recursive = TRUE)

"{myeloid}/{robj_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(path = .,
             showWarnings = FALSE,
             recursive = TRUE)
```

## Load data
```{r}
se_obj <- "{myeloid}/{robj_dir}/myeloid_annotated_level_5.rds" %>%
  glue::glue() %>%
  here::here() %>%
  readRDS(file = .)
```

## Analysis
Lets take a quick look at the seurat object
```{r}
se_obj
Seurat::Assays(se_obj)
```

Metadata...
```{r}
head(se_obj@meta.data)
```

### Marker genes
Get marker genes for each cell type:
```{r}
# Set Idents
Seurat::Idents(se_obj) <- se_obj@meta.data[, "annotation_level_4"]

myeloid_markers <- Seurat::FindAllMarkers(
  object = se_obj,
  assay = "RNA",
  slot = "data",
  logfc.threshold = 0.75,
  only.pos = TRUE)

DT::datatable(myeloid_markers, filter = "top")
```

Select those genes that are unique for each cell type to avoid overlapping
```{r}
gene_unique <- myeloid_markers %>%
  dplyr::count(gene) %>%
  dplyr::filter(n == 1) %>%
  dplyr::pull(gene)

# Keep unique genes
markers_unique <- myeloid_markers %>%
  dplyr::filter(gene %in% gene_unique)

# Check all the cell types have marker genes
markers_unique %>%                 
  dplyr::count(cluster)
```

Select the top genes to make the gene signatures according to logFC and pct.1
```{r}
markers_subset <- markers_unique %>%
  dplyr::filter(pct.1 > 0.75 & avg_log2FC > 0.5)

# Check all the cell types have marker genes
markers_subset %>%
  dplyr::count(cluster)
```

## Save Data

Save as RDS object
```{r}
"{myeloid}/{robj_dir}/myeloid_markers_signature.rds" %>%
  glue::glue() %>%
  here::here() %>%
  saveRDS(object = markers_subset, file = .)
```

Save as spreadsheet for Marco and Juan
```{r}
out_xlsx <- "{myeloid}/{robj_dir}/myeloid_markers_signature.xlsx" %>%
  glue::glue() %>%
  here::here() 

# Remove file in case it exists to overwrite it
file.remove(out_xlsx)

lapply(as.character(unique(markers_subset$cluster)), function(i) {
  print(i)
  # Extract information for the specific cell type
  tmp_df <- markers_subset %>% dplyr::filter(cluster == i)
  
  if (nrow(tmp_df) > 0) {
    xlsx::write.xlsx(
      tmp_df,
      file = out_xlsx,
      row.names = FALSE,
      sheetName = i,
      append = file.exists(out_xlsx))
      }
  })
```

## Session Info
```{r}
sessionInfo()
```
