---
title: "T cell clusters"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: Sonal 
output:
  html_document:
    toc: yes
    toc_float: yes
    number_sections: yes
    df_print: paged
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
setwd("/home/srashmi/Documents/tonsil_atlas_citeseq_vdj_20210505/")
```

# Define Parameters

## Load packages

```{r } 
library(Seurat)
library(flexclust)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(patchwork)
library(scales)
library(harmony)
```


## Define variables and functions

```{r}
path_to_save_obj <- "/home/srashmi/Documents/tonsil_atlas_citeseq_vdj_20210505/results"
path_to_save_citeseq_seurat_obj <- str_c(
  path_to_save_obj,
  "tonsil_cite_seq_annotated.rds",
  sep = "/"
)
path_to_save_seurat_obj <- str_c(
  path_to_save_obj,
  "tonsil_cite_seq_tcell_object.rds",
  sep = "/"
)
citeseq_marker <- str_c(
  path_to_save_obj,
  "tonsil_cite_seq_tcell_marker.xlsx",
  sep = "/"
)
saved_cell_cycle_obj <- str_c(
  path_to_save_obj,
  "../data/cycle.rda",
  sep = "/"
)
# Color
color <- c("black", "gray", "red", "yellow", "violet", "green4",
                   "blue", "mediumorchid2", "coral2", "blueviolet",
                   "indianred4", "deepskyblue1", "dimgray", "deeppink1",
                   "green", "lightgray", "hotpink1", "chocolate", "aquamarine", "aliceblue", "burlywood", "blueviolet")
```

## Load data

```{r data}
seurat_obj <- readRDS(path_to_save_citeseq_seurat_obj)
```

## get metadata

```{r get metadata}
metadata <- seurat_obj@meta.data 
```

# extract T cells

```{r}
b_cells <- rownames(subset(metadata, metadata$annotation %in% c("CD4_T","Cytotoxic")))
t_clusters_obj = subset(seurat_obj,cells=b_cells)
```


# Sub-clustering of T cell clusters

```{r}
t_clusters_obj@meta.data %>%
  ggplot(aes(UMAP1, UMAP2, color = annotation)) +
    geom_point(size = 0.75) +
    scale_color_manual(values = color) +
    labs(x = "UMAP1", y = "UMAP2", color = "") +
    theme_classic()
```


```{r}
metadata <- t_clusters_obj@meta.data
df = count(metadata,metadata$annotation)
colnames(df) = c("Cluster","Number_of_cells")
library(knitr)
library(kableExtra)
kable(df) %>%
  kable_styling("striped", full_width = T)

```


## ADT and RNA multimodal analysis

```{r}
DefaultAssay(t_clusters_obj) <- 'RNA'
t_clusters_obj <- NormalizeData(t_clusters_obj) %>% FindVariableFeatures() %>% ScaleData() %>% RunPCA()  %>%
  RunHarmony(reduction = "pca",dims = 1:30,group.by.vars = "gemid",assay.use = "RNA",reduction.save = "harmony_RNA")
```

## Check for Cell cycle stages variability between clusters

```{r}
load(saved_cell_cycle_obj)
t_clusters_obj <- CellCycleScoring(t_clusters_obj, 
                                 g2m.features = g2m_genes, 
                                 s.features = s_genes)
```


```{r}
DefaultAssay(t_clusters_obj) <- 'ADT'
# we will use all ADT features for dimensional reduction
# we set a dimensional reduction name to avoid overwriting the 
VariableFeatures(t_clusters_obj) <- rownames(t_clusters_obj[["ADT"]])
t_clusters_obj <- NormalizeData(t_clusters_obj, normalization.method = 'CLR', margin = 2) %>% 
  ScaleData() %>% RunPCA(reduction.name = 'apca') %>%
  RunHarmony(reduction = "apca",dims = 1:20, group.by.vars = "gemid", assay.use = "ADT", reduction.save = "harmony_protein")
```

## Multimodal Neighbor Identification

```{r}
# Identify multimodal neighbors. These will be stored in the neighbors slot, 
# and can be accessed using filtered_bcll.combined[['weighted.nn']]
# The WNN graph can be accessed at filtered_bcll.combined[["wknn"]], 
# and the SNN graph used for clustering at filtered_bcll.combined[["wsnn"]]
# Cell-specific modality weights can be accessed at filtered_bcll.combined$RNA.weight
t_clusters_obj <- FindMultiModalNeighbors(
  t_clusters_obj, reduction.list = list("harmony_RNA", "harmony_protein"),
  dims.list = list(1:30, 1:20), modality.weight.name = "RNA.weight"
)
```


## UMAP visualisation

```{r}
feat_clust <- c(0.1,0.5,1,1.5,2)
t_clusters_obj <- RunUMAP(t_clusters_obj, nn.name = "weighted.nn", reduction.name = "wnn.umap", reduction.key = "wnn_UMAP_")
t_clusters_obj <- FindClusters(t_clusters_obj, graph.name = "wsnn", algorithm = 3, resolution = feat_clust, verbose = FALSE)
```

```{r}
Seurat::DimPlot(t_clusters_obj, reduction = 'wnn.umap', label = TRUE, repel = TRUE, label.size = 2.5, group.by = c("wsnn_res.0.1","wsnn_res.0.5","wsnn_res.1","wsnn_res.1.5","wsnn_res.2")) + NoLegend()  & Seurat::NoLegend()
```

```{r}
t_clusters_obj <- FindClusters(t_clusters_obj, graph.name = "wsnn", algorithm = 3, resolution = 0.5, verbose = FALSE)
t_clusters_obj <- RunUMAP(t_clusters_obj, reduction = 'pca', dims = 1:30, assay = 'RNA', 
              reduction.name = 'rna.umap', reduction.key = 'rnaUMAP_')
t_clusters_obj <- RunUMAP(t_clusters_obj, reduction = 'apca', dims = 1:18, assay = 'ADT', 
              reduction.name = 'adt.umap', reduction.key = 'adtUMAP_')

```


```{r}
p1 <- DimPlot(t_clusters_obj, reduction = 'rna.umap', label = TRUE, 
              repel = TRUE, label.size = 2.5) + NoLegend()
p2 <- DimPlot(t_clusters_obj, reduction = 'adt.umap', label = TRUE, 
              repel = TRUE, label.size = 2.5) + NoLegend()
```

```{r}
p1
```


```{r}
p2
```

```{r}
Seurat::DimPlot(t_clusters_obj, reduction = 'wnn.umap', label = TRUE, repel = TRUE, label.size = 2.5, group.by = "wsnn_res.0.5")
```

```{r}
metadata <- t_clusters_obj@meta.data
df = count(metadata,metadata$wsnn_res.0.5)
colnames(df) = c("Cluster","Number_of_cells")
library(knitr)
library(kableExtra)
kable(df) %>%
  kable_styling("striped", full_width = T)

```

# QC

## BCR - TCR

```{r}
df <- t_clusters_obj@meta.data
with(df, table(wsnn_res.0.5, repertoire_flag, useNA = "ifany"))
```


## Check for subproject variability

```{r}
DimPlot(t_clusters_obj, reduction = "wnn.umap", group.by = 'subproject')
```

## Cell Cycle 

```{r}
# Plot the PCA colored by cell cycle phase
DimPlot(t_clusters_obj,
        reduction = "pca",
        group.by= "Phase",
        split.by = "Phase")
```

```{r}
DimPlot(t_clusters_obj, reduction = "wnn.umap", group.by = 'Phase')
```

```{r}
p1 <- FeatureScatter(t_clusters_obj, feature1 = "CD4", feature2 = "CD8", group.by
 = "wsnn_res.0.5")
p2 <- FeatureScatter(t_clusters_obj, feature1 = "CD19", feature2 = "CD3E", group.by
 = "wsnn_res.0.5")
p3 <- FeatureScatter(t_clusters_obj, feature1 = "CD79B", feature2 = "CD3E", group.by
 = "wsnn_res.0.5")
p1
p2
p3
```

## Plot mitochondrial, nGene_count, nFeature_count content across cluster

```{r}
VlnPlot(object = t_clusters_obj, features = "nCount_RNA", group.by="wsnn_res.0.5")
```

```{r}
VlnPlot(object = t_clusters_obj, features = "nFeature_RNA", group.by="wsnn_res.0.5")
```

```{r}
VlnPlot(object = t_clusters_obj, features = "nCount_ADT", group.by="wsnn_res.0.5")
```

```{r}
VlnPlot(object = t_clusters_obj, features = "nFeature_ADT", group.by="wsnn_res.0.5")
```

```{r}
VlnPlot(object = t_clusters_obj, features = "percent.mt", group.by="wsnn_res.0.5")
```

```{r}
VlnPlot(object = t_clusters_obj, features = "log10GenesPerUMI", group.by="wsnn_res.0.5")
```

```{r}
VlnPlot(object = t_clusters_obj, features = "scrublet_doublet_scores", group.by="wsnn_res.0.5")
```

```{r}
VlnPlot(object = t_clusters_obj, features = "annotation_prob", group.by="wsnn_res.0.5")
```

# Find All CITEseq Markers

```{r}
DefaultAssay(t_clusters_obj) <- "ADT"
Idents(t_clusters_obj) <- t_clusters_obj@meta.data$wsnn_res.0.5
markers <- FindAllMarkers(t_clusters_obj, assay = "ADT")
```

## Save All Markers in Excel

```{r}
markers <- markers %>%
  dplyr::arrange(cluster, desc(abs(avg_log2FC))) 
#%>% dplyr::filter(p_val_adj < 0.001 & avg_log2FC > 0.9)
markers_list <- purrr::map(levels(markers$cluster), ~ markers[markers$cluster == .x, ])
names(markers_list) <- levels(markers$cluster)
openxlsx::write.xlsx(markers_list, citeseq_marker)
DT::datatable(markers, filter ="top")
```


# Save the seurat object 

```{r}
saveRDS(t_clusters_obj, file = path_to_save_seurat_obj)
```

# Session Info

```{r}
sessionInfo()
```

