---
author: "Marc Elosua-Bayes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
params: 
    sample_id: "Default!"
    donor_id: "Default!"
title: "`r sprintf('%s Integrating Regulon Information', params$donor_id)`"
editor_options: 
  chunk_output_type: console
---

---
title: `r params$donor_id`
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message = FALSE, warning = FALSE, cache = FALSE,
                      cache.lazy = FALSE)
options(width = 1200)
```

## Introduction
In this script we are going to get the most informative genes for each regulon as obtained by the SCENIC pipeline for each cell type and map them on the tissue.

## Libraries
```{r}
library(Seurat)
library(ggpubr)
library(cowplot)
library(dplyr)
library(ggplot2)
library(stringr)
library(readr)
library(SPOTlight)
```

## Setting parameters
Loading necessary paths and parameters
```{r}
set.seed(123)
source(here::here("misc/paths.R"))
source(here::here("utils/bin.R"))

dir.create(path = here::here(sprintf("%s/%s", regulon_32, plt_dir)),
           showWarnings = FALSE,
           recursive = TRUE)

dir.create(path =here::here(sprintf("%s/%s", regulon_32, robj_dir)),
           showWarnings = FALSE,
           recursive = TRUE)
```

Extract sample id and get Donor ID
```{r}
sample_id <- params$sample_id
# sample_id <- "esvq52_nluss5"
donor_id <- id_df[id_df$gem_id == sample_id, "BCLL_ID"]
```

## Load data
We have 4 different datasets that we are going to analyze separately.
The spatial data comes from the script *03-clustering/03-clustering.Rmd*. 
Gene signatures for the regulons come from Sergio Aguilar's analysis and can be found in the following path: */scratch/devel/saguilar/PhD/GeneRegulatoryNetworks/projects/downsampling_0.75/results/FDC_top10regs_with_top15genes.rds*. I made soft links to this data so they are in *06-regulon_integration/SCENIC_signatures*.
**col_df_l1** contains the color for each cell-type and the relations between modified and unmodified names, this is generated in *misc/col_df_l1.Rmd*.
```{r}
se_obj <- readRDS(file = here::here(glue::glue("{clust_32}/{robj_dir}/clust_se_{sample_id}.RDS")))

col_df_l1 <- readRDS(file = here::here("misc/col_df_l1.rds"))
ct <- col_df_l1$cell_type
```

Read regulon signatures
```{r}
fn_ls <- list.files(here::here(glue::glue("{regulon_32}/SCENIC_signatures")),
           full.names = TRUE)

names_ls <- stringr::str_split(
  string = fn_ls,
  pattern = "/|_") %>%
  purrr::map(., 17) %>%
  unlist()

sign_df_ls <- lapply(fn_ls, function(pth) readRDS(pth))
names(sign_df_ls) <- names_ls
```

## Analysis
Show the data
```{r}
lapply(names(sign_df_ls), function(i){
  tmp_df <- sign_df_ls[[i]]
  tmp_df$`Cell Type` <- i
  return(tmp_df)
}) %>%
  dplyr::bind_rows(.) %>%
  DT::datatable(
    data = .,
    filter = "top")
```

We are going to compute the gene expression scores for each regulon gene signature
```{r}
# Iterate over the cell types
for (ct in names(sign_df_ls)) {
  print(ct)
  tmp_df <- sign_df_ls[[ct]] %>%
    dplyr::mutate(
      TF = as.character(TF),
      target = as.character(target), 
    )
    
  # Iterate over regulons
  for (tf in unique(tmp_df$TF)) {
    print(tf)
    reg_targets <- tmp_df %>%
      dplyr::filter(TF == tf) %>%
      dplyr::pull(target)
    # print(reg_targets)
    
    # Subset genes present
    reg_targets <- reg_targets[reg_targets %in% rownames(se_obj)]
    
    se_obj <- Seurat::AddModuleScore(
      object = se_obj,
      features = list(reg_targets),
      name = glue::glue("{ct}_{tf}_signature")
      # search = TRUE
      )
  }
}

colnames(se_obj@meta.data)
```

Get signature genes
```{r}
sign_mask <- stringr::str_detect(
  string = colnames(se_obj@meta.data),
  pattern = "signature")

sign_feat <- colnames(se_obj@meta.data)[sign_mask]

# saveRDS(
#   object = se_obj@meta.data[, sign_feat],
#   file = here::here(glue::glue("{regulon_32}/{robj_dir}/regulon_scores_{sample_id}.rds")))
```

Plot activity modules on the tissue
```{r}
cell_types <- stringr::str_split(
  string = sign_feat,
  pattern = "_") %>%
  purrr:::map(., 1) %>%
  unique() %>%
  unlist()

lapply(cell_types, function(ct) {
  
  sign_feat_mask <- stringr::str_detect(
    string = sign_feat,
    pattern = glue::glue("{ct}_"))
  
  sing_plt <- Seurat::SpatialFeaturePlot(
    object = se_obj,
    features = sign_feat[sign_feat_mask])
  
  print(sing_plt)
  
  cowplot::save_plot(
    filename = here::here(glue::glue("{regulon_32}/{plt_dir}/{ct}_regulon_signatures.pdf")),
    plot = sing_plt,
    base_height = 15,
    base_width = 15)

})
```

## Session Info
```{r}
sessionInfo()
```

