---
title: "Normalization and Batch Effect Correction"
author: "Paula Soler-Vila"
output: 
    BiocStyle::html_document:
      toc: true
      toc_float: true
      number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, out.width = "100%", fig.align = "center",
  message = FALSE, warning = FALSE
)
options(width = 1200)
```

# Objective

The objective of this notebook is to apply the downstream analysis on the filtered scATAC-seq aggregated data:

1. Normalization across cells and across peaks (TF-IDF) following the method proposed by Stuart & Butler et al. 2019
2. Dimensionality reduction using singular value decomposition (SVD).
3. Bacht correction to reduce the dependence between the different confounders using Harmony.
4. Non-dimensionality reduction (UMAP)

# Pre-processing

## Load packages

```{r}
library(Signac)
library(Seurat)
library(GenomicRanges)
library(future)
library(harmony)
library(EnsDb.Hsapiens.v86)
library(stringr)
library(ggpubr)

set.seed(173)
```

## Parameters

```{r}
# Paths
path_to_obj <- here::here("scATAC-seq/results/R_objects/2.tonsil_aggregated_filtered.rds")
path_to_save_obj_norm <- here::here("scATAC-seq/results/R_objects/3.tonsil_aggregated_norm.rds")
path_to_save_obj_integrated <- here::here("scATAC-seq/results/R_objects/4.tonsil_aggregated_harmony.rds")
path_tmp_dir <- here::here("scATAC-seq/2-QC/5-batch_effect_correction/1-data_asses_scATAC/tmp/")

path_to_save_dimred_uncorrect <- str_c(path_tmp_dir, "batch_uncorrected_lsi.rds", sep = "")
path_to_save_dimred_correct <- str_c(path_tmp_dir, "batch_corrected_lsi.rds", sep = "")
path_to_save_confounders_df <- str_c(path_tmp_dir, "confounders_df.rds", sep = "") 
```

## Load filtered data

```{r}
seurat <- readRDS(path_to_obj)
seurat
```

# Visualize UMAP without batch effect correction

```{r fig.wide=TRUE}
# Process Seurat object
seurat <- seurat %>%
  RunTFIDF() %>% 
  FindTopFeatures(min.cutoff = "q0") %>%
  RunSVD() %>%
  RunUMAP(reduction = "lsi", dims = 2:40)

DepthCor(seurat)

# Visualize UMAP
confounders <- c("library_name", "sex", "age_group", "technique", "hospital")
umaps_before_integration <- purrr::map(confounders, function(x) {
  p <- DimPlot(seurat, group.by = x, pt.size = 0.1)
  p
})
names(umaps_before_integration) <- confounders
print("UMAP colored by GEM:")
umaps_before_integration$library_name + NoLegend()
print("UMAP colored by sex, age group, technique and hospital:")
umaps_before_integration[2:length(umaps_before_integration)]

saveRDS(seurat, path_to_save_obj_norm)
```

# Run and visualize Harmony's integration
We used Harmony integration to reduce the dependence between the main categorical batches (such as technique, sex, age group and library) of the dataset.

```{r fig.wide=TRUE}
seurat <- RunHarmony(
  object = seurat, 
  group.by.vars = "gem_id", 
  reduction = "lsi",
  dims = 2:40,
  assay.use = "ATAC",
  project.dim = FALSE
)

seurat <- RunUMAP(seurat, dims = 2:40, reduction = "harmony")

umaps_after_integration <- purrr::map(confounders, function(x) {
  p <- DimPlot(seurat, group.by = x, pt.size = 0.1)
  p
})
names(umaps_after_integration) <- confounders
print("UMAP colored by GEM:")
umaps_after_integration$library_name + NoLegend()
print("UMAP colored by sex, age group, technique and hospital:")
umaps_after_integration[2:length(umaps_before_integration)]
```

# Spot potential doublets

## Scrublet prediction

```{r}
# Scrublet
DimPlot(seurat, group.by = "scrublet_predicted_doublet")
table(seurat$scrublet_predicted_doublet)
```

## QC metrics

```{r}
qc_vars <- c(
  "nCount_ATAC",
  "nFeature_ATAC"
)
qc_gg <- purrr::map(qc_vars, function(x) {
  p <- FeaturePlot(seurat, features = x)
  p
})
qc_gg
```

# Save

```{r}
# If it doesn't exist create temporal directory
#dir.create(path_tmp_dir, showWarnings = FALSE) 

# Save integrated Seurat object
saveRDS(seurat, path_to_save_obj_integrated)

# Save PCA matrices to compute the Local Inverse Simpson Index (LISI)
confounders_df <- seurat@meta.data[, confounders]
saveRDS(confounders_df, path_to_save_confounders_df)
saveRDS(
  seurat@reductions$lsi@cell.embeddings[, 2:40],
  path_to_save_dimred_uncorrect
)
saveRDS(
  seurat@reductions$harmony@cell.embeddings[, 2:40],
  path_to_save_dimred_correct
)
```


# Session Information

```{r}
sessionInfo()
```
