---
title: "Cytotoxic: Level 5"
author: "Ramon Massoni-Badosa"
date: "2021/05/13"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```

NOTE: from level 4/5 onwards I will execute most steps locally, as the datasets are not as large anymore.

TODO: INTEGRATE TIM3 DN!!

# Introduction

In future steps of the project, we decided it was a better option to include naive CD8 T cells, TIM3+ DN and the rest of the CD8 T cells together to show in figure 2. Thus, here we will merge, reintegrate them and recluster.

In addition, we will merge them with the CITE-seq dataset.


## Load packages

```{r}
library(Seurat)
library(SeuratWrappers)
library(harmony)
library(clusterProfiler)
library(org.Hs.eg.db)
library(tidyverse)
library(caret)
library(class)
```


## Parameters

```{r}
path_to_cd8 <- "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/results/R_objects/level_5/Cytotoxic/CD8_T/CD8_T_clustered_level_5.rds"
path_to_tim3 <- "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/results/R_objects/level_4/Cytotoxic/TIM3_DN/TIM3_DN_clustered_level_4.rds"
path_to_naive_cd8 <- "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/results/R_objects/level_4/CD4_T/naive_CD8_T_subsetted_level_4.rds"
path_to_cite_seq <- "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/CITE-seq/results/seurat_object_cite_seq_seurat_wnn.rds"


# Colors
color_palette <-  c("#1CFFCE", "#90AD1C", "#C075A6", "#85660D", "#5A5156", "#AA0DFE",   
                    "#F8A19F", "#F7E1A0", "#1C8356", "#FEAF16", "#822E1C", "#C4451C",   
                    "#1CBE4F", "#325A9B", "#F6222E", "#FE00FA", "#FBE426", "#16FF32", 
                    "black",   "#3283FE", "#B00068", "#DEA0FD", "#B10DA1", "#E4E1E3",   
                    "#90AD1C", "#FE00FA", "#85660D", "#3B00FB", "#822E1C", "coral2", 
                    "#1CFFCE", "#1CBE4F", "#3283FE", "#FBE426", "#F7E1A0", "#325A9B",   
                    "#2ED9FF", "#B5EFB5", "#5A5156", "#DEA0FD", "#FEAF16", "#683B79",   
                    "#B10DA1", "#1C7F93", "#F8A19F", "dark orange", "#FEAF16", "#FBE426",  
                    "Brown")


# Source functions
source("~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/bin/SLOcatoR_functions.R")
```


## Read data

```{r}
naive_cd8 <- readRDS(path_to_naive_cd8)
tim3 <- readRDS(path_to_tim3)
other_cd8 <- readRDS(path_to_cd8)
cite_seq <- readRDS(path_to_cite_seq)
```


# Merge and integrate naive and other CD8 T cells

```{r}
naive_cd8 <- subset(naive_cd8, hospital != "Royal London")
other_cd8 <- subset(other_cd8, hospital != "Royal London")
tim3 <- subset(tim3, hospital != "Royal London")

naive_cd8$annotation_level_4 <- "naive_CD8_T"
tim3$annotation_level_4 <- tim3$annotation_level_3
reference <- merge(x = naive_cd8, y = c(other_cd8, tim3))
reference <- NormalizeData(reference)
hvg <- find_assay_specific_features(
  reference,
  assay_var = "assay",
  n_features = 5500
)
reference <- integrate_assays(
  reference,
  assay_specific = TRUE,
  n_dim = 25,
  assay_var = "assay",
  shared_hvg = hvg
)
reference <- RunUMAP(reference, reduction = "harmony", dims = 1:25)
DimPlot(reference, cols = color_palette, pt.size = 0.75)
```


# Recluster

```{r}
reference <- FindNeighbors(reference, reduction = "harmony", dims = 1:25)
reference <- FindClusters(reference, resolution = 0.75)
DimPlot(reference, cols = color_palette)
```

We will merge some clusters together into a single naive CD8 T cell cluster, since we know it's a homogeneous population:

```{r}
reference$annotation_level_5 <- reference$seurat_clusters
reference$annotation_level_5[reference$annotation_level_5 %in% c("0", "2", "6", "7")] <- "0"
DimPlot(reference, cols = color_palette, group.by = "annotation_level_5")
```


In addition, we will break cluster 8 into 2, since one of them shows a surprising expression of TRDC:

```{r}
Idents(reference) <- "annotation_level_5"
reference <- FindSubCluster(
  reference,
  cluster = "8",
  graph.name = "RNA_snn",
  subcluster.name = "unconventional",
  resolution = 0.25
)
DimPlot(reference, group.by = "unconventional", cols = color_palette)
reference$annotation_level_5 <- reference$unconventional
Idents(reference) <- "annotation_level_5"
```


# Project CITE-seq data

```{r}
query <- subset(cite_seq, idents = c("6", "7"))
reference$UMAP1 <- Embeddings(reference, "umap")[, "UMAP_1"]
reference$UMAP2 <- Embeddings(reference, "umap")[, "UMAP_2"]
reference_diet <- DietSeurat(reference)
query_diet <- DietSeurat(query)
reference_diet$type <- "reference"
query_diet$type <- "query"
query_diet$assay <- "CITEseq"
integrated <- merge(x = reference_diet, y = query_diet)
integrated <- NormalizeData(integrated)
rm(query_diet, reference_diet, naive_cd8, other_cd8)
hvg <- find_assay_specific_features(
  integrated,
  assay_var = "assay",
  n_features = 5000
)
integrated <- integrate_assays(
  integrated,
  assay_specific = TRUE,
  n_dim = 30,
  shared_hvg = hvg
)
integrated <- RunUMAP(integrated, reduction = "harmony", dims = 1:30)
DimPlot(integrated, split.by = "assay", group.by = "assay")
data_sets <- split_training_and_test_sets(
  integrated,
  split_var = "type",
  referece_label = "reference",
  query_label = "query",
  reduction = "harmony",
  n_dims = 30
)
annotation_data  <- transfer_label(
  seurat_obj = integrated,
  training_set = data_sets$training_set,
  test_set = data_sets$test_set,
  k = 15,
  response_var = "annotation_level_5"
)
umap_coords <- transfer_umap_coords(
  seurat_obj = integrated,
  training_set = data_sets$training_set,
  test_set = data_sets$test_set,
  k = 15,
  umap1_var = "UMAP1",
  umap2_var = "UMAP2"
)
integrated$UMAP1[umap_coords$query_cells] <- umap_coords$UMAP1
integrated$UMAP2[umap_coords$query_cells] <- umap_coords$UMAP2
integrated@reductions$umap@cell.embeddings[, "UMAP_1"] <- integrated$UMAP1
integrated@reductions$umap@cell.embeddings[, "UMAP_2"] <- integrated$UMAP2
integrated$annotation_level_5[annotation_data$query_cells] <- annotation_data$annotation
integrated$annotation_prob <- NA
integrated$annotation_prob[annotation_data$query_cells] <- annotation_data$annotation_prob
DimPlot(integrated, group.by = "annotation_level_5", split.by = "assay", cols = color_palette)
```

# Find markers

```{r}
# Split
rm(cite_seq)
integrated_rna_multi <- subset(integrated, assay != "CITEseq")
integrated_cite_seq <- subset(integrated, assay == "CITEseq")


# Markers RNA
Idents(integrated_rna_multi) <- "annotation_level_5"
markers_rna <- FindAllMarkers(
  integrated_rna_multi,
  logfc.threshold = 0.75,
  only.pos = TRUE
)
sorted_levels <- gtools::mixedsort(
  unique(
    integrated_rna_multi$annotation_level_5
  )
)
markers_rna_dfs <- purrr::map(sorted_levels, function(x) {
  df <- markers_rna[markers_rna$cluster == x, ]
  df
})
names(markers_rna_dfs) <- sorted_levels


# Markers protein
Idents(integrated_cite_seq) <- "annotation_level_5"
DefaultAssay(integrated_cite_seq) <- "ADT"
markers_prot <- FindAllMarkers(
  integrated_cite_seq,
  only.pos = TRUE
)
markers_prot_dfs <- purrr::map(sorted_levels, function(x) {
  df <- markers_prot[markers_prot$cluster == x, ]
  df
})
names(markers_prot_dfs) <- sorted_levels
```


# Unconventional T cells

```{r}
unconventional_levels <- c("9", "13", "8_0", "12")
unconventional <- subset(integrated_rna_multi, idents = unconventional_levels)
unconventional <- subset(unconventional, assay == "3P")
unconventional <- unconventional %>%
  FindVariableFeatures() %>%
  ScaleData() %>%
  RunPCA() %>%
  RunHarmony(reduction = "pca", group.by.vars = "donor_id", dims = 1:20) %>%
  RunUMAP(reduction = "harmony", dims = 1:20)
hvg2 <- find_assay_specific_features(unconventional,n_features = 5000)
unconventional <- integrate_assays(
  unconventional,
  assay_specific = TRUE,
  n_dim = 20,
  shared_hvg = hvg2
)
unconventional <- RunUMAP(unconventional, dims = 1:20, reduction = "harmony")
```


# Save

```{r}
# Save Seurat objects
saveRDS(integrated_rna_multi, "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/results/R_objects/level_5/Cytotoxic/paper/CD8_T_level_5_integrated_all_scRNA_multiome.rds")
saveRDS(integrated_cite_seq, "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/results/R_objects/level_5/Cytotoxic/paper/CD8_T_level_5_integrated_all_CITEseq.rds")
saveRDS(unconventional, "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/results/R_objects/level_5/Cytotoxic/paper/unconventional_CD8_T.rds")


# Save UMAPs
umap_rna <- DimPlot(integrated_rna_multi, cols = color_palette, pt.size = 0.75)
ggsave(
  filename = "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/3-clustering/5-level_5/Cytotoxic/tmp/umap_CD8_T_level_5_integrated_all_scRNA_multiome.png",
  plot = umap_rna,
  width = 14,
  height = 12,
  units = "cm"
)
umap_unconventional <- DimPlot(
  unconventional,
  cols = color_palette,
  pt.size = 1
)
ggsave(
  filename = "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/3-clustering/5-level_5/Cytotoxic/tmp/umap_unconventional_CD8_T.png",
  plot = umap_unconventional,
  width = 14,
  height = 12,
  units = "cm"
)


# Save markers
openxlsx::write.xlsx(x = markers_rna_dfs, file = "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/3-clustering/5-level_5/Cytotoxic/tmp/markers_rna_CD8_T_level_5_integrated_all_scRNA_multiome.xlsx")
openxlsx::write.xlsx(x = markers_prot_dfs, file = "~/Desktop/CNAG/mnt_clust/tonsil_atlas/current/scRNA-seq/3-clustering/5-level_5/Cytotoxic/tmp/markers_prot_CD8_T_level_5_integrated_all_scRNA_multiome.xlsx")
```


# Session Information

```{r}
sessionInfo()
```

