---
title: "Dimensionality reduction and integration: scRNA-seq"
author: "Ramon Massoni-Badosa"
date: "`r format(Sys.time(), '%B %Y')`"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```


# Introduction

Here, we will perform dimensionality reduction for scRNA-seq data of our MCL cases.


# Pre-processing

## Load packages

```{r}
library(Seurat)
library(Signac)
library(EnsDb.Hsapiens.v86)
library(ggpubr)
library(tidyverse)
set.seed(173)
```


## Define paths and parameters

```{r}
path_to_102 <- here::here("MCL/results/R_objects/4.seurat_merged_rna_unprocessed_atac_processed_102.rds")
path_to_413 <- here::here("MCL/results/R_objects/4.seurat_merged_rna_unprocessed_atac_processed_413.rds")
path_to_save_102 <- here::here("MCL/results/R_objects/5.seurat_merged_rna_processed_atac_processed_102.rds")
path_to_save_413 <- here::here("MCL/results/R_objects/5.seurat_merged_rna_processed_atac_processed_413.rds")
```


## Read data

```{r}
seurat_102 <- readRDS(path_to_102)
seurat_102[["ATAC"]] <- NULL
seurat_413 <- readRDS(path_to_413)
seurat_413[["ATAC"]] <- NULL
seurat_list <- list("102" = seurat_102, "413" = seurat_413)
rm(seurat_102, seurat_413)
```


# Normalization and linear dimensionality reduction

```{r}
seurat_list <- map(seurat_list, function(seurat_obj) {
  seurat_obj <- seurat_obj %>%
    NormalizeData() %>%
    FindVariableFeatures(nfeatures = 2500) %>%
    ScaleData() %>%
    RunPCA()
  seurat_obj
})
VizDimLoadings(seurat_list$`102`, dims = 1:4, nfeatures = 10)
VizDimLoadings(seurat_list$`102`, dims = 5:8, nfeatures = 10)
VizDimLoadings(seurat_list$`413`, dims = 1:4, nfeatures = 10)
VizDimLoadings(seurat_list$`413`, dims = 5:8, nfeatures = 10)
```


# Non-linear dimensionality reduction (UMAP)

```{r}
seurat_list <- map(
  seurat_list,
  RunUMAP,
  reduction = "pca",
  dims = 1:30,
  reduction.key = "rnaUMAP_",
  reduction.name = "rnaUMAP"
)
map(
  seurat_list,
  DimPlot,
  reduction = "rnaUMAP",
  group.by = "library_name",
  split.by = "library_name"
)
```


Of note, we see a good mixability across samples, so we will no perform any batch effect correction method.


# Add RNA assay and UMAP reduction

## 102

```{r}
seurat_102 <- readRDS(path_to_102)
DefaultAssay(seurat_102) <- "ATAC"
seurat_102[["RNA"]] <- NULL
seurat_102[["RNA"]] <- seurat_list$`102`[["RNA"]]
seurat_102@reductions$pca <- seurat_list$`102`@reductions$pca
seurat_102@reductions$rnaUMAP <- seurat_list$`102`@reductions$rnaUMAP
saveRDS(seurat_102, path_to_save_102)
rm(seurat_102)
```


## 413

```{r}
seurat_413 <- readRDS(path_to_413)
DefaultAssay(seurat_413) <- "ATAC"
seurat_413[["RNA"]] <- NULL
seurat_413[["RNA"]] <- seurat_list$`413`[["RNA"]]
seurat_413@reductions$pca <- seurat_list$`413`@reductions$pca
seurat_413@reductions$rnaUMAP <- seurat_list$`413`@reductions$rnaUMAP
saveRDS(seurat_413, path_to_save_413)
rm(seurat_413)
```


# Session Information

```{r}
sessionInfo()
```

