---
title: "Plot Out-of-Bag Accuracy"
author: "Ramon Massoni-Badosa and Paula Soler-Vila"
date: "2021/03/09"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```


# Introduction

To define the correct number of clusters, we followed these steps:

1. Clustered cells at varying resolutions with [Seurat's](https://satijalab.org/seurat/) `FindClusters` function. Different resolutions yield different number of clusters.
2. For each resolution, we down-sampled each cluster to 3000 cells to correct for imbalances in number of cells across clusters. If a cluster had <3000, we sampled all of them.
3. We fit a [random forest](https://en.wikipedia.org/wiki/Random_forest) classifier and obtained the out-of-bag (OOB) accuracy (1 - [OOB error](https://en.wikipedia.org/wiki/Out-of-bag_error)).

We reasoned that stable clusters that represent real cell types should serve as a reference to classify unseen data with high accuracy. Thus, we choose the right resolution right before the accuracy drops markedly.

**TODO** explain t-statistic after we write the methods for it


# Pre-processing

## Load packages

```{r}
library(data.table)
library(tidyverse)
```


## Define parameters

```{r}
# Paths
path_to_oob <- here::here("scRNA-seq/3-clustering/1-level_1/tmp/oob")
path_to_utils <- here::here("scRNA-seq/bin/utils.R")


# Optimal resolutions
optimal_resolution <- 0.25
num_opt_resolutions <- 3
```


## Source functions

```{r}
source(path_to_utils)
```


## Load data

```{r}
oob_dfs_paths <- list.files(path_to_oob)
oob_dfs_paths <- str_c(path_to_oob, oob_dfs_paths, sep = "/")
oob_dfs <- purrr::map(oob_dfs_paths, readRDS)
```


# Detect the optimal resolution

```{r}
# Find the optimal resolution
oob_df <- bind_rows(oob_dfs)
oob_df$resolution <- as.numeric(as.character(oob_df$resolution))
df_optimal <- find_optimal_resolution(oob_df)
df_optimal_melt <- melt(df_optimal, id.vars = "resolution")
```


# Plot 

```{r fig.height=7}
axis_breaks <- c(90, 92.5, 95, 97.5, 100)
optimal_resolutions <- df_optimal[order(-df_optimal$optimal_values), ]$res[1:num_opt_resolutions]
levels(df_optimal_melt$variable) <- c("Out-of-Bag Accuracy (%)", "Optimal Values")
oob_gg <- df_optimal_melt %>%
  ggplot(aes(resolution, value, color = variable)) +
    geom_vline(
      xintercept = optimal_resolutions,
      colour = rep(c("black", "gray40", "gray60"), 2),
      linetype = "dashed", size = 0.5
    ) +
    geom_point() +
    geom_line(
      data = df_optimal_melt[!is.na(df_optimal_melt$value), ],
      aes(color = variable)
    ) +
    labs(x = "Clustering Resolution", y = "") +
    theme_bw() +
    theme(
      axis.title = element_text(size = 13),
      legend.position = "none",
      strip.text = element_text(size = 12)
    ) +
    facet_wrap(~variable, scales = "free_y", ncol = 1)
oob_gg
```


Based on both profiles, we will use a resolution of `r optimal_resolution`.


# Session Information

```{r}
sessionInfo()
```

