---
author: "Ramon Massoni-Badosa"
date: "2021/03/15"
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
title <- paste("Find markers (level 2):", cell_type, sep = " ")
```

---
title: "`r title`"
---

# Introduction

The objective of this notebook is to find markers for those clusters that we flagged as potential doublets. Later, we will discuss with the annotation team whether we keep them or remove.


# Pre-processing

## Load packages

```{r}
library(Seurat)
library(tidyverse)
```


## Parameters

```{r}
# Paths
path_to_dir <- here::here("scRNA-seq/results/R_objects/level_2/")
path_to_obj <- str_c(
  path_to_dir,
  cell_type,
  "/",
  cell_type,
  "_integrated_level_2.rds",
  sep = ""
)
path_to_save <- str_c(
  path_to_dir,
  cell_type,
  "/",
  cell_type,
  "_clustered_level_2.rds",
  sep = ""
)
path_to_tmp <- here::here("scRNA-seq/3-clustering/2-level_2/tmp/")
path_to_save_markers <- str_c(
  path_to_tmp,
  cell_type,
  "_markers_level_2.xlsx",
  sep = ""
)


# Functions
source(here::here("scRNA-seq/bin/utils.R"))


# Colors
color_palette <- c("black", "gray", "red", "yellow", "violet", "green4",
                   "blue", "chocolate1", "coral2", "blueviolet",
                   "brown1", "darkmagenta", "deepskyblue1", "dimgray",
                   "deeppink1", "green", "lightgray", "hotpink1",
                   "indianred4", "khaki", "mediumorchid2")


# Resolutions
resolutions <- c(
  NBC_MBC = 0.2,
  GCBC = 0.1,
  CD4_T = 0.05,
  Cytotoxic = 0.1,
  PC = 0.1,
  myeloid = 0.05,
  FDC = 0.2,
  PDC = 0.1,
  epithelial = 0.1
)
resolution <- resolutions[cell_type]
names(resolution) <- NULL


# Flagged clusters
flagged_clusters_all <- list(
  NBC_MBC = c("5"),
  GCBC = c("2", "3", "4", "5"),
  CD4_T = c("2", "4", "5", "6"),
  Cytotoxic = c("2", "5"),
  PC = c("2", "3", "4", "5"),
  myeloid = c("2", "3", "4", "6", "7", "8", "9", "10"),
  FDC = c("1", "2", "3", "5", "6", "7"),
  PDC = c("1"),
  epithelial = c("2", "3", "4", "5", "6", "7")
)
flagged_clusters <- flagged_clusters_all[[cell_type]]
```


## Load data

```{r}
# Seurat object
seurat <- readRDS(path_to_obj)
seurat
```


# Cluster

Here, we cluster with the resolution that allows us to separate our clusters of interest, which was decided by inspecting the results of previous notebooks.

```{r}
seurat <- FindClusters(seurat, resolution = resolution)
p <- DimPlot(seurat, cols = color_palette)
p
```


# Find Markers

```{r}
markers_flagged_clust <- purrr::map(flagged_clusters, function(x) {
  df <- FindMarkers(
    seurat,
    ident.1 = x,
    logfc.threshold = 0.75,
    only.pos = TRUE
  )
  fold_change_var <- str_subset(colnames(df), "FC")
  sorted_rows <- order(df[[fold_change_var]], decreasing = TRUE)
  df <- df[sorted_rows, ]
  df$gene <- rownames(df)
  df
})
names(markers_flagged_clust) <- flagged_clusters
purrr::map(markers_flagged_clust, head, 25)
```


# Save

```{r}
saveRDS(seurat, path_to_save)
dir.create(path_to_tmp, showWarnings = FALSE)
openxlsx::write.xlsx(markers_flagged_clust, file = path_to_save_markers)
```


# Session Information

```{r}
sessionInfo()
```